### 2022-11-08 scripts/MSc.Rmd
commit scripts

---
title: "`r params$doc_title`"
author: | 
        | christoph.reich@med.uni-heidelberg.de 
        | lorenzo@imbi.uni-heidelberg.de
        | 
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
params:
  date: !r Sys.Date()
  doc_title: "MSc Default Title"
output: 
  html_document:
    theme: paper
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 4
    df_print: kable
    latex_engine: xelatex
#bibliography: ../references/project01.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(scales)
library(ggthemes)
#library(biomaRt)  # different script
library(kableExtra)
library(MendelianRandomization)
library(mr.raps)
library(MRMix)
library(MRPRESSO)
library(penalized)
library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

```{r}
my_data <- as_tibble(read.delim("../data-boekstegers_felix/GST_BMI_association_results.txt"))
# GBC: new data 2022-11-04
GBC_dat <- as_tibble(read.delim("../data-boekstegers_felix/C23C24_BMI_association_results.txt"))

GBC_dat <- GBC_dat %>% 
  rename(eaf_gbc=eaf_out,
         beta_gbc=beta_out,
         se_gbc=se_out,
         p_value_gbc=p_value_out) %>% 
  select(-c(chr, pos_hg19, pos_hg38, a1, a2, exposure, outcome))
my_data <- my_data %>% left_join(GBC_dat, by =c("SNP", "dummyID"))
```

# Get gene names from rs-ids

```{r}
# see script biomart_search.R
rsIds_GeneNames <- as_tibble(readRDS("../data/biomartBMIsearch.rds"))

filter_FTO <- rsIds_GeneNames %>% 
  filter(str_detect(associated_gene, pattern = "FTO")) %>% 
  distinct(refsnp_id, .keep_all = TRUE)
```

# Simulating on empirical data

## Calculate approximate SEs for decreasing sample sizes and corresponding p-vals

Calculating p-values from a normal distribution. The se, pval and study size depend on each other, within each study. The selection of IVs should rely on the exposure study - BMI in our case. Summary statistics on genetic associations with BMI will be based on results of the GIANT consortium + GERA (n=334487 (100,418 + 234,069) individuals and 289 associated variants with p < 10-8) **(insert citation)**. We vary the sample size and calculate the number of IVs and the explained variance. *(Justo Mail 2022-09-27)*
\
Steps:
1.       Calculate more accurates SEs using the betas and Pvals provided to you (this will be the original/initial SE for n=334.487). 
2.       Calculate the approximate SEs for decreasing samples sizes
3.       Calculate the probability values considering the estimated betas and the corresponding SEs
4.       Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8) **NEW 5x10^(-6))**
5.       Calculate cumulative explained variances
6.       Represent figure similar to Fig1 in Chatterjee’s paper

For the selected IVs, please also calculate the explained variance for the confounder and outcome.

```{r z-statistic}
### justo approach email 30-10-2022  ---------------------------------------

## LOOP simulation ----------------------------------------------
n.orig <- 334487
n <- c(1000, 2000, 5000, seq(10000, 330000, by=10000), 334487)

# create vectors to store pvals_n in
colnames_pval <- c()
for (i in 1:length(n)) {
  colnames_pval[i] <- paste0("p_value_exp_sample_", format(n[i], scientific = FALSE))
}
## initialize
se_update <- c()
pvals_df <- data.frame(matrix(ncol = length(colnames_pval), nrow =  nrow(my_data)))
colnames(pvals_df) <- colnames_pval
for (i in 1:nrow(my_data)) {
  # 1) Calculate more accurates SEs using the provided betas and Pvals 
  se_update[i] <- my_data[["beta_exp"]][i]/-qnorm(p=(my_data[["p_value_exp"]][i]/2),mean=0)

  # 2) Calculate the approximate SEs for decreasing samples sizes
  ### approximation of sd (single value)
  sd <- se_update[i] * sqrt(n.orig)
  #### approximation of standard error simulating different sample size
  se_n <- sd/(sqrt(n))

  # 3) Calculate the probability values considering the estimated betas and the corresponding SEs
  for (j in 1:length(n)) {
    pvals_n <- 2*(1-pnorm(abs(my_data[["beta_exp"]][i])/abs(se_n[j]), mean=0, 
                          lower.tail = TRUE))
    pvals_df[i, j] <- pvals_n
  }
}

# 4. Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8)
pvals_df %>% 
  summarize(across(.cols = everything(), function(x) sum(x < 5*10^(-6)))) %>% 
  t() %>% 
  kable(caption = "Count of significant SNPs depending on sample size")

# pvals_df %>%
#   filter(p_value_exp_sample_100000< 5*10^(-6)) %>% 
#   count()

my_data <- as_tibble(cbind(my_data, pvals_df))

# 5. Calculate cumulative explained variances
# 6. Represent figure similar to Fig1 in Chatterjee’s paper
```

```{r calculate-explained-variance}
explained_variance_numeric <- function(eaf, beta) {
  # eaf = allele frequency
  # beta = additive genetic effect
  explained_variance_numeric <- 2*eaf *(1-eaf) * beta^2
  return(explained_variance_numeric)
}

explained_variance_binary <- function(PA,RR1,RR2,K) {
  # from So et al 2011
  # Calculates the variance in liability explained by a single biallelic loci.
  # PA: allele frequency of the risk allele (denoted A)
  # RR1: relative risk of Aa (one risk allele) compared to aa (no risk allele)
  # RR2: relative risk of AA (two risk alleles) compared to aa (no risk allele)
  # K:  overall probability of disease in population
  # Returns the variance explained (Vg) and the mean liability for each genotype (the overall liability is normalized to mean 0 and variance 1)
  Paa = (1-PA)^2
  PAa = 2*PA*(1-PA)
  PAA = PA^2
  muaa=0
  faa= K/(Paa + PAa*RR1 + PAA*RR2)
  fAa= RR1*faa
  fAA= RR2*faa 
  T = qnorm(1-faa) 
  muAa = T-qnorm(1-fAa)
  muAA = T-qnorm(1-fAA)
  mean.all= PAa*muAa+ PAA*muAA
  Vg= Paa*(muaa-mean.all)^2 + PAa*(muAa-mean.all)^2+ PAA*(muAA-mean.all)^2
  actual.Vg =  Vg/(1+Vg) 
  VR = 1-actual.Vg 
  actual.T = Paa*sqrt(VR)*qnorm(1-faa) + PAa*sqrt(VR)*qnorm(1-fAa) + PAA*sqrt(VR)*qnorm(1-fAA)
  actual.muaa = actual.T - sqrt(VR) * qnorm(1-faa)
  actual.muAa = actual.T - sqrt(VR) * qnorm(1-fAa)
  actual.muAA = actual.T - sqrt(VR) * qnorm(1-fAA)

  res <- list(Vg=actual.Vg,muaa=actual.muaa, muAa = actual.muAa, muAA=actual.muAA)
  res
}

## calculate explained variance
my_data <- my_data %>% 
  mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf_exp, beta = beta_exp), 
         explained_variance_GSD = explained_variance_binary(PA = eaf_out, 
                                                            RR1 = exp(beta_out), 
                                                            RR2 = exp(beta_out)^2, 
                                                            K = 0.1  # assumed prevalence in population
                                                            )$Vg,
         explained_variance_GBC = explained_variance_binary(PA = eaf_gbc, 
                                                            RR1 = exp(beta_gbc), 
                                                            RR2 = exp(beta_gbc)^2, 
                                                            K = 1.2/100000  # assumed prevalence in population
                                                            )$Vg
         )

## paper backing prevalence GSD
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3343155/#:~:text=Gallstones%20are%20common.,develop%20stones%20at%20some%20time.&text=The%20majority%20will%20not%20develop,cholecystitis%2C%20cholangitis%2C%20or%20pancreatitis
## paper backing prevalence GSD (ASR/100.000 = 1.2 worldwide)
# https://www.wcrf.org/cancer-trends/gallbladder-cancer-statistics/

#explained_variance_binary(PA=0.3,RR1=1.3,RR2=1.3^2,K=0.01)
```

## Plot explained variance in BMI vs -log10 P-value from Hoffmann et al.

```{r}
# labeling (only FTO label)
my_data$gene_label <- NA
my_data$gene_label[my_data$SNP %in% filter_FTO$refsnp_id] <- "FTO"

# color labeling
my_data$fto_col_annot <- "No FTO variant"
my_data$fto_col_annot[my_data$SNP %in% filter_FTO$refsnp_id] <- "FTO variant"

ggplot(data = my_data, 
       mapping = aes(x=explained_variance_BMI*100, y= -log10(p_value_exp),  
                    col=fto_col_annot, label=gene_label)
       )+
  geom_point(alpha=0.6)+
  ggrepel::geom_text_repel(show.legend = FALSE)+
  xlab("Explained variance in BMI in %")+
  ylab("-log10 P-val")+
  labs(title = "Explained variance exposure",
       col='SNP'
       ) +
  theme_few()+
  scale_colour_few() -> fig00
fig00

ggsave(filename = "../output/plots/figure00.png", plot = fig00, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```

## Plot Figure 1 from Qi&Chatterjee 2021

```{r get-data-for-fig1}
# 4.       Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8) **NEW 5x10^(-6))**
# 5.       Calculate cumulative explained variances
# initialize df
df_cum_expl_var <- tibble(n=n, 
                          cum_expl_varX = rep(NA, length(n)), 
                          cum_expl_varU = rep(NA, length(n)), 
                          cum_expl_varY = rep(NA, length(n)),
                          count_IVs = rep(NA, length(n))
)

for (i in 1:length(colnames_pval) ) {
  threshold <- 5*10^(-6)
  select_col <- colnames_pval[i]
  # calculate cum explained variance and count of IVs
  summarydat <- my_data %>% 
    filter(!!sym(select_col) < threshold) %>% 
    select(explained_variance_BMI, explained_variance_GSD, explained_variance_GBC) %>% 
    summarise(expl_variances =across(.cols = everything(), function(x) sum(x)),
              count_IVs = n())
  # fill up df
  df_cum_expl_var[i, 2] <- summarydat$expl_variances[1]
  df_cum_expl_var[i, 3] <- summarydat$expl_variances[2]
  df_cum_expl_var[i, 4] <- summarydat$expl_variances[3]
  df_cum_expl_var[i, 5] <- summarydat$count_IVs
}
```

```{r plot-fig1}
# 6.       Represent figure similar to Fig1 in Chatterjee’s paper

# Dual Y axis: we have 284 IVs, scale them to 5% (y-axis [0;5%]): 3/284 
## https://r-graph-gallery.com/line-chart-dual-Y-axis-ggplot2.html
scale_factor <- 4.5/284
df_cum_expl_var <- df_cum_expl_var %>% 
  mutate(
    cum_expl_varX = cum_expl_varX *100, # in percent 
    cum_expl_varU = cum_expl_varU *100,
    cum_expl_varY = cum_expl_varY *100,
    count_IVs_scaled = count_IVs*scale_factor)

plot_data <- df_cum_expl_var %>% 
  pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, count_IVs_scaled)) %>% 
  mutate(name = factor(name, labels = c("Number of IVs", "% var(U)", "% var(X)")))

ggplot(plot_data, aes(x=log10(n), y=value, col=name))+
  geom_point(alpha=0.6)+
  geom_line(alpha=0.6)+
  xlab("log10(N)")+
  #ylab("% variance explained by IVs")+
  labs(title = "Figure 1",
       col='', 
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  scale_y_continuous(
    # Features of the first axis
    name = "% variance explained by IVs",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
  ) + 
  theme_few()+
  scale_colour_few() ->fig01
fig01

ggsave(filename = "../output/plots/figure01.png", plot = fig01, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```

```{r updated-fig01}
plot_data1 <- df_cum_expl_var %>% 
  pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, cum_expl_varY)) %>% 
  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")))

N_annotation <- c(10000, 100000, 200000, 300000)
## plot
ggplot(plot_data1, aes(x=n, y=value, col=name))+
  geom_point(alpha=0.6)+
  geom_line(alpha=0.6)+
  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
  scale_x_continuous(name = "Sample Size",
                     labels = comma, 
                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
                     limits = c(10000, 350000))+
  #ylab("% variance explained by IVs")+
  labs(title = "Figure 1",
       col='', 
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  scale_y_continuous(
    # Features of the first axis
    name = "% variance explained by IVs",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
  ) + 
  theme_few()+
  #scale_colour_few() +
  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
  #                                         "% var(X)" = "blue", 
  #                                         "% var(Y)" = "darkblue", 
  #                                         "Number of IVs" = "black"
  #                                         ))
  #
  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                           "Number of IVs" = "black"
                                           )) -> fig01.update
fig01.update
# scale_color_manual(name = "Y series", values = c("Y1" = "darkblue", "Y2" = "red"))
ggsave(filename = "../output/plots/figure01update.png", plot = fig01.update, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```

# Different genetic architectures of the exposure

## MAF vs P-val

For the ~200 IVs, please plot the minor allele frequency (x-axis) versus -log10Pval (y-axis). Calculate the median MAF and the median –log10Pval, and draw the corresponding vertical and horizontal lines. The four areas will define four genetic architectures for the simulations (e.g. relatively rare variants with relatively large exposure effects).

```{r MAF-Pval}
ggplot(data = my_data, aes(x=eaf_exp, y=-log10(p_value_exp)))+
  geom_point(alpha=0.6)+
  geom_hline(yintercept=-log10(median(my_data$p_value_exp)), col="red", alpha=0.6)+
  geom_vline(xintercept=median(my_data$eaf_exp), col="red", alpha=0.6)+
  xlab("Effect allele frequency for exposure")+
  ylab("-log10(P-value) ")+
  labs(title = "Figure 2",
       subtitle = "Define different genetic architectures",
       #col='', 
       #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  theme_few()+
  scale_colour_few() ->fig02
fig02

ggsave(filename = "../output/plots/figure02.png", plot = fig02)
```

## MAF vs betas

```{r mafVSbetas}
plot_data_2 <- my_data %>% 
  mutate(maf_exp = ifelse(eaf_exp >0.5, 1-eaf_exp, eaf_exp)) %>% 
  select(maf_exp, beta_exp)

ggplot(data = plot_data_2, aes(x=maf_exp, y=abs(beta_exp)))+
  geom_point(alpha=0.6)+
  geom_hline(yintercept=median(abs(plot_data_2$beta_exp)), col="red", alpha=0.6)+
  geom_vline(xintercept=median(plot_data_2$maf_exp), col="red", alpha=0.6)+
  xlab("Minor allele frequency for exposure")+
  ylab("abs(beta_BMI)")+
  labs(title = "Figure 2-update",
       subtitle = "Define different genetic architectures",
       #col='', 
       #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  theme_few()+
  scale_colour_few() ->fig02.update
fig02.update
ggsave(filename = "../output/plots/figure02update.png", plot = fig02.update)
```

# Simulation Framework

- Let’s try to estimate parameters for the simulation (pi1, varx, pi2, varu, muy, vary, pi3, pi4, delta0) based on real data. Maybe you could start with N=10.000 and N=300.000 (I will help with this)

### 2022-11-08 scripts/biomart_search.R

@@ -1,0 +1,69 @@
+
+
+# get gene names from rs SNP ids
+
+library(biomaRt)
+# https://bioconductor.org/packages/release/bioc/vignettes/biomaRt/inst/doc/accessing_ensembl.html
+# https://bio340files.s3.amazonaws.com/SearchBiomart.html  https://www.youtube.com/watch?v=bMkve2uBMZo&ab_channel=profbiot
+
+# list databases (look at pulldown menu on website: choose database)
+myMarts <- listMarts()
+
+# biomart                version
+# 1 ENSEMBL_MART_ENSEMBL      Ensembl Genes 108
+# 2   ENSEMBL_MART_MOUSE      Mouse strains 108
+# 3     ENSEMBL_MART_SNP  Ensembl Variation 108
+# 4 ENSEMBL_MART_FUNCGEN Ensembl Regulation 108
+
+listEnsembl()   # the same since listMarts connects to Ensembl
+# biomart                version
+# 1         genes      Ensembl Genes 108
+# 2 mouse_strains      Mouse strains 108
+# 3          snps  Ensembl Variation 108
+# 4    regulation Ensembl Regulation 108
+
+# ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")
+
+# select database
+ensembl <- useMart(biomart = "ENSEMBL_MART_SNP")
+myDatasets <- listDatasets(ensembl)
+
+# select dataset: hsapiens
+ensembl <- useMart(biomart = "ENSEMBL_MART_SNP", 
+                   dataset = "hsapiens_snp") #  defaults: host = "https://www.ensembl.org"; path="/biomart/martservice"
+                   
+# List filters
+myFilters <- listFilters(ensembl)
+
+# make a vector for your filters (here all rs_ids)
+filter1 <- c('snp_filter')
+
+# make a vector for the values that will go into your filter
+values1 <- my_data$SNP
+
+# list attributes (columns we download from biomart)
+myAttributes <- listAttributes(ensembl)
+# make a variable for your attributes
+att1 <- c('refsnp_id', 'minor_allele', 'minor_allele_freq', 
+          'phenotype_name', 
+          'phenotype_description'  # several descriptions per SNP possible!
+          )
+
+att2 <- c('refsnp_id', 'minor_allele', 'minor_allele_freq', 
+          'clinical_significance', 'associated_gene', 
+          'phenotype_name', 
+          'ensembl_gene_stable_id', 'ensembl_gene_name'
+          )
+
[…]

### 2022-11-08 presentation/robustMR.qmd
revealjs presentation 2022-11-08

--- 2022-11-08 scripts/MSc.Rmd
@@
 ---
-title: "`r params$doc_title`"
-author: | 
-        | christoph.reich@med.uni-heidelberg.de 
-        | lorenzo@imbi.uni-heidelberg.de
-        | 
-date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
-params:
-  date: !r Sys.Date()
-  doc_title: "MSc Default Title"
-output: 
-  html_document:
-    theme: paper
-    toc: true
-    number_sections: true
-    toc_float: true
-    toc_depth: 4
-    df_print: kable
-    latex_engine: xelatex
-#bibliography: ../references/project01.bib
+title: "Assessment of simulation procedures and robust methods for Mendelian randomization analysis"
+author: "Christoph Reich"
+date: "`r Sys.Date()`"
+title-slide-attributes: 
+  data-background-image: ../templates/bg1.png
+  data-background-size: cover  # default
+  data-background-opacity: "1"  # from 0-1
+format: 
+  revealjs:
+    slide-number: true
+    scrollable: true
+    navigation-mode: vertical
+    chalkboard:   # kreidetafel
+      theme: whiteboard
+      boardmarker-width: 5
+      buttons: false
+    incremental: false  # display bullet lists all at once
+    preview-links: auto
+    logo: ../templates/mederlab_logo.png
+    css: ../templates/main.css
+    footer: "robustMRmethods"
+    theme: [simple, ../templates/custom.scss]
+    #self-contained: true  # for distribution (slows rendering), not compatible with chulkboarddic
+# reference-doc: templates/20210129_mederslidemaster.pptx  # only for pptx rendering
+bibliography: ../references/references.bib
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r}
+#| label: dependencies
+#| include: false
+#| warning: false
+
+rm(list=ls())
+
+knitr::opts_chunk$set(echo = FALSE)
 
@@
 conflict_prefer("filter", "dplyr")
+
+theme_set(ggthemes::theme_few()) 
 ```
@@
 ```{r}
+#| label: load-data
+#| include: false
+#| warning: false
+
 my_data <- as_tibble(read.delim("../data-boekstegers_felix/GST_BMI_association_results.txt"))
@@
 
-# Get gene names from rs-ids
+## Get gene names from rs-ids
 
 ```{r}
+#| label: FTO-annotation
+#| tbl-cap: "SNP associated with FTO gene"
+#| tbl-colwidths: [60,40]
+
 # see script biomart_search.R
@@
   distinct(refsnp_id, .keep_all = TRUE)
+
+t(filter_FTO) %>% kable() %>% 
+  kable_styling(font_size=18)
 ```
 
-# Simulating on empirical data
+:::: {.columns}
 
-## Calculate approximate SEs for decreasing sample sizes and corresponding p-vals
+::: {.column width="50%"}
+* from the 284 SNPs only one was associated with the FTO gene
+:::
 
-Calculating p-values from a normal distribution.
-The se, pval and study size depend on each other, within each study.
-The selection of IVs should rely on the exposure study - BMI in our case.
-Summary statistics on genetic associations with BMI will be based on results of the GIANT consortium + GERA (n=334487 (100,418 + 234,069) individuals and 289 associated variants with p < 10-8) **(insert citation)**.
-We vary the sample size and calculate the number of IVs and the explained variance.
-*(Justo Mail 2022-09-27)*
-\
-Steps:
-1.       Calculate more accurates SEs using the betas and Pvals provided to you (this will be the original/initial SE for n=334.487). 
+::: {.column width="50%"}
+![](../img/ensembl_logo.png)
+
+:::
+
+::::
+
+## Simulating on empirical data
+
+### Calculate approximate SEs for decreasing sample sizes and corresponding p-vals
+
+* se, pval and study size depend on each other, within each study
+* selection of IVs based on exposure study (BMI in our case)
+* Summary statistics on genetic associations with BMI based on results of the GIANT + GERA studies [@Hoffmann_2018] with $N=100,418 + 234,069 =334487$  and $SNPs=289$
+
+## Steps
+
+1. Calculate more accurates SEs using the betas and Pvals
 2. Calculate the approximate SEs for decreasing samples sizes
@@
 5. Calculate cumulative explained variances
-6.       Represent figure similar to Fig1 in Chatterjee’s paper
+6. Represent figure similar to Fig1 in Chatterjee’s paper @QiChatterjee_2021
 
-For the selected IVs, please also calculate the explained variance for the confounder and outcome.
+```{r}
+#| label: z-statistic
 
-```{r z-statistic}
 ### justo approach email 30-10-2022  ---------------------------------------
-
 ## LOOP simulation ----------------------------------------------
@@
 # 4. Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8)
+my_data <- as_tibble(cbind(my_data, pvals_df))
+
 pvals_df %>% 
@@
   t() %>% 
-  kable(caption = "Count of significant SNPs depending on sample size")
-
-# pvals_df %>%
-#   filter(p_value_exp_sample_100000< 5*10^(-6)) %>% 
-#   count()
-
-my_data <- as_tibble(cbind(my_data, pvals_df))
-
-# 5. Calculate cumulative explained variances
-# 6. Represent figure similar to Fig1 in Chatterjee’s paper
+  kable(caption = "Count of significant SNPs depending on sample size") %>% 
+  kable_styling(font_size=18)
 ```
 
-```{r calculate-explained-variance}
+## Explained variance
+
+* The variance in BMI explained by the IVs is calculated with the following formula: $explained\ variance = 2f(1-f)\beta^2$, where $f$ denotes the allele frequency and $\beta$ is the additive genetic effect.
+* The explained variance in liability to GSD and GBC is calculated as described by @So_2011.
+
+```{r}
+#| label: calculate-explained-variance
+
 explained_variance_numeric <- function(eaf, beta) {
@@
 ```{r}
-# labeling (only FTO label)
+#| label: explainedBMIvarianceVSpval
+#|
+
 my_data$gene_label <- NA
@@
 
+Here we show a plot similar to Figure 1 in [@QiChatterjee_2021]:
+
 ```{r get-data-for-fig1}
@@
 
-```{r plot-fig1}
+```{r}
+#| label: figure1QiChatterjee
 # 6.       Represent figure similar to Fig1 in Chatterjee’s paper
@@
   scale_colour_few() ->fig01
-fig01
 
@@
 
-# Different genetic architectures of the exposure
-
-## MAF vs P-val
-
-For the ~200 IVs, please plot the minor allele frequency (x-axis) versus -log10Pval (y-axis).
-Calculate the median MAF and the median –log10Pval, and draw the corresponding vertical and horizontal lines.
-The four areas will define four genetic architectures for the simulations (e.g.
-relatively rare variants with relatively large exposure effects).
+## Different genetic architectures of the exposure
 
@@
   scale_colour_few() ->fig02
-fig02
+#fig02
 
@@
 
-## MAF vs betas
+### MAF vs betas
 
@@
 
-# Simulation Framework
+## Simulation Framework
 
-- Let’s try to estimate parameters for the simulation (pi1, varx, pi2, varu, muy, vary, pi3, pi4, delta0) based on real data. Maybe you could start with N=10.000 and N=300.000 (I will help with this)
+* Estimate parameters for the simulation (pi1, varx, pi2, varu, muy, vary, pi3, pi4, delta0) based on real data
+* Start with $N=10.000$ and $N=300.000$
+
+![](../img/m4qi2021.png)
+
+##
+
+**References**
+
+::: {#refs}
+:::
+
+**Software**
+
+This analysis was carried out using the statistical software `r base::version$version.string` along with [RStudio](https://www.rstudio.com).
+Apart from R's base functionality, the following packages were used: <br/>
+
+```{r software, echo=FALSE}
+# or faster with function...
+installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
+```
+
+##  {background-image="../templates/bg4.png"}
+
+::: footer
+
+:::

### 2022-11-10 presentation/robustMR.qmd
added references

@@
     navigation-mode: vertical
-    chalkboard:   # kreidetafel
-      theme: whiteboard
-      boardmarker-width: 5
-      buttons: false
+    #chalkboard:   # kreidetafel
+    #  theme: whiteboard
+    #  boardmarker-width: 5
+    #  buttons: false
     incremental: false  # display bullet lists all at once
@@
     theme: [simple, ../templates/custom.scss]
-    #self-contained: true  # for distribution (slows rendering), not compatible with chulkboarddic
+    self-contained: true  # for distribution (slows rendering), not compatible with chulkboarddic
 # reference-doc: templates/20210129_mederslidemaster.pptx  # only for pptx rendering
@@
 * The variance in BMI explained by the IVs is calculated with the following formula: $explained\ variance = 2f(1-f)\beta^2$, where $f$ denotes the allele frequency and $\beta$ is the additive genetic effect.
-* The explained variance in liability to GSD and GBC is calculated as described by @So_2011.
+* The explained variance in liability to GSD and GBC is calculated as described by @So_2011
+  + Prevalence of disease $K$ must be set:
+      - [Prevalence GSD](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3343155/#:~:text=Gallstones%20are%20common.,develop%20stones%20at%20some%20time.&text=The%20majority%20will%20not%20develop,cholecystitis,%20cholangitis,%20or%20pancreatitis)
+      - [Prevalence GBC](https://www.wcrf.org/cancer-trends/gallbladder-cancer-statistics/)
 
@@
 
+---
 ##  {background-image="../templates/bg4.png"}
@@
 :::
+---

### 2022-11-10 presentation/robustMR.qmd
update presentation

@@
 
-:::: {.columns}
-
+::: columns
 ::: {.column width="50%"}
-* from the 284 SNPs only one was associated with the FTO gene
+-   from the 284 SNPs only one was associated with the FTO gene
 :::
@@
 ![](../img/ensembl_logo.png)
-
 :::
-
-::::
+:::
 
@@
 
-* se, pval and study size depend on each other, within each study
-* selection of IVs based on exposure study (BMI in our case)
-* Summary statistics on genetic associations with BMI based on results of the GIANT + GERA studies [@Hoffmann_2018] with $N=100,418 + 234,069 =334487$  and $SNPs=289$
+-   se, pval and study size depend on each other, within each study
+-   selection of IVs based on exposure study (BMI in our case)
+-   Summary statistics on genetic associations with BMI based on results of the GIANT + GERA studies [@Hoffmann_2018] with $N=100,418 + 234,069 =334487$ and $SNPs=289$
 
@@
 3.  Calculate the probability values considering the estimated betas and the corresponding SEs
-4. Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8) **NEW 5x10^(-6))**
+4.  Exclude the IV if Pval \> genome-wide significance threshold (5 × 10−8) **NEW 5x10\^(-6))**
 5.  Calculate cumulative explained variances
-6. Represent figure similar to Fig1 in Chatterjee’s paper @QiChatterjee_2021
+6.  Represent figure similar to Fig1 in Chatterjee's paper @QiChatterjee_2021
 
@@
 
-* The variance in BMI explained by the IVs is calculated with the following formula: $explained\ variance = 2f(1-f)\beta^2$, where $f$ denotes the allele frequency and $\beta$ is the additive genetic effect.
-* The explained variance in liability to GSD and GBC is calculated as described by @So_2011
-  + Prevalence of disease $K$ must be set:
+-   The variance in BMI explained by the IVs is calculated with the following formula: $explained\ variance = 2f(1-f)\beta^2$, where $f$ denotes the allele frequency and $\beta$ is the additive genetic effect.
+-   The explained variance in liability to GSD and GBC is calculated as described by @So_2011
+    -   Prevalence of disease $K$ must be set:
         -   [Prevalence GSD](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3343155/#:~:text=Gallstones%20are%20common.,develop%20stones%20at%20some%20time.&text=The%20majority%20will%20not%20develop,cholecystitis,%20cholangitis,%20or%20pancreatitis)
@@
 
-* Estimate parameters for the simulation (pi1, varx, pi2, varu, muy, vary, pi3, pi4, delta0) based on real data
-* Start with $N=10.000$ and $N=300.000$
+-   Estimate parameters for the simulation (pi1, varx, pi2, varu, muy, vary, pi3, pi4, delta0) based on real data
+-   Start with $N=10.000$ and $N=300.000$
 
 ![](../img/m4qi2021.png)
+
+## Meeting Summary I from 2022-11-09
+
+### Update previous work
+
+-   Harmonize 2-sample MR studies 
+-   Explained Variance calculation:
+    -   For BMI
+        -   Use updated formula that takes sample size into account
+    -   For GSD
+        -   Update prevalence $K$ in accordance to epidemiological data from Michael Leitzmann and Frank Lammert
+    -   For GBC
+        -   Update prevalence $K=0.05$ from [Cancer Today](https://gco.iarc.fr/today/home)
+-   Update Figure 01
+-   Highlight FTO variant and update Figure 02
+
+## Meeting Summary II from 2022-11-09
+
+### Outlook
+
+-   Calculate MR estimate of causal effect with different MR methods using:
+    -   all significant IVs [adapted $pval-threshold=5x10^{-6}$]
+    -   only FTO variant (get Wald estimate)
+
+-   Simulation study: 
+    -   Start with sample size of 50,000 and ~9 IVs (depending on how many remain after harmonizing datasets)
+    -   Simulate for all significant IVs $\beta_{X_{sim}}$ from a normal distribution $\beta_{X_{sim}} \sim N(\beta_{\hat X_{IV}}, \hat se_{X_N})$ and $\beta_{X_{sim}}$ from a normal distribution $\beta_{Y_{sim}} \sim N(\beta_{\hat Y_{IV}}, \hat se_{Y_N})$; start with $n_{sim}=100$
+    -   Simulate $maf$ from a Binomial distribution with $maf_{X_{sim}} \sim Bin(maf, 2N)$ 
+        - This I did not fully understand: We do not want to count the successes, but simulate the $maf$. Should we draw it from a normal distribution too?
 

### 2022-11-12 presentation/robustMR.qmd
updated figures, figure01 quadrant function

@@
 t(filter_FTO) %>% kable() %>% 
-  kable_styling(font_size=18)
+  kable_styling(font_size=14)
 ```
@@
     -   Start with sample size of 50,000 and ~9 IVs (depending on how many remain after harmonizing datasets)
-    -   Simulate for all significant IVs $\beta_{X_{sim}}$ from a normal distribution $\beta_{X_{sim}} \sim N(\beta_{\hat X_{IV}}, \hat se_{X_N})$ and $\beta_{X_{sim}}$ from a normal distribution $\beta_{Y_{sim}} \sim N(\beta_{\hat Y_{IV}}, \hat se_{Y_N})$; start with $n_{sim}=100$
+    -   Simulate for all significant IVs $\beta_{X_{sim}}$ from a normal distribution $\beta_{X_{sim}} \sim N(\beta_{\hat X_{IV}}, \hat se_{X_N})$ and $\beta_{Y_{sim}}$ from a normal distribution $\beta_{Y_{sim}} \sim N(\beta_{\hat Y_{IV}}, \hat se_{Y_N})$; start with $n_{sim}=100$
     -   Simulate $maf$ from a Binomial distribution with $maf_{X_{sim}} \sim Bin(maf, 2N)$ 
-        - This I did not fully understand: We do not want to count the successes, but simulate the $maf$. Should we draw it from a normal distribution too?
+        - This I did not fully understand: Should we draw the proportion $maf$ from a normal distribution (with $\mu=\hat {maf}$ and $sd=?$) too?
+    -   Choose simulated beta estimates (exposure/ outcome) and compare causal estimates of different MR methods with causal estimate from original study (n~300,000) and with biologic plausible estimate (Wald estimate from FTO gene as single IV)
+        - Calculate Bias, Variance and MSE
+
+-   From figure 2: SNPs were divided into 4 grouls -> Repeat simulation for SNPs from each group
 

### 2022-11-12 scripts/MSc.Rmd
updated figures, figure01 quadrant function

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
+figurespath <- paste0("../reports/", format(Sys.time(), "%Y%m%d"), '_figures/')
+knitr::opts_chunk$set(fig.path = figurespath)  # reference is script! also creates new dirs
 
@@
 library(MendelianRandomization)
+library(TwoSampleMR)
 library(mr.raps)
@@
 
-```{r}
+# Discussion
+
+-   2022-11-12:
+    -   8 SNPs removed after data harmonization, one SNP associated with FTO gene (none left)
+    -   in new formula for proportion of variance in phenotype explained by given SNP (PVE) we need the MAF not the EAF
+        - calculate: if $eaf>0.5$ -> $maf = 1-eaf$
+        - need to change betas as well? +-
+    -   We discussed $K=0.05$ in our last meeting for the "prevalence" of GBC after looking at [Cancer Today](https://gco.iarc.fr/today/home) 
+        - If done this way we yield a 20% proportion of variance explained in PVE. Switch back to ASR used last week!
+
+# Version Update
+
+The new report version includes:
+
+- First analysis
+
+```{r load-data}
 my_data <- as_tibble(read.delim("../data-boekstegers_felix/GST_BMI_association_results.txt"))
@@
 GBC_dat <- as_tibble(read.delim("../data-boekstegers_felix/C23C24_BMI_association_results.txt"))
-
-GBC_dat <- GBC_dat %>% 
-  rename(eaf_gbc=eaf_out,
-         beta_gbc=beta_out,
-         se_gbc=se_out,
-         p_value_gbc=p_value_out) %>% 
-  select(-c(chr, pos_hg19, pos_hg38, a1, a2, exposure, outcome))
-my_data <- my_data %>% left_join(GBC_dat, by =c("SNP", "dummyID"))
 ```
 
-# Get gene names from rs-ids
+# Data Harmonization
 
-```{r}
+We start with harmonizing our data sets for the exposure
+
+```{r data-harmonization, message=FALSE}
+##exposure_dat:
+#SNP
+#beta.exposure
+#se.exposure
+#effect_allele.exposure
+#other_allele.exposure
+#eaf.exposure
+exposure_dat <- my_data %>% 
+  select(dummyID, SNP, beta_exp, se_exp, a1, a2, eaf_exp, exposure) %>% 
+  rename(
+    id.exposure = dummyID,
+    beta.exposure = beta_exp,
+    se.exposure = se_exp,
+    effect_allele.exposure = a1,
+    other_allele.exposure = a2, 
+    eaf.exposure = eaf_exp, 
+    exposure = exposure
+  )
+##outcome_dat:
+#SNP
+#beta.outcome
+#se.outcome
+#effect_allele.outcome
+#other_allele.outcome
+#eaf.outcome
+#outcome
+outcome_dat <- GBC_dat %>% 
+  select(dummyID, SNP, beta_out, se_out, a1, a2, eaf_out, outcome) %>% 
+  rename(
+    id.outcome = dummyID, 
+    beta.outcome = beta_out,
+    se.outcome = se_out,
+    effect_allele.outcome = a1,
+    other_allele.outcome = a2, 
+    eaf.outcome = eaf_out,
+    outcome = outcome
+  )
+
+harmonised_dat <- TwoSampleMR::harmonise_data(
+  exposure_dat = exposure_dat, 
+  outcome_dat = outcome_dat, 
+  action = 2  # default
+  )
+
+sum_SNPs_remove <- sum(harmonised_dat$remove)
+sum_SNPs_palindromic <- sum(harmonised_dat$palindromic)
+sum_SNPs_ambiguous <- sum(harmonised_dat$ambiguous)
+sum_SNPs_remove <- nrow(harmonised_dat)-sum(harmonised_dat$mr_keep)
+```
+
+```{r combine-dat}
+my_data_harm <- 
+  harmonised_dat %>% 
+  rename(
+    exposureBMI = exposure, 
+    outcomeGBC = outcome,
+    # estimates exposure BMI
+    eaf.bmi=eaf.exposure,
+    beta.bmi=beta.exposure,
+    se.bmi=se.exposure,
+
+    # estimates outcome GBC
+    eaf.gbc=eaf.outcome,
+    beta.gbc=beta.outcome,  
+    se.gbc=se.outcome,
+  ) %>% 
+  left_join(my_data %>% select(SNP, p_value_exp, outcome, p_value_out, eaf_out, beta_out, se_out),by = c("SNP"="SNP")) %>% 
+  rename(
+    # BMI only need to add p-val
+    p.value.bmi = p_value_exp,
+
+    # GSD rename to "confounder"
+    confounderGSD = outcome, 
+    eaf.gsd = eaf_out,
+    beta.gsd = beta_out, 
+    se.gsd = se_out, 
+    p.value.gsd = p_value_out,
+  ) %>% 
+  select(
+    SNP, 
+    # DROP:   "effect_allele.exposure" "other_allele.exposure"  "effect_allele.outcome"  "other_allele.outcome"
+
+    # EXPOSURE
+    eaf.bmi,
+    beta.bmi,
+    se.bmi,
+    p.value.bmi,
+
+    # CONFOUNDER
+    eaf.gsd,
+    beta.gsd,
+    se.gsd,
+    p.value.gsd,
+    # OUTCOME
+    eaf.gbc,
+    beta.gbc,
+    se.gbc,
+    # pvalue missing and needs to be added from GBC_dat
+    # harmonize decision
+    palindromic, ambiguous, mr_keep
+  ) %>% 
+  left_join(GBC_dat %>% select(SNP, p_value_out), by=c("SNP"="SNP")) %>% 
+  rename(
+    p.value.gbc = p_value_out
+  ) %>% 
+  relocate(
+    p.value.gbc, .before = palindromic
+  ) %>% 
+  as_tibble()
+```
+
+After harmonizing our data sets we remove `r sum_SNPs_remove` SNPs since they are ambiguous.
+
+```{r rm-ambigous-SNPs}
+dropped_data <- my_data_harm %>% 
+  filter(mr_keep==FALSE)
+
+my_data_harm <- my_data_harm %>% 
+  filter(mr_keep==TRUE)
+```
+
+## Get gene names from rs-ids
+
+```{r highlight-fto-df}
 # see script biomart_search.R
@@
   distinct(refsnp_id, .keep_all = TRUE)
+
+t(filter_FTO) %>% kable() %>% 
+  kable_styling(font_size=10)
 ```
+
+After data harmonization we drop `r sum_SNPs_remove` IVs one of which is the SNP associated with the FTO gene.
+
+\clearpage
 
@@
 
-## Calculate approximate SEs for decreasing sample sizes and corresponding p-vals
+**Calculate approximate SEs for decreasing sample sizes and corresponding p-vals**
 
-Calculating p-values from a normal distribution.
-The se, pval and study size depend on each other, within each study.
-The selection of IVs should rely on the exposure study - BMI in our case.
-Summary statistics on genetic associations with BMI will be based on results of the GIANT consortium + GERA (n=334487 (100,418 + 234,069) individuals and 289 associated variants with p < 10-8) **(insert citation)**.
-We vary the sample size and calculate the number of IVs and the explained variance.
-*(Justo Mail 2022-09-27)*
-\
-Steps:
-1.       Calculate more accurates SEs using the betas and Pvals provided to you (this will be the original/initial SE for n=334.487). 
+-   se, pval and study size depend on each other, within each study
+-   selection of IVs based on exposure study (BMI in our case)
+-   Summary statistics on genetic associations with BMI based on results of the GIANT + GERA studies [@Hoffmann_2018] with $N=100,418 + 234,069 =334487$ and $SNPs=289$
+
+**Steps**
+
+1.  Calculate more accurate SEs using the betas and Pvals
 2.  Calculate the approximate SEs for decreasing samples sizes
 3.  Calculate the probability values considering the estimated betas and the corresponding SEs
-4.       Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8) **NEW 5x10^(-6))**
+4.  Exclude the IV if Pval $\ge$ genome-wide significance threshold (5 × 10−6)
 5.  Calculate cumulative explained variances
-6.       Represent figure similar to Fig1 in Chatterjee’s paper
-
-For the selected IVs, please also calculate the explained variance for the confounder and outcome.
+6.  Represent figure similar to Fig1 in Chatterjee's paper @QiChatterjee_2021
 
 ```{r z-statistic}
-### justo approach email 30-10-2022  ---------------------------------------
-
 ## LOOP simulation ----------------------------------------------
@@
 se_update <- c()
-pvals_df <- data.frame(matrix(ncol = length(colnames_pval), nrow =  nrow(my_data)))
+pvals_df <- data.frame(matrix(ncol = length(colnames_pval), nrow =  nrow(my_data_harm)))
 colnames(pvals_df) <- colnames_pval
-for (i in 1:nrow(my_data)) {
+for (i in 1:nrow(my_data_harm)) {
   # 1) Calculate more accurates SEs using the provided betas and Pvals 
-  se_update[i] <- my_data[["beta_exp"]][i]/-qnorm(p=(my_data[["p_value_exp"]][i]/2),mean=0)
+  se_update[i] <- my_data_harm[["beta.bmi"]][i]/-qnorm(p=(my_data_harm[["p.value.bmi"]][i]/2),mean=0)
 
@@
   for (j in 1:length(n)) {
-    pvals_n <- 2*(1-pnorm(abs(my_data[["beta_exp"]][i])/abs(se_n[j]), mean=0, 
+    pvals_n <- 2*(1-pnorm(abs(my_data_harm[["beta.bmi"]][i])/abs(se_n[j]), mean=0, 
                           lower.tail = TRUE))
@@
 
-my_data <- as_tibble(cbind(my_data, pvals_df))
+my_data_harm <- as_tibble(cbind(my_data_harm, pvals_df))
 
@@
 ```
+
+## Calculate proportion of variance in phenotype explained by a given SNP (PVE)
+
+-   The variance in BMI explained by the IVs is calculated with the following formula: $explained\ variance = 2f(1-f)\beta^2$, where $f$ denotes the allele frequency and $\beta$ is the additive genetic effect.
+-   The explained variance in liability to GSD and GBC is calculated as described by @So_2011
+    -   Prevalence of disease $K$ must be set:
+        -   [Prevalence GSD](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3343155/#:~:text=Gallstones%20are%20common.,develop%20stones%20at%20some%20time.&text=The%20majority%20will%20not%20develop,cholecystitis,%20cholangitis,%20or%20pancreatitis)
+            - @Aerts_2003
+        -   [Prevalence GBC](https://www.wcrf.org/cancer-trends/gallbladder-cancer-statistics/) 
+            - Update prevalence $K=0.05$ from [Cancer Today](https://gco.iarc.fr/today/home)
+
+We added a second formula to calculate the explained variance for numeric traits [@Teslovich_2010].
 
@@
   explained_variance_numeric <- 2*eaf *(1-eaf) * beta^2
+  return(explained_variance_numeric)
+}
+
+explained_variance_numeric2 <- function(maf, beta, se_beta, samplesize) {
+  ### Teslovich Nature 2010
+  # 1) maf = minor allele frequency
+  # 2) beta = additive genetic effect
+  # 3) se_beta = standard error of beta estimate
+  # 4) samplesize = samplesize of study
+  maf <- ifelse(maf>0.5, 1-maf, maf)
+
+  explained_variance_numeric <- (2*beta^2*maf*(1-maf))/ 
+    ( 2*beta^2*maf*(1-maf) + (se_beta^2)*2*samplesize*maf*(1-maf) )
   return(explained_variance_numeric)
@@
 ## calculate explained variance
-my_data <- my_data %>% 
-  mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf_exp, beta = beta_exp), 
-         explained_variance_GSD = explained_variance_binary(PA = eaf_out, 
-                                                            RR1 = exp(beta_out), 
-                                                            RR2 = exp(beta_out)^2, 
+my_data_harm <- my_data_harm %>% 
+  mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
+         explained_variance_BMI2 = explained_variance_numeric2(maf=eaf.bmi, 
+                                                           beta=beta.bmi, 
+                                                           se_beta=se.bmi, 
+                                                           samplesize=n[length(n)]
+
+         ),
+         explained_variance_GSD = explained_variance_binary(PA = eaf.gsd, 
+                                                            RR1 = exp(beta.gsd), 
+                                                            RR2 = exp(beta.gsd)^2, 
                                                             K = 0.1  # assumed prevalence in population
                                                             )$Vg,
-         explained_variance_GBC = explained_variance_binary(PA = eaf_gbc, 
-                                                            RR1 = exp(beta_gbc), 
-                                                            RR2 = exp(beta_gbc)^2, 
-                                                            K = 1.2/100000  # assumed prevalence in population
+         explained_variance_GBC = explained_variance_binary(PA = eaf.gbc, 
+                                                            RR1 = exp(beta.gbc), 
+                                                            RR2 = exp(beta.gbc)^2, 
+                                                            K =1.2/100000  # assumed prevalence in population
                                                             )$Vg
          )
-
-## paper backing prevalence GSD
-# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3343155/#:~:text=Gallstones%20are%20common.,develop%20stones%20at%20some%20time.&text=The%20majority%20will%20not%20develop,cholecystitis%2C%20cholangitis%2C%20or%20pancreatitis
-## paper backing prevalence GSD (ASR/100.000 = 1.2 worldwide)
-# https://www.wcrf.org/cancer-trends/gallbladder-cancer-statistics/
-
-#explained_variance_binary(PA=0.3,RR1=1.3,RR2=1.3^2,K=0.01)
 ```
@@
 
-```{r}
+```{r PVE_BMI_vs_pval, warning=FALSE}
 # labeling (only FTO label)
-my_data$gene_label <- NA
-my_data$gene_label[my_data$SNP %in% filter_FTO$refsnp_id] <- "FTO"
+my_data_harm$gene_label <- NA
+my_data_harm$gene_label[my_data_harm$SNP %in% filter_FTO$refsnp_id] <- "FTO"
 
 # color labeling
-my_data$fto_col_annot <- "No FTO variant"
-my_data$fto_col_annot[my_data$SNP %in% filter_FTO$refsnp_id] <- "FTO variant"
+my_data_harm$fto_col_annot <- "No FTO variant"
+my_data_harm$fto_col_annot[my_data_harm$SNP %in% filter_FTO$refsnp_id] <- "FTO variant"
 
-ggplot(data = my_data, 
-       mapping = aes(x=explained_variance_BMI*100, y= -log10(p_value_exp),  
+ggplot(data = my_data_harm, 
+       mapping = aes(x=explained_variance_BMI*100, y= -log10(p.value.bmi),  
                     col=fto_col_annot, label=gene_label)
@@
 
-ggsave(filename = "../output/plots/figure00.png", plot = fig00, 
+# second formula variance explained2
+ggplot(data = my_data_harm, 
+       mapping = aes(x=explained_variance_BMI2*100, y= -log10(p.value.bmi),  
+                    col=fto_col_annot, label=gene_label)
+       )+
+  geom_point(alpha=0.6)+
+  ggrepel::geom_text_repel(show.legend = FALSE)+
+  xlab("Explained variance in BMI in %")+
+  ylab("-log10 P-val")+
+  labs(title = "Explained variance exposure",
+       col='SNP'
+       ) +
+  theme_few()+
+  scale_colour_few() -> fig00_b
+fig00_b
+
+filenamefig00_a <- paste0("../output/plots/", Sys.Date(), "_figure00_a.png")
+ggsave(filename = filenamefig00_a, plot = fig00, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+
+filenamefig00_b <- paste0("../output/plots/", Sys.Date(), "_figure00_b.png")
+ggsave(filename = filenamefig00_b, plot = fig00_b, 
        width = 10, height = 6, 
@@
 
-```{r get-data-for-fig1}
+```{r get-data-for-fig1, warning=FALSE}
 # 4.       Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8) **NEW 5x10^(-6))**
@@
                           cum_expl_varX = rep(NA, length(n)), 
+                          cum_expl_varX2 = rep(NA, length(n)), 
                           cum_expl_varU = rep(NA, length(n)), 
@@
   # calculate cum explained variance and count of IVs
-  summarydat <- my_data %>% 
+  summarydat <- my_data_harm %>% 
     filter(!!sym(select_col) < threshold) %>% 
-    select(explained_variance_BMI, explained_variance_GSD, explained_variance_GBC) %>% 
+    select(explained_variance_BMI, explained_variance_BMI2, explained_variance_GSD, explained_variance_GBC) %>% 
     summarise(expl_variances =across(.cols = everything(), function(x) sum(x)),
@@
   df_cum_expl_var[i, 4] <- summarydat$expl_variances[3]
-  df_cum_expl_var[i, 5] <- summarydat$count_IVs
+  df_cum_expl_var[i, 5] <- summarydat$expl_variances[4]
+  df_cum_expl_var[i, 6] <- summarydat$count_IVs
 }
@@
 
-```{r plot-fig1}
+```{r plot-fig1, warning=FALSE}
 # 6.       Represent figure similar to Fig1 in Chatterjee’s paper
@@
     cum_expl_varX = cum_expl_varX *100, # in percent 
+    cum_expl_varX2 = cum_expl_varX2 *100,
     cum_expl_varU = cum_expl_varU *100,
@@
 
-plot_data <- df_cum_expl_var %>% 
-  pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, count_IVs_scaled)) %>% 
-  mutate(name = factor(name, labels = c("Number of IVs", "% var(U)", "% var(X)")))
-
-ggplot(plot_data, aes(x=log10(n), y=value, col=name))+
-  geom_point(alpha=0.6)+
-  geom_line(alpha=0.6)+
-  xlab("log10(N)")+
-  #ylab("% variance explained by IVs")+
-  labs(title = "Figure 1",
-       col='', 
-       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
-       ) +
-  scale_y_continuous(
-    # Features of the first axis
-    name = "% variance explained by IVs",
-    # Add a second axis and specify its features
-    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
-  ) + 
-  theme_few()+
-  scale_colour_few() ->fig01
-fig01
-
-ggsave(filename = "../output/plots/figure01.png", plot = fig01, 
-       width = 10, height = 6, 
-       units = "in"  # default
-       )
+# plot_data <- df_cum_expl_var %>% 
+#   pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, count_IVs_scaled)) %>% 
+#   mutate(name = factor(name, labels = c("Number of IVs", "% var(U)", "% var(X)")))
+# 
+# ggplot(plot_data, aes(x=log10(n), y=value, col=name))+
+#   geom_point(alpha=0.6)+
+#   geom_line(alpha=0.6)+
+#   xlab("log10(N)")+
+#   #ylab("% variance explained by IVs")+
+#   labs(title = "Figure 1",
+#        col='', 
+#        caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
+#        ) +
+#   scale_y_continuous(
+#     # Features of the first axis
+#     name = "% variance explained by IVs",
+#     # Add a second axis and specify its features
+#     sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
+#   ) + 
+#   theme_few()+
+#   scale_colour_few() ->fig01
+# fig01
+# 
+# plot_data_b <- df_cum_expl_var %>% 
+#   pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, count_IVs_scaled)) %>% 
+#   mutate(name = factor(name, labels = c("Number of IVs", "% var(U)", "% var(X)")))
+# 
+# ggplot(plot_data_b, aes(x=log10(n), y=value, col=name))+
+#   geom_point(alpha=0.6)+
+#   geom_line(alpha=0.6)+
+#   xlab("log10(N)")+
+#   #ylab("% variance explained by IVs")+
+#   labs(title = "Figure 1",
+#        col='', 
+#        caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
+#        ) +
+#   scale_y_continuous(
+#     # Features of the first axis
+#     name = "% variance explained by IVs",
+#     # Add a second axis and specify its features
+#     sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
+#   ) + 
+#   theme_few()+
+#   scale_colour_few() ->fig01_b
+# fig01_b
+# 
+# ggsave(filename = "../output/plots/figure01.png", plot = fig01_b, 
+#        width = 10, height = 6, 
+#        units = "in"  # default
+#        )
+# 
+# filenamefig01_a <- paste0("../output/plots/", Sys.Date(), "_figure01_a.png")
+# filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_b.png")
 ```
 
-```{r updated-fig01}
+```{r updated-fig01, warning=FALSE}
 plot_data1 <- df_cum_expl_var %>% 
   pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, cum_expl_varY)) %>% 
-  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")))
+  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
+         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
+         )
 
@@
   geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
-  scale_x_continuous(name = "Sample Size",
+  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                      labels = comma, 
                      breaks = N_annotation, #c( 10000, n[7:length(n)]),
-                     limits = c(10000, 350000))+
+                     limits = c(10000, 300000))+
   #ylab("% variance explained by IVs")+
@@
                                            "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
-                                           "Number of IVs" = "black"
-                                           )) -> fig01.update
+                                           "Number of IVs" = "black"),
+                     # order
+                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
+  ) -> fig01.update
+
 fig01.update
 # scale_color_manual(name = "Y series", values = c("Y1" = "darkblue", "Y2" = "red"))
-ggsave(filename = "../output/plots/figure01update.png", plot = fig01.update, 
+
+plot_data1b <- df_cum_expl_var %>% 
+  pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, cum_expl_varY)) %>% 
+  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
+         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
+         )
+## plot
+ggplot(plot_data1b, aes(x=n, y=value, col=name))+
+  geom_point(alpha=0.6)+
+  geom_line(alpha=0.6)+
+  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
+  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
+                     labels = comma, 
+                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
+                     limits = c(10000, 300000))+
+  #ylab("% variance explained by IVs")+
+  labs(title = "Figure 1",
+       col='', 
+       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
+       ) +
+  scale_y_continuous(
+    # Features of the first axis
+    name = "% variance explained by IVs",
+    # Add a second axis and specify its features
+    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
+  ) + 
+  theme_few()+
+  #scale_colour_few() +
+  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
+  #                                         "% var(X)" = "blue", 
+  #                                         "% var(Y)" = "darkblue", 
+  #                                         "Number of IVs" = "black"
+  #                                         ))
+  #
+  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
+                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
+                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
+                                           "Number of IVs" = "black"),
+                     # order
+                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
+  ) -> fig01b.update
+fig01b.update
+
+# SAVE
+filenamefig01_a <- paste0("../output/plots/", Sys.Date(), "_figure01_a.png")
+filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_b.png")
+ggsave(filename = filenamefig01_a, plot = fig01.update, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+
+ggsave(filename = filenamefig01_b, plot = fig01b.update, 
        width = 10, height = 6, 
@@
 ```
+
+Discussion: *Pleiotropy* with increasing sample size. INSIDE assumption.
 
@@
 
-```{r MAF-Pval}
-ggplot(data = my_data, aes(x=eaf_exp, y=-log10(p_value_exp)))+
+```{r MAF-Pval, warning=FALSE}
+ggplot(data = my_data_harm, aes(x=eaf.bmi, y=-log10(p.value.bmi)))+
   geom_point(alpha=0.6)+
@@
   ylab("-log10(P-value) ")+
-  labs(title = "Figure 2",
+  labs(title = "MAF vs Pval",
        subtitle = "Define different genetic architectures",
@@
 
-ggsave(filename = "../output/plots/figure02.png", plot = fig02)
+# ggsave(filename = "../output/plots/figure02.png", plot = fig02)
 ```
@@
 ```{r mafVSbetas}
-plot_data_2 <- my_data %>% 
-  mutate(maf_exp = ifelse(eaf_exp >0.5, 1-eaf_exp, eaf_exp)) %>% 
-  select(maf_exp, beta_exp)
+plot_data_2 <- my_data_harm %>% 
+  mutate(maf.bmi = ifelse(eaf.bmi >0.5, 1-eaf.bmi, eaf.bmi)) %>% 
+  select(maf.bmi, beta.bmi)
 
-ggplot(data = plot_data_2, aes(x=maf_exp, y=abs(beta_exp)))+
+ggplot(data = plot_data_2, aes(x=maf.bmi, y=abs(beta.bmi)))+
   geom_point(alpha=0.6)+
-  geom_hline(yintercept=median(abs(plot_data_2$beta_exp)), col="red", alpha=0.6)+
-  geom_vline(xintercept=median(plot_data_2$maf_exp), col="red", alpha=0.6)+
-  xlab("Minor allele frequency for exposure")+
-  ylab("abs(beta_BMI)")+
+  geom_hline(yintercept=median(abs(plot_data_2$beta.bmi)), 
+             col=ggthemes_data$few$colors$Dark[6,2][[1]], alpha=0.6, linetype=2)+
+  geom_vline(xintercept=median(plot_data_2$maf.bmi), 
+             col=ggthemes_data$few$colors$Dark[6,2][[1]], alpha=0.6, linetype=2)+
+  annotate(geom = "text", x = 0.15, y = 0.045, 
+           label="relatively rare variants\nwith relatively strong effects", 
+           color= ggthemes_data$few$colors$Dark[2,2][[1]]
+             )+
+  annotate(geom = "text", x = 0.42, y = 0.05, 
+           label="relatively common variants\nwith relatively strong effects", 
+           color=ggthemes_data$few$colors$Dark[3,2][[1]])+
+  annotate(geom = "text", x = 0.1, y = 0.015, 
+           label="rare variants\nwith weak effects", color=ggthemes_data$few$colors$Dark[4,2][[1]])+
+  annotate("segment", x=0.44,xend=0.42,y=0.025,yend=0.013,arrow=arrow(), 
+           color=ggthemes_data$few$colors$Dark[5,2][[1]])+
+  annotate(geom = "text", x = 0.44, y = 0.03, 
+           label="common variants\nwith weak effects", 
+           color=ggthemes_data$few$colors$Dark[5,2][[1]])+
+  #geom_text(x=0.1, y=0.05, label="hello", family = "Times New Roman",
+  #          size=6, col="red")+
+  xlab("Minor allele frequency")+
+  ylab("Exposure effect size")+  # abs(beta_BMI)
   labs(title = "Figure 2-update",
@@
   theme_few()+
-  scale_colour_few() ->fig02.update
+  scale_colour_few() -> fig02.update
 fig02.update
-ggsave(filename = "../output/plots/figure02update.png", plot = fig02.update)
+
+# SAVE
+filenamefig02 <- paste0("../output/plots/", Sys.Date(), "_figure02.png")
+ggsave(filename = filenamefig02, plot = fig02.update)
 ```
+
+We want to repeat Figure 01 for each of the 4 defined quadrants in the MAF vs beta plot above.
+
+```{r figure1quadrants, fig.show="hold", out.width="50%", warning=FALSE}
+
+plotfigure01 <- function(data, effect_size="strong", maf="rare"){
+  ## data: my_data_harm
+  # effect size: "strong" or "weak"
+  # maf: "rare" or "common
+  df_cum_expl_var <- tibble(n=n, 
+                            cum_expl_varX = rep(NA, length(n)), 
+                            cum_expl_varX2 = rep(NA, length(n)), 
+                            cum_expl_varU = rep(NA, length(n)), 
+                            cum_expl_varY = rep(NA, length(n)),
+                            count_IVs = rep(NA, length(n))
+  )
+  # create maf col
+  data <- data %>% 
+      mutate(maf.bmi = ifelse(eaf.bmi >0.5, 1-eaf.bmi, eaf.bmi))
+
+  ## filter for quadrant
+  if (effect_size=="strong" & maf == "rare") {
+    data <- data %>% 
+      filter(abs(beta.bmi) > median(abs(beta.bmi)) &  maf.bmi < median(maf.bmi))
+  } else if (effect_size=="strong" & maf == "common") {
+    data <- data %>% 
+      filter(abs(beta.bmi) > median(abs(beta.bmi)) &  maf.bmi > median(maf.bmi))
+  } else if (effect_size=="weak" & maf == "common") {
+    data <- data %>% 
+      filter(abs(beta.bmi) < median(abs(beta.bmi)) &  maf.bmi > median(maf.bmi))
+  } else if (effect_size=="weak" & maf == "rare") {
+    data <- data %>% 
+      filter(abs(beta.bmi) < median(abs(beta.bmi)) &  maf.bmi < median(maf.bmi))
+  }
+
+  # create data for plot
+  for (i in 1:length(colnames_pval) ) {
+    threshold <- 5*10^(-6)
+    select_col <- colnames_pval[i]
+    # calculate cum explained variance and count of IVs
+    summarydat <- data %>% 
+      filter(!!sym(select_col) < threshold) %>% 
+      select(explained_variance_BMI, explained_variance_BMI2, explained_variance_GSD, explained_variance_GBC) %>% 
+      summarise(expl_variances =across(.cols = everything(), function(x) sum(x)),
+                count_IVs = n())
+    # fill up df
+    df_cum_expl_var[i, 2] <- summarydat$expl_variances[1]
+    df_cum_expl_var[i, 3] <- summarydat$expl_variances[2]
+    df_cum_expl_var[i, 4] <- summarydat$expl_variances[3]
+    df_cum_expl_var[i, 5] <- summarydat$expl_variances[4]
+    df_cum_expl_var[i, 6] <- summarydat$count_IVs
+  }
+
+  ## plot fig 01 
+  scale_factor <- 4.5/284
+  df_cum_expl_var <- df_cum_expl_var %>% 
+    mutate(
+      cum_expl_varX = cum_expl_varX *100, # in percent 
+      cum_expl_varX2 = cum_expl_varX2 *100,
+      cum_expl_varU = cum_expl_varU *100,
+      cum_expl_varY = cum_expl_varY *100,
+      count_IVs_scaled = count_IVs*scale_factor)
+
+  plot_data1 <- df_cum_expl_var %>% 
+    pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, cum_expl_varY)) %>% 
+    mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
+           name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
+           )
+
+  N_annotation <- c(10000, 100000, 200000, 300000)
+  ## plot
+  plotsubtitle <- paste0("Relatively ", maf ," variants\nwith relatively ", effect_size, " effects")
+  ggplot(plot_data1, aes(x=n, y=value, col=name))+
+    geom_point(alpha=0.6)+
+    geom_line(alpha=0.6)+
+    geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
+    scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
+                       labels = comma, 
+                       breaks = N_annotation, #c( 10000, n[7:length(n)]),
+                       limits = c(10000, 300000))+
+    #ylab("% variance explained by IVs")+
+    labs(title = "Figure 1", subtitle = plotsubtitle,
+         col='', 
+         caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits   explained by the IVs"
+         ) +
+    scale_y_continuous(
+      # Features of the first axis
+      name = "% variance explained by IVs",
+      # Add a second axis and specify its features
+      sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
+    ) + 
+    theme_few()+
+    #scale_colour_few() +
+    #scale_color_manual(name = "", values = c("% var(U)" = "red", 
+    #                                         "% var(X)" = "blue", 
+    #                                         "% var(Y)" = "darkblue", 
+    #                                         "Number of IVs" = "black"
+    #                                         ))
+    #
+    scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
+                                             "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
+                                             "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
+                                             "Number of IVs" = "black"),
+                       # order
+                       breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
+    ) -> figure01.quadrant
+  print(figure01.quadrant)
+
+  filenamefig01_quadrant <- paste0("../output/plots/", Sys.Date(), "_figure01_", 
+                                   "EFFECT_SIZE_", effect_size,
+                                   "_VARIANTS_", maf,
+                                   ".png")
+  ggsave(filename = filenamefig01_quadrant, plot = figure01.quadrant, 
+         width = 10, height = 6, 
+         units = "in"  # default
+         )
+}
+
+# create all combis
+effect_size <- c("strong", "weak")
+maf <- c("rare", "common")
+all_quadrants <- tidyr::crossing(effect_size, maf)
+
+####### PLOT NOW  ---
+for (i in 1:nrow(all_quadrants)) {
+  plotfigure01(data = my_data_harm, 
+               effect_size = all_quadrants$effect_size[i], 
+               maf = all_quadrants$maf[i])
+}
+
+```
+
+\clearpage
 
@@
 
-- Let’s try to estimate parameters for the simulation (pi1, varx, pi2, varu, muy, vary, pi3, pi4, delta0) based on real data. Maybe you could start with N=10.000 and N=300.000 (I will help with this)
+## Get MR-estimates first

### 2022-11-12 scripts/render_param_reports_msc.R

@@ -1,0 +1,42 @@
+
+
+
+
+## render report and mail results
+library(dplyr)
+library(gmailr)
+rm(list=ls())  # not in rmd doc, otherwise params deleted
+#gm_auth_configure(path = "/Users/christophreich/Library/CloudStorage/OneDrive-bwedu/credentials.json")
+
+render_report = function(...) {
+  ## disease: acs, cad, hfref, dcm | default: dcm
+  ## analysis: selected, full | default: selected
+  
+  rmarkdown::render(
+    "scripts/MSc.Rmd",
+    params = list(
+      doc_title =  paste0("Assessment of simulation procedures and robust methods\nfor Mendelian randomization analysis")
+    ),
+    output_file = paste0(
+      format(Sys.time(), "%Y%m%d"), '_robustMRsimulation.html'
+    ),
+    output_dir = "reports",
+    envir = globalenv()
+  )
+  # save filename for sending later
+  filename.html <<-  paste0("reports/", format(Sys.time(), "%Y%m%d"), '_robustMRsimulation.html' )
+  title.mail <<- paste0("Robust MR report | Date:  ",  format(Sys.time(), "%Y-%m-%d"))
+  
+ # email <-
+ #   gmailr::gm_mime() %>%
+ #   gmailr::gm_to("reich.c@hotmail.com") %>%
+ #   gmailr::gm_from("q050cr@gmail.com") %>%
+ #   gmailr::gm_subject(paste0("Job finished: ", title.mail)) %>%
+ #   gmailr::gm_text_body("Hi Christoph,\nHere is a new report for you.\nCheers")
+ # # attachment
+ # email <- gmailr::gm_attach_file(email, filename.html)
+ # gmailr::gm_send_message(email, user_id = "me")
+}
+
+# NOW START!!!
+render_report()

### 2022-11-14 scripts/MSc.Rmd
several MR methods to calculate theta_X implemented

@@
 library(conflicted)
+source("MR_lasso.R")  # Slob&Burgess 2020
 conflict_prefer("select", "dplyr")
@@
   action = 2  # default
-  )
+  ) # # used later for analysis with the {TwoSampleMR} package
 
@@
 ## Get MR-estimates first
+
+### Using {TwoSampleMR} package
+
+```{r}
+harmonised_dat4TwoSampleMR <- harmonised_dat %>% 
+  mutate(id.exposure = "BMI", id.outcome = "GBC")  # https://mrcieu.github.io/TwoSampleMR/articles/perform_mr.html
+
+res <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR)
+res <- res %>% mutate(OR = exp(b), CI.lower= exp(b-se), CI.upper=exp(b+se))
+
+res %>% 
+  kable(digits = 3, 
+        caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI"
+        )
+
+#res.selected_methods <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR, 
+#                                        method_list = c("mr_egger_regression", "mr_ivw"))
+```
+
+### Using {MendelianRandomization} package
+
+```{r}
+mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
+est <- tibble(mr_methods=mr_methods, estimate=NA, se.estimate=NA, time=NA)
+numIV <- nrow(my_data_harm)
+
+if (numIV>2){
+        mr.obj = MendelianRandomization::mr_input(
+          # exposure
+          bx = my_data_harm$beta.bmi, 
+          bxse = my_data_harm$se.bmi, 
+          # outcome
+          by = my_data_harm$beta.gbc, 
+          byse = my_data_harm$se.gbc, 
+          snps = my_data_harm$SNP,
+          exposure = "Body mass index",
+          outcome = "Gallbladder Cancer"
+        )
+        # 1. IVW
+        T0 = proc.time()[3]
+        res = MendelianRandomization::mr_ivw(mr.obj)
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="IVW"] = res$Estimate
+        est$se.estimate[est$mr_methods=="IVW"] = res$StdError
+        est$time[est$mr_methods=="IVW"] = T1-T0
+        rm(res)
+
+        # 2. median
+        T0 = proc.time()[3]
+        res = MendelianRandomization::mr_median(mr.obj)
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="median"] = res$Estimate
+        est$se.estimate[est$mr_methods=="median"] = res$StdError
+        est$time[est$mr_methods=="median"] = T1-T0
+        rm(res)
+        # 3. mode
+        T0 = proc.time()[3]
+        res = MendelianRandomization::mr_mbe(mr.obj)
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="mode"] = res$Estimate
+        est$se.estimate[est$mr_methods=="mode"] = res$StdError
+        est$time[est$mr_methods=="mode"] = T1-T0
+        rm(res)
+        # 4.MR-PRESSO    ## CAVE: 5 minutes per round, consider only calculating for defined thresholds
+        T0 = proc.time()[3]
+        res <- MRPRESSO:: mr_presso(BetaOutcome = "beta.gbc", 
+                    BetaExposure = "beta.bmi", 
+                    SdOutcome = "se.gbc", 
+                    SdExposure = "se.bmi", 
+                    OUTLIERtest = TRUE, 
+                    DISTORTIONtest = TRUE, 
+                    seed = 20221114,
+                    data = as.data.frame(my_data_harm), ## must be a data.frame
+                    NbDistribution = 1000,   # must be changed
+                    SignifThreshold = 0.05)
+        T1 = proc.time()[3]
+        if (!is.na(res$`Main MR results`[2,"Causal Estimate"]) & !is.na(res$`Main MR results`[2,"Sd"])){
+                est$estimate[est$mr_methods=="PRESSO"] =  res$`Main MR results`[2,"Causal Estimate"]
+                est$se.estimate[est$mr_methods=="PRESSO"] = res$`Main MR results`[2,"Sd"]
+            } else{
+                est$estimate[est$mr_methods=="PRESSO"] =  res$`Main MR results`[1,"Causal Estimate"]
+                est$se.estimate[est$mr_methods=="PRESSO"] = res$`Main MR results`[1,"Sd"]
+            }
+        est$time[est$mr_methods=="PRESSO"] = T1-T0
+        rm(res)
+        # 5. robust
+        T0 = proc.time()[3]
+        res = MendelianRandomization::mr_ivw(mr.obj,"random", robust = TRUE)
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="Robust"] = res$Estimate
+        est$se.estimate[est$mr_methods=="Robust"] = res$StdError
+        est$time[est$mr_methods=="Robust"] = T1-T0
+        rm(res)
+        # 6. MR-Lasso
+        T0 = proc.time()[3]
+        res = MR_lasso(betaYG = my_data_harm$beta.gbc, betaXG = my_data_harm$beta.bmi, sebetaYG = my_data_harm$se.gbc)
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="Lasso"] = res$ThetaEstimate
+        est$se.estimate[est$mr_methods=="Lasso"] = res$ThetaSE
+        est$time[est$mr_methods=="Lasso"] = T1-T0
+        rm(res)
+        # 7. Egger
+        T0 = proc.time()[3]
+        res = MendelianRandomization::mr_egger(mr.obj)
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="egger"] = res$Estimate
+        est$se.estimate[est$mr_methods=="egger"] = res$StdError.Est
+        est$time[est$mr_methods=="egger"] = T1-T0
+        rm(res)
+        # 8. contamination mixture
+        T0 = proc.time()[3]
+        res = MendelianRandomization::mr_conmix(mr.obj)
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="conmix"] = res$Estimate
+        CIlength = res$CIUpper-res$CILower
+        if (length(CIlength)>1) print("conmix multimodal")
+        est$se.estimate[est$mr_methods=="conmix"] = sum(CIlength)/1.96/2 ## Caution: this may be problematic (Qi&Chatterjee2021)
+        est$time[est$mr_methods=="conmix"] = T1-T0
+        rm(res)
+        # 9. MRMix
+              # theta_temp_vec = seq(-0.5,0.5,by=0.01)
+        T0 = proc.time()[3]
+        res = MRMix::MRMix(betahat_x = my_data_harm$beta.bmi, 
+                          betahat_y = my_data_harm$beta.gbc, 
+                          sx =my_data_harm$se.bmi, 
+                          sy = my_data_harm$se.gbc)
+        res_se = MRMix::MRMix_se(betahat_x = my_data_harm$beta.bmi, 
+                          betahat_y = my_data_harm$beta.gbc, 
+                          sx =my_data_harm$se.bmi, 
+                          sy = my_data_harm$se.gbc, 
+                          theta = res$theta, # estimate of causal effect, assuming the summary stats are standardized, theta represents increase in mean value of Y in s.d.
+unit of Y (for cont outcomes) or log-OR of Y (for binary outcomes)
+                          pi0 = res$pi0, 
+                          sigma2 = res$sigma2
+                          )
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="MRMix"] = res$theta
+        est$se.estimate[est$mr_methods=="MRMix"] = res_se
+        est$time[est$mr_methods=="MRMix"] = T1-T0
+        rm(res, res_se)
+        # 10. MR-RAPS  
+        ## Warning: The estimated overdispersion parameter is very small. Consider using the simple model without overdispersion.
+        T0 = proc.time()[3]
+        res = mr.raps::mr.raps.overdispersed.robust(b_exp = my_data_harm$beta.bmi, 
+                                                    b_out = my_data_harm$beta.gbc, 
+                                                    se_exp = my_data_harm$se.bmi, 
+                                                    se_out = my_data_harm$se.gbc, 
+                                                    loss.function = "huber", 
+                                                    k = 1.345, 
+                                                    initialization = c("l2"), 
+                                                    suppress.warning = FALSE, 
+                                                    diagnosis = FALSE, 
+                                                    niter = 20, 
+                                                    tol = .Machine$double.eps^0.5)
+        T1 = proc.time()[3]
+        est$estimate[est$mr_methods=="RAPS"] = res$beta.hat
+        est$se.estimate[est$mr_methods=="RAPS"] = res$beta.se
+        est$time[est$mr_methods=="RAPS"] = T1-T0
+        rm(res)
+}
+
+est %>% 
+  mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
+  kable(digits=3, caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI")
+```

### 2022-11-14 scripts/MSc.Rmd
sim no_instruments, sim beta ~N(mu=beta, se=se_n)

@@
   html_document:
+    code_folding: hide
     theme: paper
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+#knitr::opts_chunk$set(echo = FALSE)
 figurespath <- paste0("../reports/", format(Sys.time(), "%Y%m%d"), '_figures/')
@@
 library(dplyr)
+library(purrr)
 library(tidyr)
@@
 library(ggplot2)
+library(ggthemes)
 library(scales)
@@
 
-```{r}
+```{r TwoSampleMR-analysis}
 harmonised_dat4TwoSampleMR <- harmonised_dat %>% 
@@
         caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI"
-        )
-
-#res.selected_methods <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR, 
-#                                        method_list = c("mr_egger_regression", "mr_ivw"))
+        ) %>% 
+  kable_classic(full_width = FALSE, html_font = "Cambria")
+```
+
+```{r scatter-TwoSampleMR, warning=FALSE, message=FALSE}
+# mr_method_list()
+res.selected_methods <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR, 
+                                        method_list = c("mr_egger_regression", "mr_ivw"))
+p1 <- mr_scatter_plot(res.selected_methods, harmonised_dat4TwoSampleMR)
+p1[[1]] #+
+  #ggthemes::theme_few()
+
+# SAVE
+filename.mr.scatter <- paste0("../output/plots/", Sys.Date(), "_figure_mrscatter.png")
+ggsave(filename = filename.mr.scatter, plot = p1[[1]], 
+         #width = 10, height = 6, 
+         #units = "in"  # default
+         )
+```
+
+```{r forestplot-TwoSampleMR, fig.width=5, fig.height=24, warning=FALSE}
+res_single <- TwoSampleMR::mr_singlesnp(harmonised_dat4TwoSampleMR, all_method = c("mr_egger_regression", "mr_ivw"))
+p2 <- mr_forest_plot(res_single,  exponentiate = FALSE)
+p2[[1]]
+
+#SAVE
+filename.mr.forest <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest.png")
+ggsave(filename = filename.mr.forest, plot = p2[[1]], 
+       width = 5, 
+       height = 24, 
+         #units = "in"  # default
+         )
+```
+
+```{r funnel-plot}
+p4 <- mr_funnel_plot(res_single)
+p4[[1]] +
+  ggthemes::theme_few()
 ```
@@
 
-```{r}
+```{r MendelianRandomization-analysis, warning=FALSE}
 mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
@@
   mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
-  kable(digits=3, caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI")
-```
+  kable(digits=3, caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI") %>% 
+  kable_classic(full_width = FALSE, html_font = "Cambria")
+```
+
+## Simulation
+
+```{r simulatesimulatesimulate}
+## try starting with n=50.000 -> selected 9 IVs
+set.seed(1234)
+
+## initialize
+nsim <- 100
+se_update_BMI <- c()
+se_update_GBC <- c()
+
+est <- tibble(no_sim = 1:nsim)
+#est <- matrix(NA, 
+#              nrow = nsim, 
+#              ncol = 2*length(n)  # theta_sim and theta_se_sim
+#              ) 
+#colnames(est) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))
+
+## START SIM
+T0 = proc.time()[3]
+for (i in 1:length(colnames_pval)) {
+  T0 = proc.time()[3]
+  threshold <- 5*10^-6
+  numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
+  if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
+    next
+  }
+
+  # FILTER DATA
+  sim_dat <- my_data_harm %>% 
+    filter(!!sym(colnames_pval[i]) < threshold) %>% 
+    select(SNP, eaf.bmi, beta.bmi, p.value.bmi, eaf.gbc, beta.gbc, p.value.gbc)
+
+  # get simulated se_{sample size}
+  for (j in 1:nrow(sim_dat)) {
+    # as done before in section "simulating on empirical data"
+    se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
+    se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
+  }
+
+  sim_dat <- sim_dat %>%
+      bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))
+
+  # simulate beta values from normal_distribution
+  for (k in 1:nsim) {
+    sim_dat <- sim_dat %>% 
+      mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
+      mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y)))
+
+    # PERFORM MR_analysis
+    mr.obj = MendelianRandomization::mr_input(
+          bx = sim_dat$beta.bmi_sim_norm, 
+          bxse = sim_dat$se_update_BMI, 
+          # outcome
+          by = sim_dat$beta.gbc_sim_norm, 
+          byse = sim_dat$se_update_GBC, 
+          snps = sim_dat$SNP,
+          exposure = "Body mass index",
+          outcome = "Gallbladder Cancer"
+        ) 
+    # CONMIX (as reference)
+    res = MendelianRandomization::mr_conmix(mr.obj)
+    colname_theta <- paste0("theta_sim_", n[i])
+    colname_theta_se <- paste0("theta_se_sim_", n[i])
+    ## store results
+    est[k, colname_theta] <- res$Estimate
+    CIlength = res$CIUpper-res$CILower
+    if (length(CIlength)>1) print("conmix multimodal")
+    est[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
+    rm(res)
+  }
+  print(paste0("|||-----------------------Run finished for sample size: ", n[i], " -----------------------|||"))
+}
+T1 = proc.time()[3]
+timediff = T1-T0
+```

### 2022-11-14 scripts/render_param_reports_msc.R

@@ -7,6 +7,7 @@ library(dplyr)
 library(gmailr)
 rm(list=ls())  # not in rmd doc, otherwise params deleted
 #gm_auth_configure(path = "/Users/christophreich/Library/CloudStorage/OneDrive-bwedu/credentials.json")
+#gm_auth()
 
 render_report = function(...) {
   ## disease: acs, cad, hfref, dcm | default: dcm

### 2022-11-14 scripts/MSc.Rmd
slight changes before nightshift

@@
 timediff = T1-T0
-```
+
+saveRDS(est, paste0(Sys.Date(), "_est.rds")
+
+```

### 2022-11-15 scripts/MSc.Rmd
update be4 nightshift

--- 2022-11-14 scripts/MSc.Rmd
@@
         | christoph.reich@med.uni-heidelberg.de 
-        | lorenzo@imbi.uni-heidelberg.de
         | 
@@
     latex_engine: xelatex
-#bibliography: ../references/project01.bib
+bibliography: ../references/references.bib
 ---
@@
 
-```{r simulatesimulatesimulate}
+```{r simulatesimulatesimulate, warning=FALSE}
 ## try starting with n=50.000 -> selected 9 IVs
-set.seed(1234)
-
-## initialize
-nsim <- 100
-se_update_BMI <- c()
-se_update_GBC <- c()
-
-est <- tibble(no_sim = 1:nsim)
-#est <- matrix(NA, 
-#              nrow = nsim, 
-#              ncol = 2*length(n)  # theta_sim and theta_se_sim
+# set.seed(1234)
+# 
+# ## initialize
+# nsim <- 25
+# est <- tibble(no_sim = 1:nsim)
+# #est <- matrix(NA, 
+# #              nrow = nsim, 
+# #              ncol = 2*length(n)  # theta_sim and theta_se_sim
+# #              ) 
+# #colnames(est) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))
+# 
+# ## START SIM
+# T0 = proc.time()[3]
+# for (i in 1:length(colnames_pval)) {
+#   threshold <- 5*10^-6
+#   se_update_BMI <- c()
+#   se_update_GBC <- c()
+#   
+#   numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
+#   
+#   if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
+#     next
+#   }
+#   
+#   # FILTER DATA
+#   sim_dat <- my_data_harm %>% 
+#     filter(!!sym(colnames_pval[i]) < threshold) %>% 
+#     select(SNP, eaf.bmi, beta.bmi, p.value.bmi, eaf.gbc, beta.gbc, p.value.gbc)
+#   
+#   # get simulated se_{sample size}
+#   for (j in 1:nrow(sim_dat)) {
+#     # as done before in section "simulating on empirical data"
+#     se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
+#     se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
+#   }
+#   
+#   sim_dat <- sim_dat %>%
+#       bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))
+#   
+#   # simulate beta values from normal_distribution
+#   for (k in 1:nsim) {
+#     sim_dat <- sim_dat %>% 
+#       mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
+#       mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y)))
+#     
+#     # PERFORM MR_analysis
+#     mr.obj = MendelianRandomization::mr_input(
+#           bx = sim_dat$beta.bmi_sim_norm, 
+#           bxse = sim_dat$se_update_BMI, 
+#           # outcome
+#           by = sim_dat$beta.gbc_sim_norm, 
+#           byse = sim_dat$se_update_GBC, 
+#           snps = sim_dat$SNP,
+#           exposure = "Body mass index",
+#           outcome = "Gallbladder Cancer"
 #         ) 
-#colnames(est) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))
-
-## START SIM
-T0 = proc.time()[3]
-for (i in 1:length(colnames_pval)) {
-  T0 = proc.time()[3]
-  threshold <- 5*10^-6
-  numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
-  if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
-    next
-  }
-
-  # FILTER DATA
-  sim_dat <- my_data_harm %>% 
-    filter(!!sym(colnames_pval[i]) < threshold) %>% 
-    select(SNP, eaf.bmi, beta.bmi, p.value.bmi, eaf.gbc, beta.gbc, p.value.gbc)
-
-  # get simulated se_{sample size}
-  for (j in 1:nrow(sim_dat)) {
-    # as done before in section "simulating on empirical data"
-    se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
-    se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
-  }
-
-  sim_dat <- sim_dat %>%
-      bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))
-
-  # simulate beta values from normal_distribution
-  for (k in 1:nsim) {
-    sim_dat <- sim_dat %>% 
-      mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
-      mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y)))
-
-    # PERFORM MR_analysis
-    mr.obj = MendelianRandomization::mr_input(
-          bx = sim_dat$beta.bmi_sim_norm, 
-          bxse = sim_dat$se_update_BMI, 
-          # outcome
-          by = sim_dat$beta.gbc_sim_norm, 
-          byse = sim_dat$se_update_GBC, 
-          snps = sim_dat$SNP,
-          exposure = "Body mass index",
-          outcome = "Gallbladder Cancer"
-        ) 
-    # CONMIX (as reference)
-    res = MendelianRandomization::mr_conmix(mr.obj)
-    colname_theta <- paste0("theta_sim_", n[i])
-    colname_theta_se <- paste0("theta_se_sim_", n[i])
-    ## store results
-    est[k, colname_theta] <- res$Estimate
-    CIlength = res$CIUpper-res$CILower
-    if (length(CIlength)>1) print("conmix multimodal")
-    est[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
-    rm(res)
-  }
-  print(paste0("|||-----------------------Run finished for sample size: ", n[i], " -----------------------|||"))
-}
-T1 = proc.time()[3]
-timediff = T1-T0
-```
+#     # CONMIX (as reference)
+#     res = MendelianRandomization::mr_conmix(mr.obj)
+#     colname_theta <- paste0("theta_sim_", n[i])
+#     colname_theta_se <- paste0("theta_se_sim_", n[i])
+#     ## store results
+#     est[k, colname_theta] <- res$Estimate
+#     CIlength = res$CIUpper-res$CILower
+#     if (length(CIlength)>1) print("conmix multimodal")
+#     est[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
+#     rm(res)
+#   }
+#   rm(se_update_BMI, se_update_GBC)
+#   print(paste0("|||-----------------------Run finished for sample size: ", n[i], " -----------------------|||"))
+# }
+# 
+# 
+# T1 = proc.time()[3]
+# timediff = T1-T0
+# 
+# 
+# saveRDS(est, paste0(Sys.Date(), "_est.rds"))
+```

### 2022-11-16 scripts/MSc.Rmd
edited Msc script

--- 2022-11-14 scripts/MSc.Rmd
@@
 
-```{r simulatesimulatesimulate}
-## try starting with n=50.000 -> selected 9 IVs
+```{r simulatesimulatesimulate, warning=FALSE}
+# try starting with n=50.000 -> selected 9 IVs
 set.seed(1234)
@@
 ## initialize
-nsim <- 100
-se_update_BMI <- c()
-se_update_GBC <- c()
-
+nsim <- 25
 est <- tibble(no_sim = 1:nsim)
@@
 for (i in 1:length(colnames_pval)) {
-  T0 = proc.time()[3]
  threshold <- 5*10^-6
+ se_update_BMI <- c()
+ se_update_GBC <- c()
+
  numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
+
  if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
@@
  }
+ rm(se_update_BMI, se_update_GBC)
  print(paste0("|||-----------------------Run finished for sample size: ", n[i], " -----------------------|||"))
 }
+
 T1 = proc.time()[3]
 timediff = T1-T0
-```
+
+saveRDS(est, paste0(Sys.Date(), "_est.rds"))
+```

### 2022-11-16 scripts/MSc.Rmd
added comments from last meeting

@@
 
+-   2022-11-15: 
+    1. Check if formula PVE numeric can be used for binary traits as well. Can the $\%var(X)$ – numeric – be compared with $\%var(U)$ $\%var(Y)$ – binary?
+    2. How is $\%var$ calculated in the MRbase package in the `mr_steiger` function (Steiger filtering)?
+    3. Prefer new formula for $\%var(X)$ calculation
+    4. FTO gene: Add SNP associated with FTO gene to analysis that was considered ambiguous. Check the beta estimate from UK Biobank (provide rsID to Felix: `rs1558902`)
+    5. Try also $K_{GBC}= 0.05/100$ for $\%var(Y)$ (before used $ASR=1.2/100.000$)
+    6. Update „Figure 1“ to 6 parts
+    7. Also run BMI -> GSD, FTO -> GSD + BMI -> GBC, FTO->GBC: How different are the MR methods in estimating the causal estimate?
+    8. Reference estimates for simulation: Wald estimate (FTO variant) + IVW total sample + conmix total sample
+    9. Plot PVE from all simulated variants as boxplots on each dot of Figure 01
+
 # Version Update
@@
 
-- First analysis
+- Issues from discussion mentioned above
 
@@
          )
+```
+
+### %var(X) BMI formula comparison
+
+```{r}
+ggplot(data=my_data_harm, aes(x=explained_variance_BMI, y=explained_variance_BMI2)) +
+  geom_point(alpha=0.6)+
+  geom_abline()+
+  theme_few()
 ```

### 2022-11-17 scripts/MSc.Rmd
added performance measures

@@
   kable_classic(full_width = FALSE, html_font = "Cambria")
+
+saveRDS(est, paste0("../output/Rdata/", Sys.Date(), "_est.rds"))
 ```
@@
 nsim <- 25
-est <- tibble(no_sim = 1:nsim)
-#est <- matrix(NA, 
+est_sim <- tibble(no_sim = 1:nsim)
+#est_sim <- matrix(NA, 
 #              nrow = nsim, 
@@
 #              ) 
-#colnames(est) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))
+#colnames(est_sim) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))
 
@@
    ## store results
-   est[k, colname_theta] <- res$Estimate
+   est_sim[k, colname_theta] <- res$Estimate
    CIlength = res$CIUpper-res$CILower
    if (length(CIlength)>1) print("conmix multimodal")
-   est[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
+   est_sim[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
    rm(res)
@@
 
-saveRDS(est, paste0(Sys.Date(), "_est.rds"))
-```
+saveRDS(est_sim, paste0("../output/Rdata/", Sys.Date(), "_est_sim.rds"))
+est_sim <- readRDS(file = "../output/Rdata/2022-11-15_est_sim.rds")
+```
+
+### Performance Metrics
+
+```{r}
+bias <- function(thetaobs,theta) mean(thetaobs)- theta
+mse <- function(thetaobs, theta) mean((thetaobs-theta)^2)
+
+bias(thetaobs = est_sim$theta_sim_50000, theta = est$estimate[est$mr_methods == "conmix"])
+mse(thetaobs = est_sim$theta_sim_50000, theta = est$estimate[est$mr_methods == "conmix"])
+```

### 2022-11-18 scripts/MSc.Rmd
performance measures bias+mse

@@
    res = MendelianRandomization::mr_conmix(mr.obj)
-   colname_theta <- paste0("theta_sim_", n[i])
-   colname_theta_se <- paste0("theta_se_sim_", n[i])
+   colname_theta <- paste0("theta_sim_", format(n[i], scientific=FALSE))
+   colname_theta_se <- paste0("theta_se_sim_", format(n[i], scientific=FALSE))
    ## store results
@@
  rm(se_update_BMI, se_update_GBC)
- print(paste0("|||-----------------------Run finished for sample size: ", n[i], " -----------------------|||"))
+ print(paste0("|||-----------------------Run finished for sample size: ", format(n[i], scientific=FALSE), " -----------------------|||"))
 }
@@
 
-```{r}
+```{r calculate-performance-metrics}
 bias <- function(thetaobs,theta) mean(thetaobs)- theta
@@
 
-bias(thetaobs = est_sim$theta_sim_50000, theta = est$estimate[est$mr_methods == "conmix"])
-mse(thetaobs = est_sim$theta_sim_50000, theta = est$estimate[est$mr_methods == "conmix"])
-```
+colname_theta <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "theta_sim")]
+colname_theta_se <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "theta_se_sim")]
+
+# filter out sample sizes
+sample_size_filtered <- as.numeric(str_extract(colname_theta, pattern = "(\\d)+"))
+
+## get performance metrics
+funs <- c(bias = bias, mse = mse)
+args <- list(theta = est$estimate[est$mr_methods == "conmix"]) # trim argument non-matching and ignored for median
+
+performance_metrics <- est_sim %>% 
+  select(!!!(colname_theta)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+```
+
+```{r plot-performance-metrics}
+#BIAS
+ggplot(performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()
+
+#MSE
+ggplot(performance_metrics, aes(x=N, y=mse))+ 
+  geom_point()
+
+# compare distribution of simulated thetas against theta_324000
+boxplot_dat <- est_sim %>% 
+  select(!!!(colname_theta)) %>% 
+  pivot_longer(cols = everything(), names_to = "theta_sim")
+
+ggplot(boxplot_dat, aes(x=theta_sim, y=value))+
+  geom_boxplot()+
+  coord_flip()
+
+```

### 2022-11-20 scripts/MSc.Rmd
performance plots
ggarrange plot

@@
     # Add a second axis and specify its features
-    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
+    sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
+    limits = c(0,5)
   ) + 
@@
       # Add a second axis and specify its features
-      sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
+      sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
+      limits = c(0,2)
     ) + 
@@
          )
+
+  # store plot in global environment
+  plotname <<- paste0("figure01_EFFECT_SIZE_", effect_size, "_VARIANTS_", maf)
+  assign(plotname, figure01.quadrant, envir = .GlobalEnv)
+  #list2env(assign(plotname, figure01.quadrant), envir = .GlobalEnv)
+
 }
@@
 
+```
+
+## FINAL PLOT
+
+```{r fig.height=10, fig.width=14, warning=FALSE}
+ggpubr::ggarrange(fig01b.update, fig02.update, 
+                  figure01_EFFECT_SIZE_strong_VARIANTS_common, figure01_EFFECT_SIZE_strong_VARIANTS_rare,
+                  figure01_EFFECT_SIZE_weak_VARIANTS_common, figure01_EFFECT_SIZE_weak_VARIANTS_rare, 
+                  ncol=2, nrow=3)
 ```
@@
 
+runSimulation <- FALSE
+
 ## START SIM
+if (runSimulation==TRUE) {
 T0 = proc.time()[3]
@@
 saveRDS(est_sim, paste0("../output/Rdata/", Sys.Date(), "_est_sim.rds"))
-est_sim <- readRDS(file = "../output/Rdata/2022-11-15_est_sim.rds")
+}
+
+# read dat
+est_sim <- readRDS(file = "../output/Rdata/2022-11-18_est_sim.rds")
 ```
@@
 
+### Performance plots
+
 ```{r plot-performance-metrics}
@@
 ggplot(performance_metrics, aes(x=N, y=bias))+ 
-  geom_point()
+  geom_point()+
+  scale_x_continuous(labels = comma)+
+  theme_few()
 
@@
 ggplot(performance_metrics, aes(x=N, y=mse))+ 
-  geom_point()
+  geom_point()+
+  scale_x_continuous(labels = comma)+
+  theme_few()
 
@@
   select(!!!(colname_theta)) %>% 
-  pivot_longer(cols = everything(), names_to = "theta_sim")
-
-ggplot(boxplot_dat, aes(x=theta_sim, y=value))+
+  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+  mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))
+
+ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
   geom_boxplot()+
-  coord_flip()
-
-```
+  geom_jitter(width=0.1,alpha=0.2)+
+  #scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "200000", "300000"), )+ 
+  ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
+  labs(title = "Estimates from beta simulation",
+       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
+       )+
+  coord_flip()+
+  theme_few()
+```

### 2022-11-20 scripts/MSc.Rmd
finalized plot 1 grid (1x2)+(1x4)

@@
 library(ggthemes)
+library(ggpubr)
 library(scales)
-library(ggthemes)
 #library(biomaRt)  # different script
@@
 ```{r fig.height=10, fig.width=14, warning=FALSE}
-ggpubr::ggarrange(fig01b.update, fig02.update, 
-                  figure01_EFFECT_SIZE_strong_VARIANTS_common, figure01_EFFECT_SIZE_strong_VARIANTS_rare,
-                  figure01_EFFECT_SIZE_weak_VARIANTS_common, figure01_EFFECT_SIZE_weak_VARIANTS_rare, 
-                  ncol=2, nrow=3)
+fig01b.update + 
+  labs(title= "Figure 1", caption = "") -> fig01b.update.arrange
+
+fig02.update+
+  labs(title= "Genetic architectures", subtitle="", caption = "") -> fig02.update.arrange
+
+figure01_EFFECT_SIZE_strong_VARIANTS_common+
+  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange
+
+figure01_EFFECT_SIZE_strong_VARIANTS_rare+
+  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange
+
+figure01_EFFECT_SIZE_weak_VARIANTS_common+
+  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange
+
+figure01_EFFECT_SIZE_weak_VARIANTS_rare+
+  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange
+
+ggpubr::ggarrange(fig01b.update.arrange, fig02.update.arrange,
+                  gfigure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
+                  figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange, 
+                  ncol=2, nrow=3, 
+                  heights = c(2, 1.5, 1.5),
+                  legend = "bottom", common.legend = TRUE, align = "hv"
+                  ) -> final_plot1.arrange
+
+figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange+
+  scale_y_continuous(name="", sec.axis = sec_axis(~./scale_factor, name=""), limits = c(0,2))
+
+    scale_y_continuous(
+      # Features of the first axis
+      name = "% variance explained by IVs",
+      # Add a second axis and specify its features
+      sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
+      limits = c(0,2)
+    ) +
+
+ggpubr::ggarrange(ggarrange(fig01b.update.arrange, fig02.update.arrange, ncol=2, labels = c("A", "B"), legend = "none", align = "hv"),
+                  ggarrange(figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
+                  figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange, 
+                  ncol=4, labels = c("C", "D", "E", "F"), legend="bottom", common.legend = TRUE),
+                  nrow=2
+                  #heights = c(2, 1.5, 1.5),
+                  #legend = "bottom", common.legend = TRUE, align = "hv"
+                  ) -> final_plot1.arrange2
+
+final_plot1.arrange
+final_plot1.arrange2
+
+filename_final_plot1 <- paste0("../output/plots/", Sys.Date(), "_final_plot1.arrange.png")
+filename_final_plot1b <- paste0("../output/plots/", Sys.Date(), "_final_plot1b.arrange.png")
+
+ggsave(filename = filename_final_plot1, plot = final_plot1.arrange, 
+         width = 14, height = 10, 
+         units = "in"  # default
+         )
+ggsave(filename = filename_final_plot1b, plot = final_plot1.arrange2, 
+         width = 14, height = 10, 
+         units = "in"  # default
+         )
+
 ```

### 2022-11-22 scripts/MSc.Rmd
outsourced fns

@@
 source("MR_lasso.R")  # Slob&Burgess 2020
+source("PVE_calculation_fns.R") 
 conflict_prefer("select", "dplyr")
@@
 ```{r calculate-explained-variance}
-explained_variance_numeric <- function(eaf, beta) {
-  # eaf = allele frequency
-  # beta = additive genetic effect
-  explained_variance_numeric <- 2*eaf *(1-eaf) * beta^2
-  return(explained_variance_numeric)
-}
-
-explained_variance_numeric2 <- function(maf, beta, se_beta, samplesize) {
-  ### Teslovich Nature 2010
-  # 1) maf = minor allele frequency
-  # 2) beta = additive genetic effect
-  # 3) se_beta = standard error of beta estimate
-  # 4) samplesize = samplesize of study
-  maf <- ifelse(maf>0.5, 1-maf, maf)
-
-  explained_variance_numeric <- (2*beta^2*maf*(1-maf))/ 
-    ( 2*beta^2*maf*(1-maf) + (se_beta^2)*2*samplesize*maf*(1-maf) )
-  return(explained_variance_numeric)
-}
-
-explained_variance_binary <- function(PA,RR1,RR2,K) {
-  # from So et al 2011
-  # Calculates the variance in liability explained by a single biallelic loci.
-  # PA: allele frequency of the risk allele (denoted A)
-  # RR1: relative risk of Aa (one risk allele) compared to aa (no risk allele)
-  # RR2: relative risk of AA (two risk alleles) compared to aa (no risk allele)
-  # K:  overall probability of disease in population
-  # Returns the variance explained (Vg) and the mean liability for each genotype (the overall liability is normalized to mean 0 and variance 1)
-  Paa = (1-PA)^2
-  PAa = 2*PA*(1-PA)
-  PAA = PA^2
-  muaa=0
-  faa= K/(Paa + PAa*RR1 + PAA*RR2)
-  fAa= RR1*faa
-  fAA= RR2*faa 
-  T = qnorm(1-faa) 
-  muAa = T-qnorm(1-fAa)
-  muAA = T-qnorm(1-fAA)
-  mean.all= PAa*muAa+ PAA*muAA
-  Vg= Paa*(muaa-mean.all)^2 + PAa*(muAa-mean.all)^2+ PAA*(muAA-mean.all)^2
-  actual.Vg =  Vg/(1+Vg) 
-  VR = 1-actual.Vg 
-  actual.T = Paa*sqrt(VR)*qnorm(1-faa) + PAa*sqrt(VR)*qnorm(1-fAa) + PAA*sqrt(VR)*qnorm(1-fAA)
-  actual.muaa = actual.T - sqrt(VR) * qnorm(1-faa)
-  actual.muAa = actual.T - sqrt(VR) * qnorm(1-fAa)
-  actual.muAA = actual.T - sqrt(VR) * qnorm(1-fAA)
-
-  res <- list(Vg=actual.Vg,muaa=actual.muaa, muAa = actual.muAa, muAA=actual.muAA)
-  res
-}
-
-## calculate explained variance
-my_data_harm <- my_data_harm %>% 
-  mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
-         explained_variance_BMI2 = explained_variance_numeric2(maf=eaf.bmi, 
-                                                           beta=beta.bmi, 
-                                                           se_beta=se.bmi, 
-                                                           samplesize=n[length(n)]
-
-         ),
-         explained_variance_GSD = explained_variance_binary(PA = eaf.gsd, 
-                                                            RR1 = exp(beta.gsd), 
-                                                            RR2 = exp(beta.gsd)^2, 
-                                                            K = 0.1  # assumed prevalence in population
-                                                            )$Vg,
-         explained_variance_GBC = explained_variance_binary(PA = eaf.gbc, 
-                                                            RR1 = exp(beta.gbc), 
-                                                            RR2 = exp(beta.gbc)^2, 
-                                                            K =1.2/100000  # assumed prevalence in population
-                                                            )$Vg
-         )
+# outsourced to "./scripts/PVE_calculation_fns.R"
 ```
@@
 ggpubr::ggarrange(fig01b.update.arrange, fig02.update.arrange,
-                  gfigure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
+                  figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
                   figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange, 
@@
 
+figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange+
+  # delete y-labels
+  scale_y_continuous(name="% variance explained by IVs", sec.axis = sec_axis(~./scale_factor, name=""), limits = c(0,2)) -> figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange
+
 figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange+
-  scale_y_continuous(name="", sec.axis = sec_axis(~./scale_factor, name=""), limits = c(0,2))
-
-    scale_y_continuous(
-      # Features of the first axis
-      name = "% variance explained by IVs",
-      # Add a second axis and specify its features
-      sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
-      limits = c(0,2)
-    ) +
+  # delete y-labels
+  scale_y_continuous(name="", sec.axis = sec_axis(~./scale_factor, name=""), limits = c(0,2)) -> figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange
+
+figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange+
+  # delete y-labels
+  scale_y_continuous(name="",sec.axis = sec_axis(~./scale_factor, name=""), limits = c(0,2)) -> figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange
+
+figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange+
+  # delete y-labels
+  scale_y_continuous(name="", sec.axis = sec_axis(~./scale_factor, name="Number of IVs"), limits = c(0,2)) -> figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange
 
@@
 saveRDS(est, paste0("../output/Rdata/", Sys.Date(), "_est.rds"))
+
 ```
@@
 ```{r simulatesimulatesimulate, warning=FALSE}
+
+# data FOR simulation in simulation script
+saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
+sim_vars <- list(n=n, colnames_pval=colnames_pval)
+saveRDS(sim_vars, paste0("../output/Rdata/", Sys.Date(), "_sim_vars.rds"))
+
 # try starting with n=50.000 -> selected 9 IVs
@@
 
- # simulate beta values from normal_distribution
+ # simulate beta values from normal_distribution # -----------------------------------------------------
  for (k in 1:nsim) {
@@
      mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
-     mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y)))
+     mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
+     mutate(#explained_variance_X_sim = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
+         explained_variance_X2_sim = explained_variance_numeric2(maf=eaf.bmi, 
+                                                           beta=beta.bmi_sim_norm, 
+                                                           se_beta=se_update_BMI, 
+                                                           samplesize=n[i]
+         ),
+        # explained_variance_GSD = explained_variance_binary(PA = eaf.gsd, 
+        #                                                    RR1 = exp(beta.gsd), 
+        #                                                    RR2 = exp(beta.gsd)^2, 
+        #                                                    K = 0.1  # assumed prevalence in population
+        #                                                    )$Vg,
+         explained_variance_Y_sim = explained_variance_binary(PA = eaf.gbc, 
+                                                            RR1 = exp(beta.gbc_sim_norm), 
+                                                            RR2 = exp(beta.gbc_sim_norm)^2, 
+                                                            K =1.2/100000  # assumed prevalence in population
+                                                            )$Vg
+         )
 
@@
    colname_theta_se <- paste0("theta_se_sim_", format(n[i], scientific=FALSE))
+   colname_pve_X <- paste0("pve_X_sim_", format(n[i], scientific=FALSE))
+   colname_pve_Y <- paste0("pve_Y_sim_", format(n[i], scientific=FALSE))
+
    ## store results
@@
    est_sim[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
+   # add PVE cols
+   est_sim[k, colname_pve_X] <- sum(sim_dat$explained_variance_X2_sim)
+   est_sim[k, colname_pve_Y] <- sum(sim_dat$explained_variance_Y_sim)
+
    rm(res)

### 2022-11-22 scripts/PVE_calculation_fns.R

@@ -1,0 +1,54 @@
+
+
+
+# PVE FORMULAS ------------------------------------------------------------
+explained_variance_numeric <- function(eaf, beta) {
+  # eaf = allele frequency
+  # beta = additive genetic effect
+  explained_variance_numeric <- 2*eaf *(1-eaf) * beta^2
+  return(explained_variance_numeric)
+}
+
+explained_variance_numeric2 <- function(maf, beta, se_beta, samplesize) {
+  ### Teslovich Nature 2010
+  # 1) maf = minor allele frequency
+  # 2) beta = additive genetic effect
+  # 3) se_beta = standard error of beta estimate
+  # 4) samplesize = samplesize of study
+  maf <- ifelse(maf>0.5, 1-maf, maf)
+  
+  explained_variance_numeric <- (2*beta^2*maf*(1-maf))/ 
+    ( 2*beta^2*maf*(1-maf) + (se_beta^2)*2*samplesize*maf*(1-maf) )
+  return(explained_variance_numeric)
+}
+
+explained_variance_binary <- function(PA,RR1,RR2,K) {
+  # from So et al 2011
+  # Calculates the variance in liability explained by a single biallelic loci.
+  # PA: allele frequency of the risk allele (denoted A)
+  # RR1: relative risk of Aa (one risk allele) compared to aa (no risk allele)
+  # RR2: relative risk of AA (two risk alleles) compared to aa (no risk allele)
+  # K:  overall probability of disease in population
+  # Returns the variance explained (Vg) and the mean liability for each genotype (the overall liability is normalized to mean 0 and variance 1)
+  Paa = (1-PA)^2
+  PAa = 2*PA*(1-PA)
+  PAA = PA^2
+  muaa=0
+  faa= K/(Paa + PAa*RR1 + PAA*RR2)
+  fAa= RR1*faa
+  fAA= RR2*faa 
+  T = qnorm(1-faa) 
+  muAa = T-qnorm(1-fAa)
+  muAA = T-qnorm(1-fAA)
+  mean.all= PAa*muAa+ PAA*muAA
+  Vg= Paa*(muaa-mean.all)^2 + PAa*(muAa-mean.all)^2+ PAA*(muAA-mean.all)^2
+  actual.Vg =  Vg/(1+Vg) 
+  VR = 1-actual.Vg 
[…]

### 2022-11-22 scripts/simstudy.R

@@ -1,0 +1,151 @@
+
+
+## ---------------------------
+##
+## Script name: Simulation of betas and sample sizes
+##
+## Author: Christoph Reich
+## Date Created: 2022-11-22
+##
+## Copyright (c) Christoph Reich, 2022
+## Email: christoph.reich@med.uni-heidelberg.de
+##
+## ---------------------------
+## Notes:
+##   
+## ---------------------------
+
+
+# set to TRUE if simulation should run when script is sourced
+runSimulation <- FALSE
+
+library(dplyr)
+library(purrr)
+library(tidyr)
+library(stringr)
+library(tibble)
+library(MendelianRandomization)
+library(TwoSampleMR)
+library(mr.raps)
+library(MRMix)
+library(MRPRESSO)
+library(penalized)
+library(conflicted)
+source("scripts/MR_lasso.R")  # Slob&Burgess 2020
+source("scripts/PVE_calculation_fns.R")
+conflict_prefer("select", "dplyr")
+conflict_prefer("filter", "dplyr")
+
+
+
+# GET DATA ----------------------------------------------------------------
+# dataa
+my_data_harm <- readRDS(file = "./output/Rdata/2022-11-22_my_data_harm.rds")
+# MR estimates from full study
+est <- readRDS(file = "./output/Rdata/2022-11-22_est.rds")
+# vars from MSc script
+sim_vars <- readRDS(file = "./output/Rdata/2022-11-22_sim_vars.rds")
+n <- sim_vars$n
+colnames_pval <- sim_vars$colnames_pval
+
+###
+# SIMULATION --------------------------------------------------------------
+###
+set.seed(1234)
+## initialize
+nsim <- 25
+est_sim <- tibble(no_sim = 1:nsim)
+
+## START SIM
+if (runSimulation==TRUE) {
+  T0 = proc.time()[3]
+  for (i in 1:length(colnames_pval)) {
+    threshold <- 5*10^-6
+    se_update_BMI <- c()
+    se_update_GBC <- c()
+    
+    numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
+    
+    if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
+      next
+    }
+    
+    # FILTER DATA
+    sim_dat <- my_data_harm %>% 
+      filter(!!sym(colnames_pval[i]) < threshold) %>% 
[…]

### 2022-11-22 scripts/MSc.Rmd
boxplots overlay

@@
 
+We simulate
+
 ```{r simulatesimulatesimulate, warning=FALSE}
@@
 
-# try starting with n=50.000 -> selected 9 IVs
-set.seed(1234)
-
-## initialize
-nsim <- 25
-est_sim <- tibble(no_sim = 1:nsim)
-#est_sim <- matrix(NA, 
-#              nrow = nsim, 
-#              ncol = 2*length(n)  # theta_sim and theta_se_sim
-#              ) 
-#colnames(est_sim) <- c(paste0("theta_sim_", n), paste0("theta_se_sim_", n))
-
-runSimulation <- FALSE
-
-## START SIM
-if (runSimulation==TRUE) {
-T0 = proc.time()[3]
-for (i in 1:length(colnames_pval)) {
- threshold <- 5*10^-6
- se_update_BMI <- c()
- se_update_GBC <- c()
-
- numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
-
- if (numIV<5) { # start with 5 IVs only (sample size ~50,000)
-   next
- }
-
- # FILTER DATA
- sim_dat <- my_data_harm %>% 
-   filter(!!sym(colnames_pval[i]) < threshold) %>% 
-   select(SNP, eaf.bmi, beta.bmi, p.value.bmi, eaf.gbc, beta.gbc, p.value.gbc)
-
- # get simulated se_{sample size}
- for (j in 1:nrow(sim_dat)) {
-   # as done before in section "simulating on empirical data"
-   se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
-   se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
- }
-
- sim_dat <- sim_dat %>%
-     bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))
-
- # simulate beta values from normal_distribution # -----------------------------------------------------
- for (k in 1:nsim) {
-   sim_dat <- sim_dat %>% 
-     mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
-     mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se_update_GBC, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
-     mutate(#explained_variance_X_sim = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
-         explained_variance_X2_sim = explained_variance_numeric2(maf=eaf.bmi, 
-                                                           beta=beta.bmi_sim_norm, 
-                                                           se_beta=se_update_BMI, 
-                                                           samplesize=n[i]
-         ),
-        # explained_variance_GSD = explained_variance_binary(PA = eaf.gsd, 
-        #                                                    RR1 = exp(beta.gsd), 
-        #                                                    RR2 = exp(beta.gsd)^2, 
-        #                                                    K = 0.1  # assumed prevalence in population
-        #                                                    )$Vg,
-         explained_variance_Y_sim = explained_variance_binary(PA = eaf.gbc, 
-                                                            RR1 = exp(beta.gbc_sim_norm), 
-                                                            RR2 = exp(beta.gbc_sim_norm)^2, 
-                                                            K =1.2/100000  # assumed prevalence in population
-                                                            )$Vg
-         )
-
-   # PERFORM MR_analysis
-   mr.obj = MendelianRandomization::mr_input(
-         bx = sim_dat$beta.bmi_sim_norm, 
-         bxse = sim_dat$se_update_BMI, 
-         # outcome
-         by = sim_dat$beta.gbc_sim_norm, 
-         byse = sim_dat$se_update_GBC, 
-         snps = sim_dat$SNP,
-         exposure = "Body mass index",
-         outcome = "Gallbladder Cancer"
-       ) 
-   # CONMIX (as reference)
-   res = MendelianRandomization::mr_conmix(mr.obj)
-   colname_theta <- paste0("theta_sim_", format(n[i], scientific=FALSE))
-   colname_theta_se <- paste0("theta_se_sim_", format(n[i], scientific=FALSE))
-   colname_pve_X <- paste0("pve_X_sim_", format(n[i], scientific=FALSE))
-   colname_pve_Y <- paste0("pve_Y_sim_", format(n[i], scientific=FALSE))
-
-   ## store results
-   est_sim[k, colname_theta] <- res$Estimate
-   CIlength = res$CIUpper-res$CILower
-   if (length(CIlength)>1) print("conmix multimodal")
-   est_sim[k, colname_theta_se] <- sum(CIlength)/1.96/2  ## Caution: this may be problematic 
-   # add PVE cols
-   est_sim[k, colname_pve_X] <- sum(sim_dat$explained_variance_X2_sim)
-   est_sim[k, colname_pve_Y] <- sum(sim_dat$explained_variance_Y_sim)
-
-   rm(res)
- }
- rm(se_update_BMI, se_update_GBC)
- print(paste0("|||-----------------------Run finished for sample size: ", format(n[i], scientific=FALSE), " -----------------------|||"))
-}
-
-T1 = proc.time()[3]
-timediff = T1-T0
-
-saveRDS(est_sim, paste0("../output/Rdata/", Sys.Date(), "_est_sim.rds"))
-}
-
-# read dat
+## SIMULATION OUTSOURCED TO "./scripts/simstudy.R"
+
+# read dat from "simstudy.R"
 est_sim <- readRDS(file = "../output/Rdata/2022-11-18_est_sim.rds")
+
+est_sim <- readRDS(file = "../output/Rdata/2022-11-22_est_sim_NSIM_5.rds")
 ```
@@
 colname_theta_se <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "theta_se_sim")]
+colname_pve_X <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_X_sim")]
+colname_pve_Y <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_Y_sim")]
 
@@
   geom_jitter(width=0.1,alpha=0.2)+
-  #scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "200000", "300000"), )+ 
+  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = "300000", y = args$theta+0.5,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+  scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "150000", "200000", "250000", "300000"), )+ 
   ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
@@
 ```
+
+```{r}
+# EXPOSURE PVE
+est_sim %>% 
+  select(!!!(colname_pve_X)) %>% 
+  pivot_longer(everything(), names_to = "PVE_X", values_to = "value_X") %>% 
+  cbind(sample_size_filtered) %>% as_tibble() -> pve_X_sim
+
+est_sim %>% 
+  select(!!!(colname_pve_Y)) %>% 
+  pivot_longer(everything(), names_to = "PVE_Y", values_to = "value_Y") %>% 
+  cbind(sample_size_filtered) %>% as_tibble() -> pve_Y_sim
+
+boxplot_overlay_dat <- pve_X_sim %>% 
+  cbind(pve_Y_sim %>% select(-sample_size_filtered)) %>% 
+  as_tibble()
+
+fig01b.update +
+  geom_boxplot(data=boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, group=sample_size_filtered))
+
+  geom_boxplot(data = test, aes(x=sample_size_filtered, y=value*100, # *100 to get percent values
+                                group=sample_size_filtered), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.6)+
+
+ggplot(test, aes(x=as.factor(sample_size_filtered), y=value))+
+  geom_boxplot()
+
+plot_data1b <- df_cum_expl_var %>% 
+  pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, cum_expl_varY)) %>% 
+  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
+         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
+         )
+## plot
+ggplot(plot_data1b, aes(x=n, y=value, col=name))+
+  geom_point(alpha=0.6)+
+  geom_line(alpha=0.6)+
+  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
+  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
+                     labels = comma, 
+                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
+                     limits = c(10000, 300000))+
+  #ylab("% variance explained by IVs")+
+  labs(title = "Figure 1",
+       col='', 
+       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
+       ) +
+  scale_y_continuous(
+    # Features of the first axis
+    name = "% variance explained by IVs",
+    # Add a second axis and specify its features
+    sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
+    limits = c(0,5)
+  ) + 
+  #geom_boxplot(data=boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, group=sample_size_filtered))
+  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, # *100 to get percent values
+                                group=sample_size_filtered), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2)+
+  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_Y*100, # *100 to get percent values
+                                group=sample_size_filtered), 
+               col=ggthemes_data$few$colors$Dark[4,2][[1]],
+               alpha=0.2, size=0.2)+
+  theme_few()+
+  #scale_colour_few() +
+  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
+  #                                         "% var(X)" = "blue", 
+  #                                         "% var(Y)" = "darkblue", 
+  #                                         "Number of IVs" = "black"
+  #                                         ))
+  #
+  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
+                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
+                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
+                                           "Number of IVs" = "black"),
+                     # order
+                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
+  )
+```

### 2022-11-22 scripts/plotfigure01fn.R

@@ -1,0 +1,166 @@
+
+
+
+
+plotfigure01 <- function(data, effect_size="all", maf="all"){
+  ## data: my_data_harm
+  # effect size: "all", "strong" or "weak"
+  # maf: "all", "rare" or "common
+  df_cum_expl_var <- tibble(n=n, 
+                            cum_expl_varX = rep(NA, length(n)), 
+                            cum_expl_varX2 = rep(NA, length(n)), 
+                            cum_expl_varU = rep(NA, length(n)), 
+                            cum_expl_varY = rep(NA, length(n)),
+                            count_IVs = rep(NA, length(n))
+  )
+  # create maf col
+  data <- data %>% 
+    mutate(maf.bmi = ifelse(eaf.bmi >0.5, 1-eaf.bmi, eaf.bmi))
+  
+  ## filter for quadrant
+  if (effect_size=="strong" & maf == "rare") {
+    data <- data %>% 
+      filter(abs(beta.bmi) > median(abs(beta.bmi)) &  maf.bmi < median(maf.bmi))
+  } else if (effect_size=="strong" & maf == "common") {
+    data <- data %>% 
+      filter(abs(beta.bmi) > median(abs(beta.bmi)) &  maf.bmi > median(maf.bmi))
+  } else if (effect_size=="weak" & maf == "common") {
+    data <- data %>% 
+      filter(abs(beta.bmi) < median(abs(beta.bmi)) &  maf.bmi > median(maf.bmi))
+  } else if (effect_size=="weak" & maf == "rare") {
+    data <- data %>% 
+      filter(abs(beta.bmi) < median(abs(beta.bmi)) &  maf.bmi < median(maf.bmi))
+  }
+  
+  # create data for plot
+  for (i in 1:length(colnames_pval) ) {
+    threshold <- 5*10^(-6)
+    select_col <- colnames_pval[i]
+    # calculate cum explained variance and count of IVs
+    summarydat <- data %>% 
+      filter(!!sym(select_col) < threshold) %>% 
+      select(explained_variance_BMI, explained_variance_BMI2, explained_variance_GSD, explained_variance_GBC) %>% 
+      summarise(expl_variances =across(.cols = everything(), function(x) sum(x)),
+                count_IVs = n())
+    # fill up df
+    df_cum_expl_var[i, 2] <- summarydat$expl_variances[1]
[…]

### 2022-11-22 scripts/simstudy.R

@@ -17,7 +17,7 @@
 
 
 # set to TRUE if simulation should run when script is sourced
-runSimulation <- FALSE
+runSimulation <- TRUE
 
 library(dplyr)
 library(purrr)
@@ -47,13 +47,14 @@ est <- readRDS(file = "./output/Rdata/2022-11-22_est.rds")
 sim_vars <- readRDS(file = "./output/Rdata/2022-11-22_sim_vars.rds")
 n <- sim_vars$n
 colnames_pval <- sim_vars$colnames_pval
+n.orig <- 334487
 
 ###
 # SIMULATION --------------------------------------------------------------
 ###
 set.seed(1234)
 ## initialize
-nsim <- 25
+nsim <- 5
 est_sim <- tibble(no_sim = 1:nsim)
 
 ## START SIM
@@ -145,7 +146,7 @@ if (runSimulation==TRUE) {
   T1 = proc.time()[3]
   timediff = T1-T0
   
-  saveRDS(est_sim, paste0("../output/Rdata/", Sys.Date(), "_est_sim.rds"))
+  saveRDS(est_sim, paste0("./output/Rdata/", Sys.Date(), "_est_sim_NSIM_", nsim,".rds"))
 }
 
 

### 2022-11-23 scripts/MSc.Rmd
added MC4R association

@@
 filter_FTO <- rsIds_GeneNames %>% 
-  filter(str_detect(associated_gene, pattern = "FTO")) %>% 
+  filter(str_detect(associated_gene, pattern = "FTO") | str_detect(associated_gene, pattern = "MC4R")) %>% 
   distinct(refsnp_id, .keep_all = TRUE)
 
-t(filter_FTO) %>% kable() %>% 
+t(filter_FTO) %>% kable() #%>% 
   kable_styling(font_size=10)
@@
 my_data_harm <- as_tibble(cbind(my_data_harm, pvals_df))
-
-# 5. Calculate cumulative explained variances
-# 6. Represent figure similar to Fig1 in Chatterjee’s paper
 ```
@@
 # outsourced to "./scripts/PVE_calculation_fns.R"
+
+## CALCULATE explained variance
+my_data_harm <- my_data_harm %>% 
+  mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
+         explained_variance_BMI2 = explained_variance_numeric2(maf=eaf.bmi, 
+                                                           beta=beta.bmi, 
+                                                           se_beta=se.bmi, 
+                                                           samplesize=n[length(n)]
+
+         ),
+         explained_variance_GSD = explained_variance_binary(PA = eaf.gsd, 
+                                                            RR1 = exp(beta.gsd), 
+                                                            RR2 = exp(beta.gsd)^2, 
+                                                            K = 0.1  # assumed prevalence in population
+                                                            )$Vg,
+         explained_variance_GBC = explained_variance_binary(PA = eaf.gbc, 
+                                                            RR1 = exp(beta.gbc), 
+                                                            RR2 = exp(beta.gbc)^2, 
+                                                            K =1.2/100000  # assumed prevalence in population
+                                                            )$Vg,
+         explained_variance_GBC2 = explained_variance_binary(PA = eaf.gbc, 
+                                                            RR1 = exp(beta.gbc), 
+                                                            RR2 = 1+(2*(exp(beta.gbc)-1)), 
+                                                            K =1.2/100000  # assumed prevalence in population
+                                                            )$Vg
+
+         )
 ```
@@
 saveRDS(est, paste0("../output/Rdata/", Sys.Date(), "_est.rds"))
-
 ```
@@
 ```{r simulatesimulatesimulate, warning=FALSE}
-
 # data FOR simulation in simulation script
@@
 
-# read dat from "simstudy.R"
-est_sim <- readRDS(file = "../output/Rdata/2022-11-18_est_sim.rds")
-
-est_sim <- readRDS(file = "../output/Rdata/2022-11-22_est_sim_NSIM_5.rds")
+# read dat from "simstudy.R"  #------------------------------------------------
+est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_NSIM_", full.names = TRUE)))
 ```
@@
 
-```{r plot-performance-metrics}
+```{r plot-performance-metrics-bias-mse}
 #BIAS
@@
   scale_x_continuous(labels = comma)+
-  theme_few()
+  theme_few() -> biasplot
 
@@
   scale_x_continuous(labels = comma)+
-  theme_few()
-
+  theme_few() -> mseplot
+
+ggarrange(biasplot, mseplot) ->bias_mse_plot
+
+filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_bias_mse.png")
+ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+```
+
+```{r plot-performance-metrics-boxplot}
 # compare distribution of simulated thetas against theta_324000
@@
   coord_flip()+
-  theme_few()
-```
-
-```{r}
+  theme_few() -> simplot_betas
+simplot_betas
+
+filename_simplot_betas <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_betas.png")
+ggsave(filename = filename_simplot_betas, plot = simplot_betas, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+```
+
+```{r simulated-pves}
 # EXPOSURE PVE
@@
 
-fig01b.update +
-  geom_boxplot(data=boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, group=sample_size_filtered))
-
-  geom_boxplot(data = test, aes(x=sample_size_filtered, y=value*100, # *100 to get percent values
-                                group=sample_size_filtered), 
-               col=ggthemes_data$few$colors$Dark[3,2][[1]],
-               alpha=0.6)+
-
-ggplot(test, aes(x=as.factor(sample_size_filtered), y=value))+
-  geom_boxplot()
-
-plot_data1b <- df_cum_expl_var %>% 
-  pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, cum_expl_varY)) %>% 
-  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
-         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
-         )
 ## plot
-ggplot(plot_data1b, aes(x=n, y=value, col=name))+
+ggplot(plot_data1b, aes(x=n, y=value, col=name))+    # plot_data1b from section Simlating on empirical data (PVE)
   geom_point(alpha=0.6)+
@@
   #ylab("% variance explained by IVs")+
-  labs(title = "Figure 1",
+  labs(title = "PVE with simulation added",
        col='', 
@@
     sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
-    limits = c(0,5)
+    limits = c(0,10)
   ) + 
@@
                      breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
-  )
-```
+  ) -> plot_pve_simulated
+
+filename_pveplot_sim <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_pve.png")
+ggsave(filename = filename_pveplot_sim, plot = plot_pve_simulated, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+```

### 2022-11-23 scripts/simstudy.R

@@ -54,7 +54,7 @@ n.orig <- 334487
 ###
 set.seed(1234)
 ## initialize
-nsim <- 5
+nsim <- 25
 est_sim <- tibble(no_sim = 1:nsim)
 
 ## START SIM

### 2022-11-23 scripts/MSc.Rmd
addded performance measures (empSE, coverage, bias-eliminated-coverage, type1error)

@@
 
-t(filter_FTO) %>% kable() #%>% 
+t(filter_FTO) %>% kable() %>% 
   kable_styling(font_size=10)
@@
     count_IVs_scaled = count_IVs*scale_factor)
-
-# plot_data <- df_cum_expl_var %>% 
-#   pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, count_IVs_scaled)) %>% 
-#   mutate(name = factor(name, labels = c("Number of IVs", "% var(U)", "% var(X)")))
-# 
-# ggplot(plot_data, aes(x=log10(n), y=value, col=name))+
-#   geom_point(alpha=0.6)+
-#   geom_line(alpha=0.6)+
-#   xlab("log10(N)")+
-#   #ylab("% variance explained by IVs")+
-#   labs(title = "Figure 1",
-#        col='', 
-#        caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
-#        ) +
-#   scale_y_continuous(
-#     # Features of the first axis
-#     name = "% variance explained by IVs",
-#     # Add a second axis and specify its features
-#     sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
-#   ) + 
-#   theme_few()+
-#   scale_colour_few() ->fig01
-# fig01
-# 
-# plot_data_b <- df_cum_expl_var %>% 
-#   pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, count_IVs_scaled)) %>% 
-#   mutate(name = factor(name, labels = c("Number of IVs", "% var(U)", "% var(X)")))
-# 
-# ggplot(plot_data_b, aes(x=log10(n), y=value, col=name))+
-#   geom_point(alpha=0.6)+
-#   geom_line(alpha=0.6)+
-#   xlab("log10(N)")+
-#   #ylab("% variance explained by IVs")+
-#   labs(title = "Figure 1",
-#        col='', 
-#        caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
-#        ) +
-#   scale_y_continuous(
-#     # Features of the first axis
-#     name = "% variance explained by IVs",
-#     # Add a second axis and specify its features
-#     sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
-#   ) + 
-#   theme_few()+
-#   scale_colour_few() ->fig01_b
-# fig01_b
-# 
-# ggsave(filename = "../output/plots/figure01.png", plot = fig01_b, 
-#        width = 10, height = 6, 
-#        units = "in"  # default
-#        )
-# 
-# filenamefig01_a <- paste0("../output/plots/", Sys.Date(), "_figure01_a.png")
-# filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_b.png")
 ```
@@
 ```{r MendelianRandomization-analysis, warning=FALSE}
-mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
-est <- tibble(mr_methods=mr_methods, estimate=NA, se.estimate=NA, time=NA)
-numIV <- nrow(my_data_harm)
-
-if (numIV>2){
-        mr.obj = MendelianRandomization::mr_input(
-          # exposure
-          bx = my_data_harm$beta.bmi, 
-          bxse = my_data_harm$se.bmi, 
-          # outcome
-          by = my_data_harm$beta.gbc, 
-          byse = my_data_harm$se.gbc, 
-          snps = my_data_harm$SNP,
-          exposure = "Body mass index",
-          outcome = "Gallbladder Cancer"
-        )
-        # 1. IVW
-        T0 = proc.time()[3]
-        res = MendelianRandomization::mr_ivw(mr.obj)
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="IVW"] = res$Estimate
-        est$se.estimate[est$mr_methods=="IVW"] = res$StdError
-        est$time[est$mr_methods=="IVW"] = T1-T0
-        rm(res)
-
-        # 2. median
-        T0 = proc.time()[3]
-        res = MendelianRandomization::mr_median(mr.obj)
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="median"] = res$Estimate
-        est$se.estimate[est$mr_methods=="median"] = res$StdError
-        est$time[est$mr_methods=="median"] = T1-T0
-        rm(res)
-        # 3. mode
-        T0 = proc.time()[3]
-        res = MendelianRandomization::mr_mbe(mr.obj)
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="mode"] = res$Estimate
-        est$se.estimate[est$mr_methods=="mode"] = res$StdError
-        est$time[est$mr_methods=="mode"] = T1-T0
-        rm(res)
-        # 4.MR-PRESSO    ## CAVE: 5 minutes per round, consider only calculating for defined thresholds
-        T0 = proc.time()[3]
-        res <- MRPRESSO:: mr_presso(BetaOutcome = "beta.gbc", 
-                    BetaExposure = "beta.bmi", 
-                    SdOutcome = "se.gbc", 
-                    SdExposure = "se.bmi", 
-                    OUTLIERtest = TRUE, 
-                    DISTORTIONtest = TRUE, 
-                    seed = 20221114,
-                    data = as.data.frame(my_data_harm), ## must be a data.frame
-                    NbDistribution = 1000,   # must be changed
-                    SignifThreshold = 0.05)
-        T1 = proc.time()[3]
-        if (!is.na(res$`Main MR results`[2,"Causal Estimate"]) & !is.na(res$`Main MR results`[2,"Sd"])){
-                est$estimate[est$mr_methods=="PRESSO"] =  res$`Main MR results`[2,"Causal Estimate"]
-                est$se.estimate[est$mr_methods=="PRESSO"] = res$`Main MR results`[2,"Sd"]
-            } else{
-                est$estimate[est$mr_methods=="PRESSO"] =  res$`Main MR results`[1,"Causal Estimate"]
-                est$se.estimate[est$mr_methods=="PRESSO"] = res$`Main MR results`[1,"Sd"]
-            }
-        est$time[est$mr_methods=="PRESSO"] = T1-T0
-        rm(res)
-        # 5. robust
-        T0 = proc.time()[3]
-        res = MendelianRandomization::mr_ivw(mr.obj,"random", robust = TRUE)
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="Robust"] = res$Estimate
-        est$se.estimate[est$mr_methods=="Robust"] = res$StdError
-        est$time[est$mr_methods=="Robust"] = T1-T0
-        rm(res)
-        # 6. MR-Lasso
-        T0 = proc.time()[3]
-        res = MR_lasso(betaYG = my_data_harm$beta.gbc, betaXG = my_data_harm$beta.bmi, sebetaYG = my_data_harm$se.gbc)
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="Lasso"] = res$ThetaEstimate
-        est$se.estimate[est$mr_methods=="Lasso"] = res$ThetaSE
-        est$time[est$mr_methods=="Lasso"] = T1-T0
-        rm(res)
-        # 7. Egger
-        T0 = proc.time()[3]
-        res = MendelianRandomization::mr_egger(mr.obj)
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="egger"] = res$Estimate
-        est$se.estimate[est$mr_methods=="egger"] = res$StdError.Est
-        est$time[est$mr_methods=="egger"] = T1-T0
-        rm(res)
-        # 8. contamination mixture
-        T0 = proc.time()[3]
-        res = MendelianRandomization::mr_conmix(mr.obj)
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="conmix"] = res$Estimate
-        CIlength = res$CIUpper-res$CILower
-        if (length(CIlength)>1) print("conmix multimodal")
-        est$se.estimate[est$mr_methods=="conmix"] = sum(CIlength)/1.96/2 ## Caution: this may be problematic (Qi&Chatterjee2021)
-        est$time[est$mr_methods=="conmix"] = T1-T0
-        rm(res)
-        # 9. MRMix
-              # theta_temp_vec = seq(-0.5,0.5,by=0.01)
-        T0 = proc.time()[3]
-        res = MRMix::MRMix(betahat_x = my_data_harm$beta.bmi, 
-                          betahat_y = my_data_harm$beta.gbc, 
-                          sx =my_data_harm$se.bmi, 
-                          sy = my_data_harm$se.gbc)
-        res_se = MRMix::MRMix_se(betahat_x = my_data_harm$beta.bmi, 
-                          betahat_y = my_data_harm$beta.gbc, 
-                          sx =my_data_harm$se.bmi, 
-                          sy = my_data_harm$se.gbc, 
-                          theta = res$theta, # estimate of causal effect, assuming the summary stats are standardized, theta represents increase in mean value of Y in s.d.
-unit of Y (for cont outcomes) or log-OR of Y (for binary outcomes)
-                          pi0 = res$pi0, 
-                          sigma2 = res$sigma2
-                          )
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="MRMix"] = res$theta
-        est$se.estimate[est$mr_methods=="MRMix"] = res_se
-        est$time[est$mr_methods=="MRMix"] = T1-T0
-        rm(res, res_se)
-        # 10. MR-RAPS  
-        ## Warning: The estimated overdispersion parameter is very small. Consider using the simple model without overdispersion.
-        T0 = proc.time()[3]
-        res = mr.raps::mr.raps.overdispersed.robust(b_exp = my_data_harm$beta.bmi, 
-                                                    b_out = my_data_harm$beta.gbc, 
-                                                    se_exp = my_data_harm$se.bmi, 
-                                                    se_out = my_data_harm$se.gbc, 
-                                                    loss.function = "huber", 
-                                                    k = 1.345, 
-                                                    initialization = c("l2"), 
-                                                    suppress.warning = FALSE, 
-                                                    diagnosis = FALSE, 
-                                                    niter = 20, 
-                                                    tol = .Machine$double.eps^0.5)
-        T1 = proc.time()[3]
-        est$estimate[est$mr_methods=="RAPS"] = res$beta.hat
-        est$se.estimate[est$mr_methods=="RAPS"] = res$beta.se
-        est$time[est$mr_methods=="RAPS"] = T1-T0
-        rm(res)
-}
+# outsourced calculation to `perform_mr.R`
+# data FOR estimates in `perform_mr.R` script and simulation in `simstudy.R` script
+saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
+
+## load results
+est <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
 
@@
   kable_classic(full_width = FALSE, html_font = "Cambria")
-
-saveRDS(est, paste0("../output/Rdata/", Sys.Date(), "_est.rds"))
 ```
@@
 ```{r simulatesimulatesimulate, warning=FALSE}
-# data FOR simulation in simulation script
-saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
+# save simulation-vars for simulation `simstudy.R` script
 sim_vars <- list(n=n, colnames_pval=colnames_pval)
 saveRDS(sim_vars, paste0("../output/Rdata/", Sys.Date(), "_sim_vars.rds"))
-
 ## SIMULATION OUTSOURCED TO "./scripts/simstudy.R"
@@
 ```{r calculate-performance-metrics}
-bias <- function(thetaobs,theta) mean(thetaobs)- theta
-mse <- function(thetaobs, theta) mean((thetaobs-theta)^2)
+# fns
+bias <- function(thetaobs,theta, ...) mean(thetaobs)- theta
+mse <- function(thetaobs, theta, ...) mean((thetaobs-theta)^2)
+emp_se <- function(nsim, thetaobs, theta, ...) sqrt( 1/(nsim-1) * sum((thetaobs-mean(thetaobs))^2) )  # == sd
 
@@
 
-## get performance metrics
-funs <- c(bias = bias, mse = mse)
-args <- list(theta = est$estimate[est$mr_methods == "conmix"]) # trim argument non-matching and ignored for median
+## get performance metrics    # https://stackoverflow.com/questions/47415072/pass-multiple-functions-to-purrrmap
+funs <- c(mean = mean, bias = bias, mse = mse, emp_se=emp_se)
+args <- list(theta = est$estimate[est$mr_methods == "conmix"], nsim=nrow(est_sim)) # trim argument non-matching and ignored for median
 
@@
   rename(N = sample_size_filtered)
+
+# est_sim %>%    # equivalent to `emp_se()` function, but `sd()` misses dotdotdot argument 
+#   select(!!!(colname_theta)) %>% 
+#   map_dbl(.f = ~sd(.))
+```
+
+```{r further-metrics}
+# coverage
+coverage <- function(thetaobs, thetaobs_se, theta, ...) {
+  lower_ci <- thetaobs - thetaobs_se
+  upper_ci <- thetaobs + thetaobs_se
+  coverage <- mean(lower_ci <= theta & upper_ci >= theta)
+  return(coverage)
+}
+coverage_bias_eliminated <- function(thetaobs, thetaobs_se, ...) {
+  lower_ci <- thetaobs - thetaobs_se
+  upper_ci <- thetaobs + thetaobs_se
+  mean_thetaobs <- mean(thetaobs)
+  coverage_bias_eliminated <- mean(lower_ci <= mean_thetaobs & upper_ci >=mean_thetaobs)
+  return(coverage_bias_eliminated)
+}
+
+type1error <- function(pvals) sum(pvals<=0.05)/length(pvals)  # no causal_effect
+
+coverage_calc <- c()
+coverage_bias_eliminated_calc <- c()
+type1error_calc <- c()
+for (i in 1:length(sample_size_filtered)) {
+  sample_size <- sample_size_filtered[i]
+  col_theta <- paste0("theta_sim_", format(sample_size, scientific = FALSE))
+  col_theta_se <- paste0("theta_se_sim_", format(sample_size, scientific = FALSE))
+  col_mr_pval <- paste0("mr_pval_", format(sample_size, scientific = FALSE))
+
+  coverage_calc[i] <- coverage(thetaobs = est_sim[col_theta], thetaobs_se = est_sim[col_theta_se], theta = est$estimate[est$mr_methods == "conmix"])
+  coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_theta]], thetaobs_se = est_sim[[col_theta_se]])
+  type1error_calc[i] <- type1error(est_sim[[col_mr_pval]])
+}
+
+performance_metrics <- performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(coverage_calc, coverage_bias_eliminated_calc, type1error_calc)) %>% 
+  as_tibble()
 ```
@@
 
-```{r plot-performance-metrics-bias-mse}
-#BIAS
-ggplot(performance_metrics, aes(x=N, y=bias))+ 
-  geom_point()+
-  scale_x_continuous(labels = comma)+
-  theme_few() -> biasplot
-
-#MSE
-ggplot(performance_metrics, aes(x=N, y=mse))+ 
-  geom_point()+
-  scale_x_continuous(labels = comma)+
-  theme_few() -> mseplot
-
-ggarrange(biasplot, mseplot) ->bias_mse_plot
-
-filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_bias_mse.png")
-ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot, 
-       width = 10, height = 6, 
-       units = "in"  # default
-       )
-```
-
-```{r plot-performance-metrics-boxplot}
+```{r boxplot-theta-sim-distribution, fig.height=6, fig.width=10}
 # compare distribution of simulated thetas against theta_324000
@@
   coord_flip()+
-  theme_few() -> simplot_betas
-simplot_betas
-
-filename_simplot_betas <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_betas.png")
-ggsave(filename = filename_simplot_betas, plot = simplot_betas, 
+  theme_few() -> simplot_thetas
+simplot_thetas
+
+### same plot as above, but from already calculated performance metrics theta_mean + theta_sd
+ggplot(performance_metrics, aes(x=N, y=mean))+ 
+  geom_point()+
+  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 300000, y = args$theta+0.25,
+         label= expression( paste("ref: ", theta[X] ))  , 
+         color= ggthemes_data$few$colors$Dark[9,2][[1]]
+           )+
+  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
+                width=3000, # x-axis dimension
+                position = "dodge")+
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Mean thetas")+
+  labs(title = "Simulated Thetas with Empirical Standard Error")+
+  theme_few() -> simplot_thetas2
+simplot_thetas2
+
+filename_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_thetas.png")
+ggsave(filename = filename_simplot_thetas, plot = simplot_thetas, 
        width = 10, height = 6, 
+       units = "in"  # default
+       )
+filename_simplot_thetas2 <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_thetas2.png")
+ggsave(filename = filename_simplot_thetas2, plot = simplot_thetas2, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+```
+
+```{r plot-performance-metrics-bias-mse, fig.height=6, fig.width=14}
+#BIAS
+ggplot(performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()+
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Bias")+
+  labs(title = "Bias with Empirical Standard Error")+
+  theme_few() -> biasplot
+
+#MSE
+ggplot(performance_metrics, aes(x=N, y=mse))+ 
+  geom_point()+
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("MSE")+
+  labs(title = "MSE")+
+  theme_few() -> mseplot
+
+ggarrange(biasplot, mseplot) ->bias_mse_plot
+bias_mse_plot
+
+filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_bias_mse.png")
+ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot, 
+       width = 14, height = 6, 
        units = "in"  # default
@@
   ) -> plot_pve_simulated
+plot_pve_simulated
 

### 2022-11-23 scripts/PVE_calculation_fns.R

@@ -1,5 +1,21 @@
 
 
+## ---------------------------
+##
+## Script name: PVE_calculation_fns.R
+##
+## Author: Christoph Reich
+## Date Created: 2022-11-23
+##
+## Copyright (c) Christoph Reich, 2022
+## Email: christoph.reich@med.uni-heidelberg.de
+##
+## ---------------------------
+## Notes:
+##   formulas to calculate the proportion of variance explained for numeric and 
+##   binary traits      
+## ---------------------------
+
 
 # PVE FORMULAS ------------------------------------------------------------
 explained_variance_numeric <- function(eaf, beta) {

### 2022-11-23 scripts/perform_mr.R

@@ -1,0 +1,173 @@
+
+
+## ---------------------------
+##
+## Script name: perform_mr.R
+##
+## Author: Christoph Reich
+## Date Created: 2022-11-23
+##
+## Copyright (c) Christoph Reich, 2022
+## Email: christoph.reich@med.uni-heidelberg.de
+##
+## ---------------------------
+## Notes:
+##   creates dataframe `est` with estimates from different robust MR methods
+##        
+## ---------------------------
+
+library(tibble)
+library(dplyr)
+# mr methods
+library(MendelianRandomization)
+library(MRMix)
+library(MRPRESSO)
+library(mr.raps)
+source("./scripts/MR_lasso.R")
+
+## load data created in MSc.rmd
+my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
+
+# perform MR analysis -----------------------------------------------------
+set.seed(123)
+
+mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
+est <- tibble(mr_methods=mr_methods, estimate=NA, se.estimate=NA, time=NA)
+numIV <- nrow(my_data_harm)
+
+if (numIV>2){
+  mr.obj = MendelianRandomization::mr_input(
+    # exposure
+    bx = my_data_harm$beta.bmi, 
+    bxse = my_data_harm$se.bmi, 
+    # outcome
+    by = my_data_harm$beta.gbc, 
+    byse = my_data_harm$se.gbc, 
+    snps = my_data_harm$SNP,
+    exposure = "Body mass index",
+    outcome = "Gallbladder Cancer"
+  )
+  # 1. IVW
+  T0 = proc.time()[3]
+  res = MendelianRandomization::mr_ivw(mr.obj)
+  T1 = proc.time()[3]
+  est$estimate[est$mr_methods=="IVW"] = res$Estimate
+  est$se.estimate[est$mr_methods=="IVW"] = res$StdError
+  est$time[est$mr_methods=="IVW"] = T1-T0
+  rm(res)
+  
+  # 2. median
+  T0 = proc.time()[3]
+  res = MendelianRandomization::mr_median(mr.obj)
+  T1 = proc.time()[3]
+  est$estimate[est$mr_methods=="median"] = res$Estimate
[…]

### 2022-11-23 scripts/simstudy.R

@@ -2,7 +2,7 @@
 
 ## ---------------------------
 ##
-## Script name: Simulation of betas and sample sizes
+## Script name: simstudy.R
 ##
 ## Author: Christoph Reich
 ## Date Created: 2022-11-22
@@ -12,7 +12,8 @@
 ##
 ## ---------------------------
 ## Notes:
-##   
+##   creates df `est_sim` with nsim* theta_estimates, nsim* standard errors of theta sims, nsim* pve for exposure and outcome
+##      cols: "no_sim", "theta_sim_SAMPLE_SIZES", "theta_se_sim_SAMPLE_SIZES", "pve_X_sim_SAMPLE_SIZES", "pve_Y_sim_SAMPLE_SIZES"
 ## ---------------------------
 
 
@@ -39,12 +40,12 @@ conflict_prefer("filter", "dplyr")
 
 
 # GET DATA ----------------------------------------------------------------
-# dataa
-my_data_harm <- readRDS(file = "./output/Rdata/2022-11-22_my_data_harm.rds")
+# data
+my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
 # MR estimates from full study
-est <- readRDS(file = "./output/Rdata/2022-11-22_est.rds")
+est <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
 # vars from MSc script
-sim_vars <- readRDS(file = "./output/Rdata/2022-11-22_sim_vars.rds")
+sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
 n <- sim_vars$n
 colnames_pval <- sim_vars$colnames_pval
 n.orig <- 334487
@@ -97,11 +98,6 @@ if (runSimulation==TRUE) {
                                                                   se_beta=se_update_BMI, 
                                                                   samplesize=n[i]
           ),
-          # explained_variance_GSD = explained_variance_binary(PA = eaf.gsd, 
-          #                                                    RR1 = exp(beta.gsd), 
-          #                                                    RR2 = exp(beta.gsd)^2, 
-          #                                                    K = 0.1  # assumed prevalence in population
-          #                                                    )$Vg,
[…]

### 2022-12-03 scripts/MSc.Rmd
mail 2022-11-26

@@
 library(ggplot2)
+library(ggdist)
+library(gghalves)
 library(ggthemes)
@@
   annotate(geom = "text", x = 0.15, y = 0.045, 
-           label="relatively rare variants\nwith relatively strong effects", 
+           label="relatively rare\nrelatively strong", 
            color= ggthemes_data$few$colors$Dark[2,2][[1]]
@@
   annotate(geom = "text", x = 0.42, y = 0.05, 
-           label="relatively common variants\nwith relatively strong effects", 
+           label="relatively common\nrelatively strong", 
            color=ggthemes_data$few$colors$Dark[3,2][[1]])+
   annotate(geom = "text", x = 0.1, y = 0.015, 
-           label="rare variants\nwith weak effects", color=ggthemes_data$few$colors$Dark[4,2][[1]])+
+           label="relatively rare\nrelatively weak", color=ggthemes_data$few$colors$Dark[4,2][[1]])+
   annotate("segment", x=0.44,xend=0.42,y=0.025,yend=0.013,arrow=arrow(), 
@@
   annotate(geom = "text", x = 0.44, y = 0.03, 
-           label="common variants\nwith weak effects", 
+           label="relatively common\nrelatively weak", 
            color=ggthemes_data$few$colors$Dark[5,2][[1]])+
@@
   ## plot
-  plotsubtitle <- paste0("Relatively ", maf ," variants\nwith relatively ", effect_size, " effects")
+  plotsubtitle <- paste0("relatively ", maf ,", relatively ", effect_size)
   ggplot(plot_data1, aes(x=n, y=value, col=name))+
@@
 fig01b.update + 
-  labs(title= "Figure 1", caption = "") -> fig01b.update.arrange
+  labs(title= "", caption = "") -> fig01b.update.arrange
 
 fig02.update+
-  labs(title= "Genetic architectures", subtitle="", caption = "") -> fig02.update.arrange
+  labs(title= "", subtitle="", caption = "") -> fig02.update.arrange
 
 figure01_EFFECT_SIZE_strong_VARIANTS_common+
-  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange
+  labs(title= "", caption = "") +
+  theme(plot.subtitle = element_text(color = ggthemes_data$few$colors$Dark[3,2][[1]])) -> figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange
 
 figure01_EFFECT_SIZE_strong_VARIANTS_rare+
-  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange
+  labs(title= "", caption = "") +
+  theme(plot.subtitle = element_text(color = ggthemes_data$few$colors$Dark[2,2][[1]])) -> 
+  figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange
 
 figure01_EFFECT_SIZE_weak_VARIANTS_common+
-  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange
+  labs(title= "", caption = "") +
+  theme(plot.subtitle = element_text(color = ggthemes_data$few$colors$Dark[5,2][[1]])) -> figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange
 
 figure01_EFFECT_SIZE_weak_VARIANTS_rare+
-  labs(title= "", caption = "") -> figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange
-
-ggpubr::ggarrange(fig01b.update.arrange, fig02.update.arrange,
+  labs(title= "", caption = "") +
+  theme(plot.subtitle = element_text(color = ggthemes_data$few$colors$Dark[4,2][[1]])) -> figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange
+
+ggpubr::ggarrange(fig02.update.arrange, fig01b.update.arrange,
                   figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
@@
 
-ggpubr::ggarrange(ggarrange(fig01b.update.arrange, fig02.update.arrange, ncol=2, labels = c("A", "B"), legend = "none", align = "hv"),
-                  ggarrange(figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
-                  figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange, 
-                  ncol=4, labels = c("C", "D", "E", "F"), legend="bottom", common.legend = TRUE),
+ggpubr::ggarrange(ggarrange(fig02.update.arrange, fig01b.update.arrange, ncol=2, labels = c("A", "B"), legend = "none", align = "hv"),
+                  ggarrange(figure01_EFFECT_SIZE_strong_VARIANTS_rare.arrange,
+                            figure01_EFFECT_SIZE_strong_VARIANTS_common.arrange, 
+                            figure01_EFFECT_SIZE_weak_VARIANTS_rare.arrange,
+                            figure01_EFFECT_SIZE_weak_VARIANTS_common.arrange, 
+                            ncol=4, labels = c("C", "D", "E", "F"), 
+                            legend="bottom", common.legend = TRUE),
                   nrow=2
@@
          )
-
 ```

### 2022-12-03 scripts/MSc.Rmd
raincloud plot added

@@
 ```
+
+```{r simulated-pves-raincloud-plot}
+ggplot(plot_data1b, aes(x=n, y=value, col=name))+    # plot_data1b from section Simlating on empirical data (PVE)
+  geom_point(alpha=0.6)+
+  geom_line(alpha=0.6)+
+  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
+  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
+                     labels = comma, 
+                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
+                     #limits = c(10000, 300000)  # throws an error when used with ggdist -> use coord_cartesian() instead!
+                     )+
+  #ylab("% variance explained by IVs")+
+  labs(title = "PVE with simulation added",
+       col='', 
+       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
+  ) +
+  scale_y_continuous(
+    # Features of the first axis
+    name = "% variance explained by IVs",
+    # Add a second axis and specify its features
+    sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
+    limits = c(0,10)
+  ) + 
+  ## Rainclouds
+  # https://z3tt.github.io/Rainclouds/
+  # exposure simulation
+  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, # *100 to get percent values
+                                               group=sample_size_filtered), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered, y=value_X*100, group=sample_size_filtered), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered-2000, y=value_X*100, group=sample_size_filtered), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  # outcome simulation
+  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_Y*100, # *100 to get percent values
+                                               group=sample_size_filtered), 
+               col=ggthemes_data$few$colors$Dark[4,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered, y=value_Y*100, group=sample_size_filtered), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[4,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered-2000, y=value_Y*100, group=sample_size_filtered), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[4,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  coord_cartesian(xlim=c(10000, 300000))+
+  theme_few()+
+  #scale_colour_few() +
+  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
+  #                                         "% var(X)" = "blue", 
+  #                                         "% var(Y)" = "darkblue", 
+  #                                         "Number of IVs" = "black"
+  #                                         ))
+  #
+  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
+                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
+                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
+                                           "Number of IVs" = "black"),
+                     # order
+                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
+  ) -> plot_pve_simulated.raincloud
+```

### 2022-12-04 scripts/MSc.Rmd
modified plot sim

@@
 
-```{r updated-fig01, warning=FALSE}
+```{r updated-fig01, fig.height=6, fig.width=10, warning=FALSE}
 plot_data1 <- df_cum_expl_var %>% 
@@
 
-```{r simulated-pves}
+```{r simulated-pves, fig.height=6, fig.width=10, warning=FALSE}
 # EXPOSURE PVE
@@
 
-```{r simulated-pves-raincloud-plot}
+```{r simulated-pves-raincloud-plot, fig.height=6, fig.width=10, warning=FALSE}
 ggplot(plot_data1b, aes(x=n, y=value, col=name))+    # plot_data1b from section Simlating on empirical data (PVE)
@@
   ) -> plot_pve_simulated.raincloud
-```
+plot_pve_simulated.raincloud
+```

### 2022-12-16 scripts/MSc.Rmd
changed standard deviation from which coefficients are drawn in simstudy.R

@@
 
+-   2022-12-05
+    1.
+How did the large CI for the simulated PVEs arise?
+The simulated beta coefficients were drawn from given betas **AND** sample-size adjusted standard errors `*get simulated se_{sample size}*`.
+Therefore, for small sample sizes we yielded large variance for the beta coefficients.
+I changed the simulation and only used the given standard errors for our simulation scenarios in "scripts/simstudy.R".
+
 # Version Update
@@
 
-# 4. Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8)
+# 4. Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8) --> updated to 5 × 10−6
 pvals_df %>% 
@@
                                                             )$Vg,
-         explained_variance_GBC = explained_variance_binary(PA = eaf.gbc, 
+         explained_variance_GBC2 = explained_variance_binary(PA = eaf.gbc, 
                                                             RR1 = exp(beta.gbc), 
@@
                                                             )$Vg,
-         explained_variance_GBC2 = explained_variance_binary(PA = eaf.gbc, 
+         explained_variance_GBC = explained_variance_binary(PA = eaf.gbc, 
                                                             RR1 = exp(beta.gbc), 
@@
          )
+
+sum(my_data_harm$explained_variance_GBC)
+sum(my_data_harm$explained_variance_GBC2)
 ```
@@
 
-```{r}
+```{r compare-formulas-numeric-pve}
 ggplot(data=my_data_harm, aes(x=explained_variance_BMI, y=explained_variance_BMI2)) +
@@
     select(explained_variance_BMI, explained_variance_BMI2, explained_variance_GSD, explained_variance_GBC) %>% 
-    summarise(expl_variances =across(.cols = everything(), function(x) sum(x)),
+    summarise(expl_variances =across(.cols = everything(), function(x) sum(x, na.rm = TRUE)),
               count_IVs = n())

### 2022-12-16 scripts/simstudy.R

@@ -75,32 +75,32 @@ if (runSimulation==TRUE) {
     # FILTER DATA
     sim_dat <- my_data_harm %>% 
       filter(!!sym(colnames_pval[i]) < threshold) %>% 
-      select(SNP, eaf.bmi, beta.bmi, p.value.bmi, eaf.gbc, beta.gbc, p.value.gbc)
+      select(SNP, eaf.bmi, beta.bmi, se.bmi, p.value.bmi, eaf.gbc, beta.gbc, se.gbc ,p.value.gbc)
     
-    # get simulated se_{sample size}
-    for (j in 1:nrow(sim_dat)) {
-      # as done before in section "simulating on empirical data"
-      se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
-      se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
-    }
-    
-    sim_dat <- sim_dat %>%
-      bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))
+    # get simulated se_{sample size} --> this yields large standard errors for small sample sizes -> use original sd
+    # for (j in 1:nrow(sim_dat)) {
+    #   # as done before in section "simulating on empirical data"
+    #   se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
+    #   se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
+    # }
+    #
+    # sim_dat <- sim_dat %>%
+    #   bind_cols(tibble(se_update_BMI=se_update_BMI, se_update_GBC=se_update_GBC))
     
     # simulate beta values from normal_distribution # -----------------------------------------------------
     for (k in 1:nsim) {
       sim_dat <- sim_dat %>% 
[…]

### 2022-12-16 scripts/MSc.Rmd
issue standard deviation for simulating from normal distribution detected

@@
 - Issues from discussion mentioned above
+- Interchanged panel A and panel B in the first version of Figure 1, used the same colours in panel B and in the legends
+- Added five estimates for "BMI -> GSD": Wald estimate (FTO), Wald estimate (MC4R variant), random IVW, MRMix, weighted mode
 
@@
 my_data_harm$gene_label <- NA
-my_data_harm$gene_label[my_data_harm$SNP %in% filter_FTO$refsnp_id] <- "FTO"
+my_data_harm$gene_label[my_data_harm$SNP %in% filter_FTO$refsnp_id] <- "FTO/MC4R"
 
 # color labeling
-my_data_harm$fto_col_annot <- "No FTO variant"
-my_data_harm$fto_col_annot[my_data_harm$SNP %in% filter_FTO$refsnp_id] <- "FTO variant"
+my_data_harm$fto_col_annot <- "No FTO/MC4R variant"
+my_data_harm$fto_col_annot[my_data_harm$SNP %in% filter_FTO$refsnp_id] <- "FTO/MC4R variant"
 

### 2022-12-16 scripts/test/MR-methods.Rmd
issue standard deviation for simulating from normal distribution detected

---
title: "MR-methods"
author: "Christoph Reich"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
# Perform MR-analysis

We first perform MR analysis.

```{r MendelianRandomization}
# create object
mr_object <- MendelianRandomization:: mr_input(
  # exposure
  bx = my_data$beta_exp, bxse = my_data$se_exp, 
  # outcome
  by = my_data$beta_out, byse = my_data$se_out, 
  exposure = "Body mass index",
  outcome = "Gallstone disease")

#perform MR-Robust
mrrobust <- mr_ivw(mr_object, "random", robust = TRUE)  # {MendelianRandomization}
#mrrobust

#perform MR-Egger
mregger <- mr_egger(mr_object)  # {MendelianRandomization}
#mregger

#perform weighted median
weighted_median <- mr_median(mr_object,  # {MendelianRandomization}
                             weighting = "weighted", 
                             iterations = 10000)
#weighted_median

#perform Mode based estimation
mode_based <- mr_mbe(mr_object,  # {MendelianRandomization}
       weighting = "weighted", 
       stderror = "delta", 
       phi = 1,
       seed = 19940407, 
       iterations = 10000, 
       distribution = "normal",
       alpha = 0.05)
#mode_based
```

```{r MRMix}
#perform MR-Mix  (Slob & Burgess squared their standard deviations`???); in vignette before analysis summary stats are standardized with `standardize` function
mrmix_est <- MRMix::MRMix(betahat_x = my_data$beta_exp, 
                          betahat_y = my_data$beta_out, 
                          sx = my_data$se_exp, 
                          sy = my_data$se_out)
#mrmix_est

mrmix_se <- MRMix::MRMix_se(betahat_x = my_data$beta_exp, 
                          betahat_y = my_data$beta_out, 
                          sx = my_data$se_exp, 
                          sy = my_data$se_out, 
                          theta = mrmix_est$theta, # estimate of causal effect, assuming the summary stats are standardized, theta represents increase in mean value of Y in s.d. unit of Y (for cont outcomes) or log-OR of Y (for binary outcomes)
                          pi0 = mrmix_est$pi0, 
                          sigma2 = mrmix_est$sigma2
                          )
#mrmix_se
```

```{r MRPRESSO}
#perform MR-PRESSO
# presso <- MRPRESSO:: mr_presso(BetaOutcome = "beta_out", 
#                     BetaExposure = "beta_exp", 
#                     SdOutcome = "se_out", 
#                     SdExposure = "se_exp", 
#                     OUTLIERtest = TRUE, 
#                     DISTORTIONtest = TRUE, 
#                     data = my_data, 
#                     NbDistribution = 1000,   # must be changed
#                     SignifThreshold = 0.05)
```

```{r MR-lasso}
#define function for MR-Lasso with heterogeneity criterion
MR_lasso<-function(betaYG, betaXG,sebetaYG){
  betaYGw = betaYG/sebetaYG # dividing the association estimates by sebetaYG is equivalent
  betaXGw = betaXG/sebetaYG # to weighting by sebetaYG^-2
  pleio = diag(rep(1, length(betaXG)))
  l1grid = c(seq(from=0.1, to=5, by=0.1), seq(from=5.2, to=10, by
  =0.2))

  # values of lambda for grid search
  l1grid_rse = NULL; l1grid_length = NULL; l1grid_beta = NULL; l1grid_se = NULL
  for (i in 1:length(l1grid)) {
    l1grid_which = which(attributes(penalized(betaYGw, pleio,
                                              betaXGw, lambda1=l1grid[i], trace=FALSE))$penalized==0)

    l1grid_rse[i] = summary(lm(betaYG[l1grid_which]~betaXG[l1grid_which]-1, 
                               weights=sebetaYG[l1grid_which]^-2))$sigma 
    l1grid_length[i] = length(l1grid_which)
    l1grid_beta[i] = lm(betaYG[l1grid_which]~betaXG[l1grid_which]-1,
                        weights=sebetaYG[l1grid_which]^-2)$coef[1]
    l1grid_se[i] = summary(lm(betaYG[l1grid_which]~betaXG[l1grid_which]-1, 
                              weights=
                                sebetaYG[l1grid_which]^-2))$coef[1,2]/
      min(summary(lm(betaYG[l1grid_which]~betaXG[l1grid_which]-1,
                     weights=sebetaYG[l1grid_which]^-2))$sigma, 1)
  }

  l1which_hetero = c(which(l1grid_rse[1:(length(l1grid)-1)]>1 & 
                             diff(l1grid_rse)>qchisq(0.95, df=1)/l1grid_length[2:length(l1grid)]), 
                     length(l1grid))[1]
  # heterogeneity criterion for choosing lambda
  l1hetero_beta = l1grid_beta[l1which_hetero] 
  l1hetero_se = l1grid_se[l1which_hetero]
  list(ThetaEstimate=l1hetero_beta, ThetaSE=l1hetero_se )
}

#perform MR-Lasso
mrlasso <- MR_lasso(betaYG = my_data$beta_out, betaXG = my_data$beta_exp, sebetaYG = my_data$se_out)
#mrlasso
```

```{r contamination-mixture}
contaminationmixture<-function(by, bx, byse){
  iters = 2001; theta = seq(from=-3, to=3, by=2/(iters-1))
  # if the causal estimate (and confidence interval) is not expected
  #     to lie between -1 and 1 then change from and to (and maybe
  #     increase iters)
  ratio = by/bx; ratio.se = abs(byse/bx); psi = 1.5*sd(ratio) 
  lik=NULL

  for (j1 in 1:iters) {
    lik.inc = exp(-(theta[j1]-ratio)^2/2/ratio.se^2) /sqrt(2*pi* ratio.se^2)
    lik.exc = exp(-ratio^2/2/(psi^2+ratio.se^2)) /(sqrt(2*pi*(psi^2+ ratio.se^2)))
    valid = (lik.inc>lik.exc)*1
    lik[j1] = prod(c(lik.inc[valid==1], lik.exc[valid==0])) 
    if (which.max(lik)==length(lik)) { valid.best = valid }
  }

  phi = ifelse(test = sum(valid.best)<1.5, 
               yes = 1, 
               no = 
                 max(
                  sqrt(sum(
                    ((ratio[valid.best==1]-weighted.mean(ratio[valid.best==1], ratio.se[valid.best==1]^-2))^2 * ratio.se[valid.best==1]^-2)) / 
                      (sum(valid.best) -1)),
                  1)
               )
  loglik = log(lik)
  whichin = which(2*loglik>(2*max(loglik)-qchisq(0.95, df=1)*phi^2))

  theta[which.max(loglik)] # estimate
  theta[whichin[1]] # lower limit of CI 
  theta[whichin[length(whichin)]] # upper limit of CI
  # return
  list(ThetaEstimate=theta[which.max(loglik)], ThetaLower=theta[ whichin[1]] , ThetaUpper= theta[whichin[length(whichin)]] )
}

#perform contamination mixture, note Slob & Burgess removed the 27th variable 
#   due to having a ratio to close to infty in their analysis
contam.mix <- contaminationmixture(by = my_data$beta_out, 
                                   bx = my_data$beta_exp, 
                                   byse = my_data$se_out)
#contam.mix
```

---

### 2022-12-16 scripts/simstudy.R

@@ -96,7 +96,7 @@ if (runSimulation==TRUE) {
           explained_variance_X2_sim = explained_variance_numeric2(maf=eaf.bmi, 
                                                                   beta=beta.bmi_sim_norm, 
                                                                   se_beta=se.bmi, 
-                                                                  samplesize=n[i]
+                                                                  samplesize=n[length(n)]  # n[i]
           ),
           explained_variance_Y_sim = explained_variance_binary(PA = eaf.gbc, 
                                                                RR1 = exp(beta.gbc_sim_norm), 

### 2022-12-16 scripts/test/calculate_explained_var_numeric.R

@@ -1,0 +1,26 @@
+
+
+explained_variance_numeric2(maf=my_data_harm$eaf.bmi, 
+                            beta=my_data_harm$beta.bmi, 
+                            se_beta=my_data_harm$se.bmi, 
+                            samplesize=n[length(n)]) %>% sum
+
+
+explained_variance_numeric2(maf=my_data_harm$eaf.bmi, 
+                            beta=my_data_harm$beta.bmi, 
+                            se_beta=my_data_harm$se.bmi, 
+                            samplesize=1000) %>% sum
+
+
+explained_variance_numeric2(maf=my_data_harm$eaf.bmi[my_data_harm$p_value_exp_sample_100000 < 5*10^(-6)], 
+                            beta=my_data_harm$beta.bmi[my_data_harm$p_value_exp_sample_100000 < 5*10^(-6)], 
+                            se_beta=my_data_harm$se.bmi[my_data_harm$p_value_exp_sample_100000 < 5*10^(-6)], 
+                            samplesize=100000) %>% sum
+
+
+explained_variance_numeric2(maf=my_data_harm$eaf.bmi[my_data_harm$p_value_exp_sample_150000 < 5*10^(-6)], 
+                            beta=my_data_harm$beta.bmi[my_data_harm$p_value_exp_sample_150000 < 5*10^(-6)], 
+                            se_beta=my_data_harm$se.bmi[my_data_harm$p_value_exp_sample_150000 < 5*10^(-6)], 
+                            samplesize=150000) %>% sum
+
+

### 2022-12-16 scripts/test/check_simulated_betas_and_PVE_plausibility.R

@@ -1,0 +1,192 @@
+
+
+
+## simcheck CORRELATION simulated values
+
+
+
+
+## ---------------------------
+##
+## Script name: simstudy.R
+##
+## Author: Christoph Reich
+## Date Created: 2022-11-22
+##
+## Copyright (c) Christoph Reich, 2022
+## Email: christoph.reich@med.uni-heidelberg.de
+##
+## ---------------------------
+## Notes:
+##   creates df `est_sim` with nsim* theta_estimates, nsim* standard errors of theta sims, nsim* pve for exposure and outcome
+##      cols: "no_sim", "theta_sim_SAMPLE_SIZES", "theta_se_sim_SAMPLE_SIZES", "pve_X_sim_SAMPLE_SIZES", "pve_Y_sim_SAMPLE_SIZES"
+## ---------------------------
+
+
+# set to TRUE if simulation should run when script is sourced
+runSimulation <- TRUE
+
+library(dplyr)
+library(purrr)
+library(tidyr)
+library(stringr)
+library(tibble)
+library(MendelianRandomization)
+library(TwoSampleMR)
+library(mr.raps)
+library(MRMix)
+library(MRPRESSO)
+library(penalized)
+library(conflicted)
+source("scripts/MR_lasso.R") # Slob&Burgess 2020
+source("scripts/PVE_calculation_fns.R")
+conflict_prefer("select", "dplyr")
+conflict_prefer("filter", "dplyr")
+
+
+
+# GET DATA ----------------------------------------------------------------
+# data
+my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
+# MR estimates from full study
+est <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
+# vars from MSc script
+sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
+n <- sim_vars$n
+colnames_pval <- sim_vars$colnames_pval
+n.orig <- 334487
+
+###
+# SIMULATION --------------------------------------------------------------
+###
+set.seed(1234)
+## initialize
+nsim <- 100
[…]

### 2022-12-16 scripts/test/check_simulated_betas_plausibility.R

@@ -1,0 +1,143 @@
+
+
+
+## simcheck CORRELATION simulated values
+
+
+
+
+## ---------------------------
+##
+## Script name: simstudy.R
+##
+## Author: Christoph Reich
+## Date Created: 2022-11-22
+##
+## Copyright (c) Christoph Reich, 2022
+## Email: christoph.reich@med.uni-heidelberg.de
+##
+## ---------------------------
+## Notes:
+##   creates df `est_sim` with nsim* theta_estimates, nsim* standard errors of theta sims, nsim* pve for exposure and outcome
+##      cols: "no_sim", "theta_sim_SAMPLE_SIZES", "theta_se_sim_SAMPLE_SIZES", "pve_X_sim_SAMPLE_SIZES", "pve_Y_sim_SAMPLE_SIZES"
+## ---------------------------
+
+
+# set to TRUE if simulation should run when script is sourced
+runSimulation <- TRUE
+
+library(dplyr)
+library(purrr)
+library(tidyr)
+library(stringr)
+library(tibble)
+library(MendelianRandomization)
+library(TwoSampleMR)
+library(mr.raps)
+library(MRMix)
+library(MRPRESSO)
+library(penalized)
+library(conflicted)
+source("scripts/MR_lasso.R") # Slob&Burgess 2020
+source("scripts/PVE_calculation_fns.R")
+conflict_prefer("select", "dplyr")
+conflict_prefer("filter", "dplyr")
+
+
+
+# GET DATA ----------------------------------------------------------------
+# data
+my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
+# MR estimates from full study
+est <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
+# vars from MSc script
+sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
+n <- sim_vars$n
+colnames_pval <- sim_vars$colnames_pval
+n.orig <- 334487
+
+###
+# SIMULATION --------------------------------------------------------------
+###
+set.seed(1234)
+## initialize
+nsim <- 100
[…]

### 2022-12-16 scripts/test/p-val-normal-distr.R

@@ -1,0 +1,21 @@
+
+
+# http://www.r-tutor.com/elementary-statistics/hypothesis-testing/two-tailed-test-population-proportion
+
+
+
+# calculate p value from normal distribution ------------------------------
+
+dnorm(6, mean=1, sd=3)
+
+# calc z-statistic
+a <- 5
+s <- 2  # standard deviation
+n <- 20  # sample size
+xbar <- 7
+z <- (xbar-a)/(s/sqrt(n))
+
+# calc p-value
+p.val <- 2*pnorm(-abs(z))
+p.val
+

### 2022-12-16 scripts/test/plausibilityCheckAfterHarmonization.R

@@ -1,0 +1,17 @@
+
+# PLAUSIBILITY CHECK AFTER DATA HARMONIZATION
+
+my_data %>% 
+  left_join(my_data_harm, by=c("SNP"="SNP") ) -> test
+
+test %>% select(beta_exp, beta.bmi, se_exp, se.bmi) -> test1
+sum(test1$beta_exp == test1$beta.bmi, na.rm = TRUE)
+sum(test1$se_exp == test1$se.bmi, na.rm = TRUE)
+## IDENTICAL
+
+test %>% 
+  left_join(GBC_dat,by=c("SNP"="SNP")) -> test2
+
+test2 %>% select(beta_out.y, beta.gbc, se_out.y, se.gbc) -> test3
+sum(test3$beta_out.y == test3$beta.gbc, na.rm = TRUE)
+sum(test3$se_out.y == test3$se.gbc, na.rm = TRUE)

### 2022-12-16 scripts/MSc.Rmd
added ambiguous FTO variant to analysis

@@
 my_data_harm <- my_data_harm %>% 
-  filter(mr_keep==TRUE)
+  filter(mr_keep==TRUE | SNP == "rs1558902")  # also include FTO SNP
 ```
@@
   mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
+         # USE THIS !
          explained_variance_BMI2 = explained_variance_numeric2(maf=eaf.bmi, 
@@
                                                             RR1 = exp(beta.gsd), 
-                                                            RR2 = exp(beta.gsd)^2, 
+                                                            RR2 = 1+(2*(exp(beta.gsd)-1)), 
                                                             K = 0.1  # assumed prevalence in population
-                                                            )$Vg,
-         explained_variance_GBC2 = explained_variance_binary(PA = eaf.gbc, 
-                                                            RR1 = exp(beta.gbc), 
-                                                            RR2 = exp(beta.gbc)^2, 
-                                                            K =1.2/100000  # assumed prevalence in population
                                                             )$Vg,
@@
                                                             RR1 = exp(beta.gbc), 
-                                                            RR2 = 1+(2*(exp(beta.gbc)-1)), 
+                                                            RR2 = 1+(2*(exp(beta.gbc)-1)), # exp(beta.gbc)^2, 
                                                             K =1.2/100000  # assumed prevalence in population
@@
          )
-
-sum(my_data_harm$explained_variance_GBC)
-sum(my_data_harm$explained_variance_GBC2)
+# sum(my_data_harm$explained_variance_GBC)
 ```
@@
 my_data_harm$gene_label[my_data_harm$SNP %in% filter_FTO$refsnp_id] <- "FTO/MC4R"
+my_data_harm$gene_label[my_data_harm$SNP == "rs1558902"] <- "FTO"
+my_data_harm$gene_label[my_data_harm$SNP == "rs6567160"] <- "MC4R"
 
@@
     select(explained_variance_BMI, explained_variance_BMI2, explained_variance_GSD, explained_variance_GBC) %>% 
+    # cumulative PVE
     summarise(expl_variances =across(.cols = everything(), function(x) sum(x, na.rm = TRUE)),
@@
 
-```{r plot-fig1, warning=FALSE}
+```{r updated-fig01, fig.height=6, fig.width=10, warning=FALSE}
 # 6.       Represent figure similar to Fig1 in Chatterjee’s paper
@@
 scale_factor <- 4.5/284
+
 df_cum_expl_var <- df_cum_expl_var %>% 
@@
     count_IVs_scaled = count_IVs*scale_factor)
-```
-
-```{r updated-fig01, fig.height=6, fig.width=10, warning=FALSE}
+
 plot_data1 <- df_cum_expl_var %>% 
@@
   ) -> fig01b.update
-fig01b.update
+fig01b.update  # USE THIS
 
@@
 plot_pve_simulated.raincloud
-```
+
+filename_pveplot_sim.raincloud <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_pve_raincloud.png")
+ggsave(filename = filename_pveplot_sim.raincloud, plot = plot_pve_simulated.raincloud, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+```

### 2022-12-16 scripts/simstudy.R

@@ -89,9 +89,13 @@ if (runSimulation==TRUE) {
     
     # simulate beta values from normal_distribution # -----------------------------------------------------
     for (k in 1:nsim) {
+      # scaling se_Y so that max(se_Y) = max(se_X)
+      factor1 <- max(sim_dat$se.bmi)/max(sim_dat$se.gbc)
+      se.gbc_narrow <- sim_dat$se.gbc * factor1
       sim_dat <- sim_dat %>% 
         mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se.bmi, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
-        mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se.gbc, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
+        # here we need to deal with standard deviation (standard errors for Y are  huge/ see mail 2022-12-16)
+        mutate(beta.gbc_sim_norm = purrr::map2_dbl(.x = beta.gbc, .y = se.gbc_narrow, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
         mutate(#explained_variance_X_sim = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
           explained_variance_X2_sim = explained_variance_numeric2(maf=eaf.bmi, 
                                                                   beta=beta.bmi_sim_norm, 

### 2022-12-20 scripts/MSc.Rmd
updated GSD simulations

@@
 bibliography: ../references/references.bib
+editor_options: 
+  chunk_output_type: inline
 ---
@@
   ) %>% 
+  mutate(
+    p.value.gsd = ifelse(p.value.gsd == "<.0001",  # convert from string
+                         yes = "0.0001", 
+                         no=p.value.gsd), 
+    p.value.gsd = as.numeric(p.value.gsd)) %>% 
   as_tibble()
@@
   geom_point(alpha=0.6)+
-  geom_hline(yintercept=-log10(median(my_data$p_value_exp)), col="red", alpha=0.6)+
-  geom_vline(xintercept=median(my_data$eaf_exp), col="red", alpha=0.6)+
+  geom_hline(yintercept=-log10(median(my_data$p_value_exp)), col=ggthemes_data$few$colors$Dark[3,2][[1]], alpha=0.6)+
+  geom_vline(xintercept=median(my_data$eaf_exp), col=ggthemes_data$few$colors$Dark[3,2][[1]], alpha=0.6)+
   xlab("Effect allele frequency for exposure")+
@@
 
-# Simulation Framework
+# A *Gallstone Disease* Simulation Framework
+
+## Get MR-estimates first
+
+### Using {MendelianRandomization} package
+
+```{r MendelianRandomization-analysis-GSD, warning=FALSE}
+# outsourced calculation to `perform_mr.R`
+# data FOR estimates in `perform_mr.R` script and simulation in `simstudy.R` script
+saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
+
+## load results
+est.bmi.gsd <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
+
+est.bmi.gsd %>% 
+  mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
+  kable(digits=3, caption = "MR Results for GSD (outcome) using genetic variants as instrumental variables for BMI") %>% 
+  kable_classic(full_width = FALSE, html_font = "Cambria")
+```
+
+### Scatter Plot
+
+```{r scatter-MendelianRandomization-GSD, fig.width=10, fig.height=6, warning=FALSE, message=FALSE}
+# create mr.object
+mr.obj = MendelianRandomization::mr_input(
+    # exposure
+    bx = my_data_harm$beta.bmi, 
+    bxse = my_data_harm$se.bmi, 
+    # outcome
+    by = my_data_harm$beta.gsd, 
+    byse = my_data_harm$se.gsd, 
+    snps = my_data_harm$SNP,
+    exposure = "Body mass index",
+    outcome = "Gallstone disease"
+  )
+
+p1.gsd <- mr_plot(mr.obj, error = TRUE, 
+                  line = "ivw", 
+                  orientate = FALSE, 
+                  interactive = FALSE, 
+                  labels = FALSE)
+
+p1.gsd$layers[[1]]$aes_params$size <- 1
+p1.gsd$layers[[1]]$aes_params$alpha <- 0.5
+
+p1.gsd$layers[[4]]$aes_params$colour <- ggthemes_data$few$colors$Dark[1,2][[1]]
+p1.gsd$layers[[5]]$aes_params$colour <- ggthemes_data$few$colors$Dark[1,2][[1]]
+p1.gsd$layers[[4]]$aes_params$alpha <- 0.3
+p1.gsd$layers[[5]]$aes_params$alpha <- 0.3
+
+p1.gsd$layers[[6]]$aes_params$colour <- ggthemes_data$few$colors$Dark[2,2][[1]]
+
+p1.gsd+
+  ggthemes::theme_few() -> p1.gsd
+
+p1.gsd
+
+# SAVE
+filename.mr.scatter.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrscatter_gsd.png")
+ggsave(filename = filename.mr.scatter.gsd, plot = p1.gsd, 
+         width = 10, height = 6, 
+         units = "in"  # default
+         )
+```
+
+### Forest Plot
+
+```{r forestplot-MendelianRandomization-GSD, fig.width=5, fig.height=24, warning=FALSE}
+p2.gsd <- mr_forest(
+  mr.obj,
+  alpha = 0.05,
+  snp_estimates = TRUE,
+  methods = "ivw",
+  ordered = TRUE
+) +
+  ggthemes::theme_few()
+
+#SAVE
+filename.mr.forest.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_gsd.png")
+ggsave(filename = filename.mr.forest.gsd, plot = p2.gsd, 
+       width = 5, 
+       height = 24, 
+         #units = "in"  # default
+         )
+```
+
+### Funnel Plot
+
+```{r funnel-plot-GSD, fig.height=3, fig.width=5}
+p4.gsd <- mr_funnel(mr.obj, CI=TRUE)+
+  ggthemes::theme_few()
+
+p4.gsd$layers[[1]]$aes_params$size <- 1
+p4.gsd$layers[[1]]$aes_params$alpha <- 0.8
+# linerange layer
+p4.gsd$layers[[2]]$aes_params$alpha <- 0.2
+# xintercept layer - estimate
+p4.gsd$layers[[4]]$aes_params$colour <- ggthemes_data$few$colors$Dark[9,2][[1]]
+# delete linerange layer
+p4.gsd$layers <- p4.gsd$layers[-2]
+
+# plot
+p4.gsd
+
+#SAVE
+filename.mr.funnel.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrfunnel_gsd.png")
+ggsave(filename = filename.mr.funnel.gsd, plot = p4.gsd, 
+       width = 8, 
+       height = 4, 
+       units = "in"  # default
+         )
+
+```
+
+\clearpage
+
+# B *Gallbladder Cancer* Simulation Framework
 
@@
 ```
+
+### Scatter Plot
 
@@
 
+### Forest Plot
+
 ```{r forestplot-TwoSampleMR, fig.width=5, fig.height=24, warning=FALSE}
@@
 
-```{r funnel-plot}
+### Funnel Plot
+
+```{r funnel-plot, fig.height=3, fig.width=5}
 p4 <- mr_funnel_plot(res_single)
 p4[[1]] +
-  ggthemes::theme_few()
+  ggthemes::theme_few() -> p4
+
+p4
+
+#SAVE
+filename.mr.funnel.gbc <- paste0("../output/plots/", Sys.Date(), "_figure_mrfunnel_gbc.png")
+ggsave(filename = filename.mr.funnel.gbc, plot = p4, 
+       width = 8, 
+       height = 4, 
+       units = "in"  # default
+         )
 ```
@@
 ```{r MendelianRandomization-analysis, warning=FALSE}
-# outsourced calculation to `perform_mr.R`
-# data FOR estimates in `perform_mr.R` script and simulation in `simstudy.R` script
-saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
-
 ## load results

### 2022-12-20 scripts/perform_mr.R

@@ -1,5 +1,4 @@
 
-
 ## ---------------------------
 ##
 ## Script name: perform_mr.R
@@ -23,6 +22,7 @@ library(MendelianRandomization)
 library(MRMix)
 library(MRPRESSO)
 library(mr.raps)
+library(penalized)
 source("./scripts/MR_lasso.R")
 
 ## load data created in MSc.rmd
@@ -171,3 +171,148 @@ if (numIV>2){
 
 ## SAVE estimates
 saveRDS(est, paste0("./output/Rdata/", Sys.Date(), "_est.rds"))
+
+
+
+# BMI -> GSD --------------------------------------------------------------
+
+mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
+est.bmi.gsd <- tibble(mr_methods=mr_methods, estimate=NA, se.estimate=NA, time=NA)
+numIV <- nrow(my_data_harm)
+
+if (numIV>2){
+  mr.obj = MendelianRandomization::mr_input(
+    # exposure
+    bx = my_data_harm$beta.bmi, 
+    bxse = my_data_harm$se.bmi, 
+    # outcome
+    by = my_data_harm$beta.gsd, 
+    byse = my_data_harm$se.gsd, 
+    snps = my_data_harm$SNP,
+    exposure = "Body mass index",
+    outcome = "Gallstone disease"
+  )
+  # 1. IVW
+  T0 = proc.time()[3]
+  res = MendelianRandomization::mr_ivw(mr.obj)
+  T1 = proc.time()[3]
+  est.bmi.gsd$estimate[est.bmi.gsd$mr_methods=="IVW"] = res$Estimate
+  est.bmi.gsd$se.estimate[est.bmi.gsd$mr_methods=="IVW"] = res$StdError
+  est.bmi.gsd$time[est.bmi.gsd$mr_methods=="IVW"] = T1-T0
+  rm(res)
+  
+  # 2. median
+  T0 = proc.time()[3]
+  res = MendelianRandomization::mr_median(mr.obj)
+  T1 = proc.time()[3]
+  est.bmi.gsd$estimate[est.bmi.gsd$mr_methods=="median"] = res$Estimate
+  est.bmi.gsd$se.estimate[est.bmi.gsd$mr_methods=="median"] = res$StdError
[…]

### 2022-12-20 scripts/simstudy.R

@@ -1,5 +1,4 @@
 
-
 ## ---------------------------
 ##
 ## Script name: simstudy.R
@@ -43,7 +42,11 @@ conflict_prefer("filter", "dplyr")
 # data
 my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
 # MR estimates from full study
+## BMI -> GSD
+est.bmi.gsd <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
+## BMI -> GBC
 est <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
+
 # vars from MSc script
 sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
 n <- sim_vars$n
@@ -64,6 +67,7 @@ if (runSimulation==TRUE) {
   for (i in 1:length(colnames_pval)) {
     threshold <- 5*10^-6
     se_update_BMI <- c()
+    se_update_GSD <- c()
     se_update_GBC <- c()
     
     numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
@@ -75,12 +79,16 @@ if (runSimulation==TRUE) {
     # FILTER DATA
     sim_dat <- my_data_harm %>% 
       filter(!!sym(colnames_pval[i]) < threshold) %>% 
-      select(SNP, eaf.bmi, beta.bmi, se.bmi, p.value.bmi, eaf.gbc, beta.gbc, se.gbc ,p.value.gbc)
+      select(SNP, 
+             eaf.bmi, beta.bmi, se.bmi, p.value.bmi, 
+             eaf.gsd, beta.gsd, se.gsd, p.value.gsd,
+             eaf.gbc, beta.gbc, se.gbc ,p.value.gbc)
     
     # get simulated se_{sample size} --> this yields large standard errors for small sample sizes -> use original sd
     # for (j in 1:nrow(sim_dat)) {
     #   # as done before in section "simulating on empirical data"
[…]

### 2022-12-20 scripts/simstudy.R
error handling of conmix method without prespecified CIs

@@ -41,11 +41,6 @@ conflict_prefer("filter", "dplyr")
 # GET DATA ----------------------------------------------------------------
 # data
 my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
-# MR estimates from full study
-## BMI -> GSD
-est.bmi.gsd <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
-## BMI -> GBC
-est <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
 
 # vars from MSc script
 sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
@@ -87,9 +82,9 @@ if (runSimulation==TRUE) {
     # get simulated se_{sample size} --> this yields large standard errors for small sample sizes -> use original sd
     # for (j in 1:nrow(sim_dat)) {
     #   # as done before in section "simulating on empirical data"
-    #   se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
-    #   se_update_GSD[j] <- abs(sim_dat[["beta.gsd"]][j]/-qnorm(p=(sim_dat[["p.value.gsd"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
-    #   se_update_GBC[j] <- abs(sim_dat[["beta.gbc"]][j]/-qnorm(p=(sim_dat[["p.value.gbc"]][i]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
+    #   se_update_BMI[j] <- abs(sim_dat[["beta.bmi"]][j]/-qnorm(p=(sim_dat[["p.value.bmi"]][j]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
+    #   se_update_GSD[j] <- abs(sim_dat[["beta.gsd"]][j]/-qnorm(p=(sim_dat[["p.value.gsd"]][j]/2),mean=0)*sqrt(n.orig)/sqrt(n[i]))
[…]

### 2022-12-20 scripts/MSc.Rmd
added wald estimates

@@
 
+# biology plausibility FTO & MC4R
+# ## The mr_ivw function implements the inverse-variance method, informally 
+# known as the "Toby Johnson" method. With a single genetic variant, this 
+# is simply the ratio method.
+fto_wald <- mr_wald_ratio(
+  my_data_harm$beta.bmi[my_data_harm$SNP == "rs1558902"], 
+  my_data_harm$beta.gsd[my_data_harm$SNP == "rs1558902"], 
+  my_data_harm$se.bmi[my_data_harm$SNP == "rs1558902"], 
+  my_data_harm$se.gsd[my_data_harm$SNP == "rs1558902"]
+)
+mc4r_wald <- mr_wald_ratio(
+  my_data_harm$beta.bmi[my_data_harm$SNP == "rs6567160"], 
+  my_data_harm$beta.gsd[my_data_harm$SNP == "rs6567160"], 
+  my_data_harm$se.bmi[my_data_harm$SNP == "rs6567160"], 
+  my_data_harm$se.gsd[my_data_harm$SNP == "rs6567160"]
+)
+# add rows
+est.bmi.gsd <- est.bmi.gsd %>% 
+  add_row(mr_methods = "FTO_Wald", estimate = fto_wald$b, se.estimate=fto_wald$se) %>% 
+  add_row(mr_methods = "MC4R_Wald", estimate = mc4r_wald$b, se.estimate=mc4r_wald$se)
+
 est.bmi.gsd %>% 
@@
 est <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
+
+# biology plausibility FTO & MC4R
+# ## The mr_ivw function implements the inverse-variance method, informally 
+# known as the "Toby Johnson" method. With a single genetic variant, this 
+# is simply the ratio method.
+fto_wald.gbc <- mr_wald_ratio(
+  my_data_harm$beta.bmi[my_data_harm$SNP == "rs1558902"], 
+  my_data_harm$beta.gbc[my_data_harm$SNP == "rs1558902"], 
+  my_data_harm$se.bmi[my_data_harm$SNP == "rs1558902"], 
+  my_data_harm$se.gbc[my_data_harm$SNP == "rs1558902"]
+)
+mc4r_wald.gbc <- mr_wald_ratio(
+  my_data_harm$beta.bmi[my_data_harm$SNP == "rs6567160"], 
+  my_data_harm$beta.gbc[my_data_harm$SNP == "rs6567160"], 
+  my_data_harm$se.bmi[my_data_harm$SNP == "rs6567160"], 
+  my_data_harm$se.gbc[my_data_harm$SNP == "rs6567160"]
+)
+# add rows
+est <- est %>% 
+  add_row(mr_methods = "FTO_Wald", estimate = fto_wald.gbc$b, se.estimate=fto_wald.gbc$se) %>% 
+  add_row(mr_methods = "MC4R_Wald", estimate = mc4r_wald.gbc$b, se.estimate=mc4r_wald.gbc$se)
 

### 2022-12-20 scripts/perform_mr.R

@@ -11,7 +11,8 @@
 ##
 ## ---------------------------
 ## Notes:
-##   creates dataframe `est` with estimates from different robust MR methods
+##   creates dataframe `est` with estimates from different robust MR methods: BMI -> GBC
+##   creates dataframe `est.bmi.gsd` with estimates from different robust MR methods: BMI -> GSD
 ##        
 ## ---------------------------
 
@@ -49,7 +50,7 @@ if (numIV>2){
   )
   # 1. IVW
   T0 = proc.time()[3]
-  res = MendelianRandomization::mr_ivw(mr.obj)
+  res = MendelianRandomization::mr_ivw(mr.obj, model = "random")
   T1 = proc.time()[3]
   est$estimate[est$mr_methods=="IVW"] = res$Estimate
   est$se.estimate[est$mr_methods=="IVW"] = res$StdError
@@ -173,7 +174,6 @@ if (numIV>2){
 saveRDS(est, paste0("./output/Rdata/", Sys.Date(), "_est.rds"))
 
 
-
 # BMI -> GSD --------------------------------------------------------------
 
 mr_methods = c("IVW", "median", "mode", "PRESSO", "Robust", "Lasso", "egger", "conmix", "MRMix", "RAPS")
@@ -194,7 +194,7 @@ if (numIV>2){
   )
   # 1. IVW
   T0 = proc.time()[3]
-  res = MendelianRandomization::mr_ivw(mr.obj)
+  res = MendelianRandomization::mr_ivw(mr.obj, model = "random")
   T1 = proc.time()[3]
   est.bmi.gsd$estimate[est.bmi.gsd$mr_methods=="IVW"] = res$Estimate
   est.bmi.gsd$se.estimate[est.bmi.gsd$mr_methods=="IVW"] = res$StdError

### 2022-12-22 scripts/MSc.Rmd
calculated performance metrics for both GSD and GBC

@@
 
-## Simulation
-
-We simulate
+# Simulation
+
+We simulate betas from a normal distribution. The script is outsourced to `simstudy.R`.
 
@@
 
-### Performance Metrics
-
-```{r calculate-performance-metrics}
-# fns
-bias <- function(thetaobs,theta, ...) mean(thetaobs)- theta
-mse <- function(thetaobs, theta, ...) mean((thetaobs-theta)^2)
-emp_se <- function(nsim, thetaobs, theta, ...) sqrt( 1/(nsim-1) * sum((thetaobs-mean(thetaobs))^2) )  # == sd
-
-colname_theta <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "theta_sim")]
-colname_theta_se <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "theta_se_sim")]
-colname_pve_X <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_X_sim")]
-colname_pve_Y <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_Y_sim")]
-
-# filter out sample sizes
-sample_size_filtered <- as.numeric(str_extract(colname_theta, pattern = "(\\d)+"))
-
-## get performance metrics    # https://stackoverflow.com/questions/47415072/pass-multiple-functions-to-purrrmap
-funs <- c(mean = mean, bias = bias, mse = mse, emp_se=emp_se)
-args <- list(theta = est$estimate[est$mr_methods == "conmix"], nsim=nrow(est_sim)) # trim argument non-matching and ignored for median
-
-performance_metrics <- est_sim %>% 
-  select(!!!(colname_theta)) %>% 
-  map_df(.f = ~ funs %>% 
-           map(exec, .x, !!!args), .id = "var") %>% 
-  cbind(sample_size_filtered) %>% 
-  rename(N = sample_size_filtered)
-
-# est_sim %>%    # equivalent to `emp_se()` function, but `sd()` misses dotdotdot argument 
-#   select(!!!(colname_theta)) %>% 
-#   map_dbl(.f = ~sd(.))
-```
-
-```{r further-metrics}
-# coverage
-coverage <- function(thetaobs, thetaobs_se, theta, ...) {
-  lower_ci <- thetaobs - thetaobs_se
-  upper_ci <- thetaobs + thetaobs_se
-  coverage <- mean(lower_ci <= theta & upper_ci >= theta)
-  return(coverage)
-}
-coverage_bias_eliminated <- function(thetaobs, thetaobs_se, ...) {
-  lower_ci <- thetaobs - thetaobs_se
-  upper_ci <- thetaobs + thetaobs_se
-  mean_thetaobs <- mean(thetaobs)
-  coverage_bias_eliminated <- mean(lower_ci <= mean_thetaobs & upper_ci >=mean_thetaobs)
-  return(coverage_bias_eliminated)
-}
-
-type1error <- function(pvals) sum(pvals<=0.05)/length(pvals)  # no causal_effect
-
-coverage_calc <- c()
-coverage_bias_eliminated_calc <- c()
-type1error_calc <- c()
-for (i in 1:length(sample_size_filtered)) {
-  sample_size <- sample_size_filtered[i]
-  col_theta <- paste0("theta_sim_", format(sample_size, scientific = FALSE))
-  col_theta_se <- paste0("theta_se_sim_", format(sample_size, scientific = FALSE))
-  col_mr_pval <- paste0("mr_pval_", format(sample_size, scientific = FALSE))
-
-  coverage_calc[i] <- coverage(thetaobs = est_sim[col_theta], thetaobs_se = est_sim[col_theta_se], theta = est$estimate[est$mr_methods == "conmix"])
-  coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_theta]], thetaobs_se = est_sim[[col_theta_se]])
-  type1error_calc[i] <- type1error(est_sim[[col_mr_pval]])
-}
-
-performance_metrics <- performance_metrics %>% 
-  select(N, everything()) %>% 
-  cbind(data.frame(coverage_calc, coverage_bias_eliminated_calc, type1error_calc)) %>% 
-  as_tibble()
-```
-
-### Performance plots
-
-```{r boxplot-theta-sim-distribution, fig.height=6, fig.width=10}
-# compare distribution of simulated thetas against theta_324000
-boxplot_dat <- est_sim %>% 
-  select(!!!(colname_theta)) %>% 
-  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
-  mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))
-
-ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
-  geom_boxplot()+
-  geom_jitter(width=0.1,alpha=0.2)+
-  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  annotate(geom = "text", x = "300000", y = args$theta+0.5,
-           label= expression( paste("ref: ", theta[X] ))  , 
-           color= ggthemes_data$few$colors$Dark[9,2][[1]]
-             )+
-  scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "150000", "200000", "250000", "300000"), )+ 
-  ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
-  labs(title = "Estimates from beta simulation",
-       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
-       )+
-  coord_flip()+
-  theme_few() -> simplot_thetas
-simplot_thetas
-
-### same plot as above, but from already calculated performance metrics theta_mean + theta_sd
-ggplot(performance_metrics, aes(x=N, y=mean))+ 
-  geom_point()+
-  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  annotate(geom = "text", x = 300000, y = args$theta+0.25,
-         label= expression( paste("ref: ", theta[X] ))  , 
-         color= ggthemes_data$few$colors$Dark[9,2][[1]]
-           )+
-  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
-                width=3000, # x-axis dimension
-                position = "dodge")+
-  scale_x_continuous(labels = comma)+
-  xlab("Size of Sample I")+
-  ylab("Mean thetas")+
-  labs(title = "Simulated Thetas with Empirical Standard Error")+
-  theme_few() -> simplot_thetas2
-simplot_thetas2
-
-filename_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_thetas.png")
-ggsave(filename = filename_simplot_thetas, plot = simplot_thetas, 
-       width = 10, height = 6, 
-       units = "in"  # default
-       )
-filename_simplot_thetas2 <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_thetas2.png")
-ggsave(filename = filename_simplot_thetas2, plot = simplot_thetas2, 
-       width = 10, height = 6, 
-       units = "in"  # default
-       )
-```
-
-```{r plot-performance-metrics-bias-mse, fig.height=6, fig.width=14}
-#BIAS
-ggplot(performance_metrics, aes(x=N, y=bias))+ 
-  geom_point()+
-  scale_x_continuous(labels = comma)+
-  xlab("Size of Sample I")+
-  ylab("Bias")+
-  labs(title = "Bias with Empirical Standard Error")+
-  theme_few() -> biasplot
-
-#MSE
-ggplot(performance_metrics, aes(x=N, y=mse))+ 
-  geom_point()+
-  scale_x_continuous(labels = comma)+
-  xlab("Size of Sample I")+
-  ylab("MSE")+
-  labs(title = "MSE")+
-  theme_few() -> mseplot
-
-ggarrange(biasplot, mseplot) ->bias_mse_plot
-bias_mse_plot
-
-filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_bias_mse.png")
-ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot, 
-       width = 14, height = 6, 
-       units = "in"  # default
-       )
-```
+## PVE with simulated data Plot
 
@@
 est_sim %>% 
+  select(!!!(colname_pve_U)) %>% 
+  pivot_longer(everything(), names_to = "PVE_U", values_to = "value_U") %>% 
+  cbind(sample_size_filtered) %>% as_tibble() -> pve_U_sim
+
+est_sim %>% 
   select(!!!(colname_pve_Y)) %>% 
@@
 boxplot_overlay_dat <- pve_X_sim %>% 
+  cbind(pve_U_sim %>% select(-sample_size_filtered)) %>% 
   cbind(pve_Y_sim %>% select(-sample_size_filtered)) %>% 
@@
                col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2)+
+  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_U*100, # *100 to get percent values
+                                group=sample_size_filtered), 
+               col=ggthemes_data$few$colors$Dark[2,2][[1]],
                alpha=0.2, size=0.2)+
@@
     side = "l", range_scale = .4, alpha = .2)+
+  # confounder simulation
+  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_U*100, # *100 to get percent values
+                                               group=sample_size_filtered), 
+               col=ggthemes_data$few$colors$Dark[2,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered, y=value_U*100, group=sample_size_filtered), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[2,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered-2000, y=value_U*100, group=sample_size_filtered), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[2,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
   # outcome simulation
@@
 ```
+
+## Performance Metrics BIAS, MSE, MEAN, EMP_SE
+
+```{r calculate-performance-metrics}
+# fns
+bias <- function(thetaobs,theta, ...) mean(thetaobs)- theta
+mse <- function(thetaobs, theta, ...) mean((thetaobs-theta)^2)
+emp_se <- function(nsim, thetaobs, theta, ...) sqrt( 1/(nsim-1) * sum((thetaobs-mean(thetaobs))^2) )  # == sd (with ...)
+
+# retrieve PVE colnames
+colname_pve_X <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_X_sim")]
+colname_pve_U <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_U_sim")]
+colname_pve_Y <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_Y_sim")]
+# retrieve IVW/ MODE/ CONMIX cols for BMI -> GSD
+colname_IVW_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GSD_")]
+colname_IVW_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GSD_")]
+colname_MODE_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GSD_")]
+colname_MODE_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GSD_")]
+colname_CONMIX_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GSD_")]
+colname_CONMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GSD_")]
+# retrieve IVW/ MODE/ CONMIX cols for BMI -> GBC
+colname_IVW_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GBC_")]
+colname_IVW_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GBC_")]
+colname_MODE_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GBC_")]
+colname_MODE_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GBC_")]
+colname_CONMIX_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GBC_")]
+colname_CONMIX_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GBC_")]
+
+# filter out sample sizes (IVW only used as example, could be any other col)
+sample_size_filtered <- as.numeric(str_extract(colname_IVW_theta_sim_BMI_GSD, pattern = "(\\d)+"))
+
+## get performance metrics    # https://stackoverflow.com/questions/47415072/pass-multiple-functions-to-purrrmap
+funs <- c(mean = mean, emp_se=emp_se, bias = bias, mse = mse)
+## args gsd
+args.gsd.ivw <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "IVW"], 
+                     nsim=nrow(est_sim)
+                     ) # trim argument non-matching and ignored for median
+args.gsd.mode <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "mode"], 
+                     nsim=nrow(est_sim)
+                     )
+args.gsd.conmix <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "conmix"], 
+                     nsim=nrow(est_sim)
+                     )
+## args gbc
+args.gbc.ivw <- list(theta = est$estimate[est$mr_methods == "IVW"], 
+                     nsim=nrow(est_sim)
+                     ) 
+args.gbc.mode <- list(theta = est$estimate[est$mr_methods == "mode"], 
+                     nsim=nrow(est_sim)
+                     )
+args.gbc.conmix <- list(theta = est$estimate[est$mr_methods == "conmix"], 
+                     nsim=nrow(est_sim)
+                     )
+## Standard Deviation
+# est_sim %>%    # equivalent to `emp_se()` function, but `sd()` misses dotdotdot argument 
+#   select(!!!(colname_theta)) %>% 
+#   map_dbl(.f = ~sd(.))
+```
+
+```{r CALC-perf-metrics-GSD}
+IVW_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_IVW_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.ivw), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+
+MODE_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_MODE_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.mode), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+
+CONMIX_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_CONMIX_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.conmix), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+```
+
+```{r CALC-perf-metrics-GBC}
+IVW_GBC_performance_metrics <- est_sim %>% 
+  select(!!!(colname_IVW_theta_sim_BMI_GBC)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gbc.ivw), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+
+MODE_GBC_performance_metrics <- est_sim %>% 
+  select(!!!(colname_MODE_theta_sim_BMI_GBC)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gbc.mode), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+
+CONMIX_GBC_performance_metrics <- est_sim %>% 
+  select(!!!(colname_CONMIX_theta_sim_BMI_GBC)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gbc.conmix), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+```
+
+## Performance Metrics COVERAGE, COVERAGE-BIAS-ELIMINATED, TYPE1ERROR
+
+```{r further-metrics}
+# coverage
+coverage <- function(thetaobs, thetaobs_se, theta, ...) {
+  lower_ci <- thetaobs - thetaobs_se
+  upper_ci <- thetaobs + thetaobs_se
+  coverage <- mean(lower_ci <= theta & upper_ci >= theta)
+  return(coverage)
+}
+# coverage bias eliminated
+coverage_bias_eliminated <- function(thetaobs, thetaobs_se, ...) {
+  lower_ci <- thetaobs - thetaobs_se
+  upper_ci <- thetaobs + thetaobs_se
+  mean_thetaobs <- mean(thetaobs)
+  coverage_bias_eliminated <- mean(lower_ci <= mean_thetaobs & upper_ci >=mean_thetaobs)
+  return(coverage_bias_eliminated)
+}
+# type1error
+type1error <- function(pvals) sum(pvals<=0.05)/length(pvals)  # no causal_effect
+
+# GSD
+IVW_GSD_coverage_calc <- c()
+IVW_GSD_coverage_bias_eliminated_calc <- c()
+IVW_GSD_type1error_calc <- c()
+MODE_GSD_coverage_calc <- c()
+MODE_GSD_coverage_bias_eliminated_calc <- c()
+MODE_GSD_type1error_calc <- c()
+CONMIX_GSD_coverage_calc <- c()
+CONMIX_GSD_coverage_bias_eliminated_calc <- c()
+CONMIX_GSD_type1error_calc <- c()
+# GBC
+IVW_GBC_coverage_calc <- c()
+IVW_GBC_coverage_bias_eliminated_calc <- c()
+IVW_GBC_type1error_calc <- c()
+MODE_GBC_coverage_calc <- c()
+MODE_GBC_coverage_bias_eliminated_calc <- c()
+MODE_GBC_type1error_calc <- c()
+CONMIX_GBC_coverage_calc <- c()
+CONMIX_GBC_coverage_bias_eliminated_calc <- c()
+CONMIX_GBC_type1error_calc <- c()
+for (i in 1:length(sample_size_filtered)) {
+  sample_size <- sample_size_filtered[i]
+  ## GSD additional metrics ----------------------------------------------------
+  # IVW
+  col_IVW_GSD_theta <- paste0("IVW_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_IVW_GSD_theta_se <- paste0("IVW_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_IVW_GSD_mr_pval <- paste0("IVW_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  IVW_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_IVW_GSD_theta], thetaobs_se = est_sim[col_IVW_GSD_theta_se], theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "IVW"])
+  IVW_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_IVW_GSD_theta]], thetaobs_se = est_sim[[col_IVW_GSD_theta_se]])
+  IVW_GSD_type1error_calc[i] <- type1error(est_sim[[col_IVW_GSD_mr_pval]])
+  # MODE
+  col_MODE_GSD_theta <- paste0("MODE_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_MODE_GSD_theta_se <- paste0("MODE_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_MODE_GSD_mr_pval <- paste0("MODE_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  MODE_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MODE_GSD_theta], 
+                                       thetaobs_se = est_sim[col_MODE_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "mode"])
+  MODE_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MODE_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_MODE_GSD_theta_se]])
+  MODE_GSD_type1error_calc[i] <- type1error(est_sim[[col_MODE_GSD_mr_pval]])
+  # CONMIX
+  col_CONMIX_GSD_theta <- paste0("CONMIX_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_CONMIX_GSD_theta_se <- paste0("CONMIX_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_CONMIX_GSD_mr_pval <- paste0("CONMIX_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  CONMIX_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_CONMIX_GSD_theta], 
+                                       thetaobs_se = est_sim[col_CONMIX_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "conmix"])
+  CONMIX_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_CONMIX_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_CONMIX_GSD_theta_se]])
+  CONMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_CONMIX_GSD_mr_pval]])
+
+  ## GBC additional metrics ----------------------------------------------------
+  # IVW
+  col_IVW_GBC_theta <- paste0("IVW_theta_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
+  col_IVW_GBC_theta_se <- paste0("IVW_theta_se_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
+  col_IVW_GBC_mr_pval <- paste0("IVW_mr_pval_BMI_GBC_", format(sample_size, scientific = FALSE))
+  # calc
+  IVW_GBC_coverage_calc[i] <- coverage(thetaobs = est_sim[col_IVW_GBC_theta], thetaobs_se = est_sim[col_IVW_GBC_theta_se], theta = est$estimate[est$mr_methods == "IVW"])
+  IVW_GBC_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_IVW_GBC_theta]], thetaobs_se = est_sim[[col_IVW_GSD_theta_se]])
+  IVW_GBC_type1error_calc[i] <- type1error(est_sim[[col_IVW_GBC_mr_pval]])
+  # MODE
+  col_MODE_GBC_theta <- paste0("MODE_theta_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
+  col_MODE_GBC_theta_se <- paste0("MODE_theta_se_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
+  col_MODE_GBC_mr_pval <- paste0("MODE_mr_pval_BMI_GBC_", format(sample_size, scientific = FALSE))
+  # calc
+  MODE_GBC_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MODE_GBC_theta], 
+                                       thetaobs_se = est_sim[col_MODE_GBC_theta_se], 
+                                       theta = est$estimate[est$mr_methods == "mode"])
+  MODE_GBC_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MODE_GBC_theta]], 
+                                                                       thetaobs_se = est_sim[[col_MODE_GBC_theta_se]])
+  MODE_GBC_type1error_calc[i] <- type1error(est_sim[[col_MODE_GBC_mr_pval]])
+  # CONMIX
+  col_CONMIX_GBC_theta <- paste0("CONMIX_theta_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
+  col_CONMIX_GBC_theta_se <- paste0("CONMIX_theta_se_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
+  col_CONMIX_GBC_mr_pval <- paste0("CONMIX_mr_pval_BMI_GBC_", format(sample_size, scientific = FALSE))
+  # calc
+  CONMIX_GBC_coverage_calc[i] <- coverage(thetaobs = est_sim[col_CONMIX_GBC_theta], 
+                                       thetaobs_se = est_sim[col_CONMIX_GBC_theta_se], 
+                                       theta = est$estimate[est$mr_methods == "conmix"])
+  CONMIX_GBC_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_CONMIX_GBC_theta]], 
+                                                                       thetaobs_se = est_sim[[col_CONMIX_GBC_theta_se]])
+  CONMIX_GBC_type1error_calc[i] <- type1error(est_sim[[col_CONMIX_GBC_mr_pval]])  
+}
+```
+
+```{r add-additional-metrics-to-df}
+## GSD
+IVW_GSD_performance_metrics <- IVW_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(IVW_GSD_coverage_calc, IVW_GSD_coverage_bias_eliminated_calc, IVW_GSD_type1error_calc)) %>% 
+  rename(coverage = IVW_GSD_coverage_calc, 
+         coverage_bias_eliminated = IVW_GSD_coverage_bias_eliminated_calc, 
+         type1error = IVW_GSD_type1error_calc) %>% 
+  as_tibble()
+MODE_GSD_performance_metrics <- MODE_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(MODE_GSD_coverage_calc, MODE_GSD_coverage_bias_eliminated_calc, MODE_GSD_type1error_calc)) %>% 
+  rename(coverage = MODE_GSD_coverage_calc, 
+         coverage_bias_eliminated = MODE_GSD_coverage_bias_eliminated_calc, 
+         type1error = MODE_GSD_type1error_calc) %>% 
+  as_tibble()
+CONMIX_GSD_performance_metrics <- CONMIX_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(CONMIX_GSD_coverage_calc, CONMIX_GSD_coverage_bias_eliminated_calc, CONMIX_GSD_type1error_calc)) %>% 
+  rename(coverage = CONMIX_GSD_coverage_calc, 
+         coverage_bias_eliminated = CONMIX_GSD_coverage_bias_eliminated_calc, 
+         type1error = CONMIX_GSD_type1error_calc) %>% 
+  as_tibble()
+
+## GBC
+IVW_GBC_performance_metrics <- IVW_GBC_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(IVW_GBC_coverage_calc, IVW_GBC_coverage_bias_eliminated_calc, IVW_GBC_type1error_calc)) %>% 
+  rename(coverage = IVW_GBC_coverage_calc, 
+         coverage_bias_eliminated = IVW_GBC_coverage_bias_eliminated_calc, 
+         type1error = IVW_GBC_type1error_calc) %>% 
+  as_tibble()
+MODE_GBC_performance_metrics <- MODE_GBC_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(MODE_GBC_coverage_calc, MODE_GBC_coverage_bias_eliminated_calc, MODE_GBC_type1error_calc)) %>% 
+  rename(coverage = MODE_GBC_coverage_calc, 
+         coverage_bias_eliminated = MODE_GBC_coverage_bias_eliminated_calc, 
+         type1error = MODE_GBC_type1error_calc) %>% 
+  as_tibble()
+CONMIX_GBC_performance_metrics <- CONMIX_GBC_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(CONMIX_GBC_coverage_calc, CONMIX_GBC_coverage_bias_eliminated_calc, CONMIX_GBC_type1error_calc)) %>% 
+  rename(coverage = CONMIX_GBC_coverage_calc, 
+         coverage_bias_eliminated = CONMIX_GBC_coverage_bias_eliminated_calc, 
+         type1error = CONMIX_GBC_type1error_calc) %>% 
+  as_tibble()
+```
+
+```{r store-performance-metrics-csv}
+## GSD
+filename_IVW_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_IVW_GSD_performance_metrics_NSIM_", nsim, ".csv")
+filename_MODE_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_MODE_GSD_performance_metrics_NSIM_", nsim, ".csv")
+filename_CONMIX_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_CONMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
+write.csv2(x = IVW_GSD_performance_metrics, file = filename_IVW_GSD_performance_metrics)
+write.csv2(x = MODE_GSD_performance_metrics, file = filename_MODE_GSD_performance_metrics)
+write.csv2(x = CONMIX_GSD_performance_metrics, file = filename_CONMIX_GSD_performance_metrics)
+
+## GBC
+filename_IVW_GBC_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_IVW_GBC_performance_metrics_NSIM_", nsim, ".csv")
+filename_MODE_GBC_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_MODE_GBC_performance_metrics_NSIM_", nsim, ".csv")
+filename_CONMIX_GBC_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_CONMIX_GBC_performance_metrics_NSIM_", nsim, ".csv")
+write.csv2(x = IVW_GBC_performance_metrics, file = filename_IVW_GBC_performance_metrics)
+write.csv2(x = MODE_GBC_performance_metrics, file = filename_MODE_GBC_performance_metrics)
+write.csv2(x = CONMIX_GBC_performance_metrics, file = filename_CONMIX_GBC_performance_metrics)
+```
+
+## Performance Plots
+
+```{r boxplot-theta-sim-distribution, fig.height=6, fig.width=10}
+# compare distribution of simulated thetas against theta_324000
+boxplot_dat <- est_sim %>% 
+  select(!!!(colname_theta)) %>% 
+  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+  mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))
+
+ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
+  geom_boxplot()+
+  geom_jitter(width=0.1,alpha=0.2)+
+  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = "300000", y = args$theta+0.5,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+  scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "150000", "200000", "250000", "300000"), )+ 
+  ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
+  labs(title = "Estimates from beta simulation",
+       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
+       )+
+  coord_flip()+
+  theme_few() -> simplot_thetas
+simplot_thetas
+
+### same plot as above, but from already calculated performance metrics theta_mean + theta_sd
+ggplot(performance_metrics, aes(x=N, y=mean))+ 
+  geom_point()+
+  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 300000, y = args$theta+0.25,
+         label= expression( paste("ref: ", theta[X] ))  , 
+         color= ggthemes_data$few$colors$Dark[9,2][[1]]
+           )+
+  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
+                width=3000, # x-axis dimension
+                position = "dodge")+
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Mean thetas")+
+  labs(title = "Simulated Thetas with Empirical Standard Error")+
+  theme_few() -> simplot_thetas2
+simplot_thetas2
+
+filename_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_thetas.png")
+ggsave(filename = filename_simplot_thetas, plot = simplot_thetas, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+filename_simplot_thetas2 <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_thetas2.png")
+ggsave(filename = filename_simplot_thetas2, plot = simplot_thetas2, 
+       width = 10, height = 6, 
+       units = "in"  # default
+       )
+```
+
+```{r plot-performance-metrics-bias-mse, fig.height=6, fig.width=14}
+#BIAS
+ggplot(performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()+
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Bias")+
+  labs(title = "Bias with Empirical Standard Error")+
+  theme_few() -> biasplot
+
+#MSE
+ggplot(performance_metrics, aes(x=N, y=mse))+ 
+  geom_point()+
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("MSE")+
+  labs(title = "MSE")+
+  theme_few() -> mseplot
+
+ggarrange(biasplot, mseplot) ->bias_mse_plot
+bias_mse_plot
+
+filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_bias_mse.png")
+ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot, 
+       width = 14, height = 6, 
+       units = "in"  # default
+       )
+```

### 2022-12-22 scripts/simstudy.R

@@ -53,7 +53,7 @@ n.orig <- 334487
 ###
 set.seed(1234)
 ## initialize
-nsim <- 25
+nsim <- 50
 est_sim <- tibble(no_sim = 1:nsim)
 
 ## START SIM
@@ -248,6 +248,7 @@ if (runSimulation==TRUE) {
       
       rm(conmix.res.gbc, mode.res.gbc, mode.res.gsd, ivw.res.gbc, ivw.res.gsd)
     }
+    
     #rm(se_update_BMI, se_update_GBC)
     print(paste0("|||-----------------------Run finished for sample size: ", format(n[i], scientific=FALSE), " -----------------------|||"))
   }  ## END OF SIMULATION

### 2022-12-22 scripts/MSc.Rmd
arranged performance plots X,U,Y

@@
 est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_NSIM_", full.names = TRUE)))
+nsim <- nrow(est_sim)
 ```
@@
 ```{r simulated-pves, fig.height=6, fig.width=10, warning=FALSE}
+# retrieve PVE colnames
+colname_pve_X <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_X_sim")]
+colname_pve_U <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_U_sim")]
+colname_pve_Y <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_Y_sim")]
+
+## for performance measures later
+# retrieve IVW/ MODE/ CONMIX cols for BMI -> GSD 
+colname_IVW_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GSD_")]
+colname_IVW_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GSD_")]
+colname_MODE_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GSD_")]
+colname_MODE_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GSD_")]
+colname_CONMIX_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GSD_")]
+colname_CONMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GSD_")]
+# retrieve IVW/ MODE/ CONMIX cols for BMI -> GBC
+colname_IVW_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GBC_")]
+colname_IVW_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GBC_")]
+colname_MODE_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GBC_")]
+colname_MODE_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GBC_")]
+colname_CONMIX_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GBC_")]
+colname_CONMIX_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GBC_")]
+
+# filter out sample sizes (IVW only used as example, could be any other col)
+sample_size_filtered <- as.numeric(str_extract(colname_IVW_theta_sim_BMI_GSD, pattern = "(\\d)+"))
+
 # EXPOSURE PVE
@@
   cbind(sample_size_filtered) %>% as_tibble() -> pve_X_sim
-
+# CONFOUNDER PVE
 est_sim %>% 
@@
   cbind(sample_size_filtered) %>% as_tibble() -> pve_U_sim
-
+# OUTCOME PVE
 est_sim %>% 
@@
 
-# retrieve PVE colnames
-colname_pve_X <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_X_sim")]
-colname_pve_U <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_U_sim")]
-colname_pve_Y <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_Y_sim")]
-# retrieve IVW/ MODE/ CONMIX cols for BMI -> GSD
-colname_IVW_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GSD_")]
-colname_IVW_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GSD_")]
-colname_MODE_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GSD_")]
-colname_MODE_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GSD_")]
-colname_CONMIX_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GSD_")]
-colname_CONMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GSD_")]
-# retrieve IVW/ MODE/ CONMIX cols for BMI -> GBC
-colname_IVW_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GBC_")]
-colname_IVW_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GBC_")]
-colname_MODE_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GBC_")]
-colname_MODE_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GBC_")]
-colname_CONMIX_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GBC_")]
-colname_CONMIX_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GBC_")]
-
-# filter out sample sizes (IVW only used as example, could be any other col)
-sample_size_filtered <- as.numeric(str_extract(colname_IVW_theta_sim_BMI_GSD, pattern = "(\\d)+"))
-
 ## get performance metrics    # https://stackoverflow.com/questions/47415072/pass-multiple-functions-to-purrrmap
@@
 
-```{r boxplot-theta-sim-distribution, fig.height=6, fig.width=10}
-# compare distribution of simulated thetas against theta_324000
-boxplot_dat <- est_sim %>% 
-  select(!!!(colname_theta)) %>% 
+### GSD simulated theta distribution
+
+```{r theta-sim-distribution-GSD-old, eval=FALSE, fig.height=6, fig.width=10, include=FALSE}
+### OLD
+# # compare distribution of simulated thetas against theta_324000
+# boxplot_dat <- est_sim %>% 
+#   select(!!!(colname_IVW_theta_sim_BMI_GSD)) %>% 
+#   pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+#   mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))
+# 
+# ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
+#   geom_boxplot()+
+#   geom_jitter(width=0.1,alpha=0.2)+
+#   geom_hline(yintercept=  args.gsd.ivw$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+#   annotate(geom = "text", x = "300000", y = args$theta+0.5,
+#            label= expression( paste("ref: ", theta[X] ))  , 
+#            color= ggthemes_data$few$colors$Dark[9,2][[1]]
+#              )+
+#   scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "150000", "200000", "250000", "300000"), )+ 
+#   ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
+#   labs(title = "Estimates from beta simulation",
+#        caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
+#        )+
+#   coord_flip()+
+#   theme_few() -> simplot_thetas
+# simplot_thetas
+```
+
+```{r theta-sim-distribution-GSD, fig.height=6, fig.width=10, warning=FALSE}
+## IVW
+IVW_BMI_GSD_geom_point_dat <- est_sim %>% 
+  select(!!!(colname_IVW_theta_sim_BMI_GSD)) %>% 
   pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
-  mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))
-
-ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
-  geom_boxplot()+
-  geom_jitter(width=0.1,alpha=0.2)+
-  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  annotate(geom = "text", x = "300000", y = args$theta+0.5,
-           label= expression( paste("ref: ", theta[X] ))  , 
-           color= ggthemes_data$few$colors$Dark[9,2][[1]]
-             )+
-  scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "150000", "200000", "250000", "300000"), )+ 
-  ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
-  labs(title = "Estimates from beta simulation",
-       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
-       )+
-  coord_flip()+
-  theme_few() -> simplot_thetas
-simplot_thetas
-
-### same plot as above, but from already calculated performance metrics theta_mean + theta_sd
-ggplot(performance_metrics, aes(x=N, y=mean))+ 
+  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+ggplot(IVW_GSD_performance_metrics, aes(x=N, y=mean))+ 
   geom_point()+
-  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  annotate(geom = "text", x = 300000, y = args$theta+0.25,
+  geom_hline(yintercept=  args.gsd.ivw$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 50000, y = args.gsd.ivw$theta+0.05,
          label= expression( paste("ref: ", theta[X] ))  , 
@@
                 position = "dodge")+
+  ## ADD rainbow plots
+  geom_boxplot(data = IVW_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = IVW_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = IVW_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  ## FINISH rainbow
   scale_x_continuous(labels = comma)+
@@
   ylab("Mean thetas")+
-  labs(title = "Simulated Thetas with Empirical Standard Error")+
-  theme_few() -> simplot_thetas2
-simplot_thetas2
-
-filename_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_thetas.png")
-ggsave(filename = filename_simplot_thetas, plot = simplot_thetas, 
-       width = 10, height = 6, 
-       units = "in"  # default
-       )
-filename_simplot_thetas2 <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_thetas2.png")
-ggsave(filename = filename_simplot_thetas2, plot = simplot_thetas2, 
-       width = 10, height = 6, 
-       units = "in"  # default
-       )
-```
-
-```{r plot-performance-metrics-bias-mse, fig.height=6, fig.width=14}
-#BIAS
-ggplot(performance_metrics, aes(x=N, y=bias))+ 
+  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "IVW estimate: BMI -> GSD")+
+  theme_few() -> IVW_GSD_simplot_thetas
+IVW_GSD_simplot_thetas
+
+## MODE
+MODE_BMI_GSD_geom_point_dat <- est_sim %>% 
+  select(!!!(colname_MODE_theta_sim_BMI_GSD)) %>% 
+  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+ggplot(MODE_GSD_performance_metrics, aes(x=N, y=mean))+ 
   geom_point()+
+  geom_hline(yintercept=  args.gsd.mode$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 50000, y = args.gsd.mode$theta+0.05,
+         label= expression( paste("ref: ", theta[X] ))  , 
+         color= ggthemes_data$few$colors$Dark[9,2][[1]]
+           )+
+  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
+                width=3000, # x-axis dimension
+                position = "dodge")+
+  ## ADD rainbow plots
+  geom_boxplot(data = MODE_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = MODE_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = MODE_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  ## FINISH rainbow
   scale_x_continuous(labels = comma)+
   xlab("Size of Sample I")+
+  ylab("Mean thetas")+
+  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "MODE estimate: BMI -> GSD")+
+  theme_few() -> MODE_GSD_simplot_thetas
+MODE_GSD_simplot_thetas
+
+## CONMIX
+CONMIX_BMI_GSD_geom_point_dat <- est_sim %>% 
+  select(!!!(colname_CONMIX_theta_sim_BMI_GSD)) %>% 
+  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=mean))+ 
+  geom_point()+
+  geom_hline(yintercept=  args.gsd.conmix$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 50000, y = args.gsd.conmix$theta+0.05,
+         label= expression( paste("ref: ", theta[X] ))  , 
+         color= ggthemes_data$few$colors$Dark[9,2][[1]]
+           )+
+  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
+                width=3000, # x-axis dimension
+                position = "dodge")+
+  ## ADD rainbow plots
+  geom_boxplot(data = CONMIX_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = CONMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = CONMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  ## FINISH rainbow
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Mean thetas")+
+  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "CONMIX estimate: BMI -> GSD")+
+  theme_few() -> CONMIX_GSD_simplot_thetas
+CONMIX_GSD_simplot_thetas
+
+# SAVE
+filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
+       width = 15, height = 6, 
+       units = "in"  # default
+       )
+filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
+       width = 15, height = 6, 
+       units = "in"  # default
+       )
+filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
+       width = 15, height = 6, 
+       units = "in"  # default
+       )
+```
+```{r theta-sim-distribution-GSD-arrange, fig.height=18, fig.width=10, warning=FALSE}
+IVW_GSD_simplot_thetas$labels$title <- ""
+MODE_GSD_simplot_thetas$labels$title <- ""
+CONMIX_GSD_simplot_thetas$labels$title <- ""
+ggarrange(IVW_GSD_simplot_thetas,MODE_GSD_simplot_thetas,CONMIX_GSD_simplot_thetas, 
+          ncol = 1, nrow = 3) -> GSD_simplot_thetas_arrange
+
+filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
+       width = 10, height = 18, 
+       units = "in"  # default
+       )
+```
+
+### GSD Bias & MSE
+
+```{r plot-performance-metrics-bias-mse-GSD, fig.height=10, fig.width=5}
+## IVW
+#BIAS
+ggplot(IVW_GSD_performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()+
+  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(-1.5,1.5))+
+  xlab("Size of Sample I")+
   ylab("Bias")+
-  labs(title = "Bias with Empirical Standard Error")+
-  theme_few() -> biasplot
-
+  labs(title = "IVW estimate: BMI -> GSD")+
+  theme_few() -> biasplot_IVW_GSD
 #MSE
-ggplot(performance_metrics, aes(x=N, y=mse))+ 
+ggplot(IVW_GSD_performance_metrics, aes(x=N, y=mse))+ 
   geom_point()+
+  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(0,1.8))+
+  xlab("Size of Sample I")+
+  ylab("MSE")+
+  labs(title = "IVW estimate: BMI -> GSD")+
+  theme_few() -> mseplot_IVW_GSD
+ggarrange(biasplot_IVW_GSD, mseplot_IVW_GSD) -> bias_mse_plot_IVW_GSD
+
+## MODE
+ggplot(MODE_GSD_performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()+
+  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(-1.5,1.5))+
+  xlab("Size of Sample I")+
+  ylab("Bias")+
+  labs(title = "MODE estimate: BMI -> GSD")+
+  theme_few() -> biasplot_MODE_GSD
+ggplot(MODE_GSD_performance_metrics, aes(x=N, y=mse))+ 
+  geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(0,1.8))+
+  xlab("Size of Sample I")+
+  ylab("MSE")+
+  labs(title = "MODE estimate: BMI -> GSD")+
+  theme_few() -> mseplot_MODE_GSD
+ggarrange(biasplot_MODE_GSD, mseplot_MODE_GSD) -> bias_mse_plot_MODE_GSD
+
+## CONMIX
+ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()+
+  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(-1.5,1.5))+
+  xlab("Size of Sample I")+
+  ylab("Bias")+
+  labs(title = "CONMIX estimate: BMI -> GSD")+
+  theme_few() -> biasplot_CONMIX_GSD
+ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=mse))+ 
+  geom_point()+
+  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(0,1.8))+
+  xlab("Size of Sample I")+
+  ylab("MSE")+
+  labs(title = "CONMIX estimate: BMI -> GSD")+
+  theme_few() -> mseplot_CONMIX_GSD
+ggarrange(biasplot_CONMIX_GSD, mseplot_CONMIX_GSD) -> bias_mse_plot_CONMIX_GSD
+
+ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD,
+          ncol = 1, nrow = 3, common.legend = TRUE) -> bias_mse_plot_GSD
+bias_mse_plot_GSD
+
+filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_NSIM_", nsim, ".png")
+ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
+       width = 7, height = 10, 
+       units = "in"  # default
+       )
+```
+
+### GBC simulated theta distribution
+
+```{r theta-sim-distribution-GBC, fig.height=6, fig.width=10, warning=FALSE}
+## IVW
+IVW_BMI_GBC_geom_point_dat <- est_sim %>% 
+  select(!!!(colname_IVW_theta_sim_BMI_GBC)) %>% 
+  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+ggplot(IVW_GBC_performance_metrics, aes(x=N, y=mean))+ 
+  geom_point()+
+  geom_hline(yintercept=  args.gbc.ivw$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 50000, y = args.gbc.ivw$theta+0.05,
+         label= expression( paste("ref: ", theta[X] ))  , 
+         color= ggthemes_data$few$colors$Dark[9,2][[1]]
+           )+
+  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
+                width=3000, # x-axis dimension
+                position = "dodge")+
+  ## ADD rainbow plots
+  geom_boxplot(data = IVW_BMI_GBC_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = IVW_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = IVW_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  ## FINISH rainbow
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Mean thetas")+
+  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "IVW estimate: BMI -> GBC")+
+  theme_few() -> IVW_GBC_simplot_thetas
+IVW_GBC_simplot_thetas
+
+## MODE
+MODE_BMI_GBC_geom_point_dat <- est_sim %>% 
+  select(!!!(colname_MODE_theta_sim_BMI_GBC)) %>% 
+  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+ggplot(MODE_GBC_performance_metrics, aes(x=N, y=mean))+ 
+  geom_point()+
+  geom_hline(yintercept=  args.gbc.mode$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 50000, y = args.gbc.mode$theta+0.05,
+         label= expression( paste("ref: ", theta[X] ))  , 
+         color= ggthemes_data$few$colors$Dark[9,2][[1]]
+           )+
+  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
+                width=3000, # x-axis dimension
+                position = "dodge")+
+  ## ADD rainbow plots
+  geom_boxplot(data = MODE_BMI_GBC_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = MODE_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = MODE_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  ## FINISH rainbow
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Mean thetas")+
+  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "MODE estimate: BMI -> GBC")+
+  theme_few() -> MODE_GBC_simplot_thetas
+MODE_GBC_simplot_thetas
+
+## CONMIX
+CONMIX_BMI_GBC_geom_point_dat <- est_sim %>% 
+  select(!!!(colname_CONMIX_theta_sim_BMI_GBC)) %>% 
+  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+ggplot(CONMIX_GBC_performance_metrics, aes(x=N, y=mean))+ 
+  geom_point()+
+  geom_hline(yintercept=  args.gbc.conmix$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 50000, y = args.gbc.conmix$theta+0.05,
+         label= expression( paste("ref: ", theta[X] ))  , 
+         color= ggthemes_data$few$colors$Dark[9,2][[1]]
+           )+
+  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
+                width=3000, # x-axis dimension
+                position = "dodge")+
+  ## ADD rainbow plots
+  geom_boxplot(data = CONMIX_BMI_GBC_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = CONMIX_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = CONMIX_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  ## FINISH rainbow
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Mean thetas")+
+  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "CONMIX estimate: BMI -> GBC")+
+  theme_few() -> CONMIX_GBC_simplot_thetas
+CONMIX_GBC_simplot_thetas
+
+# SAVE
+filename_IVW_GBC_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GBC_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_IVW_GBC_simplot_thetas, plot = IVW_GBC_simplot_thetas, 
+       width = 15, height = 6, 
+       units = "in"  # default
+       )
+filename_MODE_GBC_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GBC_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_MODE_GBC_simplot_thetas, plot = MODE_GBC_simplot_thetas, 
+       width = 15, height = 6, 
+       units = "in"  # default
+       )
+filename_CONMIX_GBC_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GBC_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_CONMIX_GBC_simplot_thetas, plot = CONMIX_GBC_simplot_thetas, 
+       width = 15, height = 6, 
+       units = "in"  # default
+       )
+```
+
+```{r theta-sim-distribution-GBC-arrange, fig.height=18, fig.width=10, warning=FALSE}
+IVW_GBC_simplot_thetas$labels$title <- ""
+MODE_GBC_simplot_thetas$labels$title <- ""
+CONMIX_GBC_simplot_thetas$labels$title <- ""
+ggarrange(IVW_GBC_simplot_thetas,MODE_GBC_simplot_thetas,CONMIX_GBC_simplot_thetas, 
+          ncol = 1, nrow = 3) -> GBC_simplot_thetas_arrange
+
+filename_GBC_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GBC_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_GBC_simplot_thetas_arrange, plot = GBC_simplot_thetas_arrange, 
+       width = 10, height = 18, 
+       units = "in"  # default
+       )
+```
+
+### GBC Bias & MSE
+
+```{r plot-performance-metrics-bias-mse-GBC, fig.height=10, fig.width=5}
+## IVW
+#BIAS
+ggplot(IVW_GBC_performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()+
+  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(-1.5,1.5))+
+  xlab("Size of Sample I")+
+  ylab("Bias")+
+  labs(title = "IVW estimate: BMI -> GBC")+
+  theme_few() -> biasplot_IVW_GBC
+#MSE
+ggplot(IVW_GBC_performance_metrics, aes(x=N, y=mse))+ 
+  geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_y_continuous(limits = c(0,1.8))+
   scale_x_continuous(labels = comma)+
@@
   ylab("MSE")+
-  labs(title = "MSE")+
-  theme_few() -> mseplot
-
-ggarrange(biasplot, mseplot) ->bias_mse_plot
-bias_mse_plot
-
-filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_bias_mse.png")
-ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot, 
-       width = 14, height = 6, 
-       units = "in"  # default
-       )
-```
+  labs(title = "IVW estimate: BMI -> GBC")+
+  theme_few() -> mseplot_IVW_GBC
+ggarrange(biasplot_IVW_GBC, mseplot_IVW_GBC) -> bias_mse_plot_IVW_GBC
+
+## MODE
+ggplot(MODE_GBC_performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()+
+  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(-1.5,1.5))+
+  xlab("Size of Sample I")+
+  ylab("Bias")+
+  labs(title = "MODE estimate: BMI -> GBC")+
+  theme_few() -> biasplot_MODE_GBC
+ggplot(MODE_GBC_performance_metrics, aes(x=N, y=mse))+ 
+  geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+  xlab("Size of Sample I")+
+  ylab("MSE")+
+  labs(title = "MODE estimate: BMI -> GBC")+
+  theme_few() -> mseplot_MODE_GBC
+ggarrange(biasplot_MODE_GBC, mseplot_MODE_GBC) -> bias_mse_plot_MODE_GBC
+
+## CONMIX
+ggplot(CONMIX_GBC_performance_metrics, aes(x=N, y=bias))+ 
+  geom_point()+
+  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(-1.5,1.5))+
+  xlab("Size of Sample I")+
+  ylab("Bias")+
+  labs(title = "CONMIX estimate: BMI -> GBC")+
+  theme_few() -> biasplot_CONMIX_GBC
+ggplot(CONMIX_GBC_performance_metrics, aes(x=N, y=mse))+ 
+  geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  scale_x_continuous(labels = comma)+
+  scale_y_continuous(limits = c(0,1.8))+
+  xlab("Size of Sample I")+
+  ylab("MSE")+
+  labs(title = "CONMIX estimate: BMI -> GBC")+
+  theme_few() -> mseplot_CONMIX_GBC
+ggarrange(biasplot_CONMIX_GBC, mseplot_CONMIX_GBC) -> bias_mse_plot_CONMIX_GBC
+
+ggarrange(bias_mse_plot_IVW_GBC, bias_mse_plot_MODE_GBC, bias_mse_plot_CONMIX_GBC,
+          ncol = 1, nrow = 3, common.legend = TRUE) -> bias_mse_plot_GBC
+bias_mse_plot_GBC
+
+filename_GBC_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GBC_simplot_bias_mse_NSIM_", nsim, ".png")
+ggsave(filename = filename_GBC_simplot_bias_mse, plot = bias_mse_plot_GBC, 
+       width = 7, height = 10, 
+       units = "in"  # default
+       )
+```

### 2022-12-22 scripts/simstudy.R
bug in conmix gsd default, fixed it with tryCatch

@@ -53,7 +53,7 @@ n.orig <- 334487
 ###
 set.seed(1234)
 ## initialize
-nsim <- 50
+nsim <- 100
 est_sim <- tibble(no_sim = 1:nsim)
 
 ## START SIM
@@ -165,7 +165,28 @@ if (runSimulation==TRUE) {
       est_sim[k, mode.colname_pvalue_gsd] <- mode.res.gsd$Pvalue
       
       # 3. CONMIX 
-      conmix.res.gsd = MendelianRandomization::mr_conmix(mr.obj.gsd)
+      conmix.res.gsd <- tryCatch(
+        expr = {
+          MendelianRandomization::mr_conmix(mr.obj.gsd)
+        },
+        error = function(e){
+          # the error we encounter is with "i=7" 
+          #     Error in seq.default(from = CIMin, to = CIMax, by = CIStep) : 
+          #         wrong sign in 'by' argument
+          message(paste0("Caught an error with mr_conmix (BMI -> GSD)!\nSample size: ", n[i], 
+                         "\nComputed CI's with predefined range of [-15;15]"))
+          message("Below is the error message from R:")
+          print(e) 
+          return(MendelianRandomization::mr_conmix(mr.obj.gbc, CIMin = -15, CIMax=15))
+        },
+        warning = function(w){
+          message('Caught an warning!')
+          print(w)
+        },
+        finally = {
+          #message('All done, quitting.')
+        }
+      )
       # colnames
       conmix.colname_theta_gsd <- paste0("CONMIX_theta_sim_BMI_GSD_", format(n[i], scientific=FALSE))
       conmix.colname_theta_se_gsd <- paste0("CONMIX_theta_se_sim_BMI_GSD_", format(n[i], scientific=FALSE))

### 2022-12-23 scripts/render_param_reports_msc.R
small changes

@@ -4,7 +4,7 @@
 
 ## render report and mail results
 library(dplyr)
-library(gmailr)
+# library(gmailr)
 rm(list=ls())  # not in rmd doc, otherwise params deleted
 #gm_auth_configure(path = "/Users/christophreich/Library/CloudStorage/OneDrive-bwedu/credentials.json")
 #gm_auth()

### 2022-12-23 scripts/MSc.Rmd
renaming figures, preparing report

@@
   t() %>% 
-  kable(caption = "Count of significant SNPs depending on sample size")
+  cbind(n) %>% 
+  base::as.data.frame() %>% 
+  tibble::rownames_to_column() %>% 
+  rename(sample_size = n, No_of_IVs=V1) %>% 
+  select(sample_size, No_of_IVs) %>% 
+  kable(caption = "Count of significant SNPs depending on sample size") %>% 
+  kable_styling(font_size=10)
 
@@
   scale_colour_few() -> fig00
-fig00
+# fig00   # we use the sample size dependent formula
 
@@
 
-filenamefig00_a <- paste0("../output/plots/", Sys.Date(), "_figure00_a.png")
+filenamefig00_a <- paste0("../output/plots/", Sys.Date(), "_explained_var_X_old_formula.png")
 ggsave(filename = filenamefig00_a, plot = fig00, 
@@
 
-filenamefig00_b <- paste0("../output/plots/", Sys.Date(), "_figure00_b.png")
+filenamefig00_b <- paste0("../output/plots/", Sys.Date(), "_explained_var_X_new_sample_size_dep.png")
 ggsave(filename = filenamefig00_b, plot = fig00_b, 
@@
 
-fig01.update
+# fig01.update   # old formula for PVE exposure X (without sample size adjusting)
+# 
 # scale_color_manual(name = "Y series", values = c("Y1" = "darkblue", "Y2" = "red"))
@@
 # SAVE
-filenamefig01_a <- paste0("../output/plots/", Sys.Date(), "_figure01_a.png")
-filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_b.png")
+filenamefig01_a <- paste0("../output/plots/", Sys.Date(), "_figure01_old_PVE_X_formula.png")
+filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_new_PVE_X_formula.png")
 ggsave(filename = filenamefig01_a, plot = fig01.update, 
@@
   ylab("Exposure effect size")+  # abs(beta_BMI)
-  labs(title = "Figure 2-update",
+  labs(title = "",
        subtitle = "Define different genetic architectures",
@@
     #ylab("% variance explained by IVs")+
-    labs(title = "Figure 1", subtitle = plotsubtitle,
+    labs(title = "", subtitle = plotsubtitle,
          col='', 
@@
 
-```{r fig.height=10, fig.width=14, warning=FALSE}
+```{r FINAL-PLOT-1-huray, fig.height=10, fig.width=14, message=FALSE, warning=FALSE}
 fig01b.update + 
@@
 
-final_plot1.arrange
-final_plot1.arrange2
-
-filename_final_plot1 <- paste0("../output/plots/", Sys.Date(), "_final_plot1.arrange.png")
-filename_final_plot1b <- paste0("../output/plots/", Sys.Date(), "_final_plot1b.arrange.png")
+# final_plot1.arrange   # OLD
+final_plot1.arrange2   # use this!
+
+filename_final_plot1 <- paste0("../output/plots/", Sys.Date(), "_final_plot1.arrange_OLD.png")
+filename_final_plot1b <- paste0("../output/plots/", Sys.Date(), "_final_plot1b.arrange_NEW.png")
 
@@
 
-```{r forestplot-MendelianRandomization-GSD, fig.width=5, fig.height=24, warning=FALSE}
+```{r forestplot-MendelianRandomization-GSD, fig.width=5, fig.height=27, warning=FALSE}
 p2.gsd <- mr_forest(
@@
 #SAVE
-filename.mr.forest.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_gsd.png")
+filename.mr.forest.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_GSD.png")
 ggsave(filename = filename.mr.forest.gsd, plot = p2.gsd, 
@@
 ```{r funnel-plot-GSD, fig.height=3, fig.width=5}
-p4.gsd <- mr_funnel(mr.obj, CI=TRUE)+
-  ggthemes::theme_few()
+p4.gsd <- mr_funnel(mr.obj, CI=TRUE)
 
@@
 # plot
+p4.gsd <- p4.gsd+
+  ggthemes::theme_few()
 p4.gsd
@@
 
-```{r forestplot-TwoSampleMR, fig.width=5, fig.height=24, warning=FALSE}
+```{r forestplot-TwoSampleMR-GBC, fig.width=5, fig.height=24, warning=FALSE}
 res_single <- TwoSampleMR::mr_singlesnp(harmonised_dat4TwoSampleMR, all_method = c("mr_egger_regression", "mr_ivw"))
@@
 #SAVE
-filename.mr.forest <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest.png")
+filename.mr.forest <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_GBC.png")
 ggsave(filename = filename.mr.forest, plot = p2[[1]], 
@@
     sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
-    limits = c(0,10)
+    limits = c(0,7.5)
   ) + 
@@
   ) -> plot_pve_simulated
-plot_pve_simulated
+#plot_pve_simulated
 
@@
 
-## Performance Metrics BIAS, MSE, MEAN, EMP_SE
+## Performance Metrics
+
+### Calculate BIAS, MSE, MEAN, EMP_SE
 
@@
 
-## Performance Metrics COVERAGE, COVERAGE-BIAS-ELIMINATED, TYPE1ERROR
+### Calculate COVERAGE, COVERAGE-BIAS-ELIMINATED, TYPE1ERROR
 
@@
 
-```{r add-additional-metrics-to-df}
-## GSD
+### GSD Performance Metrics
+
+```{r add-additional-metrics-to-df-and-print-GSD}
+## GSD --------------------------
 IVW_GSD_performance_metrics <- IVW_GSD_performance_metrics %>% 
@@
   as_tibble()
+IVW_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "IVW method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
+
 MODE_GSD_performance_metrics <- MODE_GSD_performance_metrics %>% 
   select(N, everything()) %>% 
-  cbind(data.frame(MODE_GSD_coverage_calc, MODE_GSD_coverage_bias_eliminated_calc, MODE_GSD_type1error_calc)) %>% 
+  cbind(data.frame(MODE_GSD_coverage_calc, 
+                   MODE_GSD_coverage_bias_eliminated_calc, MODE_GSD_type1error_calc)) %>% 
   rename(coverage = MODE_GSD_coverage_calc, 
@@
   as_tibble()
+
+MODE_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "MODE method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
+
 CONMIX_GSD_performance_metrics <- CONMIX_GSD_performance_metrics %>% 
   select(N, everything()) %>% 
-  cbind(data.frame(CONMIX_GSD_coverage_calc, CONMIX_GSD_coverage_bias_eliminated_calc, CONMIX_GSD_type1error_calc)) %>% 
+  cbind(data.frame(CONMIX_GSD_coverage_calc, 
+                   CONMIX_GSD_coverage_bias_eliminated_calc, CONMIX_GSD_type1error_calc)) %>% 
   rename(coverage = CONMIX_GSD_coverage_calc, 
@@
   as_tibble()
-
-## GBC
+CONMIX_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "CONMIX method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
+```
+
+### GBC Performance Metrics
+
+```{r add-additional-metrics-to-df-and-print-GBC}
+## GBC --------------------------
 IVW_GBC_performance_metrics <- IVW_GBC_performance_metrics %>% 
   select(N, everything()) %>% 
-  cbind(data.frame(IVW_GBC_coverage_calc, IVW_GBC_coverage_bias_eliminated_calc, IVW_GBC_type1error_calc)) %>% 
+  cbind(data.frame(IVW_GBC_coverage_calc, 
+                   IVW_GBC_coverage_bias_eliminated_calc, IVW_GBC_type1error_calc)) %>% 
   rename(coverage = IVW_GBC_coverage_calc, 
@@
   as_tibble()
+IVW_GBC_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "IVW method on BMI -> GBC performance metrics") %>% 
+  kable_styling(font_size=10)
+
 MODE_GBC_performance_metrics <- MODE_GBC_performance_metrics %>% 
   select(N, everything()) %>% 
-  cbind(data.frame(MODE_GBC_coverage_calc, MODE_GBC_coverage_bias_eliminated_calc, MODE_GBC_type1error_calc)) %>% 
+  cbind(data.frame(MODE_GBC_coverage_calc, 
+                   MODE_GBC_coverage_bias_eliminated_calc, MODE_GBC_type1error_calc)) %>% 
   rename(coverage = MODE_GBC_coverage_calc, 
@@
   as_tibble()
+MODE_GBC_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "MODE method on BMI -> GBC performance metrics") %>% 
+  kable_styling(font_size=10)
 CONMIX_GBC_performance_metrics <- CONMIX_GBC_performance_metrics %>% 
   select(N, everything()) %>% 
-  cbind(data.frame(CONMIX_GBC_coverage_calc, CONMIX_GBC_coverage_bias_eliminated_calc, CONMIX_GBC_type1error_calc)) %>% 
+  cbind(data.frame(CONMIX_GBC_coverage_calc, 
+                   CONMIX_GBC_coverage_bias_eliminated_calc, CONMIX_GBC_type1error_calc)) %>% 
   rename(coverage = CONMIX_GBC_coverage_calc, 
@@
   as_tibble()
+CONMIX_GBC_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "CONMIX method on BMI -> GBC performance metrics") %>% 
+  kable_styling(font_size=10)
 ```
@@
 
-```{r plot-performance-metrics-bias-mse-GSD, fig.height=10, fig.width=5}
+```{r plot-performance-metrics-bias-mse-GSD, fig.height=10, fig.width=8}
 ## IVW
@@
 
-```{r plot-performance-metrics-bias-mse-GBC, fig.height=10, fig.width=5}
+```{r plot-performance-metrics-bias-mse-GBC, fig.height=10, fig.width=8}
 ## IVW

### 2023-02-01 scripts/helper/perform_mr.R
restructuring

@@ -24,7 +24,7 @@ library(MRMix)
 library(MRPRESSO)
 library(mr.raps)
 library(penalized)
-source("./scripts/MR_lasso.R")
+source("./scripts/helper/MR_lasso.R")
 
 ## load data created in MSc.rmd
 my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))

### 2023-02-01 scripts/helper/plotfigure01fn.R

@@ -1,8 +1,9 @@
 
 
+library(tibble)
+library(ggplot2)
 
-
-plotfigure01 <- function(data, effect_size="all", maf="all"){
+plotfigure01 <- function(data, effect_size="all", maf="all", n){
   ## data: my_data_harm
   # effect size: "all", "strong" or "weak"
   # maf: "all", "rare" or "common
@@ -162,5 +163,3 @@ plotfigure01 <- function(data, effect_size="all", maf="all"){
   #list2env(assign(plotname, figure01.quadrant), envir = .GlobalEnv)
 }
 
-
-plotfigure01(my_data_harm)

### 2023-02-01 scripts/MSc.Rmd
updating simulation study

@@
 library(conflicted)
-source("MR_lasso.R")  # Slob&Burgess 2020
-source("PVE_calculation_fns.R") 
+source("helper/MR_lasso.R")  # Slob&Burgess 2020
+source("helper/PVE_calculation_fns.R") 
+source("helper/plotfigure01fn.R") 
 conflict_prefer("select", "dplyr")
@@
 
--   2022-11-12:
-    -   8 SNPs removed after data harmonization, one SNP associated with FTO gene (none left)
-    -   in new formula for proportion of variance in phenotype explained by given SNP (PVE) we need the MAF not the EAF
-        - calculate: if $eaf>0.5$ -> $maf = 1-eaf$
-        - need to change betas as well? +-
-    -   We discussed $K=0.05$ in our last meeting for the "prevalence" of GBC after looking at [Cancer Today](https://gco.iarc.fr/today/home) 
-        - If done this way we yield a 20% proportion of variance explained in PVE. Switch back to ASR used last week!
-
--   2022-11-15: 
-    1. Check if formula PVE numeric can be used for binary traits as well. Can the $\%var(X)$ – numeric – be compared with $\%var(U)$ $\%var(Y)$ – binary?
-    2. How is $\%var$ calculated in the MRbase package in the `mr_steiger` function (Steiger filtering)?
-    3. Prefer new formula for $\%var(X)$ calculation
-    4. FTO gene: Add SNP associated with FTO gene to analysis that was considered ambiguous. Check the beta estimate from UK Biobank (provide rsID to Felix: `rs1558902`)
-    5. Try also $K_{GBC}= 0.05/100$ for $\%var(Y)$ (before used $ASR=1.2/100.000$)
-    6. Update „Figure 1“ to 6 parts
-    7. Also run BMI -> GSD, FTO -> GSD + BMI -> GBC, FTO->GBC: How different are the MR methods in estimating the causal estimate?
-    8. Reference estimates for simulation: Wald estimate (FTO variant) + IVW total sample + conmix total sample
-    9. Plot PVE from all simulated variants as boxplots on each dot of Figure 01
-
--   2022-12-05
-    1.
-How did the large CI for the simulated PVEs arise?
-The simulated beta coefficients were drawn from given betas **AND** sample-size adjusted standard errors `*get simulated se_{sample size}*`.
-Therefore, for small sample sizes we yielded large variance for the beta coefficients.
-I changed the simulation and only used the given standard errors for our simulation scenarios in "scripts/simstudy.R".
+-   NA
 
@@
 
-- Issues from discussion mentioned above
-- Interchanged panel A and panel B in the first version of Figure 1, used the same colours in panel B and in the legends
-- Added five estimates for "BMI -> GSD": Wald estimate (FTO), Wald estimate (MC4R variant), random IVW, MRMix, weighted mode
+-   All genetic associations with BMI > 0 (for association statistics < 0, sign of the association statistic with GSD changed respectively)
 
@@
 GBC_dat <- as_tibble(read.delim("../data-boekstegers_felix/C23C24_BMI_association_results.txt"))
+
+# FTO data
+FTO_dat <- as_tibble(read.delim("../data-boekstegers_felix/2023-01-19_FTO/FTO_association_results.txt"))
 ```
@@
 my_data_harm <- my_data_harm %>% 
-  filter(mr_keep==TRUE | SNP == "rs1558902")  # also include FTO SNP
+  filter(mr_keep==TRUE | SNP == "rs1558902") %>%  # also include FTO SNP 
+  # change sign of beta after checking consistency of data (see mail felix 2023-01-19)
+  mutate(beta.bmi = ifelse(SNP == "rs1558902", -beta.bmi, beta.bmi))
+
+my_data_harm <- my_data_harm %>% 
+  mutate(
+    beta.gsd = ifelse(beta.bmi<0, -beta.gsd, beta.gsd),
+    beta.gbc = ifelse(beta.bmi<0, -beta.gbc, beta.gbc),
+    # change last beta.bmi, otherwise not negative any more ;)
+    beta.bmi = ifelse(beta.bmi<0, -beta.bmi, beta.bmi)
+  )
 ```
@@
   se_update[i] <- my_data_harm[["beta.bmi"]][i]/-qnorm(p=(my_data_harm[["p.value.bmi"]][i]/2),mean=0)
-
   # 2) Calculate the approximate SEs for decreasing samples sizes
@@
   se_n <- sd/(sqrt(n))
-
   # 3) Calculate the probability values considering the estimated betas and the corresponding SEs
@@
 #   count()
-
 my_data_harm <- as_tibble(cbind(my_data_harm, pvals_df))
@@
 
-## CALCULATE explained variance
+## CALCULATE explained variance for total sample size
 my_data_harm <- my_data_harm %>%
-  mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
+  mutate(
+    explained_variance_BMI = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi),
     # USE THIS !
-         explained_variance_BMI2 = explained_variance_numeric2(maf=eaf.bmi, 
-                                                           beta=beta.bmi, 
-                                                           se_beta=se.bmi, 
-                                                           samplesize=n[length(n)]
-
+    explained_variance_BMI2 = explained_variance_numeric2(
+      maf = eaf.bmi, beta = beta.bmi,
+      se_beta = se.bmi,
+      samplesize = n[length(n)]
     ),
-         explained_variance_GSD = explained_variance_binary(PA = eaf.gsd, 
+    explained_variance_GSD = explained_variance_binary(
+      PA = eaf.gsd,
       RR1 = exp(beta.gsd),
-                                                            RR2 = 1+(2*(exp(beta.gsd)-1)), 
+      RR2 = 1 + (2 * (exp(beta.gsd) - 1)),
       K = 0.1 # assumed prevalence in population
     )$Vg,
-         explained_variance_GBC = explained_variance_binary(PA = eaf.gbc, 
+    explained_variance_GBC = explained_variance_binary(
+      PA = eaf.gbc,
       RR1 = exp(beta.gbc),
-                                                            RR2 = 1+(2*(exp(beta.gbc)-1)), # exp(beta.gbc)^2, 
-                                                            K =1.2/100000  # assumed prevalence in population
+      RR2 = 1 + (2 * (exp(beta.gbc) - 1)), # exp(beta.gbc)^2,
+      K = 1.2 / 100000 # assumed prevalence in population
     )$Vg
-
-         )
+  )
+
 # sum(my_data_harm$explained_variance_GBC)
@@
 
-filenamefig00_a <- paste0("../output/plots/", Sys.Date(), "_explained_var_X_old_formula.png")
-ggsave(filename = filenamefig00_a, plot = fig00, 
-       width = 10, height = 6, 
-       units = "in"  # default
-       )
+# filenamefig00_a <- paste0("../output/plots/", Sys.Date(), "_explained_var_X_old_formula.png")
+# ggsave(filename = filenamefig00_a, plot = fig00, 
+#        width = 10, height = 6, 
+#        units = "in"  # default
+#        )
 
@@
 
+threshold <- 5*10^(-6)
 for (i in 1:length(colnames_pval) ) {
-  threshold <- 5*10^(-6)
   select_col <- colnames_pval[i]
@@
 }
+
+# for the numeric traits we have to map over sample size and recalculate se_updated
+n.orig_sqrt <- sqrt(n.orig)
+explained_variance_BMI2_sample_size <- map(n, ~explained_variance_numeric2(maf = my_data_harm$eaf.bmi, beta = my_data_harm$beta.bmi, 
+                                                                           se_beta = abs(my_data_harm$beta.bmi / -qnorm(p=(my_data_harm$p.value.bmi / 2), mean = 0) * n.orig_sqrt / sqrt(.x)), 
+                                                                           samplesize = .x))
+
+cum_expl_varX2_sample_size <- c()
+for (mysamples in 1:length(colnames_pval)) {
+  temp_dat <- my_data_harm %>% 
+    select(all_of(colnames_pval[mysamples])) %>% 
+    mutate(explained_variance_BMI2_sample_size = explained_variance_BMI2_sample_size[[mysamples]]) %>% 
+    filter(!!sym(colnames_pval[mysamples]) < threshold)
+  #print(temp_dat)
+  cum_expl_varX2_sample_size <- append(cum_expl_varX2_sample_size, sum(temp_dat$explained_variance_BMI2_sample_size))
+}
+
+df_cum_expl_var <- df_cum_expl_var %>% 
+  cbind(cum_expl_varX2_sample_size) %>% 
+  relocate(cum_expl_varX2_sample_size, .before = cum_expl_varU)
+
+ggplot(df_cum_expl_var, aes(x=n))+
+  geom_line(aes(y=cum_expl_varX2_sample_size*100), col="red")+
+  geom_line(aes(y=cum_expl_varX2*100))
 ```
@@
     cum_expl_varX = cum_expl_varX *100, # in percent 
-    cum_expl_varX2 = cum_expl_varX2 *100,
+    cum_expl_varX2 = cum_expl_varX2_sample_size *100,  #  CAVE CHANGED HERE TO INCLUDE SAMPLE SIZE DEPENDENT SE and PVE CALC!
     cum_expl_varU = cum_expl_varU *100,
@@
     count_IVs_scaled = count_IVs*scale_factor)
-
-plot_data1 <- df_cum_expl_var %>% 
-  pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, cum_expl_varY)) %>% 
-  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
-         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
-         )
-
-N_annotation <- c(10000, 100000, 200000, 300000)
-## plot
-ggplot(plot_data1, aes(x=n, y=value, col=name))+
-  geom_point(alpha=0.6)+
-  geom_line(alpha=0.6)+
-  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
-  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
-                     labels = comma, 
-                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
-                     limits = c(10000, 300000))+
-  #ylab("% variance explained by IVs")+
-  labs(title = "Figure 1",
-       col='', 
-       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
-       ) +
-  scale_y_continuous(
-    # Features of the first axis
-    name = "% variance explained by IVs",
-    # Add a second axis and specify its features
-    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
-  ) + 
-  theme_few()+
-  #scale_colour_few() +
-  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
-  #                                         "% var(X)" = "blue", 
-  #                                         "% var(Y)" = "darkblue", 
-  #                                         "Number of IVs" = "black"
-  #                                         ))
-  #
-  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
-                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
-                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
-                                           "Number of IVs" = "black"),
-                     # order
-                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
-  ) -> fig01.update
-
-# fig01.update   # old formula for PVE exposure X (without sample size adjusting)
-# 
-# scale_color_manual(name = "Y series", values = c("Y1" = "darkblue", "Y2" = "red"))
 
@@
 
-# SAVE
-filenamefig01_a <- paste0("../output/plots/", Sys.Date(), "_figure01_old_PVE_X_formula.png")
 filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_new_PVE_X_formula.png")
-ggsave(filename = filenamefig01_a, plot = fig01.update, 
-       width = 10, height = 6, 
-       units = "in"  # default
-       )
-
 ggsave(filename = filenamefig01_b, plot = fig01b.update, 
@@
   mutate(maf.bmi = ifelse(eaf.bmi >0.5, 1-eaf.bmi, eaf.bmi)) %>% 
-  select(maf.bmi, beta.bmi)
-
-ggplot(data = plot_data_2, aes(x=maf.bmi, y=abs(beta.bmi)))+
+  select(maf.bmi, beta.bmi, gene_label, fto_col_annot)
+
+ggplot(data = plot_data_2, aes(x=maf.bmi, y=abs(beta.bmi), 
+                               col=fto_col_annot, label=gene_label))+
   geom_point(alpha=0.6)+
+  ggrepel::geom_text_repel(show.legend = FALSE)+
   geom_hline(yintercept=median(abs(plot_data_2$beta.bmi)), 
@@
        subtitle = "Define different genetic architectures",
-       #col='', 
+       col='SNP', 
        #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
@@
 ```{r figure1quadrants, fig.show="hold", out.width="50%", warning=FALSE}
-
-plotfigure01 <- function(data, effect_size="strong", maf="rare"){
-  ## data: my_data_harm
-  # effect size: "strong" or "weak"
-  # maf: "rare" or "common
-  df_cum_expl_var <- tibble(n=n, 
-                            cum_expl_varX = rep(NA, length(n)), 
-                            cum_expl_varX2 = rep(NA, length(n)), 
-                            cum_expl_varU = rep(NA, length(n)), 
-                            cum_expl_varY = rep(NA, length(n)),
-                            count_IVs = rep(NA, length(n))
-  )
-  # create maf col
-  data <- data %>% 
-      mutate(maf.bmi = ifelse(eaf.bmi >0.5, 1-eaf.bmi, eaf.bmi))
-
-  ## filter for quadrant
-  if (effect_size=="strong" & maf == "rare") {
-    data <- data %>% 
-      filter(abs(beta.bmi) > median(abs(beta.bmi)) &  maf.bmi < median(maf.bmi))
-  } else if (effect_size=="strong" & maf == "common") {
-    data <- data %>% 
-      filter(abs(beta.bmi) > median(abs(beta.bmi)) &  maf.bmi > median(maf.bmi))
-  } else if (effect_size=="weak" & maf == "common") {
-    data <- data %>% 
-      filter(abs(beta.bmi) < median(abs(beta.bmi)) &  maf.bmi > median(maf.bmi))
-  } else if (effect_size=="weak" & maf == "rare") {
-    data <- data %>% 
-      filter(abs(beta.bmi) < median(abs(beta.bmi)) &  maf.bmi < median(maf.bmi))
-  }
-
-  # create data for plot
-  for (i in 1:length(colnames_pval) ) {
-    threshold <- 5*10^(-6)
-    select_col <- colnames_pval[i]
-    # calculate cum explained variance and count of IVs
-    summarydat <- data %>% 
-      filter(!!sym(select_col) < threshold) %>% 
-      select(explained_variance_BMI, explained_variance_BMI2, explained_variance_GSD, explained_variance_GBC) %>% 
-      summarise(expl_variances =across(.cols = everything(), function(x) sum(x)),
-                count_IVs = n())
-    # fill up df
-    df_cum_expl_var[i, 2] <- summarydat$expl_variances[1]
-    df_cum_expl_var[i, 3] <- summarydat$expl_variances[2]
-    df_cum_expl_var[i, 4] <- summarydat$expl_variances[3]
-    df_cum_expl_var[i, 5] <- summarydat$expl_variances[4]
-    df_cum_expl_var[i, 6] <- summarydat$count_IVs
-  }
-
-  ## plot fig 01 
-  scale_factor <- 4.5/284
-  df_cum_expl_var <- df_cum_expl_var %>% 
-    mutate(
-      cum_expl_varX = cum_expl_varX *100, # in percent 
-      cum_expl_varX2 = cum_expl_varX2 *100,
-      cum_expl_varU = cum_expl_varU *100,
-      cum_expl_varY = cum_expl_varY *100,
-      count_IVs_scaled = count_IVs*scale_factor)
-
-  plot_data1 <- df_cum_expl_var %>% 
-    pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, cum_expl_varY)) %>% 
-    mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
-           name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
-           )
-
-  N_annotation <- c(10000, 100000, 200000, 300000)
-  ## plot
-  plotsubtitle <- paste0("relatively ", maf ,", relatively ", effect_size)
-  ggplot(plot_data1, aes(x=n, y=value, col=name))+
-    geom_point(alpha=0.6)+
-    geom_line(alpha=0.6)+
-    geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
-    scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
-                       labels = comma, 
-                       breaks = N_annotation, #c( 10000, n[7:length(n)]),
-                       limits = c(10000, 300000))+
-    #ylab("% variance explained by IVs")+
-    labs(title = "", subtitle = plotsubtitle,
-         col='', 
-         caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits   explained by the IVs"
-         ) +
-    scale_y_continuous(
-      # Features of the first axis
-      name = "% variance explained by IVs",
-      # Add a second axis and specify its features
-      sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
-      limits = c(0,2)
-    ) + 
-    theme_few()+
-    #scale_colour_few() +
-    #scale_color_manual(name = "", values = c("% var(U)" = "red", 
-    #                                         "% var(X)" = "blue", 
-    #                                         "% var(Y)" = "darkblue", 
-    #                                         "Number of IVs" = "black"
-    #                                         ))
-    #
-    scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
-                                             "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
-                                             "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
-                                             "Number of IVs" = "black"),
-                       # order
-                       breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
-    ) -> figure01.quadrant
-  print(figure01.quadrant)
-
-  filenamefig01_quadrant <- paste0("../output/plots/", Sys.Date(), "_figure01_", 
-                                   "EFFECT_SIZE_", effect_size,
-                                   "_VARIANTS_", maf,
-                                   ".png")
-  ggsave(filename = filenamefig01_quadrant, plot = figure01.quadrant, 
-         width = 10, height = 6, 
-         units = "in"  # default
-         )
-
-  # store plot in global environment
-  plotname <<- paste0("figure01_EFFECT_SIZE_", effect_size, "_VARIANTS_", maf)
-  assign(plotname, figure01.quadrant, envir = .GlobalEnv)
-  #list2env(assign(plotname, figure01.quadrant), envir = .GlobalEnv)
-
-}
-
+## plotfigure sourced from helper functions ----------------------------------
 # create all combis
@@
                effect_size = all_quadrants$effect_size[i], 
-               maf = all_quadrants$maf[i])
+               maf = all_quadrants$maf[i], 
+               n=n
+               )
 }
@@
 
-## load results
+## load results from script `perform_mr.R`
 est.bmi.gsd <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
@@
 )
+
+fto_wald_ukbiobank <- mr_wald_ratio(
+  b_exp = exp(FTO_dat$beta[3])-1, # BMI UK Biobank log-transformed  (before Hoffmann GERA + GIANT)
+  b_out = FTO_dat$beta[2], # GSD UKB as before
+  se_exp = exp(FTO_dat$se[3])-1, 
+  se_out = FTO_dat$se[2]
+  )
+
+fto_wald_ukbiobank_noexp <- mr_wald_ratio(
+  b_exp = (FTO_dat$beta[3]), # BMI UK Biobank log-transformed  (before Hoffmann GERA + GIANT)
+  b_out = FTO_dat$beta[2], # GSD UKB as before
+  se_exp = (FTO_dat$se[3]), 
+  se_out = FTO_dat$se[2]
+  )
+
 mc4r_wald <- mr_wald_ratio(
@@
   add_row(mr_methods = "FTO_Wald", estimate = fto_wald$b, se.estimate=fto_wald$se) %>% 
-  add_row(mr_methods = "MC4R_Wald", estimate = mc4r_wald$b, se.estimate=mc4r_wald$se)
+  add_row(mr_methods = "MC4R_Wald", estimate = mc4r_wald$b, se.estimate=mc4r_wald$se) %>% 
+  arrange(desc(estimate))
 
@@
 res <- TwoSampleMR::mr(harmonised_dat4TwoSampleMR)
-res <- res %>% mutate(OR = exp(b), CI.lower= exp(b-se), CI.upper=exp(b+se))
+res <- res %>% mutate(OR = exp(b), CI.lower= exp(b-se), CI.upper=exp(b+se)) %>% 
+  arrange(desc(b))
 
@@
   mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
+  arrange(desc(estimate)) %>% 
   kable(digits=3, caption = "MR Results for GBC (outcome) using genetic variants as instrumental variables for BMI") %>% 
@@
 colname_pve_X <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_X_sim")]
-colname_pve_U <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_U_sim")]
-colname_pve_Y <-  colnames(est_sim)[str_detect(colnames(est_sim), pattern = "pve_Y_sim")]
+# PVE_Y, PVE_U removed on 2023-01-24 (see github)
 
@@
 colname_CONMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GSD_")]
+colname_MRMIX_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MRMIX_theta_sim_BMI_GSD_")]
+colname_MRMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MRMIX_theta_se_sim_BMI_GSD_")]
 # retrieve IVW/ MODE/ CONMIX cols for BMI -> GBC
-colname_IVW_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GBC_")]
-colname_IVW_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GBC_")]
-colname_MODE_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GBC_")]
-colname_MODE_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GBC_")]
-colname_CONMIX_theta_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GBC_")]
-colname_CONMIX_theta_se_sim_BMI_GBC <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GBC_")]
+# removed on 2023-01-24 (see github)
 
@@
   cbind(sample_size_filtered) %>% as_tibble() -> pve_X_sim
-# CONFOUNDER PVE
-est_sim %>% 
-  select(!!!(colname_pve_U)) %>% 
-  pivot_longer(everything(), names_to = "PVE_U", values_to = "value_U") %>% 
-  cbind(sample_size_filtered) %>% as_tibble() -> pve_U_sim
-# OUTCOME PVE
-est_sim %>% 
-  select(!!!(colname_pve_Y)) %>% 
-  pivot_longer(everything(), names_to = "PVE_Y", values_to = "value_Y") %>% 
-  cbind(sample_size_filtered) %>% as_tibble() -> pve_Y_sim
+# # CONFOUNDER PVE
+# est_sim %>% 
+#   select(!!!(colname_pve_U)) %>% 
+#   pivot_longer(everything(), names_to = "PVE_U", values_to = "value_U") %>% 
+#   cbind(sample_size_filtered) %>% as_tibble() -> pve_U_sim
+# # OUTCOME PVE
+# est_sim %>% 
+#   select(!!!(colname_pve_Y)) %>% 
+#   pivot_longer(everything(), names_to = "PVE_Y", values_to = "value_Y") %>% 
+#   cbind(sample_size_filtered) %>% as_tibble() -> pve_Y_sim
 
 boxplot_overlay_dat <- pve_X_sim %>% 
-  cbind(pve_U_sim %>% select(-sample_size_filtered)) %>% 
-  cbind(pve_Y_sim %>% select(-sample_size_filtered)) %>% 
+  # cbind(pve_U_sim %>% select(-sample_size_filtered)) %>% 
+  # cbind(pve_Y_sim %>% select(-sample_size_filtered)) %>% 
   as_tibble()
@@
                alpha=0.2, size=0.2)+
-  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_U*100, # *100 to get percent values
-                                group=sample_size_filtered), 
-               col=ggthemes_data$few$colors$Dark[2,2][[1]],
-               alpha=0.2, size=0.2)+
-  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_Y*100, # *100 to get percent values
-                                group=sample_size_filtered), 
-               col=ggthemes_data$few$colors$Dark[4,2][[1]],
-               alpha=0.2, size=0.2)+
+  # U/Y removed on 2023-01-24 (see github)
   theme_few()+
@@
     side = "l", range_scale = .4, alpha = .2)+
-  # confounder simulation
-  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_U*100, # *100 to get percent values
-                                               group=sample_size_filtered), 
-               col=ggthemes_data$few$colors$Dark[2,2][[1]],
-               alpha=0.2, size=0.2, outlier.shape = NA)+
-  ggdist::stat_halfeye(data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered, y=value_U*100, group=sample_size_filtered), 
-                       inherit.aes = FALSE, 
-                       slab_fill=ggthemes_data$few$colors$Dark[2,2][[1]], slab_alpha = 0.2,
-                       geom = "slabinterval",
-                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
-  gghalves::geom_half_point(
-    data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered-2000, y=value_U*100, group=sample_size_filtered), 
-    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[2,2][[1]],
-    side = "l", range_scale = .4, alpha = .2)+
-  # outcome simulation
-  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_Y*100, # *100 to get percent values
-                                               group=sample_size_filtered), 
-               col=ggthemes_data$few$colors$Dark[4,2][[1]],
-               alpha=0.2, size=0.2, outlier.shape = NA)+
-  ggdist::stat_halfeye(data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered, y=value_Y*100, group=sample_size_filtered), 
-                       inherit.aes = FALSE, 
-                       slab_fill=ggthemes_data$few$colors$Dark[4,2][[1]], slab_alpha = 0.2,
-                       geom = "slabinterval",
-                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
-  gghalves::geom_half_point(
-    data = boxplot_overlay_dat, mapping=aes(x=sample_size_filtered-2000, y=value_Y*100, group=sample_size_filtered), 
-    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[4,2][[1]],
-    side = "l", range_scale = .4, alpha = .2)+
   coord_cartesian(xlim=c(10000, 300000))+
@@
                      )
+args.gsd.mrmix <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "MRMix"], 
+                     nsim=nrow(est_sim)
+                     )
 ## args gbc
-args.gbc.ivw <- list(theta = est$estimate[est$mr_methods == "IVW"], 
-                     nsim=nrow(est_sim)
-                     ) 
-args.gbc.mode <- list(theta = est$estimate[est$mr_methods == "mode"], 
-                     nsim=nrow(est_sim)
-                     )
-args.gbc.conmix <- list(theta = est$estimate[est$mr_methods == "conmix"], 
-                     nsim=nrow(est_sim)
-                     )
+#deleted 2023-01-29
+
 ## Standard Deviation
@@
   rename(N = sample_size_filtered)
-```
-
-```{r CALC-perf-metrics-GBC}
-IVW_GBC_performance_metrics <- est_sim %>% 
-  select(!!!(colname_IVW_theta_sim_BMI_GBC)) %>% 
+
+MRMIX_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_MRMIX_theta_sim_BMI_GSD)) %>% 
   map_df(.f = ~ funs %>% 
-           map(exec, .x, !!!args.gbc.ivw), .id = "var") %>% 
+           map(exec, .x, !!!args.gsd.conmix), .id = "var") %>% 
   cbind(sample_size_filtered) %>% 
   rename(N = sample_size_filtered)
-
-MODE_GBC_performance_metrics <- est_sim %>% 
-  select(!!!(colname_MODE_theta_sim_BMI_GBC)) %>% 
-  map_df(.f = ~ funs %>% 
-           map(exec, .x, !!!args.gbc.mode), .id = "var") %>% 
-  cbind(sample_size_filtered) %>% 
-  rename(N = sample_size_filtered)
-
-CONMIX_GBC_performance_metrics <- est_sim %>% 
-  select(!!!(colname_CONMIX_theta_sim_BMI_GBC)) %>% 
-  map_df(.f = ~ funs %>% 
-           map(exec, .x, !!!args.gbc.conmix), .id = "var") %>% 
-  cbind(sample_size_filtered) %>% 
-  rename(N = sample_size_filtered)
+```
+
+```{r CALC-perf-metrics-GBC}
+# deleted 2023-01-29
 ```
@@
 CONMIX_GSD_type1error_calc <- c()
+MRMIX_GSD_coverage_calc <- c()
+MRMIX_GSD_coverage_bias_eliminated_calc <- c()
+MRMIX_GSD_type1error_calc <- c()
 # GBC
-IVW_GBC_coverage_calc <- c()
-IVW_GBC_coverage_bias_eliminated_calc <- c()
-IVW_GBC_type1error_calc <- c()
-MODE_GBC_coverage_calc <- c()
-MODE_GBC_coverage_bias_eliminated_calc <- c()
-MODE_GBC_type1error_calc <- c()
-CONMIX_GBC_coverage_calc <- c()
-CONMIX_GBC_coverage_bias_eliminated_calc <- c()
-CONMIX_GBC_type1error_calc <- c()
+## deleted 2023-01-29
 for (i in 1:length(sample_size_filtered)) {
   sample_size <- sample_size_filtered[i]
-  ## GSD additional metrics ----------------------------------------------------
-  # IVW
+  # GSD additional metrics ----------------------------------------------------
+  ## 1/ IVW
   col_IVW_GSD_theta <- paste0("IVW_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
@@
   IVW_GSD_type1error_calc[i] <- type1error(est_sim[[col_IVW_GSD_mr_pval]])
-  # MODE
+  ## 2/ MODE
   col_MODE_GSD_theta <- paste0("MODE_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
@@
   MODE_GSD_type1error_calc[i] <- type1error(est_sim[[col_MODE_GSD_mr_pval]])
-  # CONMIX
+  ## 3/ CONMIX
   col_CONMIX_GSD_theta <- paste0("CONMIX_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
@@
   CONMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_CONMIX_GSD_mr_pval]])  
+    ## 4/ MRMIX
+  col_MRMIX_GSD_theta <- paste0("MRMIX_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_MRMIX_GSD_theta_se <- paste0("MRMIX_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_MRMIX_GSD_mr_pval <- paste0("MRMIX_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  MRMIX_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MRMIX_GSD_theta], 
+                                       thetaobs_se = est_sim[col_MRMIX_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "MRMix"])
+  MRMIX_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MRMIX_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_MRMIX_GSD_theta_se]])
+  MRMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_MRMIX_GSD_mr_pval]])
 
   ## GBC additional metrics ----------------------------------------------------
-  # IVW
-  col_IVW_GBC_theta <- paste0("IVW_theta_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
-  col_IVW_GBC_theta_se <- paste0("IVW_theta_se_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
-  col_IVW_GBC_mr_pval <- paste0("IVW_mr_pval_BMI_GBC_", format(sample_size, scientific = FALSE))
-  # calc
-  IVW_GBC_coverage_calc[i] <- coverage(thetaobs = est_sim[col_IVW_GBC_theta], thetaobs_se = est_sim[col_IVW_GBC_theta_se], theta = est$estimate[est$mr_methods == "IVW"])
-  IVW_GBC_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_IVW_GBC_theta]], thetaobs_se = est_sim[[col_IVW_GSD_theta_se]])
-  IVW_GBC_type1error_calc[i] <- type1error(est_sim[[col_IVW_GBC_mr_pval]])
-  # MODE
-  col_MODE_GBC_theta <- paste0("MODE_theta_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
-  col_MODE_GBC_theta_se <- paste0("MODE_theta_se_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
-  col_MODE_GBC_mr_pval <- paste0("MODE_mr_pval_BMI_GBC_", format(sample_size, scientific = FALSE))
-  # calc
-  MODE_GBC_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MODE_GBC_theta], 
-                                       thetaobs_se = est_sim[col_MODE_GBC_theta_se], 
-                                       theta = est$estimate[est$mr_methods == "mode"])
-  MODE_GBC_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MODE_GBC_theta]], 
-                                                                       thetaobs_se = est_sim[[col_MODE_GBC_theta_se]])
-  MODE_GBC_type1error_calc[i] <- type1error(est_sim[[col_MODE_GBC_mr_pval]])
-  # CONMIX
-  col_CONMIX_GBC_theta <- paste0("CONMIX_theta_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
-  col_CONMIX_GBC_theta_se <- paste0("CONMIX_theta_se_sim_BMI_GBC_", format(sample_size, scientific = FALSE))
-  col_CONMIX_GBC_mr_pval <- paste0("CONMIX_mr_pval_BMI_GBC_", format(sample_size, scientific = FALSE))
-  # calc
-  CONMIX_GBC_coverage_calc[i] <- coverage(thetaobs = est_sim[col_CONMIX_GBC_theta], 
-                                       thetaobs_se = est_sim[col_CONMIX_GBC_theta_se], 
-                                       theta = est$estimate[est$mr_methods == "conmix"])
-  CONMIX_GBC_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_CONMIX_GBC_theta]], 
-                                                                       thetaobs_se = est_sim[[col_CONMIX_GBC_theta_se]])
-  CONMIX_GBC_type1error_calc[i] <- type1error(est_sim[[col_CONMIX_GBC_mr_pval]])  
+  # deleted 2023-01-29
 }
@@
   as_tibble()
-
 MODE_GSD_performance_metrics %>% 
@@
   kable_styling(font_size=10)
-```
-
-### GBC Performance Metrics
-
-```{r add-additional-metrics-to-df-and-print-GBC}
-## GBC --------------------------
-IVW_GBC_performance_metrics <- IVW_GBC_performance_metrics %>% 
+
+MRMIX_GSD_performance_metrics <- MRMIX_GSD_performance_metrics %>% 
   select(N, everything()) %>% 
-  cbind(data.frame(IVW_GBC_coverage_calc, 
-                   IVW_GBC_coverage_bias_eliminated_calc, IVW_GBC_type1error_calc)) %>% 
-  rename(coverage = IVW_GBC_coverage_calc, 
-         coverage_bias_eliminated = IVW_GBC_coverage_bias_eliminated_calc, 
-         type1error = IVW_GBC_type1error_calc) %>% 
+  cbind(data.frame(MRMIX_GSD_coverage_calc, 
+                   MRMIX_GSD_coverage_bias_eliminated_calc, MRMIX_GSD_type1error_calc)) %>% 
+  rename(coverage = MRMIX_GSD_coverage_calc, 
+         coverage_bias_eliminated = MRMIX_GSD_coverage_bias_eliminated_calc, 
+         type1error = MRMIX_GSD_type1error_calc) %>% 
   as_tibble()
-IVW_GBC_performance_metrics %>% 
+MRMIX_GSD_performance_metrics %>% 
   mutate(across(cols = where(is.numeric), round(3))) %>% 
-  kable(caption = "IVW method on BMI -> GBC performance metrics") %>% 
-  kable_styling(font_size=10)
-
-MODE_GBC_performance_metrics <- MODE_GBC_performance_metrics %>% 
-  select(N, everything()) %>% 
-  cbind(data.frame(MODE_GBC_coverage_calc, 
-                   MODE_GBC_coverage_bias_eliminated_calc, MODE_GBC_type1error_calc)) %>% 
-  rename(coverage = MODE_GBC_coverage_calc, 
-         coverage_bias_eliminated = MODE_GBC_coverage_bias_eliminated_calc, 
-         type1error = MODE_GBC_type1error_calc) %>% 
-  as_tibble()
-MODE_GBC_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
-  kable(caption = "MODE method on BMI -> GBC performance metrics") %>% 
-  kable_styling(font_size=10)
-CONMIX_GBC_performance_metrics <- CONMIX_GBC_performance_metrics %>% 
-  select(N, everything()) %>% 
-  cbind(data.frame(CONMIX_GBC_coverage_calc, 
-                   CONMIX_GBC_coverage_bias_eliminated_calc, CONMIX_GBC_type1error_calc)) %>% 
-  rename(coverage = CONMIX_GBC_coverage_calc, 
-         coverage_bias_eliminated = CONMIX_GBC_coverage_bias_eliminated_calc, 
-         type1error = CONMIX_GBC_type1error_calc) %>% 
-  as_tibble()
-CONMIX_GBC_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
-  kable(caption = "CONMIX method on BMI -> GBC performance metrics") %>% 
+  kable(caption = "MRMIX method on BMI -> GSD performance metrics") %>% 
   kable_styling(font_size=10)
@@
                                                Sys.Date(), "_CONMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
+filename_MRMIX_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_MRMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
 write.csv2(x = IVW_GSD_performance_metrics, file = filename_IVW_GSD_performance_metrics)
@@
 write.csv2(x = CONMIX_GSD_performance_metrics, file = filename_CONMIX_GSD_performance_metrics)
-
-## GBC
-filename_IVW_GBC_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_IVW_GBC_performance_metrics_NSIM_", nsim, ".csv")
-filename_MODE_GBC_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_MODE_GBC_performance_metrics_NSIM_", nsim, ".csv")
-filename_CONMIX_GBC_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_CONMIX_GBC_performance_metrics_NSIM_", nsim, ".csv")
-write.csv2(x = IVW_GBC_performance_metrics, file = filename_IVW_GBC_performance_metrics)
-write.csv2(x = MODE_GBC_performance_metrics, file = filename_MODE_GBC_performance_metrics)
-write.csv2(x = CONMIX_GBC_performance_metrics, file = filename_CONMIX_GBC_performance_metrics)
+write.csv2(x = MRMIX_GSD_performance_metrics, file = filename_MRMIX_GSD_performance_metrics)
 ```
@@
 
+## MRMIX
+MRMIX_BMI_GSD_geom_point_dat <- est_sim %>% 
+  select(!!!(colname_MRMIX_theta_sim_BMI_GSD)) %>% 
+  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=mean))+ 
+  geom_point()+
+  geom_hline(yintercept=  args.gsd.mrmix$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+  annotate(geom = "text", x = 50000, y = args.gsd.mrmix$theta+0.05,
+         label= expression( paste("ref: ", theta[X] ))  , 
+         color= ggthemes_data$few$colors$Dark[9,2][[1]]
+           )+
+  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
+                width=3000, # x-axis dimension
+                position = "dodge")+
+  ## ADD rainbow plots
+  geom_boxplot(data = MRMIX_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+               col=ggthemes_data$few$colors$Dark[3,2][[1]],
+               alpha=0.2, size=0.2, outlier.shape = NA)+
+  ggdist::stat_halfeye(data = MRMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                       inherit.aes = FALSE, 
+                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                       geom = "slabinterval",
+                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+  gghalves::geom_half_point(
+    data = MRMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+    side = "l", range_scale = .4, alpha = .2)+
+  ## FINISH rainbow
+  scale_x_continuous(labels = comma)+
+  xlab("Size of Sample I")+
+  ylab("Mean thetas")+
+  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "MRMIX estimate: BMI -> GSD")+
+  theme_few() -> MRMIX_GSD_simplot_thetas
+MRMIX_GSD_simplot_thetas
+
 # SAVE
@@
        )
+filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_NSIM_", nsim,".png")
+ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
+       width = 15, height = 6, 
+       units = "in"  # default
+       )
 ```
@@
 CONMIX_GSD_simplot_thetas$labels$title <- ""
-ggarrange(IVW_GSD_simplot_thetas,MODE_GSD_simplot_thetas,CONMIX_GSD_simplot_thetas, 
-          ncol = 1, nrow = 3) -> GSD_simplot_thetas_arrange
+MRMIX_GSD_simplot_thetas$labels$title <- ""
+ggarrange(IVW_GSD_simplot_thetas,MODE_GSD_simplot_thetas,CONMIX_GSD_simplot_thetas, MRMIX_GSD_simplot_thetas,
+          ncol = 1, nrow = 4) -> GSD_simplot_thetas_arrange
 
@@
 ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
-       width = 10, height = 18, 
+       width = 10, height = 22, 
        units = "in"  # default
@@
 
-ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD,
-          ncol = 1, nrow = 3, common.legend = TRUE) -> bias_mse_plot_GSD
-bias_mse_plot_GSD
-
-filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_NSIM_", nsim, ".png")
-ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
-       width = 7, height = 10, 
-       units = "in"  # default
-       )
-```
-
-### GBC simulated theta distribution
-
-```{r theta-sim-distribution-GBC, fig.height=6, fig.width=10, warning=FALSE}
-## IVW
-IVW_BMI_GBC_geom_point_dat <- est_sim %>% 
-  select(!!!(colname_IVW_theta_sim_BMI_GBC)) %>% 
-  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
-  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
-
-ggplot(IVW_GBC_performance_metrics, aes(x=N, y=mean))+ 
-  geom_point()+
-  geom_hline(yintercept=  args.gbc.ivw$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  annotate(geom = "text", x = 50000, y = args.gbc.ivw$theta+0.05,
-         label= expression( paste("ref: ", theta[X] ))  , 
-         color= ggthemes_data$few$colors$Dark[9,2][[1]]
-           )+
-  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
-                width=3000, # x-axis dimension
-                position = "dodge")+
-  ## ADD rainbow plots
-  geom_boxplot(data = IVW_BMI_GBC_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
-               col=ggthemes_data$few$colors$Dark[3,2][[1]],
-               alpha=0.2, size=0.2, outlier.shape = NA)+
-  ggdist::stat_halfeye(data = IVW_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
-                       inherit.aes = FALSE, 
-                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
-                       geom = "slabinterval",
-                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
-  gghalves::geom_half_point(
-    data = IVW_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
-    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
-    side = "l", range_scale = .4, alpha = .2)+
-  ## FINISH rainbow
-  scale_x_continuous(labels = comma)+
-  xlab("Size of Sample I")+
-  ylab("Mean thetas")+
-  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "IVW estimate: BMI -> GBC")+
-  theme_few() -> IVW_GBC_simplot_thetas
-IVW_GBC_simplot_thetas
-
-## MODE
-MODE_BMI_GBC_geom_point_dat <- est_sim %>% 
-  select(!!!(colname_MODE_theta_sim_BMI_GBC)) %>% 
-  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
-  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
-
-ggplot(MODE_GBC_performance_metrics, aes(x=N, y=mean))+ 
-  geom_point()+
-  geom_hline(yintercept=  args.gbc.mode$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  annotate(geom = "text", x = 50000, y = args.gbc.mode$theta+0.05,
-         label= expression( paste("ref: ", theta[X] ))  , 
-         color= ggthemes_data$few$colors$Dark[9,2][[1]]
-           )+
-  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
-                width=3000, # x-axis dimension
-                position = "dodge")+
-  ## ADD rainbow plots
-  geom_boxplot(data = MODE_BMI_GBC_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
-               col=ggthemes_data$few$colors$Dark[3,2][[1]],
-               alpha=0.2, size=0.2, outlier.shape = NA)+
-  ggdist::stat_halfeye(data = MODE_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
-                       inherit.aes = FALSE, 
-                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
-                       geom = "slabinterval",
-                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
-  gghalves::geom_half_point(
-    data = MODE_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
-    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
-    side = "l", range_scale = .4, alpha = .2)+
-  ## FINISH rainbow
-  scale_x_continuous(labels = comma)+
-  xlab("Size of Sample I")+
-  ylab("Mean thetas")+
-  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "MODE estimate: BMI -> GBC")+
-  theme_few() -> MODE_GBC_simplot_thetas
-MODE_GBC_simplot_thetas
-
-## CONMIX
-CONMIX_BMI_GBC_geom_point_dat <- est_sim %>% 
-  select(!!!(colname_CONMIX_theta_sim_BMI_GBC)) %>% 
-  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
-  mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
-
-ggplot(CONMIX_GBC_performance_metrics, aes(x=N, y=mean))+ 
-  geom_point()+
-  geom_hline(yintercept=  args.gbc.conmix$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  annotate(geom = "text", x = 50000, y = args.gbc.conmix$theta+0.05,
-         label= expression( paste("ref: ", theta[X] ))  , 
-         color= ggthemes_data$few$colors$Dark[9,2][[1]]
-           )+
-  geom_errorbar(aes(ymin=mean-emp_se, ymax=mean+emp_se), 
-                width=3000, # x-axis dimension
-                position = "dodge")+
-  ## ADD rainbow plots
-  geom_boxplot(data = CONMIX_BMI_GBC_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
-               col=ggthemes_data$few$colors$Dark[3,2][[1]],
-               alpha=0.2, size=0.2, outlier.shape = NA)+
-  ggdist::stat_halfeye(data = CONMIX_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
-                       inherit.aes = FALSE, 
-                       slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
-                       geom = "slabinterval",
-                       na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
-  gghalves::geom_half_point(
-    data = CONMIX_BMI_GBC_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
-    inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
-    side = "l", range_scale = .4, alpha = .2)+
-  ## FINISH rainbow
-  scale_x_continuous(labels = comma)+
-  xlab("Size of Sample I")+
-  ylab("Mean thetas")+
-  labs(title = "Simulated Thetas with Empirical Standard Error", subtitle = "CONMIX estimate: BMI -> GBC")+
-  theme_few() -> CONMIX_GBC_simplot_thetas
-CONMIX_GBC_simplot_thetas
-
-# SAVE
-filename_IVW_GBC_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GBC_simplot_thetas_NSIM_", nsim,".png")
-ggsave(filename = filename_IVW_GBC_simplot_thetas, plot = IVW_GBC_simplot_thetas, 
-       width = 15, height = 6, 
-       units = "in"  # default
-       )
-filename_MODE_GBC_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GBC_simplot_thetas_NSIM_", nsim,".png")
-ggsave(filename = filename_MODE_GBC_simplot_thetas, plot = MODE_GBC_simplot_thetas, 
-       width = 15, height = 6, 
-       units = "in"  # default
-       )
-filename_CONMIX_GBC_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GBC_simplot_thetas_NSIM_", nsim,".png")
-ggsave(filename = filename_CONMIX_GBC_simplot_thetas, plot = CONMIX_GBC_simplot_thetas, 
-       width = 15, height = 6, 
-       units = "in"  # default
-       )
-```
-
-```{r theta-sim-distribution-GBC-arrange, fig.height=18, fig.width=10, warning=FALSE}
-IVW_GBC_simplot_thetas$labels$title <- ""
-MODE_GBC_simplot_thetas$labels$title <- ""
-CONMIX_GBC_simplot_thetas$labels$title <- ""
-ggarrange(IVW_GBC_simplot_thetas,MODE_GBC_simplot_thetas,CONMIX_GBC_simplot_thetas, 
-          ncol = 1, nrow = 3) -> GBC_simplot_thetas_arrange
-
-filename_GBC_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GBC_simplot_thetas_NSIM_", nsim,".png")
-ggsave(filename = filename_GBC_simplot_thetas_arrange, plot = GBC_simplot_thetas_arrange, 
-       width = 10, height = 18, 
-       units = "in"  # default
-       )
-```
-
-### GBC Bias & MSE
-
-```{r plot-performance-metrics-bias-mse-GBC, fig.height=10, fig.width=8}
-## IVW
-#BIAS
-ggplot(IVW_GBC_performance_metrics, aes(x=N, y=bias))+ 
+## MRMIX
+ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=bias))+ 
   geom_point()+
@@
   ylab("Bias")+
-  labs(title = "IVW estimate: BMI -> GBC")+
-  theme_few() -> biasplot_IVW_GBC
-#MSE
-ggplot(IVW_GBC_performance_metrics, aes(x=N, y=mse))+ 
-  geom_point()+
-    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  scale_y_continuous(limits = c(0,1.8))+
-  scale_x_continuous(labels = comma)+
-  xlab("Size of Sample I")+
-  ylab("MSE")+
-  labs(title = "IVW estimate: BMI -> GBC")+
-  theme_few() -> mseplot_IVW_GBC
-ggarrange(biasplot_IVW_GBC, mseplot_IVW_GBC) -> bias_mse_plot_IVW_GBC
-
-## MODE
-ggplot(MODE_GBC_performance_metrics, aes(x=N, y=bias))+ 
-  geom_point()+
-  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  scale_x_continuous(labels = comma)+
-  scale_y_continuous(limits = c(-1.5,1.5))+
-  xlab("Size of Sample I")+
-  ylab("Bias")+
-  labs(title = "MODE estimate: BMI -> GBC")+
-  theme_few() -> biasplot_MODE_GBC
-ggplot(MODE_GBC_performance_metrics, aes(x=N, y=mse))+ 
+  labs(title = "MRMIX estimate: BMI -> GSD")+
+  theme_few() -> biasplot_MRMIX_GSD
+ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=mse))+ 
   geom_point()+
@@
   ylab("MSE")+
-  labs(title = "MODE estimate: BMI -> GBC")+
-  theme_few() -> mseplot_MODE_GBC
-ggarrange(biasplot_MODE_GBC, mseplot_MODE_GBC) -> bias_mse_plot_MODE_GBC
-
-## CONMIX
-ggplot(CONMIX_GBC_performance_metrics, aes(x=N, y=bias))+ 
-  geom_point()+
-  geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  scale_x_continuous(labels = comma)+
-  scale_y_continuous(limits = c(-1.5,1.5))+
-  xlab("Size of Sample I")+
-  ylab("Bias")+
-  labs(title = "CONMIX estimate: BMI -> GBC")+
-  theme_few() -> biasplot_CONMIX_GBC
-ggplot(CONMIX_GBC_performance_metrics, aes(x=N, y=mse))+ 
-  geom_point()+
-    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-  scale_x_continuous(labels = comma)+
-  scale_y_continuous(limits = c(0,1.8))+
-  xlab("Size of Sample I")+
-  ylab("MSE")+
-  labs(title = "CONMIX estimate: BMI -> GBC")+
-  theme_few() -> mseplot_CONMIX_GBC
-ggarrange(biasplot_CONMIX_GBC, mseplot_CONMIX_GBC) -> bias_mse_plot_CONMIX_GBC
-
-ggarrange(bias_mse_plot_IVW_GBC, bias_mse_plot_MODE_GBC, bias_mse_plot_CONMIX_GBC,
-          ncol = 1, nrow = 3, common.legend = TRUE) -> bias_mse_plot_GBC
-bias_mse_plot_GBC
-
-filename_GBC_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GBC_simplot_bias_mse_NSIM_", nsim, ".png")
-ggsave(filename = filename_GBC_simplot_bias_mse, plot = bias_mse_plot_GBC, 
-       width = 7, height = 10, 
+  labs(title = "MRMIX estimate: BMI -> GSD")+
+  theme_few() -> mseplot_MRMIX_GSD
+ggarrange(biasplot_MRMIX_GSD, mseplot_MRMIX_GSD) -> bias_mse_plot_MRMIX_GSD
+
+ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD, bias_mse_plot_MRMIX_GSD
+          ncol = 1, nrow = 4, common.legend = TRUE) -> bias_mse_plot_GSD
+bias_mse_plot_GSD
+
+filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_NSIM_", nsim, ".png")
+ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
+       width = 7, height = 12, 
        units = "in"  # default
@@
 ```
+
+**Deleted all code for GBC simulation downstream analysis since no longer needed for thesis.**
+
+# Parallelize performance estimates for different "TRUE" estimates
+
+We use 5 BMI -> GSD reference estimates: IVW random model, MRMix, weighted median, Wald estimate MC4R variant, Wald estimate FTO variant
+
+```{r}
+selected_true_methods <- c("IVW", "MRMix", "median", "MC4R_Wald", "FTO_Wald")
+true_estimates <- est.bmi.gsd %>% 
+  filter(mr_methods %in% selected_true_methods)
+```
+
+```{r parallel-performance}
+
+funs <- c(mean = mean, emp_se=emp_se, bias = bias, mse = mse)  # as above
+
+for (true_method in 1:length(selected_true_methods)) {
+  print(true_method)
+
+  # define reference estimate for all 5 chosen methods in loop 
+  args.gsd.method <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]], 
+                          nsim=nrow(est_sim)
+  )
+
+  IVW_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_IVW_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, IVW_GSD_mean=mean, IVW_GSD_emp_se=emp_se, IVW_GSD_bias=bias, IVW_GSD_mse=mse)
+
+  MODE_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_MODE_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, MODE_GSD_mean=mean, MODE_GSD_emp_se=emp_se, MODE_GSD_bias=bias, MODE_GSD_mse=mse)
+
+  CONMIX_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_CONMIX_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, CONMIX_GSD_mean=mean, CONMIX_GSD_emp_se=emp_se, CONMIX_GSD_bias=bias, CONMIX_GSD_mse=mse)
+
+  MRMIX_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_MRMIX_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, MRMIX_GSD_mean=mean, MRMIX_GSD_emp_se=emp_se, MRMIX_GSD_bias=bias, MRMIX_GSD_mse=mse)
+
+  ## further metrics
+  IVW_GSD_coverage_calc <- c()
+  IVW_GSD_coverage_bias_eliminated_calc <- c()
+  IVW_GSD_type1error_calc <- c()
+  MODE_GSD_coverage_calc <- c()
+  MODE_GSD_coverage_bias_eliminated_calc <- c()
+  MODE_GSD_type1error_calc <- c()
+  CONMIX_GSD_coverage_calc <- c()
+  CONMIX_GSD_coverage_bias_eliminated_calc <- c()
+  CONMIX_GSD_type1error_calc <- c()
+  MRMIX_GSD_coverage_calc <- c()
+  MRMIX_GSD_coverage_bias_eliminated_calc <- c()
+  MRMIX_GSD_type1error_calc <- c()
+
+  for (i in 1:length(sample_size_filtered)) {
+    sample_size <- sample_size_filtered[i]
+    # GSD additional metrics ----------------------------------------------------
+    ## 1/ IVW
+    col_IVW_GSD_theta <- paste0("IVW_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_IVW_GSD_theta_se <- paste0("IVW_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_IVW_GSD_mr_pval <- paste0("IVW_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    IVW_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_IVW_GSD_theta], thetaobs_se = est_sim[col_IVW_GSD_theta_se], theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    IVW_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_IVW_GSD_theta]], thetaobs_se = est_sim[[col_IVW_GSD_theta_se]])
+    IVW_GSD_type1error_calc[i] <- type1error(est_sim[[col_IVW_GSD_mr_pval]])
+    ## 2/ MODE
+    col_MODE_GSD_theta <- paste0("MODE_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_MODE_GSD_theta_se <- paste0("MODE_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_MODE_GSD_mr_pval <- paste0("MODE_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    MODE_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MODE_GSD_theta], 
+                                         thetaobs_se = est_sim[col_MODE_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    MODE_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MODE_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_MODE_GSD_theta_se]])
+    MODE_GSD_type1error_calc[i] <- type1error(est_sim[[col_MODE_GSD_mr_pval]])
+    ## 3/ CONMIX
+    col_CONMIX_GSD_theta <- paste0("CONMIX_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_CONMIX_GSD_theta_se <- paste0("CONMIX_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_CONMIX_GSD_mr_pval <- paste0("CONMIX_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    CONMIX_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_CONMIX_GSD_theta], 
+                                         thetaobs_se = est_sim[col_CONMIX_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    CONMIX_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_CONMIX_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_CONMIX_GSD_theta_se]])
+    CONMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_CONMIX_GSD_mr_pval]])  
+      ## 4/ MRMIX
+    col_MRMIX_GSD_theta <- paste0("MRMIX_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_MRMIX_GSD_theta_se <- paste0("MRMIX_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_MRMIX_GSD_mr_pval <- paste0("MRMIX_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    MRMIX_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MRMIX_GSD_theta], 
+                                         thetaobs_se = est_sim[col_MRMIX_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    MRMIX_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MRMIX_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_MRMIX_GSD_theta_se]])
+    MRMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_MRMIX_GSD_mr_pval]])  
+  }
+
+  ## combine metrics
+  IVW_GSD_performance_metrics <- IVW_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(IVW_GSD_coverage_calc, IVW_GSD_coverage_bias_eliminated_calc, IVW_GSD_type1error_calc)) %>% 
+    as_tibble()
+  MODE_GSD_performance_metrics <- MODE_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(MODE_GSD_coverage_calc, 
+                     MODE_GSD_coverage_bias_eliminated_calc, MODE_GSD_type1error_calc)) %>% 
+    as_tibble()
+  CONMIX_GSD_performance_metrics <- CONMIX_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(CONMIX_GSD_coverage_calc, 
+                     CONMIX_GSD_coverage_bias_eliminated_calc, CONMIX_GSD_type1error_calc)) %>% 
+    as_tibble()
+  MRMIX_GSD_performance_metrics <- MRMIX_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(MRMIX_GSD_coverage_calc, 
+                     MRMIX_GSD_coverage_bias_eliminated_calc, MRMIX_GSD_type1error_calc)) %>% 
+    as_tibble()
+  MRMIX_GSD_performance_metrics %>% 
+    mutate(across(cols = where(is.numeric), round(3))) %>% 
+    kable(caption = "MRMIX method on BMI -> GSD performance metrics") %>% 
+    kable_styling(font_size=10)
+
+}
+
+```

### 2023-02-01 scripts/simstudy.R

@@ -11,10 +11,16 @@
 ##
 ## ---------------------------
 ## Notes:
-##   creates df `est_sim` with nsim* theta_estimates, nsim* standard errors of theta sims, nsim* pve for exposure and outcome
-##      cols: "no_sim", "theta_sim_SAMPLE_SIZES", "theta_se_sim_SAMPLE_SIZES", "pve_X_sim_SAMPLE_SIZES", "pve_Y_sim_SAMPLE_SIZES"
+##   creates df `est_sim` with nsim* theta_estimates, nsim* standard errors of theta sims, nsim* pve for exposure
+##      cols: "no_sim", "theta_sim_SAMPLE_SIZES", "theta_se_sim_SAMPLE_SIZES", "pve_X_sim_SAMPLE_SIZES"
+##
+##   implemented methods in simulation so far:
+##          1. random IVW
+##          2. weighted mode
+##          3. CONMIX
+##          4. MRMix
+##
 ## ---------------------------
-
 
 # set to TRUE if simulation should run when script is sourced
 runSimulation <- TRUE
@@ -31,12 +37,10 @@ library(MRMix)
 library(MRPRESSO)
 library(penalized)
 library(conflicted)
-source("scripts/MR_lasso.R")  # Slob&Burgess 2020
-source("scripts/PVE_calculation_fns.R")
+source("scripts/helper/MR_lasso.R") # Slob&Burgess 2020
+source("scripts/helper/PVE_calculation_fns.R")
 conflict_prefer("select", "dplyr")
 conflict_prefer("filter", "dplyr")
-
-
 
 # GET DATA ----------------------------------------------------------------
 # data
@@ -53,230 +57,141 @@ n.orig <- 334487
 ###
 set.seed(1234)
 ## initialize
-nsim <- 100
+nsim <- 50
+threshold <- 5 * 10^-6
+n.orig_sqrt <- sqrt(n.orig)
+
 est_sim <- tibble(no_sim = 1:nsim)
 
 ## START SIM
-if (runSimulation==TRUE) {
-  T0 = proc.time()[3]
+if (runSimulation == TRUE) {
+  T0 <- proc.time()[3]
   for (i in 1:length(colnames_pval)) {
-    threshold <- 5*10^-6
-    se_update_BMI <- c()
-    se_update_GSD <- c()
-    se_update_GBC <- c()
-    
     numIV <- sum(my_data_harm[[colnames_pval[i]]] < threshold)
-    
[…]

### 2023-02-01 scripts/helper/simulation_helpers.R
improve performance with purrr

@@ -1,0 +1,131 @@
+
+library(dplyr)
+library(purrr)
+library(tidyr)
+library(stringr)
+library(tibble)
+library(MendelianRandomization)
+library(TwoSampleMR)
+library(mr.raps)
+library(MRMix)
+library(MRPRESSO)
+library(penalized)
+library(conflicted)
+source("scripts/helper/MR_lasso.R") # Slob&Burgess 2020
+source("scripts/helper/PVE_calculation_fns.R")
+conflict_prefer("select", "dplyr")
+conflict_prefer("filter", "dplyr")
+
+sim_function <- function(dat, index) {
+  if(any(is.na(dat))) return(NULL)
+  return_vector <- c()
+  dat <- dat %>% 
+    mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
+    mutate(explained_variance_X2_sim = explained_variance_numeric2(maf = eaf.bmi, beta = beta.bmi_sim_norm, se_beta = se_update_BMI, samplesize = n[index]))
+  # PVE 
+  return_vector <- sum(dat$explained_variance_X2_sim)
+  
+  # PERFORM MR_analysis 
+  ## A/ GSD  ---------------------------------------------------------------
+  mr.obj.gsd = MendelianRandomization::mr_input(
+    bx = dat$beta.bmi_sim_norm, 
+    bxse = dat$se_update_BMI, 
+    # outcome
+    by = dat$beta.gsd, 
+    byse = dat$se.gsd, 
+    snps = dat$SNP,
+    exposure = "Body mass index",
+    outcome = "Gallstone Disease"
+  ) 
+  
+  # 1. random IVW
+  ivw.res.gsd = MendelianRandomization::mr_ivw(mr.obj.gsd, model = "random")
+  # store results
+  return_vector <- append(return_vector,ivw.res.gsd$Estimate)
+  return_vector <- append(return_vector,ivw.res.gsd$StdError)
+  return_vector <- append(return_vector, ivw.res.gsd$Pvalue)
+  
+  # 2. weighted mode
+  mode.res.gsd = MendelianRandomization::mr_mbe(mr.obj.gsd)
+  # store results
+  return_vector <- append(return_vector, mode.res.gsd$Estimate)
+  return_vector <- append(return_vector, mode.res.gsd$StdError)
[…]

### 2023-02-01 scripts/purrr-simstudy.R

@@ -1,0 +1,95 @@
+
+
+## ---------------------------
+##
+## Script name: purrr-simstudy.R
+##
+## Author: Christoph Reich
+## Date Created: 2022-11-22
+##
+## Copyright (c) Christoph Reich, 2022
+## Email: christoph.reich@med.uni-heidelberg.de
+##
+## ---------------------------
+## Notes:
+##   creates df `est_sim` with nsim* theta_estimates, nsim* standard errors of theta sims, nsim* pve for exposure
+##      cols: "no_sim", "theta_sim_SAMPLE_SIZES", "theta_se_sim_SAMPLE_SIZES", "pve_X_sim_SAMPLE_SIZES"
+##
+##   implemented methods in simulation so far:
+##          1. random IVW
+##          2. weighted mode
+##          3. CONMIX
+##          4. MRMix
+##
+## ---------------------------
+
+# set to TRUE if simulation should run when script is sourced
+rm(list=ls())
+runSimulation <- TRUE
+nsim <- 5
+
+library(dplyr)
+library(purrr)
+library(tidyr)
+library(stringr)
+library(tibble)
+library(MendelianRandomization)
+library(TwoSampleMR)
+library(mr.raps)
+library(MRMix)
+library(MRPRESSO)
+library(penalized)
+library(conflicted)
+source("scripts/helper/MR_lasso.R") # Slob&Burgess 2020
+source("scripts/helper/PVE_calculation_fns.R")
+source("scripts/helper/simulation_helpers.R")
+conflict_prefer("select", "dplyr")
+conflict_prefer("filter", "dplyr")
+
+# GET DATA ----------------------------------------------------------------
+# data
+my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
+# vars from MSc script
+sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
+n <- sim_vars$n
+colnames_pval <- sim_vars$colnames_pval
+n.orig <- 334487
+
+## PARAMS for simulation
+set.seed(1234)
+threshold <- 5 * 10^-6
+n.orig_sqrt <- sqrt(n.orig)
+
+# space to populate
+est_sim_temp <- tibble(no_sim = 1:nsim)
[…]

### 2023-02-01 scripts/MSc.Rmd
updated purrr simulation, checked implementation speed

@@
 # for the numeric traits we have to map over sample size and recalculate se_updated
-n.orig_sqrt <- sqrt(n.orig)
 explained_variance_BMI2_sample_size <- map(n, ~explained_variance_numeric2(maf = my_data_harm$eaf.bmi, beta = my_data_harm$beta.bmi, 
-                                                                           se_beta = abs(my_data_harm$beta.bmi / -qnorm(p=(my_data_harm$p.value.bmi / 2), mean = 0) * n.orig_sqrt / sqrt(.x)), 
+                                                                           se_beta = abs(my_data_harm$beta.bmi / -qnorm(p=(my_data_harm$p.value.bmi / 2), mean = 0) * sqrt(n.orig) / sqrt(.x)), 
                                                                            samplesize = .x))
@@
     select(all_of(colnames_pval[mysamples])) %>% 
-    mutate(explained_variance_BMI2_sample_size = explained_variance_BMI2_sample_size[[mysamples]]) %>% 
-    filter(!!sym(colnames_pval[mysamples]) < threshold)
+    filter(!!sym(colnames_pval[mysamples]) < threshold) %>% 
+    mutate(explained_variance_BMI2_sample_size = explained_variance_BMI2_sample_size[[mysamples]])
   #print(temp_dat)

### 2023-02-01 scripts/helper/simulation_helpers.R

@@ -18,6 +18,10 @@ conflict_prefer("filter", "dplyr")
 
 sim_function <- function(dat, index) {
   if(any(is.na(dat))) return(NULL)
+  
+  sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
+  n <- sim_vars$n
+  
   return_vector <- c()
   dat <- dat %>% 
     mutate(beta.bmi_sim_norm = purrr::map2_dbl(.x = beta.bmi, .y = se_update_BMI, .f = ~rnorm(n = 1, mean = .x, sd = .y))) %>% 
@@ -105,27 +109,31 @@ sim_function <- function(dat, index) {
   return(return_vector)
 }
 
+
+
+# POPULATE est_sim_temp --------------------------------------------------------
 populate_est_sim <- function(nsim_list, nsim, no_cases) {
   if(is.null(unlist(nsim_list))) return(NULL)
+  est_sim_temp <- tibble(no_sim = 1:nsim)
   for (nsims in 1:length(nsim_list)) {
     # cum_PVE_X
-    est_sim[nsims, paste0("pve_X_sim_", format(no_cases, scientific = FALSE))] <- nsim_list[[nsims]][1] 
+    est_sim_temp[nsims, paste0("pve_X_sim_", format(no_cases, scientific = FALSE))] <- nsim_list[[nsims]][1] 
     # 1. random IVW
-    est_sim[nsims, paste0("IVW_theta_sim_BMI_GSD_", format(no_cases, scientific = FALSE))] <- nsim_list[[nsims]][2] 
-    est_sim[nsims, paste0("IVW_theta_se_sim_BMI_GSD_", format(no_cases, scientific = FALSE))] <- nsim_list[[nsims]][3] 
-    est_sim[nsims, paste0("IVW_mr_pval_BMI_GSD_", format(no_cases, scientific = FALSE))] <- nsim_list[[nsims]][4] 
+    est_sim_temp[nsims, paste0("IVW_theta_sim_BMI_GSD_", format(no_cases, scientific = FALSE))] <- nsim_list[[nsims]][2] 
+    est_sim_temp[nsims, paste0("IVW_theta_se_sim_BMI_GSD_", format(no_cases, scientific = FALSE))] <- nsim_list[[nsims]][3] 
[…]

### 2023-02-01 scripts/purrr-simstudy.R

@@ -26,7 +26,7 @@
 # set to TRUE if simulation should run when script is sourced
 rm(list=ls())
 runSimulation <- TRUE
-nsim <- 5
+nsim <- 100
 
 library(dplyr)
 library(purrr)
@@ -61,9 +61,7 @@ threshold <- 5 * 10^-6
 n.orig_sqrt <- sqrt(n.orig)
 
 # space to populate
-est_sim_temp <- tibble(no_sim = 1:nsim)
 est_sim <- tibble(no_sim = 1:nsim)
-
 # create dataframes for each specified sample size ------------------------
 gen_dfs <- map(1:length(colnames_pval), ~{
   threshold <- 5*10^-6
@@ -80,16 +78,19 @@ gen_dfs <- map(1:length(colnames_pval), ~{
 # run simulation ----------------------------------------------------------
 T0 <- proc.time()[3]
 for (dataset in 1:length(gen_dfs)) {
+  # initialize in loop/ discard at the end
+  est_sim_temp <- tibble(no_sim = 1:nsim)
   nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs[[dataset]], index = dataset))  # returns nsim* results for MR simulations PER sample size 
   est_sim_temp <- populate_est_sim(nsim_list = nsim_list_dataset, nsim = nsim, no_cases = n[dataset])
   if(is.null(est_sim_temp)) next
   est_sim <- est_sim %>% cbind(est_sim_temp[-1])
-  rm(est_sim_temp)
+  rm(est_sim_temp, nsim_list_dataset)
   #
   print(paste0("|||-----------------------Run finished for sample size: ", format(n[dataset], scientific = FALSE), " -----------------------|||"))
 }
 T1 <- proc.time()[3]
 timediff_purrr <- T1 - T0
+print(paste0("!!!!!--------------------------- ", round(timediff_purrr/60,2), " minutes passed for the simulation scenario! --------------------------------------!!!!!"))
 
 saveRDS(est_sim, paste0("./output/Rdata/", Sys.Date(), "_est_sim_purrr_NSIM_", nsim, ".rds"))
 

### 2023-02-01 scripts/simstudy.R

@@ -24,6 +24,7 @@
 
 # set to TRUE if simulation should run when script is sourced
 runSimulation <- TRUE
+nsim <- 5
 
 library(dplyr)
 library(purrr)
@@ -57,7 +58,6 @@ n.orig <- 334487
 ###
 set.seed(1234)
 ## initialize
-nsim <- 50
 threshold <- 5 * 10^-6
 n.orig_sqrt <- sqrt(n.orig)
 
@@ -191,7 +191,7 @@ if (runSimulation == TRUE) {
 
   T1 <- proc.time()[3]
   timediff <- T1 - T0
+  print(paste0("!!!!!--------------------------- ", round(timediff/60,2), " minutes passed for the simulation scenario! --------------------------------------!!!!!"))
 
-  saveRDS(est_sim, paste0("./output/Rdata/", Sys.Date(), "_est_sim_NSIM_", nsim, ".rds"))
-  print(timediff)
+  saveRDS(est_sim, paste0("./output/Rdata/", Sys.Date(), "_est_sim_forloop_NSIM_", nsim, ".rds"))
 }

### 2023-02-08 scripts/MSc.Rmd
added loop for performance measure tables and figures

--- 2023-02-01 scripts/MSc.Rmd
@@
 library(penalized)
+# library(rsimsum)   do not use
 library(conflicted)
@@
 conflict_prefer("filter", "dplyr")
+# conflict_prefer("kable", "kableExtra")
 ```
@@
 
-t(filter_FTO) %>% kable() %>% 
+t(filter_FTO) %>%  kableExtra::kable() %>% 
   kable_styling(font_size=10)
@@
 ggplot(df_cum_expl_var, aes(x=n))+
-  geom_line(aes(y=cum_expl_varX2_sample_size*100), col="red")+
+  geom_line(aes(y=cum_expl_varX2_sample_size*100), col="red")+  # slightly higher PVE if samplesize mapped into PVE_X formula
   geom_line(aes(y=cum_expl_varX2*100))
@@
 scale_factor <- 4.5/284
+N_annotation <- c(10000, 100000, 200000, 300000)
 
@@
 ## load results from script `perform_mr.R`
-est.bmi.gsd <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
+est.bmi.gsd <- readRDS(dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
 
@@
 # read dat from "simstudy.R"  #------------------------------------------------
-est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_NSIM_", full.names = TRUE)))
+#est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_NSIM_", full.names = TRUE)))
+# from "purrr-simstudy.R"
+est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_purrr_NSIM_", full.names = TRUE)))
 nsim <- nrow(est_sim)
@@
   cbind(sample_size_filtered) %>% as_tibble() -> pve_X_sim
-# # CONFOUNDER PVE
-# est_sim %>% 
-#   select(!!!(colname_pve_U)) %>% 
-#   pivot_longer(everything(), names_to = "PVE_U", values_to = "value_U") %>% 
-#   cbind(sample_size_filtered) %>% as_tibble() -> pve_U_sim
-# # OUTCOME PVE
-# est_sim %>% 
-#   select(!!!(colname_pve_Y)) %>% 
-#   pivot_longer(everything(), names_to = "PVE_Y", values_to = "value_Y") %>% 
-#   cbind(sample_size_filtered) %>% as_tibble() -> pve_Y_sim
+
+# mean PVE per sample size
+est_sim %>% 
+  select(!!!(colname_pve_X)) %>% 
+  map(~mean(.x)) %>% 
+  unlist() -> PVE_X_sim_mean
 
@@
 
+```{r}
+# print("95% Confidence Interval: [{:.2f}, {:.2f}]".format(lower_bound, upper_bound))
+```
+
 ```{r further-metrics}
@@
 coverage <- function(thetaobs, thetaobs_se, theta, ...) {
-  lower_ci <- thetaobs - thetaobs_se
-  upper_ci <- thetaobs + thetaobs_se
+  # https://www.r-bloggers.com/2021/11/calculate-confidence-intervals-in-r/
+  lower_ci <- thetaobs - 1.96*thetaobs_se
+  upper_ci <- thetaobs + 1.96*thetaobs_se
   coverage <- mean(lower_ci <= theta & upper_ci >= theta)
@@
 }
+
 # coverage bias eliminated
 coverage_bias_eliminated <- function(thetaobs, thetaobs_se, ...) {
-  lower_ci <- thetaobs - thetaobs_se
-  upper_ci <- thetaobs + thetaobs_se
+  lower_ci <- thetaobs - 1.96*thetaobs_se
+  upper_ci <- thetaobs + 1.96*thetaobs_se
   mean_thetaobs <- mean(thetaobs)
@@
   MRMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_MRMIX_GSD_mr_pval]])  
-
-  ## GBC additional metrics ----------------------------------------------------
-  # deleted 2023-01-29
 }
@@
 
+```{r}
+#s1 <- simsum(data = est_sim, estvarname = "IVW_theta_sim_BMI_GSD_40000", se = "IVW_theta_se_sim_BMI_GSD_40000", true = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "IVW"] )
+
+## stats param
+# • nsim,the number of replications with non-missing point estimates and standard error.
+# • thetamean, average point estimate.
+# • thetamedian, median point estimate.
+# • se2mean, average variance.
+# • se2median, median variance.
+# • bias, bias in point estimate.
+# • empse, empirical standard error.
+# • mse, mean squared error.
+# • relprec, percentage gain in precision relative to the reference method.
+# • modelse, model-based standard error.
+# • relerror, relative percentage error in standard error.
+# • cover, coverage of a nominal level\
+# • becover, bias corrected coverage of a nominal level\
+# • power, power of a (1 - level)\
+
+#statsparam <- c("nsim", "thetamean", "se2mean", "bias", "empse", "mse", "cover", "becover", "power")
+#simsum_summary <- summary(s1, stats = statsparam)
+```
+
+---
+
 ## Performance Plots
@@
 ### GSD simulated theta distribution
-
-```{r theta-sim-distribution-GSD-old, eval=FALSE, fig.height=6, fig.width=10, include=FALSE}
-### OLD
-# # compare distribution of simulated thetas against theta_324000
-# boxplot_dat <- est_sim %>% 
-#   select(!!!(colname_IVW_theta_sim_BMI_GSD)) %>% 
-#   pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
-#   mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))
-# 
-# ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
-#   geom_boxplot()+
-#   geom_jitter(width=0.1,alpha=0.2)+
-#   geom_hline(yintercept=  args.gsd.ivw$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-#   annotate(geom = "text", x = "300000", y = args$theta+0.5,
-#            label= expression( paste("ref: ", theta[X] ))  , 
-#            color= ggthemes_data$few$colors$Dark[9,2][[1]]
-#              )+
-#   scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "150000", "200000", "250000", "300000"), )+ 
-#   ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
-#   labs(title = "Estimates from beta simulation",
-#        caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
-#        )+
-#   coord_flip()+
-#   theme_few() -> simplot_thetas
-# simplot_thetas
-```
 
@@
 
-ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD, bias_mse_plot_MRMIX_GSD
+ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD, bias_mse_plot_MRMIX_GSD,
           ncol = 1, nrow = 4, common.legend = TRUE) -> bias_mse_plot_GSD
@@
 
+---
+
 # Parallelize performance estimates for different "TRUE" estimates
@@
 We use 5 BMI -> GSD reference estimates: IVW random model, MRMix, weighted median, Wald estimate MC4R variant, Wald estimate FTO variant
+
+## Calculate performance measures
 
@@
 ```{r parallel-performance}
-
-funs <- c(mean = mean, emp_se=emp_se, bias = bias, mse = mse)  # as above
-
 for (true_method in 1:length(selected_true_methods)) {
-  print(true_method)
-
-  # define reference estimate for all 5 chosen methods in loop 
+  # define reference estimate for ALL 5 chosen methods in loop 
   args.gsd.method <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]], 
@@
     as_tibble()
-  MRMIX_GSD_performance_metrics %>% 
-    mutate(across(cols = where(is.numeric), round(3))) %>% 
-    kable(caption = "MRMIX method on BMI -> GSD performance metrics") %>% 
-    kable_styling(font_size=10)
-
+
+  # STORE
+  filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_IVW_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = IVW_GSD_performance_metrics, file = filename.IVW_table.performance)
+  filename.MODE_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_MODE_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = MODE_GSD_performance_metrics, file = filename.MODE_table.performance)
+  filename.CONMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_CONMIX_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = CONMIX_GSD_performance_metrics, file = filename.CONMIX_table.performance)
+  filename.MRMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_MRMIX_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
+
+  ###
+  ## PLOT  -------------------------------------------------------------------
+  IVW_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_IVW_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(IVW_GSD_performance_metrics, aes(x=N, y=IVW_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    # geom_errorbar(aes(ymin=IVW_GSD_mean-1.96*IVW_GSD_emp_se, ymax=IVW_GSD_mean+1.96*IVW_GSD_emp_se), 
+    #               width=3000, # x-axis dimension
+    #               position = "dodge")+
+    ## ADD rainbow plots
+    geom_boxplot(data = IVW_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = IVW_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = IVW_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("IVW estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> IVW_GSD_simplot_thetas
+
+  ## MODE
+  MODE_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_MODE_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(MODE_GSD_performance_metrics, aes(x=N, y=MODE_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    # geom_errorbar(aes(ymin=MODE_GSD_mean-1.96*MODE_GSD_emp_se, ymax=MODE_GSD_mean+1.96*MODE_GSD_emp_se), 
+    #               width=3000, # x-axis dimension
+    #               position = "dodge")+
+    ## ADD rainbow plots
+    geom_boxplot(data = MODE_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = MODE_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = MODE_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("MODE estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> MODE_GSD_simplot_thetas
+
+  ## CONMIX
+  CONMIX_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_CONMIX_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=CONMIX_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    ## ADD rainbow plots
+    geom_boxplot(data = CONMIX_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = CONMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = CONMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("CONMIX estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> CONMIX_GSD_simplot_thetas
+
+  ## MRMIX
+  MRMIX_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_MRMIX_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=MRMIX_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    ## ADD rainbow plots
+    geom_boxplot(data = MRMIX_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = MRMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = MRMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("MRMIX estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> MRMIX_GSD_simplot_thetas
+
+  # SAVE
+  filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+  filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+  filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+  filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MRMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+
+  IVW_GSD_simplot_thetas$labels$title <- ""
+  MODE_GSD_simplot_thetas$labels$title <- ""
+  CONMIX_GSD_simplot_thetas$labels$title <- ""
+  MRMIX_GSD_simplot_thetas$labels$title <- ""
+  ggarrange(IVW_GSD_simplot_thetas,MODE_GSD_simplot_thetas,CONMIX_GSD_simplot_thetas, MRMIX_GSD_simplot_thetas,
+            ncol = 1, nrow = 4) -> GSD_simplot_thetas_arrange
+  GSD_simplot_thetas_arrange
+
+  filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
+         width = 10, height = 22, 
+         units = "in"  # default
+         )
+
+  ### GSD Bias & MSE
+  #BIAS
+  ggplot(IVW_GSD_performance_metrics, aes(x=N, y=IVW_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "IVW estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_IVW_GSD
+  #MSE
+  ggplot(IVW_GSD_performance_metrics, aes(x=N, y=IVW_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "IVW estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_IVW_GSD
+  ggarrange(biasplot_IVW_GSD, mseplot_IVW_GSD) -> bias_mse_plot_IVW_GSD
+
+  ## MODE
+  ggplot(MODE_GSD_performance_metrics, aes(x=N, y=MODE_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "MODE estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_MODE_GSD
+  ggplot(MODE_GSD_performance_metrics, aes(x=N, y=MODE_GSD_mse))+ 
+    geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "MODE estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_MODE_GSD
+  ggarrange(biasplot_MODE_GSD, mseplot_MODE_GSD) -> bias_mse_plot_MODE_GSD
+
+  ## CONMIX
+  ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=CONMIX_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "CONMIX estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_CONMIX_GSD
+  ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=CONMIX_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "CONMIX estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_CONMIX_GSD
+  ggarrange(biasplot_CONMIX_GSD, mseplot_CONMIX_GSD) -> bias_mse_plot_CONMIX_GSD
+
+  ## MRMIX
+  ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=MRMIX_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "MRMIX estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_MRMIX_GSD
+  ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=MRMIX_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "MRMIX estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_MRMIX_GSD
+  ggarrange(biasplot_MRMIX_GSD, mseplot_MRMIX_GSD) -> bias_mse_plot_MRMIX_GSD
+
+  ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD, bias_mse_plot_MRMIX_GSD,
+            ncol = 1, nrow = 4, common.legend = TRUE) -> bias_mse_plot_GSD
+  bias_mse_plot_GSD
+
+  filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], "_NSIM_",nsim, ".png")
+  ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
+         width = 7, height = 12, 
+         units = "in"  # default
+         )
 }
-
-```
+```

### 2023-02-08 scripts/purrr-simstudy.R
small change

@@ -80,7 +80,12 @@ T0 <- proc.time()[3]
 for (dataset in 1:length(gen_dfs)) {
   # initialize in loop/ discard at the end
   est_sim_temp <- tibble(no_sim = 1:nsim)
-  nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs[[dataset]], index = dataset))  # returns nsim* results for MR simulations PER sample size 
+  ## now purrr
+  # returns nsim* results for MR simulations PER sample size
+  nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs[[dataset]], index = dataset), 
+                           .progress=list(clear=FALSE, 
+                                          name=paste0("Progress of simulation of sample size: ", format(n[dataset], scientific = FALSE)))
+  )
   est_sim_temp <- populate_est_sim(nsim_list = nsim_list_dataset, nsim = nsim, no_cases = n[dataset])
   if(is.null(est_sim_temp)) next
   est_sim <- est_sim %>% cbind(est_sim_temp[-1])
@@ -94,3 +99,5 @@ print(paste0("!!!!!--------------------------- ", round(timediff_purrr/60,2), " 
 
 saveRDS(est_sim, paste0("./output/Rdata/", Sys.Date(), "_est_sim_purrr_NSIM_", nsim, ".rds"))
 
+## with nsim=100 and the purrr script on the M1 Mac ∞ 5h
+#[1] "!!!!!--------------------------- 299.61 minutes passed for the simulation scenario! --------------------------------------!!!!!"

### 2023-03-01 scripts/helper/simulation_helpers.R
added all 10 robust MR methods

@@ -11,7 +11,7 @@ library(MRMix)
 library(MRPRESSO)
 library(penalized)
 library(conflicted)
-source("scripts/helper/MR_lasso.R") # Slob&Burgess 2020
+# source("scripts/helper/MR_lasso.R") # Slob&Burgess 2020, now implemented in `MendelianRandomization`
 source("scripts/helper/PVE_calculation_fns.R")
 conflict_prefer("select", "dplyr")
 conflict_prefer("filter", "dplyr")
@@ -106,6 +106,53 @@ sim_function <- function(dat, index) {
   return_vector <- append(return_vector, mrmix.res_se.gsd)
   return_vector <- append(return_vector, mrmix.res.gsd$pvalue_theta)
   
+  ## updated 2023-03-01
+  # 5. weighted median
+  median.res.gsd = MendelianRandomization::mr_median(mr.obj.gsd)
+  return_vector <- append(return_vector, median.res.gsd$Estimate)
+  return_vector <- append(return_vector, median.res.gsd$StdError)
+  return_vector <- append(return_vector, median.res.gsd$Pvalue)
+  
+  # 6. Egger regression
+  egger.res.gsd <- MendelianRandomization::mr_egger(mr.obj.gsd)
+  return_vector <- append(return_vector,egger.res.gsd$Estimate)
+  return_vector <- append(return_vector, egger.res.gsd$StdError.Est)
+  return_vector <- append(return_vector, egger.res.gsd$Pvalue.Est)
+  
+  # 7. MR-robust
+  robust.res.gsd <- MendelianRandomization::mr_ivw(mr.obj.gsd, "random", robust = TRUE) 
+  return_vector <- append(return_vector,robust.res.gsd$Estimate)
+  return_vector <- append(return_vector, robust.res.gsd$StdError)
+  return_vector <- append(return_vector, robust.res.gsd$Pvalue)
+  
+  # 8. MR-PRESSO
+  presso.df = data.frame(bx = dat$beta.bmi_sim_norm, by = dat$beta.gsd, bxse = dat$se_update_BMI, byse = dat$se.gsd)
+  presso.res.gsd = MRPRESSO::mr_presso(BetaOutcome = "by", BetaExposure = "bx", SdOutcome = "byse", SdExposure = "bxse", 
[…]

### 2023-03-01 scripts/purrr-simstudy.R

@@ -40,7 +40,6 @@ library(MRMix)
 library(MRPRESSO)
 library(penalized)
 library(conflicted)
-source("scripts/helper/MR_lasso.R") # Slob&Burgess 2020
 source("scripts/helper/PVE_calculation_fns.R")
 source("scripts/helper/simulation_helpers.R")
 conflict_prefer("select", "dplyr")

### 2023-03-02 scripts/helper/simulation_helpers.R
updated simstudy to include diff genetic architectures

@@ -18,7 +18,7 @@ conflict_prefer("filter", "dplyr")
 
 sim_function <- function(dat, index) {
   if(any(is.na(dat))) return(NULL)
-  
+  if(nrow(dat)<=3) return(NULL)  # cannot perform robust methods with one IV, "mr_presso method requires data on >3 variants."
   sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
   n <- sim_vars$n
   
@@ -65,7 +65,7 @@ sim_function <- function(dat, index) {
       # the error we encounter is with "i=7" 
       #     Error in seq.default(from = CIMin, to = CIMax, by = CIStep) : 
       #         wrong sign in 'by' argument
-      message(paste0("Caught an error with mr_conmix (BMI -> GSD)!\nSample size: ", n[i], 
+      message(paste0("Caught an error with mr_conmix (BMI -> GSD)!\nSample size: ", n[dataset], 
                      "\nComputed CI's with predefined range of [-15;15]"))
       message("Below is the error message from R:")
       print(e) 
@@ -148,14 +148,39 @@ sim_function <- function(dat, index) {
   return_vector <-  append(return_vector, raps.res.gsd$beta.p.value)
   
   # 10. MR-Lasso
-  lasso.res.gsd <- MendelianRandomization::mr_lasso(object = mr.obj.gsd)
-  return_vector <- append(return_vector, lasso.res.gsd$Estimate)
-  return_vector <- append(return_vector, lasso.res.gsd$StdError)
-  return_vector <- append(return_vector, lasso.res.gsd$Pvalue)
-  
+  lasso.res.gsd <- tryCatch(
+    expr = {
+      MendelianRandomization::mr_lasso(object = mr.obj.gsd)
+    },
+    error = function(e){
+      # the error we encounter is with "i=7" 
+      #     Error in seq.default(from = CIMin, to = CIMax, by = CIStep) : 
+      #         wrong sign in 'by' argument
+      message(paste0("Caught an error with mr_lasso (BMI -> GSD)!\nSample size: ", n[dataset]))
+      message("Below is the error message from R:")
+      print(e) 
+      return(NA)
+    },
+    warning = function(w){
+      message('Caught an warning!')
+      print(w)
+    },
+    finally = {
+      #message('All done, quitting.')
+    }
+  )
+  if (is.na(lasso.res.gsd)) {
[…]

### 2023-03-02 scripts/purrr-simstudy.R

@@ -15,19 +15,10 @@
 ##   creates df `est_sim` with nsim* theta_estimates, nsim* standard errors of theta sims, nsim* pve for exposure
 ##      cols: "no_sim", "theta_sim_SAMPLE_SIZES", "theta_se_sim_SAMPLE_SIZES", "pve_X_sim_SAMPLE_SIZES"
 ##
-##   implemented methods in simulation so far:
-##          1. random IVW
-##          2. weighted mode
-##          3. CONMIX
-##          4. MRMix
 ##
 ## ---------------------------
 
-# set to TRUE if simulation should run when script is sourced
 rm(list=ls())
-runSimulation <- TRUE
-nsim <- 100
-
 library(dplyr)
 library(purrr)
 library(tidyr)
@@ -44,6 +35,16 @@ source("scripts/helper/PVE_calculation_fns.R")
 source("scripts/helper/simulation_helpers.R")
 conflict_prefer("select", "dplyr")
 conflict_prefer("filter", "dplyr")
+
+# set to TRUE if simulation should run when script is sourced
+runSimulation <- TRUE
+nsim <- 2
+
+runGeneticArchitectures <- TRUE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
+# analysis set - create all combis
+effect_size <- c("strong", "weak")
+maf <- c("rare", "common")
+all_quadrants <- tidyr::crossing(effect_size, maf)
 
 # GET DATA ----------------------------------------------------------------
 # data
@@ -71,26 +72,30 @@ gen_dfs <- map(1:length(colnames_pval), ~{
   sim_dat <- my_data_harm %>% 
     filter(!!sym(colnames_pval[.x]) < threshold) %>% 
     select(SNP, eaf.bmi, beta.bmi, se.bmi, p.value.bmi, eaf.gsd, beta.gsd, se.gsd, p.value.gsd) %>% 
-    mutate(se_update_BMI = abs(beta.bmi / -qnorm(p=(p.value.bmi / 2), mean = 0) * sqrt(n.orig) / sqrt(n[.x])))
+    mutate(se_update_BMI = abs(beta.bmi / -qnorm(p=(p.value.bmi / 2), mean = 0) * sqrt(n.orig) / sqrt(n[.x]))) %>% 
+    mutate(maf.bmi = ifelse(eaf.bmi >0.5, 1-eaf.bmi, eaf.bmi))
 })
 
[…]

### 2023-03-02 scripts/purrr-simstudy.R
updated to save results

@@ -109,6 +109,7 @@ saveRDS(est_sim, paste0("./output/Rdata/", Sys.Date(), "_est_sim_purrr_NSIM_", n
 
 safe_filter <- safely(dplyr::filter, otherwise = NA, quiet = TRUE)  # needed because also NAs in `gen_dfs`
 est_sim_architectures <- list()
+T0_architecture <- proc.time()[3]
 if (runGeneticArchitectures == TRUE) {
   for (architecture in 1:nrow(all_quadrants)) {
     effect_size <- all_quadrants[[architecture, 1]]
@@ -148,8 +149,11 @@ if (runGeneticArchitectures == TRUE) {
   }
 }
 
-est_sim_temp <- populate_est_sim(nsim_list = nsim_list_dataset, nsim = nsim, no_cases = n[dataset])
-if(is.null(est_sim_temp)) next
-est_sim <- est_sim %>% cbind(est_sim_temp[-1])
+T1_architecture <- proc.time()[3]
+timediff_purrr_archi <- T1_architecture - T0_architecture
+print(paste0("!!!!!--------------------------- ", round(timediff_purrr_archi/60,2), " minutes passed for the simulation scenario! --------------------------------------!!!!!"))
+
+saveRDS(est_sim_architectures, paste0("./output/Rdata/", Sys.Date(), "_est_sim_architectures_purrr_NSIM_", nsim, ".rds"))
+
 
 

### 2023-03-02 scripts/purrr-simstudy.R
nsim10

@@ -38,7 +38,7 @@ conflict_prefer("filter", "dplyr")
 
 # set to TRUE if simulation should run when script is sourced
 runSimulation <- TRUE
-nsim <- 2
+nsim <- 10
 
 runGeneticArchitectures <- TRUE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 # analysis set - create all combis
@@ -103,7 +103,7 @@ print(paste0("!!!!!--------------------------- ", round(timediff_purrr/60,2), " 
 
 saveRDS(est_sim, paste0("./output/Rdata/", Sys.Date(), "_est_sim_purrr_NSIM_", nsim, ".rds"))
 
-## with nsim=100 and the purrr script on the M1 Mac ∞ 5h
+## with nsim=100 and the purrr script on the M1 Mac ∞ 5h (only 4 methods at that time)
 #[1] "!!!!!--------------------------- 299.61 minutes passed for the simulation scenario! --------------------------------------!!!!!"
 
 

### 2023-03-02 scripts/helper/simulation_helpers.R
error in helper function's index

@@ -65,7 +65,7 @@ sim_function <- function(dat, index) {
       # the error we encounter is with "i=7" 
       #     Error in seq.default(from = CIMin, to = CIMax, by = CIStep) : 
       #         wrong sign in 'by' argument
-      message(paste0("Caught an error with mr_conmix (BMI -> GSD)!\nSample size: ", n[dataset], 
+      message(paste0("Caught an error with mr_conmix (BMI -> GSD)!\nSample size: ", n[index], 
                      "\nComputed CI's with predefined range of [-15;15]"))
       message("Below is the error message from R:")
       print(e) 
@@ -156,7 +156,7 @@ sim_function <- function(dat, index) {
       # the error we encounter is with "i=7" 
       #     Error in seq.default(from = CIMin, to = CIMax, by = CIStep) : 
       #         wrong sign in 'by' argument
-      message(paste0("Caught an error with mr_lasso (BMI -> GSD)!\nSample size: ", n[dataset]))
+      message(paste0("Caught an error with mr_lasso (BMI -> GSD)!\nSample size: ", n[index]))
       message("Below is the error message from R:")
       print(e) 
       return(NA)

### 2023-03-02 scripts/MSc.Rmd
updated plots and performance measures for all 10 methods

@@
 ## for performance measures later
-# retrieve IVW/ MODE/ CONMIX cols for BMI -> GSD 
+# retrieve IVW/ MODE/ CONMIX/... cols for BMI -> GSD 
+#1
 colname_IVW_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GSD_")]
 colname_IVW_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GSD_")]
+#2
 colname_MODE_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GSD_")]
 colname_MODE_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GSD_")]
+#3
 colname_CONMIX_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GSD_")]
 colname_CONMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GSD_")]
+#4
 colname_MRMIX_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MRMIX_theta_sim_BMI_GSD_")]
 colname_MRMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MRMIX_theta_se_sim_BMI_GSD_")]
+#5
+colname_MEDIAN_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MEDIAN_theta_sim_BMI_GSD_")]
+colname_MEDIAN_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MEDIAN_theta_se_sim_BMI_GSD_")]
+#6
+colname_EGGER_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "EGGER_theta_sim_BMI_GSD_")]
+colname_EGGER_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "EGGER_theta_se_sim_BMI_GSD_")]
+#7
+colname_ROBUST_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "ROBUST_theta_sim_BMI_GSD_")]
+colname_ROBUST_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "ROBUST_theta_se_sim_BMI_GSD_")]
+#8
+colname_PRESSO_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "PRESSO_theta_sim_BMI_GSD_")]
+colname_PRESSO_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "PRESSO_theta_se_sim_BMI_GSD_")]
+#9
+colname_RAPS_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "RAPS_theta_sim_BMI_GSD_")]
+colname_RAPS_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "RAPS_theta_se_sim_BMI_GSD_")]
+#10
+colname_LASSO_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "LASSO_theta_sim_BMI_GSD_")]
+colname_LASSO_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "LASSO_theta_se_sim_BMI_GSD_")]
+
 # retrieve IVW/ MODE/ CONMIX cols for BMI -> GBC
@@
 ## args gsd
+#1
 args.gsd.ivw <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "IVW"], 
@@
                      ) # trim argument non-matching and ignored for median
+#2
 args.gsd.mode <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "mode"], 
@@
                      )
+#3
 args.gsd.conmix <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "conmix"], 
@@
                      )
+#4
 args.gsd.mrmix <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "MRMix"], 
+                     nsim=nrow(est_sim)
+                     )
+#5
+args.gsd.median <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "median"], 
+                     nsim=nrow(est_sim)
+                     )
+#6
+args.gsd.egger <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "egger"], 
+                     nsim=nrow(est_sim)
+                     )
+#7
+args.gsd.robust <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "Robust"], 
+                     nsim=nrow(est_sim)
+                     )
+#8
+args.gsd.presso <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "PRESSO"], 
+                     nsim=nrow(est_sim)
+                     )
+#9
+args.gsd.raps <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "RAPS"], 
+                     nsim=nrow(est_sim)
+                     )
+#10
+args.gsd.lasso <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "Lasso"], 
                      nsim=nrow(est_sim)
@@
 ```{r CALC-perf-metrics-GSD}
+#1
 IVW_GSD_performance_metrics <- est_sim %>% 
@@
   rename(N = sample_size_filtered)
-
+#2
 MODE_GSD_performance_metrics <- est_sim %>% 
@@
   rename(N = sample_size_filtered)
-
+#3
 CONMIX_GSD_performance_metrics <- est_sim %>% 
@@
   rename(N = sample_size_filtered)
-
+#4
 MRMIX_GSD_performance_metrics <- est_sim %>% 
@@
   map_df(.f = ~ funs %>% 
-           map(exec, .x, !!!args.gsd.conmix), .id = "var") %>% 
+           map(exec, .x, !!!args.gsd.mrmix), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+#5
+MEDIAN_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_MEDIAN_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.median), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+#6
+EGGER_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_EGGER_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.egger), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+#7
+ROBUST_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_ROBUST_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.robust), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+#8
+PRESSO_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_PRESSO_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.presso), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+#9
+RAPS_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_RAPS_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.raps), .id = "var") %>% 
+  cbind(sample_size_filtered) %>% 
+  rename(N = sample_size_filtered)
+#10
+LASSO_GSD_performance_metrics <- est_sim %>% 
+  select(!!!(colname_LASSO_theta_sim_BMI_GSD)) %>% 
+  map_df(.f = ~ funs %>% 
+           map(exec, .x, !!!args.gsd.lasso), .id = "var") %>% 
   cbind(sample_size_filtered) %>% 
@@
 # GSD
+#1
 IVW_GSD_coverage_calc <- c()
@@
 IVW_GSD_type1error_calc <- c()
+#2
 MODE_GSD_coverage_calc <- c()
@@
 MODE_GSD_type1error_calc <- c()
+#3
 CONMIX_GSD_coverage_calc <- c()
@@
 CONMIX_GSD_type1error_calc <- c()
+#4
 MRMIX_GSD_coverage_calc <- c()
@@
 MRMIX_GSD_type1error_calc <- c()
-# GBC
-## deleted 2023-01-29
+#5
+MEDIAN_GSD_coverage_calc <- c()
+MEDIAN_GSD_coverage_bias_eliminated_calc <- c()
+MEDIAN_GSD_type1error_calc <- c()
+#6
+EGGER_GSD_coverage_calc <- c()
+EGGER_GSD_coverage_bias_eliminated_calc <- c()
+EGGER_GSD_type1error_calc <- c()
+#7
+ROBUST_GSD_coverage_calc <- c()
+ROBUST_GSD_coverage_bias_eliminated_calc <- c()
+ROBUST_GSD_type1error_calc <- c()
+#8
+PRESSO_GSD_coverage_calc <- c()
+PRESSO_GSD_coverage_bias_eliminated_calc <- c()
+PRESSO_GSD_type1error_calc <- c()
+#9
+RAPS_GSD_coverage_calc <- c()
+RAPS_GSD_coverage_bias_eliminated_calc <- c()
+RAPS_GSD_type1error_calc <- c()
+#10
+LASSO_GSD_coverage_calc <- c()
+LASSO_GSD_coverage_bias_eliminated_calc <- c()
+LASSO_GSD_type1error_calc <- c()
+
 for (i in 1:length(sample_size_filtered)) {
@@
   MRMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_MRMIX_GSD_mr_pval]])
+  ## 5/ MEDIAN
+  col_MEDIAN_GSD_theta <- paste0("MEDIAN_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_MEDIAN_GSD_theta_se <- paste0("MEDIAN_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_MEDIAN_GSD_mr_pval <- paste0("MEDIAN_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  MEDIAN_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MEDIAN_GSD_theta], 
+                                       thetaobs_se = est_sim[col_MEDIAN_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "median"])
+  MEDIAN_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MEDIAN_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_MEDIAN_GSD_theta_se]])
+  MEDIAN_GSD_type1error_calc[i] <- type1error(est_sim[[col_MEDIAN_GSD_mr_pval]])
+  ## 6/ EGGER
+  col_EGGER_GSD_theta <- paste0("EGGER_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_EGGER_GSD_theta_se <- paste0("EGGER_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_EGGER_GSD_mr_pval <- paste0("EGGER_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  EGGER_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_EGGER_GSD_theta], 
+                                       thetaobs_se = est_sim[col_EGGER_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "egger"])
+  EGGER_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_EGGER_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_EGGER_GSD_theta_se]])
+  EGGER_GSD_type1error_calc[i] <- type1error(est_sim[[col_EGGER_GSD_mr_pval]])  
+  ## 7/ ROBUST
+  col_ROBUST_GSD_theta <- paste0("ROBUST_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_ROBUST_GSD_theta_se <- paste0("ROBUST_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_ROBUST_GSD_mr_pval <- paste0("ROBUST_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  ROBUST_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_ROBUST_GSD_theta], 
+                                       thetaobs_se = est_sim[col_ROBUST_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "Robust"])
+  ROBUST_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_ROBUST_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_ROBUST_GSD_theta_se]])
+  ROBUST_GSD_type1error_calc[i] <- type1error(est_sim[[col_ROBUST_GSD_mr_pval]])  
+  ## 8/ PRESSO
+  col_PRESSO_GSD_theta <- paste0("PRESSO_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_PRESSO_GSD_theta_se <- paste0("PRESSO_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_PRESSO_GSD_mr_pval <- paste0("PRESSO_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  PRESSO_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_PRESSO_GSD_theta], 
+                                       thetaobs_se = est_sim[col_PRESSO_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "PRESSO"])
+  PRESSO_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_PRESSO_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_PRESSO_GSD_theta_se]])
+  PRESSO_GSD_type1error_calc[i] <- type1error(est_sim[[col_PRESSO_GSD_mr_pval]])
+  ## 9/ RAPS
+  col_RAPS_GSD_theta <- paste0("RAPS_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_RAPS_GSD_theta_se <- paste0("RAPS_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_RAPS_GSD_mr_pval <- paste0("RAPS_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  RAPS_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_RAPS_GSD_theta], 
+                                       thetaobs_se = est_sim[col_RAPS_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "RAPS"])
+  RAPS_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_RAPS_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_RAPS_GSD_theta_se]])
+  RAPS_GSD_type1error_calc[i] <- type1error(est_sim[[col_RAPS_GSD_mr_pval]])  
+  ## 10/ LASSO
+  col_LASSO_GSD_theta <- paste0("LASSO_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_LASSO_GSD_theta_se <- paste0("LASSO_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+  col_LASSO_GSD_mr_pval <- paste0("LASSO_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+  # calc
+  LASSO_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_LASSO_GSD_theta], 
+                                       thetaobs_se = est_sim[col_LASSO_GSD_theta_se], 
+                                       theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "Lasso"])
+  LASSO_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_LASSO_GSD_theta]], 
+                                                                       thetaobs_se = est_sim[[col_LASSO_GSD_theta_se]])
+  LASSO_GSD_type1error_calc[i] <- type1error(est_sim[[col_LASSO_GSD_mr_pval]])  
 }
@@
 ## GSD --------------------------
+#1
 IVW_GSD_performance_metrics <- IVW_GSD_performance_metrics %>% 
@@
   kable_styling(font_size=10)
-
+#2
 MODE_GSD_performance_metrics <- MODE_GSD_performance_metrics %>% 
@@
   kable_styling(font_size=10)
-
+#3
 CONMIX_GSD_performance_metrics <- CONMIX_GSD_performance_metrics %>% 
@@
   kable_styling(font_size=10)
-
+#4
 MRMIX_GSD_performance_metrics <- MRMIX_GSD_performance_metrics %>% 
@@
   kable_styling(font_size=10)
+#5
+MEDIAN_GSD_performance_metrics <- MEDIAN_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(MEDIAN_GSD_coverage_calc, 
+                   MEDIAN_GSD_coverage_bias_eliminated_calc, MEDIAN_GSD_type1error_calc)) %>% 
+  rename(coverage = MEDIAN_GSD_coverage_calc, 
+         coverage_bias_eliminated = MEDIAN_GSD_coverage_bias_eliminated_calc, 
+         type1error = MEDIAN_GSD_type1error_calc) %>% 
+  as_tibble()
+MEDIAN_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "MEDIAN method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
+#6
+EGGER_GSD_performance_metrics <- EGGER_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(EGGER_GSD_coverage_calc, 
+                   EGGER_GSD_coverage_bias_eliminated_calc, EGGER_GSD_type1error_calc)) %>% 
+  rename(coverage = EGGER_GSD_coverage_calc, 
+         coverage_bias_eliminated = EGGER_GSD_coverage_bias_eliminated_calc, 
+         type1error = EGGER_GSD_type1error_calc) %>% 
+  as_tibble()
+EGGER_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "EGGER method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
+#7
+ROBUST_GSD_performance_metrics <- ROBUST_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(ROBUST_GSD_coverage_calc, 
+                   ROBUST_GSD_coverage_bias_eliminated_calc, ROBUST_GSD_type1error_calc)) %>% 
+  rename(coverage = ROBUST_GSD_coverage_calc, 
+         coverage_bias_eliminated = ROBUST_GSD_coverage_bias_eliminated_calc, 
+         type1error = ROBUST_GSD_type1error_calc) %>% 
+  as_tibble()
+ROBUST_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "ROBUST method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
+#8
+PRESSO_GSD_performance_metrics <- PRESSO_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(PRESSO_GSD_coverage_calc, 
+                   PRESSO_GSD_coverage_bias_eliminated_calc, PRESSO_GSD_type1error_calc)) %>% 
+  rename(coverage = PRESSO_GSD_coverage_calc, 
+         coverage_bias_eliminated = PRESSO_GSD_coverage_bias_eliminated_calc, 
+         type1error = PRESSO_GSD_type1error_calc) %>% 
+  as_tibble()
+PRESSO_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "PRESSO method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
+#9
+RAPS_GSD_performance_metrics <- RAPS_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(RAPS_GSD_coverage_calc, 
+                   RAPS_GSD_coverage_bias_eliminated_calc, RAPS_GSD_type1error_calc)) %>% 
+  rename(coverage = RAPS_GSD_coverage_calc, 
+         coverage_bias_eliminated = RAPS_GSD_coverage_bias_eliminated_calc, 
+         type1error = RAPS_GSD_type1error_calc) %>% 
+  as_tibble()
+RAPS_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "RAPS method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
+#10
+LASSO_GSD_performance_metrics <- LASSO_GSD_performance_metrics %>% 
+  select(N, everything()) %>% 
+  cbind(data.frame(LASSO_GSD_coverage_calc, 
+                   LASSO_GSD_coverage_bias_eliminated_calc, LASSO_GSD_type1error_calc)) %>% 
+  rename(coverage = LASSO_GSD_coverage_calc, 
+         coverage_bias_eliminated = LASSO_GSD_coverage_bias_eliminated_calc, 
+         type1error = LASSO_GSD_type1error_calc) %>% 
+  as_tibble()
+LASSO_GSD_performance_metrics %>% 
+  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  kable(caption = "LASSO method on BMI -> GSD performance metrics") %>% 
+  kable_styling(font_size=10)
 ```
@@
                                                Sys.Date(), "_MRMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
+#5
+filename_MEDIAN_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_MEDIAN_GSD_performance_metrics_NSIM_", nsim, ".csv")
+#6
+filename_EGGER_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_EGGER_GSD_performance_metrics_NSIM_", nsim, ".csv")
+#7
+filename_ROBUST_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_ROBUST_GSD_performance_metrics_NSIM_", nsim, ".csv")
+#8
+filename_PRESSO_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_PRESSO_GSD_performance_metrics_NSIM_", nsim, ".csv")
+#9
+filename_RAPS_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_RAPS_GSD_performance_metrics_NSIM_", nsim, ".csv")
+#10
+filename_LASSO_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+                                               Sys.Date(), "_LASSO_GSD_performance_metrics_NSIM_", nsim, ".csv")
 write.csv2(x = IVW_GSD_performance_metrics, file = filename_IVW_GSD_performance_metrics)
@@
 write.csv2(x = MRMIX_GSD_performance_metrics, file = filename_MRMIX_GSD_performance_metrics)
+#5
+write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename_MEDIAN_GSD_performance_metrics)
+#6
+write.csv2(x = EGGER_GSD_performance_metrics, file = filename_EGGER_GSD_performance_metrics)
+#7
+write.csv2(x = ROBUST_GSD_performance_metrics, file = filename_ROBUST_GSD_performance_metrics)
+#8
+write.csv2(x = PRESSO_GSD_performance_metrics, file = filename_PRESSO_GSD_performance_metrics)
+#9
+write.csv2(x = RAPS_GSD_performance_metrics, file = filename_RAPS_GSD_performance_metrics)
+#10
+write.csv2(x = LASSO_GSD_performance_metrics, file = filename_LASSO_GSD_performance_metrics)
 ```
@@
 ```
+
 ```{r theta-sim-distribution-GSD-arrange, fig.height=18, fig.width=10, warning=FALSE}
@@
   )
-
+  #1
   IVW_GSD_performance_metrics <- est_sim %>% 
@@
     rename(N = sample_size_filtered, IVW_GSD_mean=mean, IVW_GSD_emp_se=emp_se, IVW_GSD_bias=bias, IVW_GSD_mse=mse)
-
+  #2
   MODE_GSD_performance_metrics <- est_sim %>% 
@@
     rename(N = sample_size_filtered, MODE_GSD_mean=mean, MODE_GSD_emp_se=emp_se, MODE_GSD_bias=bias, MODE_GSD_mse=mse)
-
+  #3
   CONMIX_GSD_performance_metrics <- est_sim %>% 
@@
     rename(N = sample_size_filtered, CONMIX_GSD_mean=mean, CONMIX_GSD_emp_se=emp_se, CONMIX_GSD_bias=bias, CONMIX_GSD_mse=mse)
-
+  #4
   MRMIX_GSD_performance_metrics <- est_sim %>% 
@@
     rename(N = sample_size_filtered, MRMIX_GSD_mean=mean, MRMIX_GSD_emp_se=emp_se, MRMIX_GSD_bias=bias, MRMIX_GSD_mse=mse)
+  #5
+  MEDIAN_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_MEDIAN_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, MEDIAN_GSD_mean=mean, MEDIAN_GSD_emp_se=emp_se, MEDIAN_GSD_bias=bias, MEDIAN_GSD_mse=mse)
+  #6
+  EGGER_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_EGGER_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, EGGER_GSD_mean=mean, EGGER_GSD_emp_se=emp_se, EGGER_GSD_bias=bias, EGGER_GSD_mse=mse)
+  #7
+  ROBUST_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_ROBUST_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, ROBUST_GSD_mean=mean, ROBUST_GSD_emp_se=emp_se, ROBUST_GSD_bias=bias, ROBUST_GSD_mse=mse)
+  #8
+  PRESSO_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_PRESSO_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, PRESSO_GSD_mean=mean, PRESSO_GSD_emp_se=emp_se, PRESSO_GSD_bias=bias, PRESSO_GSD_mse=mse)
+  #9
+  RAPS_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_RAPS_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, RAPS_GSD_mean=mean, RAPS_GSD_emp_se=emp_se, RAPS_GSD_bias=bias, RAPS_GSD_mse=mse)
+  #10
+  LASSO_GSD_performance_metrics <- est_sim %>% 
+    select(!!!(colname_LASSO_theta_sim_BMI_GSD)) %>% 
+    map_df(.f = ~ funs %>% 
+             map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+    cbind(sample_size_filtered) %>% 
+    rename(N = sample_size_filtered, LASSO_GSD_mean=mean, LASSO_GSD_emp_se=emp_se, LASSO_GSD_bias=bias, LASSO_GSD_mse=mse)
 
@@
   MRMIX_GSD_type1error_calc <- c()
+  #5
+  MEDIAN_GSD_coverage_calc <- c()
+  MEDIAN_GSD_coverage_bias_eliminated_calc <- c()
+  MEDIAN_GSD_type1error_calc <- c()
+  #6
+  EGGER_GSD_coverage_calc <- c()
+  EGGER_GSD_coverage_bias_eliminated_calc <- c()
+  EGGER_GSD_type1error_calc <- c()
+  #7
+  ROBUST_GSD_coverage_calc <- c()
+  ROBUST_GSD_coverage_bias_eliminated_calc <- c()
+  ROBUST_GSD_type1error_calc <- c()
+  #8
+  PRESSO_GSD_coverage_calc <- c()
+  PRESSO_GSD_coverage_bias_eliminated_calc <- c()
+  PRESSO_GSD_type1error_calc <- c()
+  #9
+  RAPS_GSD_coverage_calc <- c()
+  RAPS_GSD_coverage_bias_eliminated_calc <- c()
+  RAPS_GSD_type1error_calc <- c()
+  #10
+  LASSO_GSD_coverage_calc <- c()
+  LASSO_GSD_coverage_bias_eliminated_calc <- c()
+  LASSO_GSD_type1error_calc <- c()
 
@@
     MRMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_MRMIX_GSD_mr_pval]])  
+    ## 5/ MEDIAN
+    col_MEDIAN_GSD_theta <- paste0("MEDIAN_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_MEDIAN_GSD_theta_se <- paste0("MEDIAN_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_MEDIAN_GSD_mr_pval <- paste0("MEDIAN_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    MEDIAN_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MEDIAN_GSD_theta], 
+                                         thetaobs_se = est_sim[col_MEDIAN_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    MEDIAN_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MEDIAN_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_MEDIAN_GSD_theta_se]])
+    MEDIAN_GSD_type1error_calc[i] <- type1error(est_sim[[col_MEDIAN_GSD_mr_pval]])  
+    ## 6/ EGGER
+    col_EGGER_GSD_theta <- paste0("EGGER_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_EGGER_GSD_theta_se <- paste0("EGGER_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_EGGER_GSD_mr_pval <- paste0("EGGER_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    EGGER_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_EGGER_GSD_theta], 
+                                         thetaobs_se = est_sim[col_EGGER_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    EGGER_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_EGGER_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_EGGER_GSD_theta_se]])
+    EGGER_GSD_type1error_calc[i] <- type1error(est_sim[[col_EGGER_GSD_mr_pval]])
+    ## 7/ ROBUST
+    col_ROBUST_GSD_theta <- paste0("ROBUST_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_ROBUST_GSD_theta_se <- paste0("ROBUST_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_ROBUST_GSD_mr_pval <- paste0("ROBUST_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    ROBUST_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_ROBUST_GSD_theta], 
+                                         thetaobs_se = est_sim[col_ROBUST_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    ROBUST_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_ROBUST_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_ROBUST_GSD_theta_se]])
+    ROBUST_GSD_type1error_calc[i] <- type1error(est_sim[[col_ROBUST_GSD_mr_pval]])
+    ## 8/ PRESSO
+    col_PRESSO_GSD_theta <- paste0("PRESSO_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_PRESSO_GSD_theta_se <- paste0("PRESSO_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_PRESSO_GSD_mr_pval <- paste0("PRESSO_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    PRESSO_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_PRESSO_GSD_theta], 
+                                         thetaobs_se = est_sim[col_PRESSO_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    PRESSO_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_PRESSO_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_PRESSO_GSD_theta_se]])
+    PRESSO_GSD_type1error_calc[i] <- type1error(est_sim[[col_PRESSO_GSD_mr_pval]])
+    ## 9/ RAPS
+    col_RAPS_GSD_theta <- paste0("RAPS_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_RAPS_GSD_theta_se <- paste0("RAPS_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_RAPS_GSD_mr_pval <- paste0("RAPS_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    RAPS_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_RAPS_GSD_theta], 
+                                         thetaobs_se = est_sim[col_RAPS_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    RAPS_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_RAPS_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_RAPS_GSD_theta_se]])
+    RAPS_GSD_type1error_calc[i] <- type1error(est_sim[[col_RAPS_GSD_mr_pval]])
+    ## 10/LASSO
+    col_LASSO_GSD_theta <- paste0("LASSO_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_LASSO_GSD_theta_se <- paste0("LASSO_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+    col_LASSO_GSD_mr_pval <- paste0("LASSO_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+    # calc
+    LASSO_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_LASSO_GSD_theta], 
+                                         thetaobs_se = est_sim[col_LASSO_GSD_theta_se], 
+                                         theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+    LASSO_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_LASSO_GSD_theta]], 
+                                                                         thetaobs_se = est_sim[[col_LASSO_GSD_theta_se]])
+    LASSO_GSD_type1error_calc[i] <- type1error(est_sim[[col_LASSO_GSD_mr_pval]])
+    #-----------
   }
@@
     as_tibble()
+  #5
+  MEDIAN_GSD_performance_metrics <- MEDIAN_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(MEDIAN_GSD_coverage_calc, 
+                     MEDIAN_GSD_coverage_bias_eliminated_calc, MEDIAN_GSD_type1error_calc)) %>% 
+    as_tibble()
+  #6
+  EGGER_GSD_performance_metrics <- EGGER_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(EGGER_GSD_coverage_calc, 
+                     EGGER_GSD_coverage_bias_eliminated_calc, EGGER_GSD_type1error_calc)) %>% 
+    as_tibble()
+  #7
+  ROBUST_GSD_performance_metrics <- ROBUST_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(ROBUST_GSD_coverage_calc, 
+                     ROBUST_GSD_coverage_bias_eliminated_calc, ROBUST_GSD_type1error_calc)) %>% 
+    as_tibble()
+  #8
+  PRESSO_GSD_performance_metrics <- PRESSO_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(PRESSO_GSD_coverage_calc, 
+                     PRESSO_GSD_coverage_bias_eliminated_calc, PRESSO_GSD_type1error_calc)) %>% 
+    as_tibble()
+  #9
+  RAPS_GSD_performance_metrics <- RAPS_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(RAPS_GSD_coverage_calc, 
+                     RAPS_GSD_coverage_bias_eliminated_calc, RAPS_GSD_type1error_calc)) %>% 
+    as_tibble()
+  #10
+  LASSO_GSD_performance_metrics <- LASSO_GSD_performance_metrics %>% 
+    select(N, everything()) %>% 
+    cbind(data.frame(LASSO_GSD_coverage_calc, 
+                     LASSO_GSD_coverage_bias_eliminated_calc, LASSO_GSD_type1error_calc)) %>% 
+    as_tibble()
 
@@
   write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
+  #5
+  filename.MEDIAN_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_MEDIAN_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename.MEDIAN_table.performance)
+  #6
+  filename.EGGER_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_EGGER_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = EGGER_GSD_performance_metrics, file = filename.EGGER_table.performance)
+  #7
+  filename.ROBUST_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_ROBUST_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = ROBUST_GSD_performance_metrics, file = filename.ROBUST_table.performance)
+  #8
+  filename.PRESSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_PRESSO_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = PRESSO_GSD_performance_metrics, file = filename.PRESSO_table.performance)
+  #9
+  filename.RAPS_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_RAPS_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = RAPS_GSD_performance_metrics, file = filename.RAPS_table.performance)
+  #10
+  filename.LASSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                      "_LASSO_GSD_performance_metrics_TRUE_EST_", 
+                                      selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
 
@@
   ## PLOT  -------------------------------------------------------------------
+  ###
   IVW_BMI_GSD_geom_point_dat <- est_sim %>% 
@@
 
-  ## MODE
+  ## 2 MODE
   MODE_BMI_GSD_geom_point_dat <- est_sim %>% 
@@
 
-  ## CONMIX
+  ## 3 CONMIX
   CONMIX_BMI_GSD_geom_point_dat <- est_sim %>% 
@@
 
-  ## MRMIX
+  ## 4 MRMIX
   MRMIX_BMI_GSD_geom_point_dat <- est_sim %>% 
@@
 
-  # SAVE
+  ## 5 MEDIAN
+  MEDIAN_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_MEDIAN_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(MEDIAN_GSD_performance_metrics, aes(x=N, y=MEDIAN_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    ## ADD rainbow plots
+    geom_boxplot(data = MEDIAN_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = MEDIAN_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = MEDIAN_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("MEDIAN estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> MEDIAN_GSD_simplot_thetas
+
+  ## 6 EGGER
+  EGGER_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_EGGER_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(EGGER_GSD_performance_metrics, aes(x=N, y=EGGER_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    ## ADD rainbow plots
+    geom_boxplot(data = EGGER_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = EGGER_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = EGGER_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("EGGER estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> EGGER_GSD_simplot_thetas
+  ## 7 ROBUST
+  ROBUST_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_ROBUST_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(ROBUST_GSD_performance_metrics, aes(x=N, y=ROBUST_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    ## ADD rainbow plots
+    geom_boxplot(data = ROBUST_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = ROBUST_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = ROBUST_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("ROBUST estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> ROBUST_GSD_simplot_thetas
+  ## 8 PRESSO
+  PRESSO_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_PRESSO_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(PRESSO_GSD_performance_metrics, aes(x=N, y=PRESSO_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    ## ADD rainbow plots
+    geom_boxplot(data = PRESSO_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = PRESSO_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = PRESSO_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("PRESSO estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> PRESSO_GSD_simplot_thetas
+  ## 9 RAPS
+  RAPS_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_RAPS_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(RAPS_GSD_performance_metrics, aes(x=N, y=RAPS_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    ## ADD rainbow plots
+    geom_boxplot(data = RAPS_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = RAPS_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = RAPS_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("RAPS estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> RAPS_GSD_simplot_thetas
+  ## 10 LASSO
+  LASSO_BMI_GSD_geom_point_dat <- est_sim %>% 
+    select(!!!(colname_LASSO_theta_sim_BMI_GSD)) %>% 
+    pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+    mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+  ggplot(LASSO_GSD_performance_metrics, aes(x=N, y=LASSO_GSD_mean))+ 
+    geom_point()+
+    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+           label= expression( paste("ref: ", theta[X] ))  , 
+           color= ggthemes_data$few$colors$Dark[9,2][[1]]
+             )+
+    ## ADD rainbow plots
+    geom_boxplot(data = LASSO_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                 col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                 alpha=0.2, size=0.2, outlier.shape = NA)+
+    ggdist::stat_halfeye(data = LASSO_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                         inherit.aes = FALSE, 
+                         slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                         geom = "slabinterval",
+                         na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+    gghalves::geom_half_point(
+      data = LASSO_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+      inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+      side = "l", range_scale = .4, alpha = .2)+
+    ## FINISH rainbow
+    scale_x_continuous(labels = comma)+
+    xlab("Size of Sample I")+
+    ylab("Mean thetas")+
+    scale_y_continuous(limits = c(0.25,0.8))+
+    labs(title = "Simulated Thetas with Empirical Standard Error", 
+         subtitle = paste0("LASSO estimate: BMI -> GSD\nSelected Reference Method: ", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> LASSO_GSD_simplot_thetas
+
+  ####-------------
+  # SAVE  --------------------------------
+  #1
   filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
@@
          )
+  #2
   filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
@@
          )
+  #3
   filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
@@
          )
+  #4
   filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MRMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
@@
          )
+  #5
+  filename_MEDIAN_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MEDIAN_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_MEDIAN_GSD_simplot_thetas, plot = MEDIAN_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+  #6
+  filename_EGGER_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_EGGER_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_EGGER_GSD_simplot_thetas, plot = EGGER_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+  #7
+  filename_ROBUST_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_ROBUST_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_ROBUST_GSD_simplot_thetas, plot = ROBUST_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+  #8
+  filename_PRESSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_PRESSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_PRESSO_GSD_simplot_thetas, plot = PRESSO_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+  #9
+  filename_RAPS_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_RAPS_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_RAPS_GSD_simplot_thetas, plot = RAPS_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
+  #10
+  filename_LASSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_LASSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  ggsave(filename = filename_LASSO_GSD_simplot_thetas, plot = LASSO_GSD_simplot_thetas, 
+         width = 15, height = 6, 
+         units = "in"  # default
+         )
 
@@
   MRMIX_GSD_simplot_thetas$labels$title <- ""
-  ggarrange(IVW_GSD_simplot_thetas,MODE_GSD_simplot_thetas,CONMIX_GSD_simplot_thetas, MRMIX_GSD_simplot_thetas,
-            ncol = 1, nrow = 4) -> GSD_simplot_thetas_arrange
+  #5
+  MEDIAN_GSD_simplot_thetas$labels$title <- ""
+  #6
+  EGGER_GSD_simplot_thetas$labels$title <- ""
+  #7
+  ROBUST_GSD_simplot_thetas$labels$title <- ""
+  #8
+  PRESSO_GSD_simplot_thetas$labels$title <- ""
+  #9
+  RAPS_GSD_simplot_thetas$labels$title <- ""
+  #10
+  LASSO_GSD_simplot_thetas$labels$title <- ""
+  ggarrange(IVW_GSD_simplot_thetas,MODE_GSD_simplot_thetas,CONMIX_GSD_simplot_thetas, MRMIX_GSD_simplot_thetas, MEDIAN_GSD_simplot_thetas,
+            EGGER_GSD_simplot_thetas, ROBUST_GSD_simplot_thetas, PRESSO_GSD_simplot_thetas, RAPS_GSD_simplot_thetas, LASSO_GSD_simplot_thetas,
+            ncol = 1, nrow = 8) -> GSD_simplot_thetas_arrange
   GSD_simplot_thetas_arrange
@@
   ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
-         width = 10, height = 22, 
+         width = 10, height = 38, 
          units = "in"  # default
@@
 
-  ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD, bias_mse_plot_MRMIX_GSD,
+  ## 5 MEDIAN
+  ggplot(MEDIAN_GSD_performance_metrics, aes(x=N, y=MEDIAN_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "MEDIAN estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_MEDIAN_GSD
+  ggplot(MEDIAN_GSD_performance_metrics, aes(x=N, y=MEDIAN_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "MEDIAN estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_MEDIAN_GSD
+  ggarrange(biasplot_MEDIAN_GSD, mseplot_MEDIAN_GSD) -> bias_mse_plot_MEDIAN_GSD
+  ## 6 EGGER
+  ggplot(EGGER_GSD_performance_metrics, aes(x=N, y=EGGER_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "EGGER estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_EGGER_GSD
+  ggplot(EGGER_GSD_performance_metrics, aes(x=N, y=EGGER_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "EGGER estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_EGGER_GSD
+  ggarrange(biasplot_EGGER_GSD, mseplot_EGGER_GSD) -> bias_mse_plot_EGGER_GSD
+  ## 7 ROBUST
+  ggplot(ROBUST_GSD_performance_metrics, aes(x=N, y=ROBUST_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "ROBUST estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_ROBUST_GSD
+  ggplot(ROBUST_GSD_performance_metrics, aes(x=N, y=ROBUST_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "ROBUST estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_ROBUST_GSD
+  ggarrange(biasplot_ROBUST_GSD, mseplot_ROBUST_GSD) -> bias_mse_plot_ROBUST_GSD
+  ## 8 PRESSO
+  ggplot(PRESSO_GSD_performance_metrics, aes(x=N, y=PRESSO_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "PRESSO estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_PRESSO_GSD
+  ggplot(PRESSO_GSD_performance_metrics, aes(x=N, y=PRESSO_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "PRESSO estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_PRESSO_GSD
+  ggarrange(biasplot_PRESSO_GSD, mseplot_PRESSO_GSD) -> bias_mse_plot_PRESSO_GSD
+  ## 9 RAPS
+  ggplot(RAPS_GSD_performance_metrics, aes(x=N, y=RAPS_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "RAPS estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_RAPS_GSD
+  ggplot(RAPS_GSD_performance_metrics, aes(x=N, y=RAPS_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "RAPS estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_RAPS_GSD
+  ggarrange(biasplot_RAPS_GSD, mseplot_RAPS_GSD) -> bias_mse_plot_RAPS_GSD
+  ## LASSO
+  ggplot(LASSO_GSD_performance_metrics, aes(x=N, y=LASSO_GSD_bias))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(-1.5,1.5))+
+    xlab("Size of Sample I")+
+    ylab("Bias")+
+    labs(title = "LASSO estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> biasplot_LASSO_GSD
+  ggplot(LASSO_GSD_performance_metrics, aes(x=N, y=LASSO_GSD_mse))+ 
+    geom_point()+
+    geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+    scale_x_continuous(labels = comma)+
+    scale_y_continuous(limits = c(0,1.8))+
+    xlab("Size of Sample I")+
+    ylab("MSE")+
+    labs(title = "LASSO estimate: BMI -> GSD",
+         caption = paste0("Selected Reference Method:\n", 
+                           selected_true_methods[true_method], 
+                           " estimate calculated with total sample size"))+
+    theme_few() -> mseplot_LASSO_GSD
+  ggarrange(biasplot_LASSO_GSD, mseplot_LASSO_GSD) -> bias_mse_plot_LASSO_GSD
+
+  ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD, bias_mse_plot_MRMIX_GSD, bias_mse_plot_MEDIAN_GSD,
+            bias_mse_plot_EGGER_GSD, bias_mse_plot_ROBUST_GSD, bias_mse_plot_PRESSO_GSD, bias_mse_plot_RAPS_GSD, bias_mse_plot_LASSO_GSD,
             ncol = 1, nrow = 4, common.legend = TRUE) -> bias_mse_plot_GSD

### 2023-03-03 scripts/rscript4simstudy.R
Rscript for simulation process

@@ -1,0 +1,24 @@
+
+# callr -------------------------------------------------------------------
+callr::rscript( # https://callr.r-lib.org/reference/rscript.html
+  script = "scripts/purrr-simstudy.R",
+  cmdargs = character(),
+  libpath = .libPaths()[1],
+  #repos = default_repos(),
+  stdout = paste0(Sys.Date(), "_slurm.out"),  # Optionally a file name to send the standard output to
+  stderr = paste0(Sys.Date(), "_error.out"),
+  #poll_connection = TRUE,
+  echo = FALSE,  # default=FALSE
+  show = TRUE,
+  #callback = NULL,
+  #block_callback = NULL,
+  #spinner = FALSE,
+  #system_profile = FALSE,
+  #user_profile = "project",
+  #env = rcmd_safe_env(),
+  timeout = as.difftime(12, units = "hours"),  # default=Inf
+  #wd = "~/bestageing2022",  # defaults to the current working directory.
+  fail_on_status = TRUE,  # default=FALSE
+  color = TRUE,
+  #...
+)

### 2023-03-13 scripts/MSc.Rmd
combine performance stats into list of dataframes per scenario

@@
 knitr::opts_chunk$set(fig.path = figurespath)  # reference is script! also creates new dirs
+
+SAVE.files <- FALSE
 
@@
 
+if (SAVE.files ==TRUE ){
   filenamefig00_b <- paste0("../output/plots/", Sys.Date(), "_explained_var_X_new_sample_size_dep.png")
@@
        )
+}
+
 ```
@@
 
+if (SAVE.files ==TRUE ){
   filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_new_PVE_X_formula.png")
@@
          )
+}
 ```
@@
 # SAVE
+if (SAVE.files ==TRUE ){
   filenamefig02 <- paste0("../output/plots/", Sys.Date(), "_figure02.png")
   ggsave(filename = filenamefig02, plot = fig02.update)
+}
 ```
@@
 final_plot1.arrange2   # use this!
-
+if (SAVE.files ==TRUE ){
   filename_final_plot1 <- paste0("../output/plots/", Sys.Date(), "_final_plot1.arrange_OLD.png")
@@
            )
+}
 ```
@@
 # data FOR estimates in `perform_mr.R` script and simulation in `simstudy.R` script
+
+if (SAVE.files ==TRUE ){
   saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
+}
 
@@
 # SAVE
+if (SAVE.files ==TRUE ){
   filename.mr.scatter.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrscatter_gsd.png")
@@
 #SAVE
+if (SAVE.files ==TRUE ){
   filename.mr.forest.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_GSD.png")
@@
 #SAVE
+if (SAVE.files ==TRUE ){
   filename.mr.funnel.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrfunnel_gsd.png")
@@
            )
+}
 
@@
 # SAVE
+if (SAVE.files ==TRUE ){
   filename.mr.scatter <- paste0("../output/plots/", Sys.Date(), "_figure_mrscatter.png")
@@
 #SAVE
+if (SAVE.files ==TRUE ){
   filename.mr.forest <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_GBC.png")
@@
 #SAVE
+if (SAVE.files ==TRUE ){
   filename.mr.funnel.gbc <- paste0("../output/plots/", Sys.Date(), "_figure_mrfunnel_gbc.png")
@@
 sim_vars <- list(n=n, colnames_pval=colnames_pval)
+if (SAVE.files ==TRUE ){
   saveRDS(sim_vars, paste0("../output/Rdata/", Sys.Date(), "_sim_vars.rds"))
+}
 ## SIMULATION OUTSOURCED TO "./scripts/simstudy.R"
@@
 est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_purrr_NSIM_", full.names = TRUE)))
+est_sim_architectures <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_architectures_purrr_NSIM_", full.names = TRUE)))
 nsim <- nrow(est_sim)
@@
 
-# EXPOSURE PVE
+# EXPOSURE PVE ------------------------------------------------------------------------------------------------------
 est_sim %>% 
@@
 
+if (SAVE.files ==TRUE ){
   filename_pveplot_sim <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_pve.png")
@@
 
+if (SAVE.files ==TRUE ){
   filename_pveplot_sim.raincloud <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_pve_raincloud.png")
@@
 ```{r store-performance-metrics-csv}
-## GSD
-filename_IVW_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_IVW_GSD_performance_metrics_NSIM_", nsim, ".csv")
-filename_MODE_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_MODE_GSD_performance_metrics_NSIM_", nsim, ".csv")
-filename_CONMIX_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_CONMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
-filename_MRMIX_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_MRMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
-#5
-filename_MEDIAN_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_MEDIAN_GSD_performance_metrics_NSIM_", nsim, ".csv")
-#6
-filename_EGGER_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_EGGER_GSD_performance_metrics_NSIM_", nsim, ".csv")
-#7
-filename_ROBUST_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_ROBUST_GSD_performance_metrics_NSIM_", nsim, ".csv")
-#8
-filename_PRESSO_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_PRESSO_GSD_performance_metrics_NSIM_", nsim, ".csv")
-#9
-filename_RAPS_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_RAPS_GSD_performance_metrics_NSIM_", nsim, ".csv")
-#10
-filename_LASSO_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-                                               Sys.Date(), "_LASSO_GSD_performance_metrics_NSIM_", nsim, ".csv")
-write.csv2(x = IVW_GSD_performance_metrics, file = filename_IVW_GSD_performance_metrics)
-write.csv2(x = MODE_GSD_performance_metrics, file = filename_MODE_GSD_performance_metrics)
-write.csv2(x = CONMIX_GSD_performance_metrics, file = filename_CONMIX_GSD_performance_metrics)
-write.csv2(x = MRMIX_GSD_performance_metrics, file = filename_MRMIX_GSD_performance_metrics)
-#5
-write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename_MEDIAN_GSD_performance_metrics)
-#6
-write.csv2(x = EGGER_GSD_performance_metrics, file = filename_EGGER_GSD_performance_metrics)
-#7
-write.csv2(x = ROBUST_GSD_performance_metrics, file = filename_ROBUST_GSD_performance_metrics)
-#8
-write.csv2(x = PRESSO_GSD_performance_metrics, file = filename_PRESSO_GSD_performance_metrics)
-#9
-write.csv2(x = RAPS_GSD_performance_metrics, file = filename_RAPS_GSD_performance_metrics)
-#10
-write.csv2(x = LASSO_GSD_performance_metrics, file = filename_LASSO_GSD_performance_metrics)
+## GSD  we loop in the next section and save the files there with all references!
+# filename_IVW_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_IVW_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# filename_MODE_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_MODE_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# filename_CONMIX_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_CONMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# filename_MRMIX_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_MRMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# #5
+# filename_MEDIAN_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_MEDIAN_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# #6
+# filename_EGGER_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_EGGER_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# #7
+# filename_ROBUST_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_ROBUST_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# #8
+# filename_PRESSO_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_PRESSO_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# #9
+# filename_RAPS_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_RAPS_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# #10
+# filename_LASSO_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
+#                                                Sys.Date(), "_LASSO_GSD_performance_metrics_NSIM_", nsim, ".csv")
+# write.csv2(x = IVW_GSD_performance_metrics, file = filename_IVW_GSD_performance_metrics)
+# write.csv2(x = MODE_GSD_performance_metrics, file = filename_MODE_GSD_performance_metrics)
+# write.csv2(x = CONMIX_GSD_performance_metrics, file = filename_CONMIX_GSD_performance_metrics)
+# write.csv2(x = MRMIX_GSD_performance_metrics, file = filename_MRMIX_GSD_performance_metrics)
+# #5
+# write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename_MEDIAN_GSD_performance_metrics)
+# #6
+# write.csv2(x = EGGER_GSD_performance_metrics, file = filename_EGGER_GSD_performance_metrics)
+# #7
+# write.csv2(x = ROBUST_GSD_performance_metrics, file = filename_ROBUST_GSD_performance_metrics)
+# #8
+# write.csv2(x = PRESSO_GSD_performance_metrics, file = filename_PRESSO_GSD_performance_metrics)
+# #9
+# write.csv2(x = RAPS_GSD_performance_metrics, file = filename_RAPS_GSD_performance_metrics)
+# #10
+# write.csv2(x = LASSO_GSD_performance_metrics, file = filename_LASSO_GSD_performance_metrics)
 ```
@@
 
-# SAVE
-filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_NSIM_", nsim,".png")
-ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
-       width = 15, height = 6, 
-       units = "in"  # default
-       )
-filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_NSIM_", nsim,".png")
-ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
-       width = 15, height = 6, 
-       units = "in"  # default
-       )
-filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_NSIM_", nsim,".png")
-ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
-       width = 15, height = 6, 
-       units = "in"  # default
-       )
-filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_NSIM_", nsim,".png")
-ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
-       width = 15, height = 6, 
-       units = "in"  # default
-       )
+# # SAVE   # we loop in the next section and save the ggarrange plots for inspection
+# filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_NSIM_", nsim,".png")
+# ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
+#        width = 15, height = 6, 
+#        units = "in"  # default
+#        )
+# filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_NSIM_", nsim,".png")
+# ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
+#        width = 15, height = 6, 
+#        units = "in"  # default
+#        )
+# filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_NSIM_", nsim,".png")
+# ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
+#        width = 15, height = 6, 
+#        units = "in"  # default
+#        )
+# filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_NSIM_", nsim,".png")
+# ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
+#        width = 15, height = 6, 
+#        units = "in"  # default
+#        )
 ```
@@
 
+if (SAVE.files ==TRUE ){
   filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_NSIM_", nsim,".png")
@@
 
+if (SAVE.files ==TRUE ){
   filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_NSIM_", nsim, ".png")
@@
 
+  # #----
+  # COMBINE TO PRESENT RESULTS ----------------------------------------------------
+  # #----
+  df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method])
+  assign(x = df_name, value = tibble(N=n))
+  # scenarios <- c("all_instruments", "common_strong", "common_weak", "rare_strong", "rare_weak")
+  assign(x = df_name, value = eval(as.symbol(df_name)) %>% 
+    left_join(IVW_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(MODE_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(CONMIX_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(MRMIX_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(MEDIAN_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(EGGER_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(ROBUST_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(PRESSO_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(RAPS_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+    left_join(LASSO_GSD_performance_metrics %>% select(-var), by = c("N"="N"))
+  )
+  #eval(as.symbol(df_name))
+
   # STORE
+  if (SAVE.files ==TRUE ){
     filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
@@
     write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
+  }
 
@@
   # SAVE  --------------------------------
+  if (SAVE.files ==TRUE ){
     #1
@@
            )
+  }
 
@@
             EGGER_GSD_simplot_thetas, ROBUST_GSD_simplot_thetas, PRESSO_GSD_simplot_thetas, RAPS_GSD_simplot_thetas, LASSO_GSD_simplot_thetas,
-            ncol = 1, nrow = 8) -> GSD_simplot_thetas_arrange
+            ncol = 1, nrow = 10) -> GSD_simplot_thetas_arrange
   GSD_simplot_thetas_arrange
 
+  if (SAVE.files ==TRUE ){
     filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
@@
             bias_mse_plot_EGGER_GSD, bias_mse_plot_ROBUST_GSD, bias_mse_plot_PRESSO_GSD, bias_mse_plot_RAPS_GSD, bias_mse_plot_LASSO_GSD,
-            ncol = 1, nrow = 4, common.legend = TRUE) -> bias_mse_plot_GSD
+            ncol = 1, nrow = 10, common.legend = TRUE) -> bias_mse_plot_GSD
   bias_mse_plot_GSD
 
+  if (SAVE.files ==TRUE ){
     filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], "_NSIM_",nsim, ".png")
     ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
-         width = 7, height = 12, 
+           width = 7, height = 24, 
            units = "in"  # default
@@
   }
+}
+
 ```
+
+## Now loop over all genetic architecture simulations
+
+```{r parallel-performance-genetic-architectures}
+effect_size <- c("strong", "weak")
+maf <- c("rare", "common")
+all_quadrants <- tidyr::crossing(effect_size, maf)
+
+# create copy of est_sim so the genetic architectures can be assigned to this variable in the loop 
+# without losing the simulations on all variants not considering genetic architecture
+est_sim_all_variants <- est_sim
+
+# now we do not need to rewrite the code from above but can assign each genetic architecture df to  est_sim
+# and only need to change the names of the files we store for later usage
+for (my_architecture in 1:length(est_sim_architectures)) {
+
+  est_sim <- est_sim_architectures[[my_architecture]]
+  name_GenArch <- gsub(pattern = "_est_sim", replacement = "", names(est_sim_architectures)[my_architecture])
+
+  # fewer variants -> no results for smaller sample sizes -> reassign sample sizes and colnames
+
+  #1
+  colname_IVW_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_sim_BMI_GSD_")]
+  colname_IVW_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "IVW_theta_se_sim_BMI_GSD_")]
+  #2
+  colname_MODE_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_sim_BMI_GSD_")]
+  colname_MODE_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MODE_theta_se_sim_BMI_GSD_")]
+  #3
+  colname_CONMIX_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_sim_BMI_GSD_")]
+  colname_CONMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "CONMIX_theta_se_sim_BMI_GSD_")]
+  #4
+  colname_MRMIX_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MRMIX_theta_sim_BMI_GSD_")]
+  colname_MRMIX_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MRMIX_theta_se_sim_BMI_GSD_")]
+  #5
+  colname_MEDIAN_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MEDIAN_theta_sim_BMI_GSD_")]
+  colname_MEDIAN_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "MEDIAN_theta_se_sim_BMI_GSD_")]
+  #6
+  colname_EGGER_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "EGGER_theta_sim_BMI_GSD_")]
+  colname_EGGER_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "EGGER_theta_se_sim_BMI_GSD_")]
+  #7
+  colname_ROBUST_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "ROBUST_theta_sim_BMI_GSD_")]
+  colname_ROBUST_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "ROBUST_theta_se_sim_BMI_GSD_")]
+  #8
+  colname_PRESSO_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "PRESSO_theta_sim_BMI_GSD_")]
+  colname_PRESSO_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "PRESSO_theta_se_sim_BMI_GSD_")]
+  #9
+  colname_RAPS_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "RAPS_theta_sim_BMI_GSD_")]
+  colname_RAPS_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "RAPS_theta_se_sim_BMI_GSD_")]
+  #10
+  colname_LASSO_theta_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "LASSO_theta_sim_BMI_GSD_")]
+  colname_LASSO_theta_se_sim_BMI_GSD <- colnames(est_sim)[str_detect(colnames(est_sim), pattern = "LASSO_theta_se_sim_BMI_GSD_")]
+
+  # retrieve IVW/ MODE/ CONMIX cols for BMI -> GBC
+  # removed on 2023-01-24 (see github)
+
+  # filter out sample sizes (IVW only used as example, could be any other col)
+  sample_size_filtered <- as.numeric(str_extract(colname_IVW_theta_sim_BMI_GSD, pattern = "(\\d)+"))
+
+  ### NOW START LOOP --------------------------------------------------------
+  for (true_method in 1:length(selected_true_methods)) {
+    # define reference estimate for ALL 5 chosen methods in loop 
+    args.gsd.method <- list(theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]], 
+                            nsim=nrow(est_sim)
+    )
+    #1
+    IVW_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_IVW_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, IVW_GSD_mean=mean, IVW_GSD_emp_se=emp_se, IVW_GSD_bias=bias, IVW_GSD_mse=mse)
+    #2
+    MODE_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_MODE_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, MODE_GSD_mean=mean, MODE_GSD_emp_se=emp_se, MODE_GSD_bias=bias, MODE_GSD_mse=mse)
+    #3
+    CONMIX_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_CONMIX_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, CONMIX_GSD_mean=mean, CONMIX_GSD_emp_se=emp_se, CONMIX_GSD_bias=bias, CONMIX_GSD_mse=mse)
+    #4
+    MRMIX_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_MRMIX_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, MRMIX_GSD_mean=mean, MRMIX_GSD_emp_se=emp_se, MRMIX_GSD_bias=bias, MRMIX_GSD_mse=mse)
+    #5
+    MEDIAN_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_MEDIAN_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, MEDIAN_GSD_mean=mean, MEDIAN_GSD_emp_se=emp_se, MEDIAN_GSD_bias=bias, MEDIAN_GSD_mse=mse)
+    #6
+    EGGER_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_EGGER_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, EGGER_GSD_mean=mean, EGGER_GSD_emp_se=emp_se, EGGER_GSD_bias=bias, EGGER_GSD_mse=mse)
+    #7
+    ROBUST_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_ROBUST_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, ROBUST_GSD_mean=mean, ROBUST_GSD_emp_se=emp_se, ROBUST_GSD_bias=bias, ROBUST_GSD_mse=mse)
+    #8
+    PRESSO_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_PRESSO_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, PRESSO_GSD_mean=mean, PRESSO_GSD_emp_se=emp_se, PRESSO_GSD_bias=bias, PRESSO_GSD_mse=mse)
+    #9
+    RAPS_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_RAPS_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, RAPS_GSD_mean=mean, RAPS_GSD_emp_se=emp_se, RAPS_GSD_bias=bias, RAPS_GSD_mse=mse)
+    #10
+    LASSO_GSD_performance_metrics <- est_sim %>% 
+      select(!!!(colname_LASSO_theta_sim_BMI_GSD)) %>% 
+      map_df(.f = ~ funs %>% 
+               map(exec, .x, !!!args.gsd.method), .id = "var") %>% 
+      cbind(sample_size_filtered) %>% 
+      rename(N = sample_size_filtered, LASSO_GSD_mean=mean, LASSO_GSD_emp_se=emp_se, LASSO_GSD_bias=bias, LASSO_GSD_mse=mse)
+
+    ## further metrics
+    IVW_GSD_coverage_calc <- c()
+    IVW_GSD_coverage_bias_eliminated_calc <- c()
+    IVW_GSD_type1error_calc <- c()
+    MODE_GSD_coverage_calc <- c()
+    MODE_GSD_coverage_bias_eliminated_calc <- c()
+    MODE_GSD_type1error_calc <- c()
+    CONMIX_GSD_coverage_calc <- c()
+    CONMIX_GSD_coverage_bias_eliminated_calc <- c()
+    CONMIX_GSD_type1error_calc <- c()
+    MRMIX_GSD_coverage_calc <- c()
+    MRMIX_GSD_coverage_bias_eliminated_calc <- c()
+    MRMIX_GSD_type1error_calc <- c()
+    #5
+    MEDIAN_GSD_coverage_calc <- c()
+    MEDIAN_GSD_coverage_bias_eliminated_calc <- c()
+    MEDIAN_GSD_type1error_calc <- c()
+    #6
+    EGGER_GSD_coverage_calc <- c()
+    EGGER_GSD_coverage_bias_eliminated_calc <- c()
+    EGGER_GSD_type1error_calc <- c()
+    #7
+    ROBUST_GSD_coverage_calc <- c()
+    ROBUST_GSD_coverage_bias_eliminated_calc <- c()
+    ROBUST_GSD_type1error_calc <- c()
+    #8
+    PRESSO_GSD_coverage_calc <- c()
+    PRESSO_GSD_coverage_bias_eliminated_calc <- c()
+    PRESSO_GSD_type1error_calc <- c()
+    #9
+    RAPS_GSD_coverage_calc <- c()
+    RAPS_GSD_coverage_bias_eliminated_calc <- c()
+    RAPS_GSD_type1error_calc <- c()
+    #10
+    LASSO_GSD_coverage_calc <- c()
+    LASSO_GSD_coverage_bias_eliminated_calc <- c()
+    LASSO_GSD_type1error_calc <- c()
+
+    for (i in 1:length(sample_size_filtered)) {
+      sample_size <- sample_size_filtered[i]
+      # GSD additional metrics ----------------------------------------------------
+      ## 1/ IVW
+      col_IVW_GSD_theta <- paste0("IVW_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_IVW_GSD_theta_se <- paste0("IVW_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_IVW_GSD_mr_pval <- paste0("IVW_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      IVW_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_IVW_GSD_theta], thetaobs_se = est_sim[col_IVW_GSD_theta_se], theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      IVW_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_IVW_GSD_theta]], thetaobs_se = est_sim[[col_IVW_GSD_theta_se]])
+      IVW_GSD_type1error_calc[i] <- type1error(est_sim[[col_IVW_GSD_mr_pval]])
+      ## 2/ MODE
+      col_MODE_GSD_theta <- paste0("MODE_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_MODE_GSD_theta_se <- paste0("MODE_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_MODE_GSD_mr_pval <- paste0("MODE_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      MODE_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MODE_GSD_theta], 
+                                           thetaobs_se = est_sim[col_MODE_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      MODE_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MODE_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_MODE_GSD_theta_se]])
+      MODE_GSD_type1error_calc[i] <- type1error(est_sim[[col_MODE_GSD_mr_pval]])
+      ## 3/ CONMIX
+      col_CONMIX_GSD_theta <- paste0("CONMIX_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_CONMIX_GSD_theta_se <- paste0("CONMIX_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_CONMIX_GSD_mr_pval <- paste0("CONMIX_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      CONMIX_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_CONMIX_GSD_theta], 
+                                           thetaobs_se = est_sim[col_CONMIX_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      CONMIX_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_CONMIX_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_CONMIX_GSD_theta_se]])
+      CONMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_CONMIX_GSD_mr_pval]])  
+      ## 4/ MRMIX
+      col_MRMIX_GSD_theta <- paste0("MRMIX_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_MRMIX_GSD_theta_se <- paste0("MRMIX_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_MRMIX_GSD_mr_pval <- paste0("MRMIX_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      MRMIX_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MRMIX_GSD_theta], 
+                                           thetaobs_se = est_sim[col_MRMIX_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      MRMIX_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MRMIX_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_MRMIX_GSD_theta_se]])
+      MRMIX_GSD_type1error_calc[i] <- type1error(est_sim[[col_MRMIX_GSD_mr_pval]])  
+      ## 5/ MEDIAN
+      col_MEDIAN_GSD_theta <- paste0("MEDIAN_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_MEDIAN_GSD_theta_se <- paste0("MEDIAN_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_MEDIAN_GSD_mr_pval <- paste0("MEDIAN_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      MEDIAN_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_MEDIAN_GSD_theta], 
+                                           thetaobs_se = est_sim[col_MEDIAN_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      MEDIAN_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_MEDIAN_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_MEDIAN_GSD_theta_se]])
+      MEDIAN_GSD_type1error_calc[i] <- type1error(est_sim[[col_MEDIAN_GSD_mr_pval]])  
+      ## 6/ EGGER
+      col_EGGER_GSD_theta <- paste0("EGGER_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_EGGER_GSD_theta_se <- paste0("EGGER_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_EGGER_GSD_mr_pval <- paste0("EGGER_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      EGGER_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_EGGER_GSD_theta], 
+                                           thetaobs_se = est_sim[col_EGGER_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      EGGER_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_EGGER_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_EGGER_GSD_theta_se]])
+      EGGER_GSD_type1error_calc[i] <- type1error(est_sim[[col_EGGER_GSD_mr_pval]])
+      ## 7/ ROBUST
+      col_ROBUST_GSD_theta <- paste0("ROBUST_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_ROBUST_GSD_theta_se <- paste0("ROBUST_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_ROBUST_GSD_mr_pval <- paste0("ROBUST_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      ROBUST_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_ROBUST_GSD_theta], 
+                                           thetaobs_se = est_sim[col_ROBUST_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      ROBUST_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_ROBUST_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_ROBUST_GSD_theta_se]])
+      ROBUST_GSD_type1error_calc[i] <- type1error(est_sim[[col_ROBUST_GSD_mr_pval]])
+      ## 8/ PRESSO
+      col_PRESSO_GSD_theta <- paste0("PRESSO_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_PRESSO_GSD_theta_se <- paste0("PRESSO_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_PRESSO_GSD_mr_pval <- paste0("PRESSO_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      PRESSO_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_PRESSO_GSD_theta], 
+                                           thetaobs_se = est_sim[col_PRESSO_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      PRESSO_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_PRESSO_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_PRESSO_GSD_theta_se]])
+      PRESSO_GSD_type1error_calc[i] <- type1error(est_sim[[col_PRESSO_GSD_mr_pval]])
+      ## 9/ RAPS
+      col_RAPS_GSD_theta <- paste0("RAPS_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_RAPS_GSD_theta_se <- paste0("RAPS_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_RAPS_GSD_mr_pval <- paste0("RAPS_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      RAPS_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_RAPS_GSD_theta], 
+                                           thetaobs_se = est_sim[col_RAPS_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      RAPS_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_RAPS_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_RAPS_GSD_theta_se]])
+      RAPS_GSD_type1error_calc[i] <- type1error(est_sim[[col_RAPS_GSD_mr_pval]])
+      ## 10/LASSO
+      col_LASSO_GSD_theta <- paste0("LASSO_theta_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_LASSO_GSD_theta_se <- paste0("LASSO_theta_se_sim_BMI_GSD_", format(sample_size, scientific = FALSE))
+      col_LASSO_GSD_mr_pval <- paste0("LASSO_mr_pval_BMI_GSD_", format(sample_size, scientific = FALSE))
+      # calc
+      LASSO_GSD_coverage_calc[i] <- coverage(thetaobs = est_sim[col_LASSO_GSD_theta], 
+                                           thetaobs_se = est_sim[col_LASSO_GSD_theta_se], 
+                                           theta = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == selected_true_methods[true_method]])
+      LASSO_GSD_coverage_bias_eliminated_calc[i] <- coverage_bias_eliminated(thetaobs = est_sim[[col_LASSO_GSD_theta]], 
+                                                                           thetaobs_se = est_sim[[col_LASSO_GSD_theta_se]])
+      LASSO_GSD_type1error_calc[i] <- type1error(est_sim[[col_LASSO_GSD_mr_pval]])
+      #-----------
+    }
+
+    ## combine metrics
+    IVW_GSD_performance_metrics <- IVW_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(IVW_GSD_coverage_calc, IVW_GSD_coverage_bias_eliminated_calc, IVW_GSD_type1error_calc)) %>% 
+      as_tibble()
+    MODE_GSD_performance_metrics <- MODE_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(MODE_GSD_coverage_calc, 
+                       MODE_GSD_coverage_bias_eliminated_calc, MODE_GSD_type1error_calc)) %>% 
+      as_tibble()
+    CONMIX_GSD_performance_metrics <- CONMIX_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(CONMIX_GSD_coverage_calc, 
+                       CONMIX_GSD_coverage_bias_eliminated_calc, CONMIX_GSD_type1error_calc)) %>% 
+      as_tibble()
+    MRMIX_GSD_performance_metrics <- MRMIX_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(MRMIX_GSD_coverage_calc, 
+                       MRMIX_GSD_coverage_bias_eliminated_calc, MRMIX_GSD_type1error_calc)) %>% 
+      as_tibble()
+    #5
+    MEDIAN_GSD_performance_metrics <- MEDIAN_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(MEDIAN_GSD_coverage_calc, 
+                       MEDIAN_GSD_coverage_bias_eliminated_calc, MEDIAN_GSD_type1error_calc)) %>% 
+      as_tibble()
+    #6
+    EGGER_GSD_performance_metrics <- EGGER_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(EGGER_GSD_coverage_calc, 
+                       EGGER_GSD_coverage_bias_eliminated_calc, EGGER_GSD_type1error_calc)) %>% 
+      as_tibble()
+    #7
+    ROBUST_GSD_performance_metrics <- ROBUST_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(ROBUST_GSD_coverage_calc, 
+                       ROBUST_GSD_coverage_bias_eliminated_calc, ROBUST_GSD_type1error_calc)) %>% 
+      as_tibble()
+    #8
+    PRESSO_GSD_performance_metrics <- PRESSO_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(PRESSO_GSD_coverage_calc, 
+                       PRESSO_GSD_coverage_bias_eliminated_calc, PRESSO_GSD_type1error_calc)) %>% 
+      as_tibble()
+    #9
+    RAPS_GSD_performance_metrics <- RAPS_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(RAPS_GSD_coverage_calc, 
+                       RAPS_GSD_coverage_bias_eliminated_calc, RAPS_GSD_type1error_calc)) %>% 
+      as_tibble()
+    #10
+    LASSO_GSD_performance_metrics <- LASSO_GSD_performance_metrics %>% 
+      select(N, everything()) %>% 
+      cbind(data.frame(LASSO_GSD_coverage_calc, 
+                       LASSO_GSD_coverage_bias_eliminated_calc, LASSO_GSD_type1error_calc)) %>% 
+      as_tibble()
+
+    # #----
+    # COMBINE TO PRESENT RESULTS Scenarios 2-5 ----------------------------------------------------
+    # #----
+    df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method], "_GenArch_", name_GenArch)
+    assign(x = df_name, value = tibble(N=n))
+
+    # scenarios <- c("all_instruments", "common_strong", "common_weak", "rare_strong", "rare_weak")
+    assign(x = df_name, value = eval(as.symbol(df_name)) %>% 
+      left_join(IVW_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(MODE_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(CONMIX_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(MRMIX_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(MEDIAN_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(EGGER_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(ROBUST_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(PRESSO_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(RAPS_GSD_performance_metrics %>% select(-var), by = c("N"="N")) %>% 
+      left_join(LASSO_GSD_performance_metrics %>% select(-var), by = c("N"="N"))
+    )
+    #eval(as.symbol(df_name))
+
+    # STORE
+    if (SAVE.files ==TRUE ){
+      filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_IVW_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = IVW_GSD_performance_metrics, file = filename.IVW_table.performance)
+      filename.MODE_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_MODE_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = MODE_GSD_performance_metrics, file = filename.MODE_table.performance)
+      filename.CONMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_CONMIX_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = CONMIX_GSD_performance_metrics, file = filename.CONMIX_table.performance)
+      filename.MRMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_MRMIX_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
+      #5
+      filename.MEDIAN_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_MEDIAN_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename.MEDIAN_table.performance)
+      #6
+      filename.EGGER_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_EGGER_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = EGGER_GSD_performance_metrics, file = filename.EGGER_table.performance)
+      #7
+      filename.ROBUST_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_ROBUST_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = ROBUST_GSD_performance_metrics, file = filename.ROBUST_table.performance)
+      #8
+      filename.PRESSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_PRESSO_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = PRESSO_GSD_performance_metrics, file = filename.PRESSO_table.performance)
+      #9
+      filename.RAPS_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_RAPS_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = RAPS_GSD_performance_metrics, file = filename.RAPS_table.performance)
+      #10
+      filename.LASSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+                                          "_LASSO_GSD_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch,
+                                          "_NSIM_", nsim, ".csv")
+      write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
+    }
+
+    ###
+    ## PLOT  -------------------------------------------------------------------
+    ###
+    IVW_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_IVW_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(IVW_GSD_performance_metrics, aes(x=N, y=IVW_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      # geom_errorbar(aes(ymin=IVW_GSD_mean-1.96*IVW_GSD_emp_se, ymax=IVW_GSD_mean+1.96*IVW_GSD_emp_se), 
+      #               width=3000, # x-axis dimension
+      #               position = "dodge")+
+      ## ADD rainbow plots
+      geom_boxplot(data = IVW_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = IVW_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = IVW_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("IVW estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> IVW_GSD_simplot_thetas
+
+    ## 2 MODE
+    MODE_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_MODE_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(MODE_GSD_performance_metrics, aes(x=N, y=MODE_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      # geom_errorbar(aes(ymin=MODE_GSD_mean-1.96*MODE_GSD_emp_se, ymax=MODE_GSD_mean+1.96*MODE_GSD_emp_se), 
+      #               width=3000, # x-axis dimension
+      #               position = "dodge")+
+      ## ADD rainbow plots
+      geom_boxplot(data = MODE_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = MODE_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = MODE_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("MODE estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> MODE_GSD_simplot_thetas
+
+    ## 3 CONMIX
+    CONMIX_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_CONMIX_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=CONMIX_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      ## ADD rainbow plots
+      geom_boxplot(data = CONMIX_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = CONMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = CONMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("CONMIX estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> CONMIX_GSD_simplot_thetas
+
+    ## 4 MRMIX
+    MRMIX_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_MRMIX_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=MRMIX_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      ## ADD rainbow plots
+      geom_boxplot(data = MRMIX_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = MRMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = MRMIX_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("MRMIX estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> MRMIX_GSD_simplot_thetas
+
+    ## 5 MEDIAN
+    MEDIAN_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_MEDIAN_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(MEDIAN_GSD_performance_metrics, aes(x=N, y=MEDIAN_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      ## ADD rainbow plots
+      geom_boxplot(data = MEDIAN_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = MEDIAN_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = MEDIAN_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("MEDIAN estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> MEDIAN_GSD_simplot_thetas
+
+    ## 6 EGGER
+    EGGER_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_EGGER_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(EGGER_GSD_performance_metrics, aes(x=N, y=EGGER_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      ## ADD rainbow plots
+      geom_boxplot(data = EGGER_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = EGGER_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = EGGER_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("EGGER estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> EGGER_GSD_simplot_thetas
+    ## 7 ROBUST
+    ROBUST_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_ROBUST_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(ROBUST_GSD_performance_metrics, aes(x=N, y=ROBUST_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      ## ADD rainbow plots
+      geom_boxplot(data = ROBUST_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = ROBUST_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = ROBUST_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("ROBUST estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> ROBUST_GSD_simplot_thetas
+    ## 8 PRESSO
+    PRESSO_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_PRESSO_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(PRESSO_GSD_performance_metrics, aes(x=N, y=PRESSO_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      ## ADD rainbow plots
+      geom_boxplot(data = PRESSO_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = PRESSO_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = PRESSO_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("PRESSO estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> PRESSO_GSD_simplot_thetas
+    ## 9 RAPS
+    RAPS_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_RAPS_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(RAPS_GSD_performance_metrics, aes(x=N, y=RAPS_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      ## ADD rainbow plots
+      geom_boxplot(data = RAPS_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = RAPS_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = RAPS_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("RAPS estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> RAPS_GSD_simplot_thetas
+    ## 10 LASSO
+    LASSO_BMI_GSD_geom_point_dat <- est_sim %>% 
+      select(!!!(colname_LASSO_theta_sim_BMI_GSD)) %>% 
+      pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
+      mutate(theta_sim_numeric = as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")))
+
+    ggplot(LASSO_GSD_performance_metrics, aes(x=N, y=LASSO_GSD_mean))+ 
+      geom_point()+
+      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+             label= expression( paste("ref: ", theta[X] ))  , 
+             color= ggthemes_data$few$colors$Dark[9,2][[1]]
+               )+
+      ## ADD rainbow plots
+      geom_boxplot(data = LASSO_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                   col=ggthemes_data$few$colors$Dark[3,2][[1]],
+                   alpha=0.2, size=0.2, outlier.shape = NA)+
+      ggdist::stat_halfeye(data = LASSO_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
+                           inherit.aes = FALSE, 
+                           slab_fill=ggthemes_data$few$colors$Dark[3,2][[1]], slab_alpha = 0.2,
+                           geom = "slabinterval",
+                           na.rm=TRUE, trim=TRUE, adjust = .5, width = 10000, .width = 0, justification = -.3, point_colour = NA)+
+      gghalves::geom_half_point(
+        data = LASSO_BMI_GSD_geom_point_dat, mapping=aes(x=theta_sim_numeric-2000, y=value, group=theta_sim_numeric), 
+        inherit.aes = FALSE, size=0.2, colour=ggthemes_data$few$colors$Dark[3,2][[1]],
+        side = "l", range_scale = .4, alpha = .2)+
+      ## FINISH rainbow
+      scale_x_continuous(labels = comma)+
+      xlab("Size of Sample I")+
+      ylab("Mean thetas")+
+      scale_y_continuous(limits = c(0.25,0.8))+
+      labs(title = "Simulated Thetas with Empirical Standard Error", 
+           subtitle = paste0("LASSO estimate: BMI -> GSD\nSelected Reference Method: ", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> LASSO_GSD_simplot_thetas
+
+    ####-------------
+    # SAVE  --------------------------------
+    # #1
+    # filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
+    #                                           "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #2
+    # filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
+    #                                            "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #3
+    # filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
+    #                                              "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #4
+    # filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MRMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
+    #                                             "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #5
+    # filename_MEDIAN_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MEDIAN_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
+    #                                              "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_MEDIAN_GSD_simplot_thetas, plot = MEDIAN_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #6
+    # filename_EGGER_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_EGGER_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
+    #                                             "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_EGGER_GSD_simplot_thetas, plot = EGGER_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #7
+    # filename_ROBUST_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_ROBUST_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
+    #                                              "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_ROBUST_GSD_simplot_thetas, plot = ROBUST_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #8
+    # filename_PRESSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_PRESSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
+    #                                              "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_PRESSO_GSD_simplot_thetas, plot = PRESSO_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #9
+    # filename_RAPS_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_RAPS_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
+    #                                            "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_RAPS_GSD_simplot_thetas, plot = RAPS_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+    # #10
+    # filename_LASSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_LASSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
+    #                                             "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+    # ggsave(filename = filename_LASSO_GSD_simplot_thetas, plot = LASSO_GSD_simplot_thetas, 
+    #        width = 15, height = 6, 
+    #        units = "in"  # default
+    #        )
+
+    IVW_GSD_simplot_thetas$labels$title <- ""
+    MODE_GSD_simplot_thetas$labels$title <- ""
+    CONMIX_GSD_simplot_thetas$labels$title <- ""
+    MRMIX_GSD_simplot_thetas$labels$title <- ""
+    #5
+    MEDIAN_GSD_simplot_thetas$labels$title <- ""
+    #6
+    EGGER_GSD_simplot_thetas$labels$title <- ""
+    #7
+    ROBUST_GSD_simplot_thetas$labels$title <- ""
+    #8
+    PRESSO_GSD_simplot_thetas$labels$title <- ""
+    #9
+    RAPS_GSD_simplot_thetas$labels$title <- ""
+    #10
+    LASSO_GSD_simplot_thetas$labels$title <- ""
+    ggarrange(IVW_GSD_simplot_thetas,MODE_GSD_simplot_thetas,CONMIX_GSD_simplot_thetas, MRMIX_GSD_simplot_thetas, MEDIAN_GSD_simplot_thetas,
+              EGGER_GSD_simplot_thetas, ROBUST_GSD_simplot_thetas, PRESSO_GSD_simplot_thetas, RAPS_GSD_simplot_thetas, LASSO_GSD_simplot_thetas,
+              ncol = 1, nrow = 10) -> GSD_simplot_thetas_arrange
+    GSD_simplot_thetas_arrange
+
+    if (SAVE.files ==TRUE ){
+      filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
+                                                    "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+      ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
+             width = 10, height = 38, 
+             units = "in"  # default
+             )
+    }
+
+    ### GSD Bias & MSE
+    #BIAS
+    ggplot(IVW_GSD_performance_metrics, aes(x=N, y=IVW_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "IVW estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_IVW_GSD
+    #MSE
+    ggplot(IVW_GSD_performance_metrics, aes(x=N, y=IVW_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "IVW estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_IVW_GSD
+    ggarrange(biasplot_IVW_GSD, mseplot_IVW_GSD) -> bias_mse_plot_IVW_GSD
+
+    ## MODE
+    ggplot(MODE_GSD_performance_metrics, aes(x=N, y=MODE_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "MODE estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_MODE_GSD
+    ggplot(MODE_GSD_performance_metrics, aes(x=N, y=MODE_GSD_mse))+ 
+      geom_point()+
+        geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "MODE estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_MODE_GSD
+    ggarrange(biasplot_MODE_GSD, mseplot_MODE_GSD) -> bias_mse_plot_MODE_GSD
+
+    ## CONMIX
+    ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=CONMIX_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "CONMIX estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_CONMIX_GSD
+    ggplot(CONMIX_GSD_performance_metrics, aes(x=N, y=CONMIX_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "CONMIX estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_CONMIX_GSD
+    ggarrange(biasplot_CONMIX_GSD, mseplot_CONMIX_GSD) -> bias_mse_plot_CONMIX_GSD
+
+    ## MRMIX
+    ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=MRMIX_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "MRMIX estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_MRMIX_GSD
+    ggplot(MRMIX_GSD_performance_metrics, aes(x=N, y=MRMIX_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "MRMIX estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_MRMIX_GSD
+    ggarrange(biasplot_MRMIX_GSD, mseplot_MRMIX_GSD) -> bias_mse_plot_MRMIX_GSD
+
+    ## 5 MEDIAN
+    ggplot(MEDIAN_GSD_performance_metrics, aes(x=N, y=MEDIAN_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "MEDIAN estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_MEDIAN_GSD
+    ggplot(MEDIAN_GSD_performance_metrics, aes(x=N, y=MEDIAN_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "MEDIAN estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_MEDIAN_GSD
+    ggarrange(biasplot_MEDIAN_GSD, mseplot_MEDIAN_GSD) -> bias_mse_plot_MEDIAN_GSD
+    ## 6 EGGER
+    ggplot(EGGER_GSD_performance_metrics, aes(x=N, y=EGGER_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "EGGER estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_EGGER_GSD
+    ggplot(EGGER_GSD_performance_metrics, aes(x=N, y=EGGER_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "EGGER estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_EGGER_GSD
+    ggarrange(biasplot_EGGER_GSD, mseplot_EGGER_GSD) -> bias_mse_plot_EGGER_GSD
+    ## 7 ROBUST
+    ggplot(ROBUST_GSD_performance_metrics, aes(x=N, y=ROBUST_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "ROBUST estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_ROBUST_GSD
+    ggplot(ROBUST_GSD_performance_metrics, aes(x=N, y=ROBUST_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "ROBUST estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_ROBUST_GSD
+    ggarrange(biasplot_ROBUST_GSD, mseplot_ROBUST_GSD) -> bias_mse_plot_ROBUST_GSD
+    ## 8 PRESSO
+    ggplot(PRESSO_GSD_performance_metrics, aes(x=N, y=PRESSO_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "PRESSO estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_PRESSO_GSD
+    ggplot(PRESSO_GSD_performance_metrics, aes(x=N, y=PRESSO_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "PRESSO estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_PRESSO_GSD
+    ggarrange(biasplot_PRESSO_GSD, mseplot_PRESSO_GSD) -> bias_mse_plot_PRESSO_GSD
+    ## 9 RAPS
+    ggplot(RAPS_GSD_performance_metrics, aes(x=N, y=RAPS_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "RAPS estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_RAPS_GSD
+    ggplot(RAPS_GSD_performance_metrics, aes(x=N, y=RAPS_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "RAPS estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_RAPS_GSD
+    ggarrange(biasplot_RAPS_GSD, mseplot_RAPS_GSD) -> bias_mse_plot_RAPS_GSD
+    ## LASSO
+    ggplot(LASSO_GSD_performance_metrics, aes(x=N, y=LASSO_GSD_bias))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(-1.5,1.5))+
+      xlab("Size of Sample I")+
+      ylab("Bias")+
+      labs(title = "LASSO estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> biasplot_LASSO_GSD
+    ggplot(LASSO_GSD_performance_metrics, aes(x=N, y=LASSO_GSD_mse))+ 
+      geom_point()+
+      geom_hline(yintercept=  0, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
+      scale_x_continuous(labels = comma)+
+      scale_y_continuous(limits = c(0,1.8))+
+      xlab("Size of Sample I")+
+      ylab("MSE")+
+      labs(title = "LASSO estimate: BMI -> GSD",
+           caption = paste0("Selected Reference Method:\n", 
+                             selected_true_methods[true_method], 
+                             " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
+      theme_few() -> mseplot_LASSO_GSD
+    ggarrange(biasplot_LASSO_GSD, mseplot_LASSO_GSD) -> bias_mse_plot_LASSO_GSD
+
+    ggarrange(bias_mse_plot_IVW_GSD, bias_mse_plot_MODE_GSD, bias_mse_plot_CONMIX_GSD, bias_mse_plot_MRMIX_GSD, bias_mse_plot_MEDIAN_GSD,
+              bias_mse_plot_EGGER_GSD, bias_mse_plot_ROBUST_GSD, bias_mse_plot_PRESSO_GSD, bias_mse_plot_RAPS_GSD, bias_mse_plot_LASSO_GSD,
+              ncol = 1, nrow = 10, common.legend = TRUE) -> bias_mse_plot_GSD
+    bias_mse_plot_GSD
+
+    if (SAVE.files ==TRUE ){
+      filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+      ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
+             width = 7, height = 24, 
+             units = "in"  # default
+             )
+    }
+  }
+}
+
+```
+
+# Create final tables and figures
+
+## Tables
+
+For all scenarios (a. all instruments, b. common+strong, c. common+weak, d. rare+strong, e. rare+weak) we need to describe our results for a smaller set of sample sizes.
+
+```{r}
+df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method])
+df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method], "_GenArch_", name_GenArch)
+
+df_metrics_all_scenarios <- list()
+for(true_method in seq_along(selected_true_methods)) {
+  # scenario 1) all IVs
+  df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method])
+  df_metrics_all_scenarios[[true_method]] <- eval(as.symbol(df_name)) %>% 
+    rename_with(.fn = ~ paste0(.x, "_ALL_INSTR"), .cols = -"N")
+
+  # scenarios 2)-5) GenArchs
+  for (my_architecture in seq_along(est_sim_architectures)) {
+    name_GenArch <- gsub(pattern = "_est_sim", replacement = "", names(est_sim_architectures)[my_architecture])
+    df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method], "_GenArch_", name_GenArch)
+
+    df_metrics_all_scenarios[[true_method]] <- df_metrics_all_scenarios[[true_method]] %>% 
+      left_join(eval(as.symbol(df_name)) %>% 
+                  rename_with(.fn = ~ paste0(.x, "_", name_GenArch), .cols = -"N"), by=c("N"="N"))
+  }
+}
+# name list
+names(df_metrics_all_scenarios) <- c("scenario_ALL_INSTR", "STRONG_COMMON", "STRONG_RARE", "WEAK_COMMON", "WEAK_RARE")
+```

### 2023-03-13 scripts/MSc.Rmd
summary gt tables

@@
 
-  # retrieve IVW/ MODE/ CONMIX cols for BMI -> GBC
-  # removed on 2023-01-24 (see github)
-
   # filter out sample sizes (IVW only used as example, could be any other col)
@@
 
+\clearpage
+
 # Create final tables and figures
@@
 
-For all scenarios (a. all instruments, b. common+strong, c. common+weak, d. rare+strong, e. rare+weak) we need to describe our results for a smaller set of sample sizes.
+For all scenarios (a.
+all instruments, b.
+common+strong, c.
+common+weak, d.
+rare+strong, e.
+rare+weak) we need to describe our results for a smaller set of sample sizes.
+We create a list that holds a dataframe for each reference method (1:5).
+Each dataframe contains results from all scenarios.
 
 ```{r}
+# we have assigned dataframes in the loop above!
 df_name <- paste0("df_combined_method_metrics_TRUE_EST_", selected_true_methods[true_method])
@@
 }
+
 # name list
-names(df_metrics_all_scenarios) <- c("scenario_ALL_INSTR", "STRONG_COMMON", "STRONG_RARE", "WEAK_COMMON", "WEAK_RARE")
+names(df_metrics_all_scenarios) <- paste0("ref_", selected_true_methods)
 ```
+
+```{r gt-summary-tabs}
+# https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r/
+library(gt)
+df.temp.long <- list()
+df.temp.wide <- list()
+
+sample_size_subset <- c(50000, 100000, 200000, 300000)
+
+for(ref.method in seq_along(df_metrics_all_scenarios)) {
+
+  ## MEAN ## ---------------------------------------------------------------
+  # later plot N=50,000 all methods and different scenarios
+
+  ## CREATE LONG FORMAT
+  df_metrics_all_scenarios[[ref.method]] %>% 
+    filter(N %in% sample_size_subset) %>% 
+    select(N, contains("_mean_") & 
+             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
+                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
+           ) %>% 
+    #mutate(N = format(N, scientific=FALSE)) %>% 
+    pivot_longer(cols=!N, names_to="Method", values_to="Mean")  %>% 
+    separate(
+      col="Method", into = c("Method", "Scenario"), sep = "_GSD_mean_"
+    ) -> df.temp.long[[ref.method]]
+
+  ## EMP SE ## ---------------------------------------------------------------
+  df_metrics_all_scenarios[[ref.method]] %>% 
+    filter(N %in% sample_size_subset) %>% 
+    select(N, contains("_emp_se_") & 
+             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
+                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
+           ) %>% 
+    #mutate(N = format(N, scientific=FALSE)) %>% 
+    pivot_longer(cols=!N, names_to="Method", values_to="emp_se")  %>% 
+    separate(
+      col="Method", into = c("Method", "Scenario"), sep = "_GSD_emp_se_"
+    )  -> temp.dat
+
+  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
+    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))
+
+  ## BIAS ## ---------------------------------------------------------------
+  df_metrics_all_scenarios[[ref.method]] %>% 
+    filter(N %in% sample_size_subset) %>% 
+    select(N, contains("GSD_bias_") & 
+             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
+                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
+           ) %>% 
+    #mutate(N = format(N, scientific=FALSE)) %>% 
+    pivot_longer(cols=!N, names_to="Method", values_to="bias")  %>% 
+    separate(
+      col="Method", into = c("Method", "Scenario"), sep = "_GSD_bias_"
+    )  -> temp.dat
+
+  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
+    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))
+
+  ## MSE ## ---------------------------------------------------------------
+  df_metrics_all_scenarios[[ref.method]] %>% 
+    filter(N %in% sample_size_subset) %>% 
+    select(N, contains("_mse_") & 
+             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
+                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
+           ) %>% 
+    #mutate(N = format(N, scientific=FALSE)) %>% 
+    pivot_longer(cols=!N, names_to="Method", values_to="mse")  %>% 
+    separate(
+      col="Method", into = c("Method", "Scenario"), sep = "_GSD_mse_"
+    )  -> temp.dat
+
+  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
+    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))
+
+  ## COVERAGE ## ---------------------------------------------------------------
+  df_metrics_all_scenarios[[ref.method]] %>% 
+    filter(N %in% sample_size_subset) %>% 
+    select(N, contains("GSD_coverage_calc_") & 
+             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
+                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
+           ) %>% 
+    #mutate(N = format(N, scientific=FALSE)) %>% 
+    pivot_longer(cols=!N, names_to="Method", values_to="coverage")  %>% 
+    separate(
+      col="Method", into = c("Method", "Scenario"), sep = "_GSD_coverage_calc_"
+    )  -> temp.dat
+
+  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
+    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))
+
+  ## TYPE1ERROR ## ---------------------------------------------------------------
+  df_metrics_all_scenarios[[ref.method]] %>% 
+    filter(N %in% sample_size_subset) %>% 
+    select(N, contains("GSD_type1error_calc_") & 
+             (contains("_ALL_INSTR") | contains("_STRONG_COMMON") | 
+                contains("STRONG_RARE") | contains("WEAK_COMMON") | contains("WEAK_RARE"))
+           ) %>% 
+    #mutate(N = format(N, scientific=FALSE)) %>% 
+    pivot_longer(cols=!N, names_to="Method", values_to="type1error")  %>% 
+    separate(
+      col="Method", into = c("Method", "Scenario"), sep = "_GSD_type1error_calc_"
+    )  -> temp.dat
+
+  df.temp.long[[ref.method]] <- df.temp.long[[ref.method]] %>% 
+    left_join(temp.dat, by=c("Method"="Method", "Scenario"="Scenario", "N"="N"))
+
+  #--
+  ### CREATE WIDE FORMAT -------------------------------------------------------
+  #--
+
+  ## only one variable 
+  #df.temp.long[[ref.method]] %>% 
+  #  select(N, Method, Scenario, Mean) %>% 
+  #  pivot_wider(names_from = "Scenario", values_from = "Mean", names_glue = "Mean_{Scenario}")
+
+  ## several variables
+  df.temp.long[[ref.method]] %>% 
+    select(N, Method, Scenario, Mean, emp_se) %>% 
+    pivot_wider(names_from = Scenario, values_from = c(Mean, emp_se),  
+                #names_vary = "slowest",  # naming scheme
+                names_glue = "{.value}_{Scenario}") -> df.temp.wide[[ref.method]]
+
+}
+
+# name list
+names(df.temp.long) <- paste0("ref_", selected_true_methods)
+
+### {gt} manual stub and grouping of rows  ###
+### https://gt.rstudio.com/articles/intro-creating-gt-tables.html 
+### https://themockup.blog/static/resources/gt-cookbook.html#introduction
+
+#df.temp.wide[[ref.method]] %>% 
+  #gt(groupname_col="N") %>% 
+  #gt(rowname_col = "N") %>% 
+  # tab_stubhead(label = "Sample Size") %>% 
+  # tab_row_group(
+  #   label = "Sample Size = 50,000", 
+  #   rows = 1:10
+  # ) %>% 
+  # tab_row_group(
+  #   label = "Sample Size = 100,000", 
+  #   rows = 11:20
+  # )
+
+gt.tables.mean.se <- list()
+gt.tables.performance2 <- list()
+
+## loop to create tables
+for (ref.method in seq_along(selected_true_methods)) {
+
+  ## TABLE 01: MEAN +- SE
+  df.temp.long[[ref.method]] %>% 
+    select(N, Method, Scenario, Mean, emp_se) %>% 
+    # first pivot wider
+    pivot_wider(names_from = Scenario, values_from = c(Mean, emp_se),  
+                  #names_vary = "slowest",  # naming scheme
+                  names_glue = "{.value}_{Scenario}") %>% 
+    mutate(N = paste0("Sample Size = ", format(N, scientific=FALSE))) %>% 
+    gt(groupname_col="N") %>% 
+    tab_header(
+      title = md("**Mean estimate with corresponding empirical standard errors**"),
+      subtitle = md(paste0("Reference estimate: ", selected_true_methods[ref.method]))
+    ) %>% 
+    tab_spanner(
+      label = "Mean [\u00b1 Empirical Standard Error]", 
+      columns = 3:7
+    ) %>% 
+    cols_label(
+      Mean_ALL_INSTR = md("**Scenario 1**"),
+      Mean_STRONG_COMMON = md("**Scenario 2**"),
+      Mean_STRONG_RARE = md("**Scenario 3**"),
+      Mean_WEAK_COMMON = md("**Scenario 4**"),
+      Mean_WEAK_RARE = md("**Scenario 5**"),
+    ) %>% 
+    cols_merge(
+      columns = c(Mean_ALL_INSTR, emp_se_ALL_INSTR),
+      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
+    ) %>% 
+    cols_merge(
+      columns = c(Mean_STRONG_COMMON, emp_se_STRONG_COMMON),
+      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
+    ) %>% 
+    cols_merge(
+      columns = c(Mean_STRONG_RARE, emp_se_STRONG_RARE),
+      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
+    ) %>% 
+    cols_merge(
+      columns = c(Mean_WEAK_COMMON, emp_se_WEAK_COMMON),
+      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
+    ) %>% 
+    cols_merge(
+      columns = c(Mean_WEAK_RARE, emp_se_WEAK_RARE),
+      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
+    ) %>% 
+    fmt_number(columns = 3:12,decimals = 3, use_seps=TRUE) %>% 
+    tab_source_note(
+      source_note = md("*Scenario 1* corresponds to all genetic instruments being used; 
+                       *Scenario 2* uses relatively strong and common variants;
+                       *Scenario 3* uses relatively strong and rare variants; 
+                       *Scenario 4* uses relatively weak and common variants;
+                       *Scenario 5* uses relatively weak and rare variants")
+    ) %>% 
+    opt_stylize(style=1, color="gray") -> gt.tables.mean.se[[ref.method]]
+
+  if (SAVE.files ==TRUE ){
+    filename.gt.mean.se <- paste0("../output/gt-tables/", Sys.Date(), 
+                                          "_mean_emp_se_performance_metrics_TRUE_EST_", 
+                                          selected_true_methods[ref.method], 
+                                          "_NSIM_", nsim)
+
+    gt.tables.mean.se[[ref.method]] %>% 
+      gtsave(paste0(filename.gt.mean.se, ".docx"))   # alternative .rtf, .tex, .png, .html
+    gt.tables.mean.se[[ref.method]] %>% 
+      gtsave(paste0(filename.gt.mean.se, ".html")) 
+  }
+
+  # TABLE 02: BIAS, MSE, COVERAGE, TYPE1ERROR
+  df.temp.long[[ref.method]] %>%
+    select(N, Method, Scenario, bias, mse, coverage, type1error) %>% 
+    # first pivot wider
+    pivot_wider(names_from = Scenario, values_from = c(bias, mse, coverage, type1error),  
+                  #names_vary = "slowest",  # naming scheme
+                  names_glue = "{.value}_{Scenario}") %>% 
+    mutate(N = paste0("Sample Size = ", format(N, scientific=FALSE))) %>% 
+    gt(groupname_col="N") %>% 
+    tab_header(
+      title = md("**Performance measures**"),
+      subtitle = md(paste0("Reference estimate: ", selected_true_methods[ref.method]))
+    ) %>% 
+    tab_spanner(
+      label = "Bias", 
+      columns = 3:7
+    ) %>% 
+    tab_spanner(
+      label = "MSE", 
+      columns = 8:12
+    ) %>% 
+    tab_spanner(
+      label = "Coverage", 
+      columns = 13:17
+    ) %>% 
+    tab_spanner(
+      label = "Type 1 Error", 
+      columns = 18:22
+    ) %>% 
+    cols_label(
+      bias_ALL_INSTR = md("**Scenario 1**"),
+      bias_STRONG_COMMON = md("**Scenario 2**"),
+      bias_STRONG_RARE = md("**Scenario 3**"),
+      bias_WEAK_COMMON = md("**Scenario 4**"),
+      bias_WEAK_RARE = md("**Scenario 5**"),
+    ) %>% 
+    cols_label(
+      mse_ALL_INSTR = md("**Scenario 1**"),
+      mse_STRONG_COMMON = md("**Scenario 2**"),
+      mse_STRONG_RARE = md("**Scenario 3**"),
+      mse_WEAK_COMMON = md("**Scenario 4**"),
+      mse_WEAK_RARE = md("**Scenario 5**"),
+    ) %>% 
+    cols_label(
+      coverage_ALL_INSTR = md("**Scenario 1**"),
+      coverage_STRONG_COMMON = md("**Scenario 2**"),
+      coverage_STRONG_RARE = md("**Scenario 3**"),
+      coverage_WEAK_COMMON = md("**Scenario 4**"),
+      coverage_WEAK_RARE = md("**Scenario 5**"),
+    ) %>% 
+    cols_label(
+      type1error_ALL_INSTR = md("**Scenario 1**"),
+      type1error_STRONG_COMMON = md("**Scenario 2**"),
+      type1error_STRONG_RARE = md("**Scenario 3**"),
+      type1error_WEAK_COMMON = md("**Scenario 4**"),
+      type1error_WEAK_RARE = md("**Scenario 5**"),
+    ) %>% 
+    fmt_number(columns = 3:22,decimals = 3, use_seps=TRUE) %>% 
+    tab_source_note(
+      source_note = md("*Scenario 1* corresponds to all genetic instruments being used; 
+                       *Scenario 2* uses relatively strong and common variants;
+                       *Scenario 3* uses relatively strong and rare variants; 
+                       *Scenario 4* uses relatively weak and common variants;
+                       *Scenario 5* uses relatively weak and rare variants")
+    ) %>% 
+    opt_stylize(style=1, color="gray") -> gt.tables.performance2[[ref.method]]
+
+    if (SAVE.files ==TRUE ){
+      filename.gt.performance2 <- paste0("../output/gt-tables/", Sys.Date(), 
+                                            "_performance2_metrics_TRUE_EST_", 
+                                            selected_true_methods[ref.method], 
+                                            "_NSIM_", nsim)
+
+      gt.tables.performance2[[ref.method]] %>% 
+        gtsave(paste0(filename.gt.performance2, ".docx"))   # alternative .rtf, .tex, .png, .html
+      gt.tables.performance2[[ref.method]] %>% 
+        gtsave(paste0(filename.gt.performance2, ".html")) 
+    }
+
+  ## --
+  # CREATE PLOTS -------------------------------------------------------------
+  ## --
+
+  df.temp.long[[ref.method]] %>%
+    mutate(N=paste0("Sample Size: ", format(N, scientific=FALSE))) %>% 
+    ggplot(aes(x=Method, y=mse, fill=Scenario )) +
+    geom_bar(stat="identity", position="dodge") +
+    facet_wrap(~N) + 
+    coord_flip()+
+    ggthemes::theme_few()+
+    ggthemes::scale_fill_few() -> dodged.barplot.mse
+
+}
+
+```

### 2023-03-13 scripts/MSc.Rmd
added summary plots faceted by scenario

@@
 if (SAVE.files ==TRUE ){
-  filenamefig00_b <- paste0("../output/plots/", Sys.Date(), "_explained_var_X_new_sample_size_dep.png")
+  filenamefig00_b <- paste0("../output/plots/pve/", Sys.Date(), "_explained_var_X_new_sample_size_dep.svg")
   ggsave(filename = filenamefig00_b, plot = fig00_b, 
@@
 
-if (SAVE.files ==TRUE ){
-  filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_new_PVE_X_formula.png")
-  ggsave(filename = filenamefig01_b, plot = fig01b.update, 
-         width = 10, height = 6, 
-         units = "in"  # default
-         )
-}
+# if (SAVE.files ==TRUE ){
+#   filenamefig01_b <- paste0("../output/plots/", Sys.Date(), "_figure01_new_PVE_X_formula.png")
+#   ggsave(filename = filenamefig01_b, plot = fig01b.update, 
+#          width = 10, height = 6, 
+#          units = "in"  # default
+#          )
+# }
 ```
@@
 
-# SAVE
-if (SAVE.files ==TRUE ){
-  filenamefig02 <- paste0("../output/plots/", Sys.Date(), "_figure02.png")
-  ggsave(filename = filenamefig02, plot = fig02.update)
-}
+# # SAVE
+# if (SAVE.files ==TRUE ){
+#   filenamefig02 <- paste0("../output/plots/", Sys.Date(), "_figure02.png")
+#   ggsave(filename = filenamefig02, plot = fig02.update)
+# }
 ```
@@
 if (SAVE.files ==TRUE ){
-  filename_final_plot1 <- paste0("../output/plots/", Sys.Date(), "_final_plot1.arrange_OLD.png")
-  filename_final_plot1b <- paste0("../output/plots/", Sys.Date(), "_final_plot1b.arrange_NEW.png")
-
-  ggsave(filename = filename_final_plot1, plot = final_plot1.arrange, 
-           width = 14, height = 10, 
-           units = "in"  # default
-           )
+  filename_final_plot1 <- paste0("../output/plots/pve/", Sys.Date(), "_final_plot1.arrange_OLD.svg")
+  filename_final_plot1b <- paste0("../output/plots/pve/", Sys.Date(), "_final_plot1b.arrange_NEW.svg")
+
+  # ggsave(filename = filename_final_plot1, plot = final_plot1.arrange, 
+  #          width = 14, height = 10, 
+  #          units = "in"  # default
+  #          )
   ggsave(filename = filename_final_plot1b, plot = final_plot1.arrange2, 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.scatter.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrscatter_gsd.png")
+  filename.mr.scatter.gsd <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrscatter_gsd.svg")
   ggsave(filename = filename.mr.scatter.gsd, plot = p1.gsd, 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.forest.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_GSD.png")
+  filename.mr.forest.gsd <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrforest_GSD.svg")
   ggsave(filename = filename.mr.forest.gsd, plot = p2.gsd, 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.funnel.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrfunnel_gsd.png")
+  filename.mr.funnel.gsd <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrfunnel_gsd.svg")
   ggsave(filename = filename.mr.funnel.gsd, plot = p4.gsd, 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.scatter <- paste0("../output/plots/", Sys.Date(), "_figure_mrscatter.png")
+  filename.mr.scatter <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrscatter.svg")
   ggsave(filename = filename.mr.scatter, plot = p1[[1]], 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.forest <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_GBC.png")
+  filename.mr.forest <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrforest_GBC.svg")
   ggsave(filename = filename.mr.forest, plot = p2[[1]], 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.funnel.gbc <- paste0("../output/plots/", Sys.Date(), "_figure_mrfunnel_gbc.png")
+  filename.mr.funnel.gbc <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrfunnel_gbc.svg")
   ggsave(filename = filename.mr.funnel.gbc, plot = p4, 
@@
 
-if (SAVE.files ==TRUE ){
-  filename_pveplot_sim <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_pve.png")
-  ggsave(filename = filename_pveplot_sim, plot = plot_pve_simulated, 
-         width = 10, height = 6, 
-         units = "in"  # default
-         )
-}
+# if (SAVE.files ==TRUE ){
+#   filename_pveplot_sim <- paste0("../output/plots/pve/", Sys.Date(), "_figure_simplot_pve.png")
+#   ggsave(filename = filename_pveplot_sim, plot = plot_pve_simulated, 
+#          width = 10, height = 6, 
+#          units = "in"  # default
+#          )
+# }
 ```
@@
 if (SAVE.files ==TRUE ){
-  filename_pveplot_sim.raincloud <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_pve_raincloud.png")
+  filename_pveplot_sim.raincloud <- paste0("../output/plots/pve/", Sys.Date(), "_figure_simplot_pve_raincloud.svg")
   ggsave(filename = filename_pveplot_sim.raincloud, plot = plot_pve_simulated.raincloud, 
@@
 # type1error
-type1error <- function(pvals) sum(pvals<=0.05)/length(pvals)  # no causal_effect
+type1error <- function(pvals) sum(pvals<=0.05)/length(pvals)  # no causal_effect detected
 
@@
 
-```{r store-performance-metrics-csv}
-## GSD  we loop in the next section and save the files there with all references!
-# filename_IVW_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_IVW_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# filename_MODE_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_MODE_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# filename_CONMIX_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_CONMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# filename_MRMIX_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_MRMIX_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# #5
-# filename_MEDIAN_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_MEDIAN_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# #6
-# filename_EGGER_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_EGGER_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# #7
-# filename_ROBUST_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_ROBUST_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# #8
-# filename_PRESSO_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_PRESSO_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# #9
-# filename_RAPS_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_RAPS_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# #10
-# filename_LASSO_GSD_performance_metrics <- paste0("../output/performance_metrics/", 
-#                                                Sys.Date(), "_LASSO_GSD_performance_metrics_NSIM_", nsim, ".csv")
-# write.csv2(x = IVW_GSD_performance_metrics, file = filename_IVW_GSD_performance_metrics)
-# write.csv2(x = MODE_GSD_performance_metrics, file = filename_MODE_GSD_performance_metrics)
-# write.csv2(x = CONMIX_GSD_performance_metrics, file = filename_CONMIX_GSD_performance_metrics)
-# write.csv2(x = MRMIX_GSD_performance_metrics, file = filename_MRMIX_GSD_performance_metrics)
-# #5
-# write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename_MEDIAN_GSD_performance_metrics)
-# #6
-# write.csv2(x = EGGER_GSD_performance_metrics, file = filename_EGGER_GSD_performance_metrics)
-# #7
-# write.csv2(x = ROBUST_GSD_performance_metrics, file = filename_ROBUST_GSD_performance_metrics)
-# #8
-# write.csv2(x = PRESSO_GSD_performance_metrics, file = filename_PRESSO_GSD_performance_metrics)
-# #9
-# write.csv2(x = RAPS_GSD_performance_metrics, file = filename_RAPS_GSD_performance_metrics)
-# #10
-# write.csv2(x = LASSO_GSD_performance_metrics, file = filename_LASSO_GSD_performance_metrics)
-```
-
 ```{r}
-#s1 <- simsum(data = est_sim, estvarname = "IVW_theta_sim_BMI_GSD_40000", se = "IVW_theta_se_sim_BMI_GSD_40000", true = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "IVW"] )
+library(rsimsum)
+s1 <- simsum(data = est_sim, estvarname = "IVW_theta_sim_BMI_GSD_150000", se = "IVW_theta_se_sim_BMI_GSD_150000", true = est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "IVW"] )
 
@@
 # • power, power of a (1 - level)\
-
-#statsparam <- c("nsim", "thetamean", "se2mean", "bias", "empse", "mse", "cover", "becover", "power")
-#simsum_summary <- summary(s1, stats = statsparam)
+statsparam <- c("nsim", "thetamean", "se2mean", "bias", "empse", "mse", "cover", "becover", "power")
+simsum_summary <- summary(s1, stats = statsparam)
 ```
@@
 
-if (SAVE.files ==TRUE ){
-  filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_NSIM_", nsim,".png")
-  ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
-         width = 10, height = 22, 
-         units = "in"  # default
-         )
-}
+# if (SAVE.files ==TRUE ){
+#   filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_NSIM_", nsim,".png")
+#   ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
+#          width = 10, height = 22, 
+#          units = "in"  # default
+#          )
+# }
 ```
@@
 
-if (SAVE.files ==TRUE ){
-  filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_NSIM_", nsim, ".png")
-  ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
-         width = 7, height = 12, 
-         units = "in"  # default
-         )
-}
+# if (SAVE.files ==TRUE ){
+#   filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_NSIM_", nsim, ".png")
+#   ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
+#          width = 7, height = 12, 
+#          units = "in"  # default
+#          )
+# }
 ```
@@
 
-  # STORE
-  if (SAVE.files ==TRUE ){
-    filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_IVW_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = IVW_GSD_performance_metrics, file = filename.IVW_table.performance)
-    filename.MODE_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_MODE_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = MODE_GSD_performance_metrics, file = filename.MODE_table.performance)
-    filename.CONMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_CONMIX_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = CONMIX_GSD_performance_metrics, file = filename.CONMIX_table.performance)
-    filename.MRMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_MRMIX_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
-    #5
-    filename.MEDIAN_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_MEDIAN_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename.MEDIAN_table.performance)
-    #6
-    filename.EGGER_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_EGGER_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = EGGER_GSD_performance_metrics, file = filename.EGGER_table.performance)
-    #7
-    filename.ROBUST_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_ROBUST_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = ROBUST_GSD_performance_metrics, file = filename.ROBUST_table.performance)
-    #8
-    filename.PRESSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_PRESSO_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = PRESSO_GSD_performance_metrics, file = filename.PRESSO_table.performance)
-    #9
-    filename.RAPS_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_RAPS_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = RAPS_GSD_performance_metrics, file = filename.RAPS_table.performance)
-    #10
-    filename.LASSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                        "_LASSO_GSD_performance_metrics_TRUE_EST_", 
-                                        selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-    write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
-  }
+  # # STORE
+  # if (SAVE.files ==TRUE ){
+  #   filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_IVW_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = IVW_GSD_performance_metrics, file = filename.IVW_table.performance)
+  #   filename.MODE_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_MODE_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = MODE_GSD_performance_metrics, file = filename.MODE_table.performance)
+  #   filename.CONMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_CONMIX_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = CONMIX_GSD_performance_metrics, file = filename.CONMIX_table.performance)
+  #   filename.MRMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_MRMIX_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
+  #   #5
+  #   filename.MEDIAN_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_MEDIAN_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename.MEDIAN_table.performance)
+  #   #6
+  #   filename.EGGER_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_EGGER_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = EGGER_GSD_performance_metrics, file = filename.EGGER_table.performance)
+  #   #7
+  #   filename.ROBUST_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_ROBUST_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = ROBUST_GSD_performance_metrics, file = filename.ROBUST_table.performance)
+  #   #8
+  #   filename.PRESSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_PRESSO_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = PRESSO_GSD_performance_metrics, file = filename.PRESSO_table.performance)
+  #   #9
+  #   filename.RAPS_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_RAPS_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = RAPS_GSD_performance_metrics, file = filename.RAPS_table.performance)
+  #   #10
+  #   filename.LASSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+  #                                       "_LASSO_GSD_performance_metrics_TRUE_EST_", 
+  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
+  #   write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
+  # }
 
@@
 
-  ####-------------
-  # SAVE  --------------------------------
-  if (SAVE.files ==TRUE ){
-    #1
-    filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #2
-    filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #3
-    filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #4
-    filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MRMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #5
-    filename_MEDIAN_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MEDIAN_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_MEDIAN_GSD_simplot_thetas, plot = MEDIAN_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #6
-    filename_EGGER_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_EGGER_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_EGGER_GSD_simplot_thetas, plot = EGGER_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #7
-    filename_ROBUST_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_ROBUST_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_ROBUST_GSD_simplot_thetas, plot = ROBUST_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #8
-    filename_PRESSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_PRESSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_PRESSO_GSD_simplot_thetas, plot = PRESSO_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #9
-    filename_RAPS_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_RAPS_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_RAPS_GSD_simplot_thetas, plot = RAPS_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-    #10
-    filename_LASSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_LASSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-    ggsave(filename = filename_LASSO_GSD_simplot_thetas, plot = LASSO_GSD_simplot_thetas, 
-           width = 15, height = 6, 
-           units = "in"  # default
-           )
-  }
+  # ####-------------
+  # # SAVE  --------------------------------
+  # if (SAVE.files ==TRUE ){
+  #   #1
+  #   filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #2
+  #   filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #3
+  #   filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #4
+  #   filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MRMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #5
+  #   filename_MEDIAN_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MEDIAN_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_MEDIAN_GSD_simplot_thetas, plot = MEDIAN_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #6
+  #   filename_EGGER_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_EGGER_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_EGGER_GSD_simplot_thetas, plot = EGGER_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #7
+  #   filename_ROBUST_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_ROBUST_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_ROBUST_GSD_simplot_thetas, plot = ROBUST_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #8
+  #   filename_PRESSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_PRESSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_PRESSO_GSD_simplot_thetas, plot = PRESSO_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #9
+  #   filename_RAPS_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_RAPS_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_RAPS_GSD_simplot_thetas, plot = RAPS_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  #   #10
+  #   filename_LASSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_LASSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+  #   ggsave(filename = filename_LASSO_GSD_simplot_thetas, plot = LASSO_GSD_simplot_thetas, 
+  #          width = 15, height = 6, 
+  #          units = "in"  # default
+  #          )
+  # }
 
@@
   if (SAVE.files ==TRUE ){
-    filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
+    filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/theta_sims/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".svg")
     ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
@@
   if (SAVE.files ==TRUE ){
-    filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], "_NSIM_",nsim, ".png")
+    filename_simplot_bias_mse <- paste0("../output/plots/bias-mse/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], "_NSIM_",nsim, ".svg")
     ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
@@
 
-    # STORE
-    if (SAVE.files ==TRUE ){
-      filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_IVW_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = IVW_GSD_performance_metrics, file = filename.IVW_table.performance)
-      filename.MODE_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_MODE_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = MODE_GSD_performance_metrics, file = filename.MODE_table.performance)
-      filename.CONMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_CONMIX_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = CONMIX_GSD_performance_metrics, file = filename.CONMIX_table.performance)
-      filename.MRMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_MRMIX_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
-      #5
-      filename.MEDIAN_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_MEDIAN_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename.MEDIAN_table.performance)
-      #6
-      filename.EGGER_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_EGGER_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = EGGER_GSD_performance_metrics, file = filename.EGGER_table.performance)
-      #7
-      filename.ROBUST_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_ROBUST_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = ROBUST_GSD_performance_metrics, file = filename.ROBUST_table.performance)
-      #8
-      filename.PRESSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_PRESSO_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = PRESSO_GSD_performance_metrics, file = filename.PRESSO_table.performance)
-      #9
-      filename.RAPS_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_RAPS_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = RAPS_GSD_performance_metrics, file = filename.RAPS_table.performance)
-      #10
-      filename.LASSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-                                          "_LASSO_GSD_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch,
-                                          "_NSIM_", nsim, ".csv")
-      write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
-    }
+    # # STORE
+    # if (SAVE.files ==TRUE ){
+    #   filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_IVW_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = IVW_GSD_performance_metrics, file = filename.IVW_table.performance)
+    #   filename.MODE_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_MODE_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = MODE_GSD_performance_metrics, file = filename.MODE_table.performance)
+    #   filename.CONMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_CONMIX_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = CONMIX_GSD_performance_metrics, file = filename.CONMIX_table.performance)
+    #   filename.MRMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_MRMIX_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
+    #   #5
+    #   filename.MEDIAN_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_MEDIAN_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename.MEDIAN_table.performance)
+    #   #6
+    #   filename.EGGER_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_EGGER_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = EGGER_GSD_performance_metrics, file = filename.EGGER_table.performance)
+    #   #7
+    #   filename.ROBUST_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_ROBUST_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = ROBUST_GSD_performance_metrics, file = filename.ROBUST_table.performance)
+    #   #8
+    #   filename.PRESSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_PRESSO_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = PRESSO_GSD_performance_metrics, file = filename.PRESSO_table.performance)
+    #   #9
+    #   filename.RAPS_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_RAPS_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = RAPS_GSD_performance_metrics, file = filename.RAPS_table.performance)
+    #   #10
+    #   filename.LASSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
+    #                                       "_LASSO_GSD_performance_metrics_TRUE_EST_", 
+    #                                       selected_true_methods[true_method], 
+    #                                       "_GenArch_", name_GenArch,
+    #                                       "_NSIM_", nsim, ".csv")
+    #   write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
+    # }
 
@@
     if (SAVE.files ==TRUE ){
-      filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
-                                                    "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+      filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/theta-sims/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
+                                                    "_GenArch_", name_GenArch, "_NSIM_", nsim,".svg")
       ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
@@
     if (SAVE.files ==TRUE ){
-      filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], 
-                                          "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
+      filename_simplot_bias_mse <- paste0("../output/plots/bias-mse/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], 
+                                          "_GenArch_", name_GenArch, "_NSIM_", nsim,".svg")
       ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
@@
 
+}
+
+```
+
+## Summary Plots
+
+```{r}
+dodged.barplot.mse <- list()
+dodged.barplot.bias <- list()
+dodged.barplot.coverage <- list()
+dodged.barplot.power<- list()
+
+for(ref.method in seq_along(df_metrics_all_scenarios)) {
   ## --
@@
 
+  ## MSE PLOT FACET WRAP ------------------------------------------------------
+
   df.temp.long[[ref.method]] %>%
     mutate(N=paste0("Sample Size: ", format(N, scientific=FALSE))) %>% 
+    mutate(label.too.large = ifelse(mse>0.5, "MSE \u2265 0.5", NA)) %>% 
     ggplot(aes(x=Method, y=mse, fill=Scenario )) +
-    geom_bar(stat="identity", position="dodge") +
+    geom_bar(stat="identity", position=position_dodge2(reverse=TRUE)) +   # reverse dodge
+    geom_text(aes(x= Method, y= 0.4, label = label.too.large),# position = position_dodge(),
+              #size = 2.25, angle = 90
+              )+
     facet_wrap(~N) + 
+    xlab("Method")+
+    ylab(paste0("Mean Squared Error with nsim=", nsim))+
+    labs(title = paste0("Selected Reference Method: ", 
+                            selected_true_methods[ref.method])
+         )+
+    #ylim(0, 0.5)+
+    coord_flip(ylim=c(0,0.5))+
+    ggthemes::theme_few()+
+    ggthemes::scale_fill_few() -> dodged.barplot.mse[[ref.method]]
+
+  if (SAVE.files ==TRUE ){
+    ## arranged plots are toooo small 
+    filename.dodged.barplot.mse <- paste0("../output/plots/summary-plots/", Sys.Date(), "_GSD_dodged_barplot_mse_REF_METHOD_",
+                                          selected_true_methods[ref.method], "_NSIM_", nsim)
+    ggsave(filename = paste0(filename.dodged.barplot.mse, ".svg"), plot =  dodged.barplot.mse[[ref.method]], 
+             width = 12, height = 8, 
+             units = "in"  # default
+             )
+  }
+
+  ## BIAS PLOT FACET WRAP ------------------------------------------------------
+
+  df.temp.long[[ref.method]] %>%
+    mutate(N=paste0("Sample Size: ", format(N, scientific=FALSE))) %>% 
+    ggplot(aes(x=Method, y=bias, fill=Scenario )) +
+    geom_bar(stat="identity", position=position_dodge2(reverse=TRUE)) +   # reverse dodge
+    geom_hline(yintercept = 0,  col=ggthemes_data$few$colors$Dark[3,2][[1]], alpha=0.6)+
+    facet_wrap(~N) + 
+    xlab("Method")+
+    ylab(paste0("Bias with nsim=", nsim))+
+    labs(title = paste0("Selected Reference Method: ", 
+                            selected_true_methods[ref.method])
+         )+
     coord_flip()+
     ggthemes::theme_few()+
-    ggthemes::scale_fill_few() -> dodged.barplot.mse
-
+    ggthemes::scale_fill_few() -> dodged.barplot.bias[[ref.method]]
+
+  if (SAVE.files ==TRUE ){
+    ## arranged plots are toooo small 
+    filename.dodged.barplot.bias <- paste0("../output/plots/summary-plots/", Sys.Date(), "_GSD_dodged_barplot_bias_REF_METHOD_",
+                                          selected_true_methods[ref.method], "_NSIM_", nsim)
+    ggsave(filename = paste0(filename.dodged.barplot.bias, ".svg"), plot =  dodged.barplot.bias[[ref.method]], 
+             width = 12, height = 8, 
+             units = "in"  # default
+             )
   }
 
+  ## COVERAGE PLOT FACET WRAP ------------------------------------------------------
+  df.temp.long[[ref.method]] %>%
+    mutate(N=paste0("Sample Size: ", format(N, scientific=FALSE))) %>% 
+    ggplot(aes(x=Method, y=coverage, fill=Scenario )) +
+    geom_bar(stat="identity", position=position_dodge2(reverse=TRUE)) +   # reverse dodge
+    geom_hline(yintercept = 0.95,  col=ggthemes_data$few$colors$Dark[3,2][[1]], alpha=0.6)+
+    facet_wrap(~N) + 
+    xlab("Method")+
+    ylab(paste0("Coverage with nsim=", nsim))+
+    labs(title = paste0("Selected Reference Method: ", 
+                            selected_true_methods[ref.method])
+         )+
+    coord_flip()+
+    ggthemes::theme_few()+
+    ggthemes::scale_fill_few() -> dodged.barplot.coverage[[ref.method]]
+
+  if (SAVE.files ==TRUE ){
+    ## arranged plots are toooo small 
+    filename.dodged.barplot.coverage <- paste0("../output/plots/summary-plots/", Sys.Date(), "_GSD_dodged_barplot_coverage_REF_METHOD_",
+                                          selected_true_methods[ref.method], "_NSIM_", nsim)
+    ggsave(filename = paste0(filename.dodged.barplot.coverage, ".svg"), plot =  dodged.barplot.coverage[[ref.method]], 
+             width = 12, height = 8, 
+             units = "in"  # default
+             )
+  }
+
+  ## Power of 5% level test (not TYPE1ERROR) PLOT FACET WRAP ------------------------------------------------------
+  df.temp.long[[ref.method]] %>%
+    mutate(N=paste0("Sample Size: ", format(N, scientific=FALSE))) %>% 
+    ggplot(aes(x=Method, y=type1error, fill=Scenario )) +
+    geom_bar(stat="identity", position=position_dodge2(reverse=TRUE)) +   # reverse dodge
+    geom_hline(yintercept = 0.95,  col=ggthemes_data$few$colors$Dark[3,2][[1]], alpha=0.6)+
+    facet_wrap(~N) + 
+    xlab("Method")+
+    ylab(paste0("Power of 5% Level Test with nsim=", nsim))+
+    labs(title = paste0("Selected Reference Method: ", 
+                            selected_true_methods[ref.method])
+         )+
+    coord_flip()+
+    ggthemes::theme_few()+
+    ggthemes::scale_fill_few() -> dodged.barplot.power[[ref.method]]
+
+  if (SAVE.files ==TRUE ){
+    ## arranged plots are toooo small 
+    filename.dodged.barplot.power <- paste0("../output/plots/summary-plots/", Sys.Date(), "_GSD_dodged_barplot_power_REF_METHOD_",
+                                          selected_true_methods[ref.method], "_NSIM_", nsim)
+    ggsave(filename = paste0(filename.dodged.barplot.power, ".svg"), plot =  dodged.barplot.power[[ref.method]], 
+             width = 12, height = 8, 
+             units = "in"  # default
+             )
+  }
+}
 ```

### 2023-03-14 scripts/helper/simulation_helpers.R
added argument to run MR-PRESSO only if needed

@@ -16,7 +16,10 @@ source("scripts/helper/PVE_calculation_fns.R")
 conflict_prefer("select", "dplyr")
 conflict_prefer("filter", "dplyr")
 
-sim_function <- function(dat, index) {
+sim_function <- function(dat, index, runMRpresso) {
+  ## if runMRpresso == FALSE, only NAs will returned
+  
+  
   if(any(is.na(dat))) return(NULL)
   if(nrow(dat)<=3) return(NULL)  # cannot perform robust methods with one IV, "mr_presso method requires data on >3 variants."
   sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
@@ -126,19 +129,25 @@ sim_function <- function(dat, index) {
   return_vector <- append(return_vector, robust.res.gsd$Pvalue)
   
   # 8. MR-PRESSO
-  presso.df = data.frame(bx = dat$beta.bmi_sim_norm, by = dat$beta.gsd, bxse = dat$se_update_BMI, byse = dat$se.gsd)
-  presso.res.gsd = MRPRESSO::mr_presso(BetaOutcome = "by", BetaExposure = "bx", SdOutcome = "byse", SdExposure = "bxse", 
-                  OUTLIERtest = TRUE, DISTORTIONtest = TRUE, data = presso.df, NbDistribution = 500, SignifThreshold = 0.05)
-  if (!is.na(presso.res.gsd$`Main MR results`[2,"Causal Estimate"]) & !is.na(presso.res.gsd$`Main MR results`[2,"Sd"])){
-    return_vector <-  append(return_vector, presso.res.gsd$`Main MR results`[2,"Causal Estimate"])
-    return_vector <- append(return_vector, presso.res.gsd$`Main MR results`[2,"Sd"])
-    return_vector <- append(return_vector, presso.res.gsd$`Main MR results`[2,"P-value"])
-  } else{  # if no outlier-corrected analysis was calculated
-    return_vector <-  append(return_vector, presso.res.gsd$`Main MR results`[1,"Causal Estimate"])
-    return_vector <- append(return_vector, presso.res.gsd$`Main MR results`[1,"Sd"])
-    return_vector <- append(return_vector, presso.res.gsd$`Main MR results`[1,"P-value"])
[…]

### 2023-03-14 scripts/purrr-simstudy.R

@@ -38,7 +38,8 @@ conflict_prefer("filter", "dplyr")
 
 # set to TRUE if simulation should run when script is sourced
 runSimulation <- TRUE
-nsim <- 10
+runMRpresso <- FALSE  # runs foreeeever ;)
+nsim <- 2
 
 runGeneticArchitectures <- TRUE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 # analysis set - create all combis
@@ -85,7 +86,7 @@ if (runSimulation == TRUE) {
     
     ## now purrr
     # returns nsim* results for MR simulations PER sample size
-    nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs[[dataset]], index = dataset), 
+    nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs[[dataset]], index = dataset, runMRpresso = runMRpresso),      # RUN MR PRESSO???
                              .progress=list(clear=FALSE, 
                                             name=paste0("Progress of simulation of sample size: ", format(n[dataset], scientific = FALSE)))
     )

### 2023-03-14 scripts/purrr-simstudy.R
added subsetSampleSizes as argument

@@ -38,6 +38,7 @@ conflict_prefer("filter", "dplyr")
 
 # set to TRUE if simulation should run when script is sourced
 runSimulation <- TRUE
+subsetSampleSizes <- TRUE
 runMRpresso <- FALSE  # runs foreeeever ;)
 nsim <- 2
 
@@ -52,7 +53,14 @@ all_quadrants <- tidyr::crossing(effect_size, maf)
 my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
 # vars from MSc script
 sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
-n <- sim_vars$n
+
+# only simulate defined set of smaller sample sizes?
+if (subsetSampleSizes == TRUE) {
+  n <- c(50000, 100000, 150000, 200000, 300000)
+} else{
+  n <- sim_vars$n
+}
+
 colnames_pval <- sim_vars$colnames_pval
 n.orig <- 334487
 

### 2023-03-14 scripts/purrr-simstudy.R
fine tuned sim script to run on smaller sample size set ;)

@@ -55,14 +55,16 @@ my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern
 sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
 
 # only simulate defined set of smaller sample sizes?
-if (subsetSampleSizes == TRUE) {
-  n <- c(50000, 100000, 150000, 200000, 300000)
-} else{
-  n <- sim_vars$n
-}
-
+n <- sim_vars$n
 colnames_pval <- sim_vars$colnames_pval
 n.orig <- 334487
+
+## subset?
+if (subsetSampleSizes == TRUE) {
+  n <- as.integer(n[c(8, 13, 18, 23, 33)])
+  colnames_pval <- colnames_pval[c(8, 13, 18, 23, 33)]
+}
+
 
 ## PARAMS for simulation
 set.seed(1234)

### 2023-03-14 scripts/purrr-simstudy.R
nsim 100

@@ -40,9 +40,9 @@ conflict_prefer("filter", "dplyr")
 runSimulation <- TRUE
 subsetSampleSizes <- TRUE
 runMRpresso <- FALSE  # runs foreeeever ;)
-nsim <- 2
-
+nsim <- 100
 runGeneticArchitectures <- TRUE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
+
 # analysis set - create all combis
 effect_size <- c("strong", "weak")
 maf <- c("rare", "common")
@@ -60,9 +60,10 @@ colnames_pval <- sim_vars$colnames_pval
 n.orig <- 334487
 
 ## subset?
+subset.index <- c(8, 13, 18, 23, 33)
 if (subsetSampleSizes == TRUE) {
-  n <- as.integer(n[c(8, 13, 18, 23, 33)])
-  colnames_pval <- colnames_pval[c(8, 13, 18, 23, 33)]
+  n <- as.integer(n[subset.index])
+  colnames_pval <- colnames_pval[subset.index]
 }
 
 
@@ -146,7 +147,7 @@ if (runGeneticArchitectures == TRUE) {
       
       ## now purrr
       # returns nsim* results for MR simulations PER sample size
-      nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs_filter[[dataset]], index = dataset), 
+      nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs_filter[[dataset]], index = dataset, runMRpresso = runMRpresso),      # RUN MR PRESSO???
                                .progress=list(clear=FALSE, 
                                               name=paste0("Progress of simulation of sample size: ", format(n[dataset], scientific = FALSE)))
       )
@@ -165,6 +166,4 @@ timediff_purrr_archi <- T1_architecture - T0_architecture
 print(paste0("!!!!!--------------------------- ", round(timediff_purrr_archi/60,2), " minutes passed for the simulation scenario! --------------------------------------!!!!!"))
 
 saveRDS(est_sim_architectures, paste0("./output/Rdata/", Sys.Date(), "_est_sim_architectures_purrr_NSIM_", nsim, ".rds"))
-
-
 

### 2023-03-16 scripts/MSc.Rmd
small changes, dir structure

@@
 if (SAVE.files ==TRUE ){
-  saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
+  saveRDS(my_data_harm, paste0("../output/Rdata/harmonized-dat/", Sys.Date(), "_my_data_harm.rds"))
 }
@@
 ## load results from script `perform_mr.R`
-est.bmi.gsd <- readRDS(dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
+est.bmi.gsd <- readRDS(dplyr::last(list.files("../output/Rdata/mr-results/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
 
@@
 # from "purrr-simstudy.R"
-est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_purrr_NSIM_", full.names = TRUE)))
-est_sim_architectures <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_architectures_purrr_NSIM_", full.names = TRUE)))
+est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata/sim-results/", pattern = "_est_sim_purrr_NSIM_", full.names = TRUE)))
+est_sim_architectures <- readRDS(dplyr::last(list.files(path = "../output/Rdata/sim-results/", pattern = "_est_sim_architectures_purrr_NSIM_", full.names = TRUE)))
 nsim <- nrow(est_sim)

### 2023-03-16 scripts/perform_mr.R

@@ -27,7 +27,7 @@ library(penalized)
 source("./scripts/helper/MR_lasso.R")
 
 ## load data created in MSc.rmd
-my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
+my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/harmonized-dat/", pattern = "my_data_harm.rds", full.names = TRUE)))
 
 # perform MR analysis -----------------------------------------------------
 set.seed(123)
@@ -171,7 +171,7 @@ if (numIV>2){
 }
 
 ## SAVE estimates
-saveRDS(est, paste0("./output/Rdata/", Sys.Date(), "_est.rds"))
+saveRDS(est, paste0("./output/Rdata/mr-results/", Sys.Date(), "_est.rds"))
 
 
 # BMI -> GSD --------------------------------------------------------------
@@ -315,4 +315,4 @@ if (numIV>2){
 }
 
 ## SAVE estimates
-saveRDS(est.bmi.gsd, paste0("./output/Rdata/", Sys.Date(), "_est_BMI_GSD.rds"))
+saveRDS(est.bmi.gsd, paste0("./output/Rdata/mr-results/", Sys.Date(), "_est_BMI_GSD.rds"))

### 2023-03-16 scripts/purrr-simstudy.R

@@ -50,7 +50,7 @@ all_quadrants <- tidyr::crossing(effect_size, maf)
 
 # GET DATA ----------------------------------------------------------------
 # data
-my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "my_data_harm.rds", full.names = TRUE)))
+my_data_harm <- readRDS(file = dplyr::last(list.files("./output/Rdata/harmonized-dat/", pattern = "my_data_harm.rds", full.names = TRUE)))
 # vars from MSc script
 sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
 
@@ -113,7 +113,7 @@ T1 <- proc.time()[3]
 timediff_purrr <- T1 - T0
 print(paste0("!!!!!--------------------------- ", round(timediff_purrr/60,2), " minutes passed for the simulation scenario! --------------------------------------!!!!!"))
 
-saveRDS(est_sim, paste0("./output/Rdata/", Sys.Date(), "_est_sim_purrr_NSIM_", nsim, ".rds"))
+saveRDS(est_sim, paste0("./output/Rdata/sim-results/", Sys.Date(), "_est_sim_purrr_NSIM_", nsim, ".rds"))
 
 ## with nsim=100 and the purrr script on the M1 Mac ∞ 5h (only 4 methods at that time)
 #[1] "!!!!!--------------------------- 299.61 minutes passed for the simulation scenario! --------------------------------------!!!!!"
@@ -165,5 +165,5 @@ T1_architecture <- proc.time()[3]
 timediff_purrr_archi <- T1_architecture - T0_architecture
 print(paste0("!!!!!--------------------------- ", round(timediff_purrr_archi/60,2), " minutes passed for the simulation scenario! --------------------------------------!!!!!"))
 
-saveRDS(est_sim_architectures, paste0("./output/Rdata/", Sys.Date(), "_est_sim_architectures_purrr_NSIM_", nsim, ".rds"))
+saveRDS(est_sim_architectures, paste0("./output/Rdata/sim-results/", Sys.Date(), "_est_sim_architectures_purrr_NSIM_", nsim, ".rds"))
 

### 2023-03-16 scripts/purrr-simstudy.R
run sim n500 but no genArchs

@@ -40,8 +40,8 @@ conflict_prefer("filter", "dplyr")
 runSimulation <- TRUE
 subsetSampleSizes <- TRUE
 runMRpresso <- FALSE  # runs foreeeever ;)
-nsim <- 100
-runGeneticArchitectures <- TRUE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
+nsim <- 500
+runGeneticArchitectures <- FALSE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 
 # analysis set - create all combis
 effect_size <- c("strong", "weak")

### 2023-03-16 scripts/helper/simulation_helpers.R
fixed a sample size index bug

@@ -16,14 +16,23 @@ source("scripts/helper/PVE_calculation_fns.R")
 conflict_prefer("select", "dplyr")
 conflict_prefer("filter", "dplyr")
 
-sim_function <- function(dat, index, runMRpresso) {
+sim_function <- function(dat, index, runMRpresso, subsetSampleSizes) {
+  ## index used to address specific samplesize in PVE calculation
   ## if runMRpresso == FALSE, only NAs will returned
+  ## subsetSampleSizes ... use smaller set of sample sizes
   
   
   if(any(is.na(dat))) return(NULL)
   if(nrow(dat)<=3) return(NULL)  # cannot perform robust methods with one IV, "mr_presso method requires data on >3 variants."
   sim_vars <- readRDS(file = dplyr::last(list.files("./output/Rdata/", pattern = "_sim_vars.rds", full.names = TRUE)))
   n <- sim_vars$n
+  
+  ## subset?
+  subset.index <- c(8, 13, 18, 23, 33)
+  if (subsetSampleSizes == TRUE) {
+    n <- as.integer(n[subset.index])
+    colnames_pval <- colnames_pval[subset.index]
+  }
   
   return_vector <- c()
   dat <- dat %>% 

### 2023-03-16 scripts/purrr-simstudy.R

@@ -40,7 +40,7 @@ conflict_prefer("filter", "dplyr")
 runSimulation <- TRUE
 subsetSampleSizes <- TRUE
 runMRpresso <- FALSE  # runs foreeeever ;)
-nsim <- 500
+nsim <- 3
 runGeneticArchitectures <- FALSE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 
 # analysis set - create all combis
@@ -97,7 +97,7 @@ if (runSimulation == TRUE) {
     
     ## now purrr
     # returns nsim* results for MR simulations PER sample size
-    nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs[[dataset]], index = dataset, runMRpresso = runMRpresso),      # RUN MR PRESSO???
+    nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs[[dataset]], index = dataset, runMRpresso = runMRpresso, subsetSampleSizes=subsetSampleSizes),      # RUN MR PRESSO???
                              .progress=list(clear=FALSE, 
                                             name=paste0("Progress of simulation of sample size: ", format(n[dataset], scientific = FALSE)))
     )
@@ -147,7 +147,7 @@ if (runGeneticArchitectures == TRUE) {
       
       ## now purrr
       # returns nsim* results for MR simulations PER sample size
-      nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs_filter[[dataset]], index = dataset, runMRpresso = runMRpresso),      # RUN MR PRESSO???
+      nsim_list_dataset <- map(1:nsim, ~sim_function(dat = gen_dfs_filter[[dataset]], index = dataset, runMRpresso = runMRpresso, subsetSampleSizes=subsetSampleSizes),      # RUN MR PRESSO???
                                .progress=list(clear=FALSE, 
                                               name=paste0("Progress of simulation of sample size: ", format(n[dataset], scientific = FALSE)))
       )

### 2023-03-16 scripts/purrr-simstudy.R
runSim-n100

@@ -40,7 +40,7 @@ conflict_prefer("filter", "dplyr")
 runSimulation <- TRUE
 subsetSampleSizes <- TRUE
 runMRpresso <- FALSE  # runs foreeeever ;)
-nsim <- 3
+nsim <- 100
 runGeneticArchitectures <- FALSE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 
 # analysis set - create all combis

### 2023-03-16 scripts/purrr-simstudy.R
runsim

@@ -41,7 +41,7 @@ runSimulation <- TRUE
 subsetSampleSizes <- TRUE
 runMRpresso <- FALSE  # runs foreeeever ;)
 nsim <- 100
-runGeneticArchitectures <- FALSE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
+runGeneticArchitectures <- TRUE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 
 # analysis set - create all combis
 effect_size <- c("strong", "weak")

### 2023-03-16 scripts/helper/simulation_helpers.R
bug with colnames_pval object fixed

@@ -31,7 +31,7 @@ sim_function <- function(dat, index, runMRpresso, subsetSampleSizes) {
   subset.index <- c(8, 13, 18, 23, 33)
   if (subsetSampleSizes == TRUE) {
     n <- as.integer(n[subset.index])
-    colnames_pval <- colnames_pval[subset.index]
+    #colnames_pval <- colnames_pval[subset.index]  # not imported colnames_pval into function !
   }
   
   return_vector <- c()

### 2023-04-04 scripts/MSc.Rmd
small dir changes

@@
 ## load results
-est <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
+est <- readRDS(file = dplyr::last(list.files("../output/Rdata/mr-results/", pattern = "_est.rds", full.names = TRUE)))
 

### 2023-04-04 scripts/purrr-simstudy.R
small changes

@@ -40,7 +40,7 @@ conflict_prefer("filter", "dplyr")
 runSimulation <- TRUE
 subsetSampleSizes <- TRUE
 runMRpresso <- FALSE  # runs foreeeever ;)
-nsim <- 500
+nsim <- 1000
 runGeneticArchitectures <- FALSE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 
 # analysis set - create all combis
@@ -161,9 +161,11 @@ if (runGeneticArchitectures == TRUE) {
   }
 }
 
-T1_architecture <- proc.time()[3]
-timediff_purrr_archi <- T1_architecture - T0_architecture
-print(paste0("!!!!!--------------------------- ", round(timediff_purrr_archi/60,2), " minutes passed for the simulation scenario! --------------------------------------!!!!!"))
-
-saveRDS(est_sim_architectures, paste0("./output/Rdata/sim-results/", Sys.Date(), "_est_sim_architectures_purrr_NSIM_", nsim, ".rds"))
+if (runGeneticArchitectures == TRUE) {
+  T1_architecture <- proc.time()[3]
+  timediff_purrr_archi <- T1_architecture - T0_architecture
+  print(paste0("!!!!!--------------------------- ", round(timediff_purrr_archi/60,2), " minutes passed for the simulation scenario! --------------------------------------!!!!!"))
+  
+  saveRDS(est_sim_architectures, paste0("./output/Rdata/sim-results/", Sys.Date(), "_est_sim_architectures_purrr_NSIM_", nsim, ".rds"))
+}
 

### 2023-04-18 scripts/MSc.Rmd
small changes

--- 2023-03-13 scripts/MSc.Rmd
@@
     latex_engine: xelatex
-bibliography: ../references/references.bib
+bibliography: ./references/references.bib
 editor_options: 
@@
 #knitr::opts_chunk$set(echo = FALSE)
-figurespath <- paste0("../reports/", format(Sys.time(), "%Y%m%d"), '_figures/')
+figurespath <- paste0("./reports/", format(Sys.time(), "%Y%m%d"), '_figures/')
 knitr::opts_chunk$set(fig.path = figurespath)  # reference is script! also creates new dirs
@@
 library(ggpubr)
+library(gt)
 library(scales)
@@
 conflict_prefer("filter", "dplyr")
+conflicts_prefer(kableExtra::kable)
 # conflict_prefer("kable", "kableExtra")
@@
 ```{r load-data}
-my_data <- as_tibble(read.delim("../data-boekstegers_felix/GST_BMI_association_results.txt"))
+my_data <- as_tibble(read.delim("./data-boekstegers_felix/GST_BMI_association_results.txt"))
 # GBC: new data 2022-11-04
-GBC_dat <- as_tibble(read.delim("../data-boekstegers_felix/C23C24_BMI_association_results.txt"))
+GBC_dat <- as_tibble(read.delim("./data-boekstegers_felix/C23C24_BMI_association_results.txt"))
 
 # FTO data
-FTO_dat <- as_tibble(read.delim("../data-boekstegers_felix/2023-01-19_FTO/FTO_association_results.txt"))
+FTO_dat <- as_tibble(read.delim("./data-boekstegers_felix/2023-01-19_FTO/FTO_association_results.txt"))
 ```
@@
 # see script biomart_search.R
-rsIds_GeneNames <- as_tibble(readRDS("../data/biomartBMIsearch.rds"))
+rsIds_GeneNames <- as_tibble(readRDS("./data/biomartBMIsearch.rds"))
 
@@
 if (SAVE.files ==TRUE ){
-  filenamefig00_b <- paste0("../output/plots/pve/", Sys.Date(), "_explained_var_X_new_sample_size_dep.svg")
+  filenamefig00_b <- paste0("./output/plots/pve/", Sys.Date(), "_explained_var_X_new_sample_size_dep.svg")
   ggsave(filename = filenamefig00_b, plot = fig00_b, 
@@
 if (SAVE.files ==TRUE ){
-  filename_final_plot1 <- paste0("../output/plots/pve/", Sys.Date(), "_final_plot1.arrange_OLD.svg")
-  filename_final_plot1b <- paste0("../output/plots/pve/", Sys.Date(), "_final_plot1b.arrange_NEW.svg")
+  filename_final_plot1 <- paste0("./output/plots/pve/", Sys.Date(), "_final_plot1.arrange_OLD.svg")
+  filename_final_plot1b <- paste0("./output/plots/pve/", Sys.Date(), "_final_plot1b.arrange_NEW.svg")
 
@@
 if (SAVE.files ==TRUE ){
-  saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
+  saveRDS(my_data_harm, paste0("./output/Rdata/harmonized-dat/", Sys.Date(), "_my_data_harm.rds"))
 }
@@
 ## load results from script `perform_mr.R`
-est.bmi.gsd <- readRDS(dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
+est.bmi.gsd <- readRDS(dplyr::last(list.files("./output/Rdata/mr-results/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))
 
@@
   mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
-  kable(digits=3, caption = "MR Results for GSD (outcome) using genetic variants as instrumental variables for BMI") %>% 
-  kable_classic(full_width = FALSE, html_font = "Cambria")
+  mutate(across(where(is.numeric), round, 3)) %>% 
+  arrange(tolower(mr_methods)) %>% 
+  relocate(time, .after = last_col()) -> est.bmi.gsd.table
+
+# est.bmi.gsd.table %>% 
+#   kable(digits=3, caption = "MR Results for GSD (outcome) using genetic variants as instrumental variables for BMI") %>% 
+#   kable_classic(full_width = FALSE, html_font = "Cambria")
+
+est.bmi.gsd.table %>% 
+  gt() %>% 
+  cols_merge(
+      columns = c(estimate, se.estimate),
+      pattern = "{1} [\u00b1 {2}]"  # +- sign : "\u00b1"
+    ) %>% 
+  cols_merge(
+      columns = c(OR, CI.lower, CI.upper),
+      pattern = "{1} [{2}; {3}]"  
+    ) %>% 
+  cols_label(
+    mr_methods = md("Methods"),
+    estimate = md("Estimate [\u00b1 StdErr]"),
+    OR = md("OR [95% CI]"),
+    time=md("Computational time [sec]")
+  ) %>% 
+  fmt_number(decimals = 3, use_seps=TRUE) %>% 
+  tab_source_note(
+    source_note = md("MR Results for GSD (outcome) using genetic variants as instrumental variables for BMI")
+  ) %>% 
+  opt_stylize(style=1, color="gray") -> gt.tables.est.bmi.gsd
+
+if (SAVE.files ==TRUE ){
+ filename.gt.est.bmi.gsd <- paste0("./output/gt-tables/mr-estimates/", Sys.Date(), 
+                                       "_est_bmi_gsd")
+
+ gt.tables.est.bmi.gsd %>% 
+   gtsave(paste0(filename.gt.est.bmi.gsd, ".docx"))   # alternative .rtf, .tex, .png, .html
+ gt.tables.est.bmi.gsd %>% 
+   gtsave(paste0(filename.gt.est.bmi.gsd, ".html")) 
+}
 ```
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.scatter.gsd <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrscatter_gsd.svg")
+  filename.mr.scatter.gsd <- paste0("./output/plots/mr-analysis/", Sys.Date(), "_figure_mrscatter_gsd.svg")
   ggsave(filename = filename.mr.scatter.gsd, plot = p1.gsd, 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.forest.gsd <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrforest_GSD.svg")
+  filename.mr.forest.gsd <- paste0("./output/plots/mr-analysis/", Sys.Date(), "_figure_mrforest_GSD.svg")
   ggsave(filename = filename.mr.forest.gsd, plot = p2.gsd, 
@@
 
+ggarrange(p1.gsd, p4.gsd, nrow = 1, labels = c("A", "B")) -> gsd_scatter_funnel_arrange
+
 #SAVE
 if (SAVE.files ==TRUE ){
-  filename.mr.funnel.gsd <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrfunnel_gsd.svg")
+  filename.mr.funnel.gsd <- paste0("./output/plots/mr-analysis/", Sys.Date(), "_figure_mrfunnel_gsd.svg")
+  filename.mr.scatterfunnel.gsd <- paste0("./output/plots/mr-analysis/", Sys.Date(), "_figure_scatter_mrfunnel_arrange_gsd.svg")
   ggsave(filename = filename.mr.funnel.gsd, plot = p4.gsd, 
@@
            )
+
+  ggsave(filename.mr.scatterfunnel.gsd, gsd_scatter_funnel_arrange, width=8, height=4, units="in")
 }
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.scatter <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrscatter.svg")
+  filename.mr.scatter <- paste0("./output/plots/mr-analysis/", Sys.Date(), "_figure_mrscatter.svg")
   ggsave(filename = filename.mr.scatter, plot = p1[[1]], 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.forest <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrforest_GBC.svg")
+  filename.mr.forest <- paste0("./output/plots/mr-analysis/", Sys.Date(), "_figure_mrforest_GBC.svg")
   ggsave(filename = filename.mr.forest, plot = p2[[1]], 
@@
 if (SAVE.files ==TRUE ){
-  filename.mr.funnel.gbc <- paste0("../output/plots/mr-analysis/", Sys.Date(), "_figure_mrfunnel_gbc.svg")
+  filename.mr.funnel.gbc <- paste0("./output/plots/mr-analysis/", Sys.Date(), "_figure_mrfunnel_gbc.svg")
   ggsave(filename = filename.mr.funnel.gbc, plot = p4, 
@@
 ## load results
-est <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
+est <- readRDS(file = dplyr::last(list.files("./output/Rdata/mr-results/", pattern = "_est.rds", full.names = TRUE)))
 
@@
 if (SAVE.files ==TRUE ){
-  saveRDS(sim_vars, paste0("../output/Rdata/", Sys.Date(), "_sim_vars.rds"))
+  saveRDS(sim_vars, paste0("./output/Rdata/", Sys.Date(), "_sim_vars.rds"))
 }
@@
 # from "purrr-simstudy.R"
-est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_purrr_NSIM_", full.names = TRUE)))
-est_sim_architectures <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_architectures_purrr_NSIM_", full.names = TRUE)))
+est_sim <- readRDS(dplyr::last(list.files(path = "./output/Rdata/sim-results/", pattern = "_est_sim_purrr_NSIM_", full.names = TRUE)))
+est_sim_architectures <- readRDS(dplyr::last(list.files(path = "./output/Rdata/sim-results/", pattern = "_est_sim_architectures_purrr_NSIM_", full.names = TRUE)))
 nsim <- nrow(est_sim)
+nsim_architectures <- nrow(est_sim_architectures[[1]])
 ```
@@
 if (SAVE.files ==TRUE ){
-  filename_pveplot_sim.raincloud <- paste0("../output/plots/pve/", Sys.Date(), "_figure_simplot_pve_raincloud.svg")
+  filename_pveplot_sim.raincloud <- paste0("./output/plots/pve/", Sys.Date(), "_figure_simplot_pve_raincloud.svg")
   ggsave(filename = filename_pveplot_sim.raincloud, plot = plot_pve_simulated.raincloud, 
@@
 IVW_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "IVW method on BMI -> GSD performance metrics") %>% 
@@
 MODE_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "MODE method on BMI -> GSD performance metrics") %>% 
@@
 CONMIX_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "CONMIX method on BMI -> GSD performance metrics") %>% 
@@
 MRMIX_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "MRMIX method on BMI -> GSD performance metrics") %>% 
@@
 MEDIAN_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "MEDIAN method on BMI -> GSD performance metrics") %>% 
@@
 EGGER_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "EGGER method on BMI -> GSD performance metrics") %>% 
@@
 ROBUST_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "ROBUST method on BMI -> GSD performance metrics") %>% 
@@
 PRESSO_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "PRESSO method on BMI -> GSD performance metrics") %>% 
@@
 RAPS_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "RAPS method on BMI -> GSD performance metrics") %>% 
@@
 LASSO_GSD_performance_metrics %>% 
-  mutate(across(cols = where(is.numeric), round(3))) %>% 
+  mutate(across(where(is.numeric), round, 3)) %>% 
   kable(caption = "LASSO method on BMI -> GSD performance metrics") %>% 
@@
 
-  # # STORE
-  # if (SAVE.files ==TRUE ){
-  #   filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_IVW_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = IVW_GSD_performance_metrics, file = filename.IVW_table.performance)
-  #   filename.MODE_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_MODE_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = MODE_GSD_performance_metrics, file = filename.MODE_table.performance)
-  #   filename.CONMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_CONMIX_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = CONMIX_GSD_performance_metrics, file = filename.CONMIX_table.performance)
-  #   filename.MRMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_MRMIX_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
-  #   #5
-  #   filename.MEDIAN_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_MEDIAN_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename.MEDIAN_table.performance)
-  #   #6
-  #   filename.EGGER_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_EGGER_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = EGGER_GSD_performance_metrics, file = filename.EGGER_table.performance)
-  #   #7
-  #   filename.ROBUST_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_ROBUST_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = ROBUST_GSD_performance_metrics, file = filename.ROBUST_table.performance)
-  #   #8
-  #   filename.PRESSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_PRESSO_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = PRESSO_GSD_performance_metrics, file = filename.PRESSO_table.performance)
-  #   #9
-  #   filename.RAPS_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_RAPS_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = RAPS_GSD_performance_metrics, file = filename.RAPS_table.performance)
-  #   #10
-  #   filename.LASSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-  #                                       "_LASSO_GSD_performance_metrics_TRUE_EST_", 
-  #                                       selected_true_methods[true_method], "_NSIM_", nsim, ".csv")
-  #   write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
-  # }
-
   ###
@@
 
-  # ####-------------
-  # # SAVE  --------------------------------
-  # if (SAVE.files ==TRUE ){
-  #   #1
-  #   filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #2
-  #   filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #3
-  #   filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #4
-  #   filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MRMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #5
-  #   filename_MEDIAN_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MEDIAN_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_MEDIAN_GSD_simplot_thetas, plot = MEDIAN_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #6
-  #   filename_EGGER_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_EGGER_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_EGGER_GSD_simplot_thetas, plot = EGGER_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #7
-  #   filename_ROBUST_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_ROBUST_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_ROBUST_GSD_simplot_thetas, plot = ROBUST_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #8
-  #   filename_PRESSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_PRESSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_PRESSO_GSD_simplot_thetas, plot = PRESSO_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #9
-  #   filename_RAPS_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_RAPS_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_RAPS_GSD_simplot_thetas, plot = RAPS_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  #   #10
-  #   filename_LASSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_LASSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".png")
-  #   ggsave(filename = filename_LASSO_GSD_simplot_thetas, plot = LASSO_GSD_simplot_thetas, 
-  #          width = 15, height = 6, 
-  #          units = "in"  # default
-  #          )
-  # }
-
   IVW_GSD_simplot_thetas$labels$title <- ""
@@
   if (SAVE.files ==TRUE ){
-    filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/theta_sims/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".svg")
+    filename_GSD_simplot_thetas_arrange <- paste0("./output/plots/theta_sims/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], "_NSIM_", nsim,".svg")
     ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
-           width = 10, height = 38, 
+           width = 10, height = 38, device = "svg",
            units = "in"  # default
@@
   if (SAVE.files ==TRUE ){
-    filename_simplot_bias_mse <- paste0("../output/plots/bias-mse/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], "_NSIM_",nsim, ".svg")
+    filename_simplot_bias_mse <- paste0("./output/plots/bias-mse/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], "_NSIM_",nsim, ".svg")
     ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot_GSD, 
@@
 
-    # # STORE
-    # if (SAVE.files ==TRUE ){
-    #   filename.IVW_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_IVW_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = IVW_GSD_performance_metrics, file = filename.IVW_table.performance)
-    #   filename.MODE_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_MODE_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = MODE_GSD_performance_metrics, file = filename.MODE_table.performance)
-    #   filename.CONMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_CONMIX_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = CONMIX_GSD_performance_metrics, file = filename.CONMIX_table.performance)
-    #   filename.MRMIX_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_MRMIX_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = MRMIX_GSD_performance_metrics, file = filename.MRMIX_table.performance)
-    #   #5
-    #   filename.MEDIAN_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_MEDIAN_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = MEDIAN_GSD_performance_metrics, file = filename.MEDIAN_table.performance)
-    #   #6
-    #   filename.EGGER_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_EGGER_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = EGGER_GSD_performance_metrics, file = filename.EGGER_table.performance)
-    #   #7
-    #   filename.ROBUST_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_ROBUST_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = ROBUST_GSD_performance_metrics, file = filename.ROBUST_table.performance)
-    #   #8
-    #   filename.PRESSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_PRESSO_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = PRESSO_GSD_performance_metrics, file = filename.PRESSO_table.performance)
-    #   #9
-    #   filename.RAPS_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_RAPS_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = RAPS_GSD_performance_metrics, file = filename.RAPS_table.performance)
-    #   #10
-    #   filename.LASSO_table.performance <- paste0("../output/performance_metrics/", Sys.Date(), 
-    #                                       "_LASSO_GSD_performance_metrics_TRUE_EST_", 
-    #                                       selected_true_methods[true_method], 
-    #                                       "_GenArch_", name_GenArch,
-    #                                       "_NSIM_", nsim, ".csv")
-    #   write.csv2(x = LASSO_GSD_performance_metrics, file = filename.LASSO_table.performance)
-    # }
-
     ###
@@
 
-    ####-------------
-    # SAVE  --------------------------------
-    # #1
-    # filename_IVW_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_IVW_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
-    #                                           "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_IVW_GSD_simplot_thetas, plot = IVW_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #2
-    # filename_MODE_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MODE_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
-    #                                            "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_MODE_GSD_simplot_thetas, plot = MODE_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #3
-    # filename_CONMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_CONMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
-    #                                              "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_CONMIX_GSD_simplot_thetas, plot = CONMIX_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #4
-    # filename_MRMIX_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MRMIX_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
-    #                                             "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_MRMIX_GSD_simplot_thetas, plot = MRMIX_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #5
-    # filename_MEDIAN_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_MEDIAN_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
-    #                                              "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_MEDIAN_GSD_simplot_thetas, plot = MEDIAN_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #6
-    # filename_EGGER_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_EGGER_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method], 
-    #                                             "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_EGGER_GSD_simplot_thetas, plot = EGGER_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #7
-    # filename_ROBUST_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_ROBUST_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
-    #                                              "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_ROBUST_GSD_simplot_thetas, plot = ROBUST_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #8
-    # filename_PRESSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_PRESSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
-    #                                              "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_PRESSO_GSD_simplot_thetas, plot = PRESSO_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #9
-    # filename_RAPS_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_RAPS_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
-    #                                            "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_RAPS_GSD_simplot_thetas, plot = RAPS_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-    # #10
-    # filename_LASSO_GSD_simplot_thetas <- paste0("../output/plots/", Sys.Date(), "_LASSO_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
-    #                                             "_GenArch_", name_GenArch, "_NSIM_", nsim,".png")
-    # ggsave(filename = filename_LASSO_GSD_simplot_thetas, plot = LASSO_GSD_simplot_thetas, 
-    #        width = 15, height = 6, 
-    #        units = "in"  # default
-    #        )
-
     IVW_GSD_simplot_thetas$labels$title <- ""
@@
     if (SAVE.files ==TRUE ){
-      filename_GSD_simplot_thetas_arrange <- paste0("../output/plots/theta-sims/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
+      filename_GSD_simplot_thetas_arrange <- paste0("./output/plots/theta-sims/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
                                                     "_GenArch_", name_GenArch, "_NSIM_", nsim,".svg")
@@
     if (SAVE.files ==TRUE ){
-      filename_simplot_bias_mse <- paste0("../output/plots/bias-mse/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], 
+      filename_simplot_bias_mse <- paste0("./output/plots/bias-mse/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], 
                                           "_GenArch_", name_GenArch, "_NSIM_", nsim,".svg")
@@
 }
-
 ```
@@
 
-### {gt} manual stub and grouping of rows  ###
-### https://gt.rstudio.com/articles/intro-creating-gt-tables.html 
-### https://themockup.blog/static/resources/gt-cookbook.html#introduction
-
-#df.temp.wide[[ref.method]] %>% 
-  #gt(groupname_col="N") %>% 
-  #gt(rowname_col = "N") %>% 
-  # tab_stubhead(label = "Sample Size") %>% 
-  # tab_row_group(
-  #   label = "Sample Size = 50,000", 
-  #   rows = 1:10
-  # ) %>% 
-  # tab_row_group(
-  #   label = "Sample Size = 100,000", 
-  #   rows = 11:20
-  # )
-
 gt.tables.mean.se <- list()
@@
 
-  ## TABLE 01: MEAN +- SE
+  ## TABLE 01: MEAN +- SE -----------------------------------------------------
   df.temp.long[[ref.method]] %>% 
@@
   if (SAVE.files ==TRUE ){
-    filename.gt.mean.se <- paste0("../output/gt-tables/", Sys.Date(), 
+    filename.gt.mean.se <- paste0("./output/gt-tables/sim-results/mean-empse/", Sys.Date(), 
                                           "_mean_emp_se_performance_metrics_TRUE_EST_", 
@@
       gtsave(paste0(filename.gt.mean.se, ".html")) 
-  }
-
-  # TABLE 02: BIAS, MSE, COVERAGE, TYPE1ERROR
+    gt.tables.mean.se[[ref.method]] %>% 
+      gtsave(paste0(filename.gt.mean.se, ".tex")) 
+  }
+
+  # TABLE 02: BIAS, MSE, COVERAGE, TYPE1ERROR ---------------------------------
   df.temp.long[[ref.method]] %>%
@@
     if (SAVE.files ==TRUE ){
-      filename.gt.performance2 <- paste0("../output/gt-tables/", Sys.Date(), 
+      filename.gt.performance2 <- paste0("./output/gt-tables/sim-results/", Sys.Date(), 
                                             "_performance2_metrics_TRUE_EST_", 
@@
         gtsave(paste0(filename.gt.performance2, ".html")) 
-    }
-
+      gt.tables.performance2[[ref.method]] %>% 
+        gtsave(paste0(filename.gt.performance2, ".tex")) 
+    }
 }
@@
 ```
+
+\clearpage
 
@@
 dodged.barplot.mse <- list()
+dodged.barplot.mse.scenario <- list()  # store plots for 1 ref-estimate, all Scenarios
+dodged.barplot.mse.references <- list()  # store plots for ALL_INSTR for all different ref-estimates
 dodged.barplot.bias <- list()
+dodged.barplot.bias.scenario <- list()
+dodged.barplot.bias.references <- list()
 dodged.barplot.coverage <- list()
 dodged.barplot.power<- list()
+
+scenarios <- c("ALL_INSTR", "STRONG_COMMON", "STRONG_RARE", "WEAK_COMMON", "WEAK_RARE")
 
@@
 
+  # plot per scenario
+  for(scenario in seq_along(scenarios)) {
+
+    # customize axis 
+    if (scenarios[scenario] != "WEAK_RARE") {
+      my.y.lim <- 0.15
+    } else {
+      my.y.lim <- 0.3
+    }
+
+    df.temp.long[[ref.method]] %>%
+      mutate(N=paste0("Sample Size: ", format(N, scientific=FALSE))) %>% 
+      mutate(label.too.large = ifelse(mse>my.y.lim, paste0("MSE \u2265 ", my.y.lim) , NA)) %>% 
+      filter(Scenario == scenarios[scenario]) %>% 
+      ## DROP NA (e.g. no sample sizes 50.000)
+      drop_na(Mean) %>%  
+      ggplot(aes(x=Method, y=mse)) +
+      geom_bar(stat="identity", position=position_dodge2(reverse=TRUE),
+               #color=ggthemes_data$few$colors$Dark[2,2][[1]],
+               ) +   # reverse dodge
+      geom_text(aes(x= Method, y= my.y.lim, label = label.too.large,
+                    colour=label.too.large), #position = position_dodge(),
+                nudge_y = -0.1, #size = 2.25, angle = 90
+                color=ggthemes_data$few$colors$Dark[4,2][[1]],
+                fontface="bold"
+                )+
+      facet_wrap(~N, nrow = 4) + 
+      xlab("Method")+
+      ylab(paste0("Mean Squared Error with nsim=", nsim))+
+      labs(title = paste0("Selected Reference Method: ", 
+                              selected_true_methods[ref.method]),
+           subtitle = paste0("Genetic Architecture: ", scenarios[scenario])
+           )+
+      #ylim(0, 0.5)+
+      coord_flip(ylim=c(0, my.y.lim))+
+      ggthemes::theme_few()+  # remove legend
+      ggthemes::scale_fill_few() -> dodged.barplot.mse.scenario[[scenario]]
+
+      if (SAVE.files ==TRUE ){
+        ## arranged plots are toooo small 
+        filename.dodged.barplot.mse.scenario <- paste0("./output/plots/summary-plots/scenario_",scenarios[scenario],"/",
+                                              Sys.Date(), "_mse_GSD_dodged_barplot_REF_METHOD_",
+                                              selected_true_methods[ref.method], "_NSIM_", nsim)
+        ggsave(filename = paste0(filename.dodged.barplot.mse.scenario, ".svg"), 
+               plot = dodged.barplot.mse.scenario[[scenario]], 
+               width = 5, height = 8, 
+               units = "in"  # default
+               )
+      }
+  }
+  dodged.barplot.mse.references[ref.method] <- dodged.barplot.mse.scenario[1]
+
+  # scenarios without all instruments in arranged plot
+  arrangedplot.mse.scenarios <- 
+    ggarrange(plotlist = dodged.barplot.mse.scenario[-1], labels = c("A", "B", "C", "D", "E", "F"))
+  if (SAVE.files ==TRUE ){
+        ## arranged plots are toooo small 
+        filename.arrangedplot.mse.scenarios <- paste0("./output/plots/summary-plots/arranged_scenarios/",
+                                              Sys.Date(), "_mse_GSD_dodged_barplot_REF_METHOD_",
+                                              selected_true_methods[ref.method], "_NSIM_", nsim)
+        ggsave(filename = paste0(filename.arrangedplot.mse.scenarios, ".svg"), 
+               plot = arrangedplot.mse.scenarios, 
+               width = 10, height = 12, 
+               units = "in"  # default
+               )
+  }
+
+  # with all scenarios in 1 plot stacked
   df.temp.long[[ref.method]] %>%
@@
     ## arranged plots are toooo small 
-    filename.dodged.barplot.mse <- paste0("../output/plots/summary-plots/", Sys.Date(), "_GSD_dodged_barplot_mse_REF_METHOD_",
+    filename.dodged.barplot.mse <- paste0("./output/plots/summary-plots/MSE_stacked/", Sys.Date(),
+                                          "_mse_GSD_dodged_barplot_REF_METHOD_",
                                           selected_true_methods[ref.method], "_NSIM_", nsim)
@@
 
+  # plot per scenario
+  for(scenario in seq_along(scenarios)) {
+
+    # customize axis 
+    if (scenarios[scenario] != "WEAK_RARE") {
+      my.y.lim <- 0.15
+    } else {
+      my.y.lim <- 0.3
+    }
+
+    df.temp.long[[ref.method]] %>%
+      mutate(N=paste0("Sample Size: ", format(N, scientific=FALSE))) %>% 
+      filter(Scenario == scenarios[scenario]) %>% 
+      ## DROP NA (e.g. no sample sizes 50.000)
+      drop_na(Mean) %>%  
+      ggplot(aes(x=Method, y=bias)) +
+      geom_bar(stat="identity", position=position_dodge2(reverse=TRUE),
+               #color=ggthemes_data$few$colors$Dark[2,2][[1]],
+               ) +   # reverse dodge
+      geom_hline(yintercept = 0,  col=ggthemes_data$few$colors$Dark[3,2][[1]], alpha=0.6)+
+      facet_wrap(~N, nrow = 4) + 
+      xlab("Method")+
+      ylab(paste0("Bias with nsim=", nsim))+
+      labs(title = paste0("Selected Reference Method: ", 
+                              selected_true_methods[ref.method]),
+           subtitle = paste0("Genetic Architecture: ", scenarios[scenario])
+           )+
+      #ylim(0, 0.5)+
+      coord_flip()+
+      ggthemes::theme_few()+  # remove legend
+      ggthemes::scale_fill_few() -> dodged.barplot.bias.scenario[[scenario]]
+
+      if (SAVE.files ==TRUE ){
+        ## arranged plots are toooo small 
+        filename.dodged.barplot.bias.scenario <- paste0("./output/plots/summary-plots/scenario_",scenarios[scenario],"/",
+                                              Sys.Date(), "_bias__GSD_dodged_barplot_REF_METHOD_",
+                                              selected_true_methods[ref.method], "_NSIM_", nsim)
+        ggsave(filename = paste0(filename.dodged.barplot.bias.scenario, ".svg"), 
+               plot = dodged.barplot.bias.scenario[[scenario]], 
+               width = 5, height = 8, 
+               units = "in"  # default
+               )
+      }
+  }
+  dodged.barplot.bias.references[ref.method] <- dodged.barplot.bias.scenario[1]
+
+   # scenarios without all instruments in arranged plot
+  arrangedplot.bias.scenarios <- 
+    ggarrange(plotlist = dodged.barplot.bias.scenario[-1], labels = c("A", "B", "C", "D", "E", "F"))
+  if (SAVE.files ==TRUE ){
+        ## arranged plots are toooo small 
+        filename.arrangedplot.bias.scenarios <- paste0("./output/plots/summary-plots/arranged_scenarios/",
+                                              Sys.Date(), "_bias_GSD_dodged_barplot_REF_METHOD_",
+                                              selected_true_methods[ref.method], "_NSIM_", nsim)
+        ggsave(filename = paste0(filename.arrangedplot.bias.scenarios, ".svg"), 
+               plot = arrangedplot.bias.scenarios, 
+               width = 10, height = 12, 
+               units = "in"  # default
+               )
+  }
+
+  # stacked
   df.temp.long[[ref.method]] %>%
@@
     ## arranged plots are toooo small 
-    filename.dodged.barplot.bias <- paste0("../output/plots/summary-plots/", Sys.Date(), "_GSD_dodged_barplot_bias_REF_METHOD_",
+    filename.dodged.barplot.bias <- paste0("./output/plots/summary-plots/BIAS_stacked/", Sys.Date(),
+                                           "_bias_GSD_dodged_barplot_REF_METHOD_",
                                           selected_true_methods[ref.method], "_NSIM_", nsim)
@@
   ## COVERAGE PLOT FACET WRAP ------------------------------------------------------
+  # stacked
   df.temp.long[[ref.method]] %>%
@@
     ## arranged plots are toooo small 
-    filename.dodged.barplot.coverage <- paste0("../output/plots/summary-plots/", Sys.Date(), "_GSD_dodged_barplot_coverage_REF_METHOD_",
+    filename.dodged.barplot.coverage <- paste0("./output/plots/summary-plots/COVERAGE_stacked/", Sys.Date(),
+                                               "_coverage_GSD_dodged_barplot_REF_METHOD_",
                                           selected_true_methods[ref.method], "_NSIM_", nsim)
@@
   ## Power of 5% level test (not TYPE1ERROR) PLOT FACET WRAP ------------------------------------------------------
+  # stacked
   df.temp.long[[ref.method]] %>%
@@
     ## arranged plots are toooo small 
-    filename.dodged.barplot.power <- paste0("../output/plots/summary-plots/", Sys.Date(), "_GSD_dodged_barplot_power_REF_METHOD_",
+    filename.dodged.barplot.power <- paste0("./output/plots/summary-plots/POWER_stacked/", Sys.Date(), "_power_GSD_dodged_barplot_REF_METHOD_",
                                           selected_true_methods[ref.method], "_NSIM_", nsim)
@@
 }
+
+  # scenarios without all instruments in arranged plot
+
+dodged.barplot.mse.references.update <- map(seq_along(dodged.barplot.mse.references), ~ {
+  dodged.barplot.mse.references[[.x]][["labels"]][["subtitle"]] <- NULL
+  return(dodged.barplot.mse.references[[.x]])
+} )
+
+arrangedplot.mse.references <- 
+    ggarrange(plotlist = dodged.barplot.mse.references.update, labels = c("A", "B", "C", "D", "E", "F"), ncol=3, nrow=2)
+
+if (SAVE.files ==TRUE ){
+    ## arranged plots are toooo small 
+    filename.dodged.barplot.mse <- paste0("./output/plots/summary-plots/arranged_references/", Sys.Date(), "_mse_GSD_dodged_barplot_REF_METHOD_",
+                                          selected_true_methods[ref.method], "_NSIM_", nsim)
+    ggsave(filename = paste0(filename.dodged.barplot.mse, ".svg"), plot =  arrangedplot.mse.references, 
+             width = 14, height = 14, 
+             units = "in"  # default
+             )
+}
+
+dodged.barplot.bias.references.update <- map(seq_along(dodged.barplot.bias.references), ~ {
+  dodged.barplot.bias.references[[.x]][["labels"]][["subtitle"]] <- NULL
+  return(dodged.barplot.bias.references[[.x]])
+} )
+
+arrangedplot.bias.references <- 
+    ggarrange(plotlist = dodged.barplot.bias.references.update, labels = c("A", "B", "C", "D", "E", "F"), ncol=3, nrow=2)
+
+if (SAVE.files ==TRUE ){
+    ## arranged plots are toooo small 
+    filename.dodged.barplot.bias <- paste0("./output/plots/summary-plots/arranged_references/", Sys.Date(), "_bias_GSD_dodged_barplot_REF_METHOD_",
+                                          selected_true_methods[ref.method], "_NSIM_", nsim)
+    ggsave(filename = paste0(filename.dodged.barplot.bias, ".svg"), plot =  arrangedplot.bias.references, 
+             width = 14, height = 14, 
+             units = "in"  # default
+             )
+}
+
 ```

### 2023-04-18 scripts/helper/plotfigure01fn.R
changed plots to exclude GBC

@@ -3,7 +3,7 @@
 library(tibble)
 library(ggplot2)
 
-plotfigure01 <- function(data, effect_size="all", maf="all", n){
+plotfigure01.withGBC <- function(data, effect_size="all", maf="all", n){
   ## data: my_data_harm
   # effect size: "all", "strong" or "weak"
   # maf: "all", "rare" or "common
@@ -83,7 +83,7 @@ plotfigure01 <- function(data, effect_size="all", maf="all", n){
       #ylab("% variance explained by IVs")+
       labs(title = "Figure 1", subtitle = plotsubtitle,
            col='', 
-           caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits   explained by the IVs"
+           caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
       scale_y_continuous(
         # Features of the first axis
@@ -148,7 +148,159 @@ plotfigure01 <- function(data, effect_size="all", maf="all", n){
   
   print(figure01.quadrant)
   
-  filenamefig01_quadrant <- paste0("../output/plots/", Sys.Date(), "_figure01_", 
+  filenamefig01_quadrant <- paste0("./output/plots/pve/plot-fn-pve/", Sys.Date(), "_figure01_", 
+                                   "EFFECT_SIZE_", effect_size,
+                                   "_VARIANTS_", maf,
+                                   ".png")
+  ggsave(filename = filenamefig01_quadrant, plot = figure01.quadrant, 
+         width = 10, height = 6, 
+         units = "in"  # default
+  )
+  
+  # store plot in global environment
+  plotname <<- paste0("figure01_EFFECT_SIZE_", effect_size, "_VARIANTS_", maf)
+  assign(plotname, figure01.quadrant, envir = .GlobalEnv)
+  #list2env(assign(plotname, figure01.quadrant), envir = .GlobalEnv)
+}
+
+## without GBC -------------------------------------------------------------
+
+plotfigure01 <- function(data, effect_size="all", maf="all", n){
+  ## data: my_data_harm
+  # effect size: "all", "strong" or "weak"
+  # maf: "all", "rare" or "common
+  df_cum_expl_var <- tibble(n=n, 
+                            cum_expl_varX = rep(NA, length(n)), 
+                            cum_expl_varX2 = rep(NA, length(n)), 
[…]

### 2023-04-18 scripts/MSc.Rmd
changed to GSD only, no GBC

@@
 plot_data1b <- df_cum_expl_var %>% 
-  pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, cum_expl_varY)) %>% 
-  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
-         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
+  pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, 
+                        #cum_expl_varY
+                        )) %>% 
+  mutate(name = factor(name, labels = c("% var(Y)", "% var(X)")) #, "% var(Y)")),   # change made on 2023-04-18
+         #name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
          )
@@
   #
-  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
+  scale_color_manual(name = "", values = c(#"% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                            "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
@@
                      # order
-                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
+                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)")  #, "% var(U)")
   ) -> fig01b.update

### 2023-04-19 scripts/MSc.Rmd
small changes

@@
 ```
+
+\clearpage
+
+# References
+
+<div id="refs"></div>
+
+\clearpage
+
+# Software
+
+This analysis was carried out using the statistical software `r base::version$version.string`.
+
+Apart from R's base functionality, the following packages were used: <br/>
+
+```{r software, echo=FALSE}
+# or faster with function...
+installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
+# map(names(sessionInfo()$otherPkgs), ~citation(.x))
+```

### 2023-04-19 scripts/MSc.Rmd
run nsim100 with PRESSO and GENARCH

@@
 nsim_architectures <- nrow(est_sim_architectures[[1]])
+
+all(is.na(est_sim$PRESSO_theta_se_sim_BMI_GSD_200000))
+all(is.na(est_sim_architectures[[1]]$PRESSO_mr_pval_BMI_GSD_200000))
+est_sim_PRESSO <- readRDS("./output/Rdata/_old/2022-12-23_est_sim_NSIM_100.rds")
+
 ```

### 2023-04-19 scripts/purrr-simstudy.R

@@ -39,9 +39,9 @@ conflict_prefer("filter", "dplyr")
 # set to TRUE if simulation should run when script is sourced
 runSimulation <- TRUE
 subsetSampleSizes <- TRUE
-runMRpresso <- FALSE  # runs foreeeever ;)
-nsim <- 1000
-runGeneticArchitectures <- FALSE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
+runMRpresso <- TRUE  # runs foreeeever ;)
+nsim <- 100
+runGeneticArchitectures <- TRUE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 
 # analysis set - create all combis
 effect_size <- c("strong", "weak")

### 2023-04-21 scripts/purrr-simstudy.R
nsim-500

@@ -40,7 +40,7 @@ conflict_prefer("filter", "dplyr")
 runSimulation <- TRUE
 subsetSampleSizes <- TRUE
 runMRpresso <- TRUE  # runs foreeeever ;)
-nsim <- 100
+nsim <- 500
 runGeneticArchitectures <- TRUE  # if set to TRUE, simulations run on 4 different genetic architectures (split: median maf, median beta)
 
 # analysis set - create all combis

### 2023-04-28 scripts/MSc.Rmd
slight changes in dirs and figure dimensions

--- 2023-04-19 scripts/MSc.Rmd
@@
     latex_engine: xelatex
-bibliography: ./references/references.bib
+bibliography: ../references/references.bib
 editor_options: 
@@
 ```{r setup, include=FALSE}
-#knitr::opts_chunk$set(echo = FALSE)
-figurespath <- paste0("./reports/", format(Sys.time(), "%Y%m%d"), '_figures/')
-knitr::opts_chunk$set(fig.path = figurespath)  # reference is script! also creates new dirs
-
-SAVE.files <- FALSE
+###
+# to knit the doc go to 'render_param_reports_msc.R'
+###
+
+# Set the working directory to the directory containing the R Markdown file
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+
+figurespath <- paste0("../reports/", format(Sys.time(), "%Y%m%d"), '_figures/')
+knitr::opts_chunk$set(echo = FALSE, 
+                      fig.path = figurespath,  # still a rmd relative file structure here
+                      dev = "svg")
+
+SAVE.files <- TRUE
 
@@
 ```
+
+![AG Meder](../img/meder_presentation_heart.png)
 
@@
 # from "purrr-simstudy.R"
-est_sim <- readRDS(dplyr::last(list.files(path = "./output/Rdata/sim-results/", pattern = "_est_sim_purrr_NSIM_", full.names = TRUE)))
-est_sim_architectures <- readRDS(dplyr::last(list.files(path = "./output/Rdata/sim-results/", pattern = "_est_sim_architectures_purrr_NSIM_", full.names = TRUE)))
+
+# est_sim <- readRDS(dplyr::last(list.files(path = "./output/Rdata/sim-results/", pattern = "_est_sim_purrr_NSIM_", full.names = TRUE)))
+# est_sim_architectures <- readRDS(dplyr::last(list.files(path = "./output/Rdata/sim-results/", pattern = "_est_sim_architectures_purrr_NSIM_", full.names = TRUE)))
+
+# # ANALYSIS SET 2023-04-20 NSIM = 100
+# est_sim <- readRDS(file = "./output/Rdata/sim-results/2023-04-20-nsim100/2023-04-20_est_sim_purrr_NSIM_100.rds")
+# est_sim_architectures <- readRDS(file = "./output/Rdata/sim-results/2023-04-20-nsim100/2023-04-20_est_sim_architectures_purrr_NSIM_100.rds")
+
+# ANALYSIS SET 2023-04-27 NSIM = 500
+est_sim <- readRDS(file = "./output/Rdata/sim-results/2023-04-27-nsim500/2023-04-24_est_sim_purrr_NSIM_500.rds")
+est_sim_architectures <- readRDS(file = "./output/Rdata/sim-results/2023-04-27-nsim500/2023-04-27_est_sim_architectures_purrr_NSIM_500.rds")
+
 nsim <- nrow(est_sim)
 nsim_architectures <- nrow(est_sim_architectures[[1]])
+
+# all(is.na(est_sim$PRESSO_theta_se_sim_BMI_GSD_200000))
+# all(is.na(est_sim_architectures[[1]]$PRESSO_mr_pval_BMI_GSD_200000))
+# est_sim_PRESSO <- readRDS("./output/Rdata/_old/2022-12-23_est_sim_NSIM_100.rds")
+
 ```
@@
 
----
-
 # Parallelize performance estimates for different "TRUE" estimates
@@
 
-```{r parallel-performance}
+```{r parallel-performance, warning=FALSE}
 for (true_method in 1:length(selected_true_methods)) {
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("IVW estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("IVW estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+                           ))+
     theme_few() -> IVW_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("MODE estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("MODE estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> MODE_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("CONMIX estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("CONMIX estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> CONMIX_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("MRMIX estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("MRMIX estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> MRMIX_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("MEDIAN estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("MEDIAN estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> MEDIAN_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("EGGER estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("EGGER estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> EGGER_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("ROBUST estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("ROBUST estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> ROBUST_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("PRESSO estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("PRESSO estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> PRESSO_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("RAPS estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("RAPS estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> RAPS_GSD_simplot_thetas
@@
     labs(title = "Simulated Thetas with Empirical Standard Error", 
-         subtitle = paste0("LASSO estimate: BMI -> GSD\nSelected Reference Method: ", 
-                           selected_true_methods[true_method], 
-                           " estimate calculated with total sample size"))+
+         subtitle = paste0("LASSO estimate: BMI -> GSD"
+                           #"\nSelected Reference Method: ", 
+                           #selected_true_methods[true_method], 
+                           #" estimate calculated with total sample size")
+         )
+    ) +
     theme_few() -> LASSO_GSD_simplot_thetas
@@
 
-```{r parallel-performance-genetic-architectures}
+```{r parallel-performance-genetic-architectures, warning=FALSE}
 effect_size <- c("strong", "weak")
@@
     if (SAVE.files ==TRUE ){
-      filename_GSD_simplot_thetas_arrange <- paste0("./output/plots/theta-sims/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
+      filename_GSD_simplot_thetas_arrange <- paste0("./output/plots/theta_sims/", Sys.Date(), "_GSD_simplot_thetas_REF_METHOD_", selected_true_methods[true_method],
                                                     "_GenArch_", name_GenArch, "_NSIM_", nsim,".svg")
@@
     if (SAVE.files ==TRUE ){
-      filename_simplot_bias_mse <- paste0("./output/plots/bias-mse/", Sys.Date(), "_GSD_simplot_bias_mse_REF_METHOD_", selected_true_methods[true_method], 
+      filename_simplot_bias_mse <- paste0("./output/plots/bias-mse/", Sys.Date(), "_GSD_bias_mse_REF_METHOD_",
+                                          selected_true_methods[true_method], 
                                           "_GenArch_", name_GenArch, "_NSIM_", nsim,".svg")
@@
 
-## loop to create tables
-for (ref.method in seq_along(selected_true_methods)) {
-
 ## TABLE 01: MEAN +- SE -----------------------------------------------------
-  df.temp.long[[ref.method]] %>% 
+# same for all references
+df.temp.long[[1]] %>% 
   select(N, Method, Scenario, Mean, emp_se) %>% 
@@
     title = md("**Mean estimate with corresponding empirical standard errors**"),
-      subtitle = md(paste0("Reference estimate: ", selected_true_methods[ref.method]))
+    #subtitle = md(paste0("Reference estimate: ", selected_true_methods[ref.method]))
   ) %>% 
@@
   ) %>% 
-    opt_stylize(style=1, color="gray") -> gt.tables.mean.se[[ref.method]]
+  opt_stylize(style=1, color="gray") -> gt.tables.mean.se[[1]]
+
+print(gt.tables.mean.se[[1]])
 
@@
   filename.gt.mean.se <- paste0("./output/gt-tables/sim-results/mean-empse/", Sys.Date(), 
-                                          "_mean_emp_se_performance_metrics_TRUE_EST_", 
-                                          selected_true_methods[ref.method], 
-                                          "_NSIM_", nsim)
-
-    gt.tables.mean.se[[ref.method]] %>% 
+                                        "_mean_emp_se_performance_metrics_NSIM_", nsim)
+
+  gt.tables.mean.se[[1]] %>% 
     gtsave(paste0(filename.gt.mean.se, ".docx"))   # alternative .rtf, .tex, .png, .html
-    gt.tables.mean.se[[ref.method]] %>% 
+  gt.tables.mean.se[[1]] %>% 
     gtsave(paste0(filename.gt.mean.se, ".html")) 
-    gt.tables.mean.se[[ref.method]] %>% 
+  gt.tables.mean.se[[1]] %>% 
     gtsave(paste0(filename.gt.mean.se, ".tex")) 
 }
+
+## loop to create tables with reference change
+for (ref.method in seq_along(selected_true_methods)) {
 
@@
     tab_spanner(
-      label = "Type 1 Error", 
+      label = "Power", 
       columns = 18:22
@@
 
+  print(gt.tables.performance2[[ref.method]])
+
   if (SAVE.files ==TRUE ){
@@
 
-```{r}
+```{r summary-plots-all-1, figures-side, fig.show='hold', out.width="100%", warning=FALSE, message=FALSE}
 dodged.barplot.mse <- list()
@@
 
+    plot(dodged.barplot.mse.scenario[[scenario]])
+
       if (SAVE.files ==TRUE ){
@@
     ggarrange(plotlist = dodged.barplot.mse.scenario[-1], labels = c("A", "B", "C", "D", "E", "F"))
+  plot(arrangedplot.mse.scenarios)
   if (SAVE.files ==TRUE ){
@@
     ggthemes::scale_fill_few() -> dodged.barplot.mse[[ref.method]]
+
+  plot(dodged.barplot.mse[[ref.method]])
 
@@
 
+    plot(dodged.barplot.bias.scenario[[scenario]])
+
       if (SAVE.files ==TRUE ){
@@
     ggarrange(plotlist = dodged.barplot.bias.scenario[-1], labels = c("A", "B", "C", "D", "E", "F"))
+
+  plot(arrangedplot.bias.scenarios)
+
   if (SAVE.files ==TRUE ){
@@
     ggthemes::scale_fill_few() -> dodged.barplot.bias[[ref.method]]
+
+  plot(dodged.barplot.bias[[ref.method]])
 
@@
 
+  plot(dodged.barplot.coverage[[ref.method]])
+
   if (SAVE.files ==TRUE ){
@@
 
+  plot(dodged.barplot.power[[ref.method]])
+
   if (SAVE.files ==TRUE ){
@@
 
+plot(arrangedplot.mse.references)
+
 if (SAVE.files ==TRUE ){
     ## arranged plots are toooo small 
-    filename.dodged.barplot.mse <- paste0("./output/plots/summary-plots/arranged_references/", Sys.Date(), "_mse_GSD_dodged_barplot_REF_METHOD_",
+    filename.dodged.barplot.mse <- paste0("./output/plots/summary-plots/arranged_references/", 
+                                          Sys.Date(), "_mse_GSD_dodged_barplot_REF_METHOD_",
                                           selected_true_methods[ref.method], "_NSIM_", nsim)
@@
     ggarrange(plotlist = dodged.barplot.bias.references.update, labels = c("A", "B", "C", "D", "E", "F"), ncol=3, nrow=2)
+
+plot(arrangedplot.bias.references)
 

### 2023-05-09 scripts/MSc.Rmd
updated tables, highlighting best methods

@@
 ```{r parallel-performance, warning=FALSE}
+
+arrangedplot.thetas.all <- list()
+
 for (true_method in 1:length(selected_true_methods)) {
@@
   ###
+
+  if (selected_true_methods[true_method] == "MRMix") {
+    color.dash <- "black"
+  } else if (selected_true_methods[true_method] == "FTO_Wald"){
+    color.dash <- "blue"
+  } else {
+    color.dash <- ggthemes_data$few$colors$Dark[9,2][[1]]
+  }
+  base_size_theta_plot <- 6
+
   IVW_BMI_GSD_geom_point_dat <- est_sim %>% 
@@
     geom_point()+
-    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
-           label= expression( paste("ref: ", theta[X] ))  , 
-           color= ggthemes_data$few$colors$Dark[9,2][[1]]
-             )+
+    geom_hline(yintercept=  args.gsd.method$theta, col=color.dash, alpha=0.6, linetype=2, size=0.8)+
+    # annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+    #        label= expression( paste("ref: ", theta[X] ))  , 
+    #        color= ggthemes_data$few$colors$Dark[9,2][[1]]
+    #          )+
     # geom_errorbar(aes(ymin=IVW_GSD_mean-1.96*IVW_GSD_emp_se, ymax=IVW_GSD_mean+1.96*IVW_GSD_emp_se), 
@@
                            ))+
-    theme_few() -> IVW_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> IVW_GSD_simplot_thetas
 
@@
     geom_point()+
-    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
-           label= expression( paste("ref: ", theta[X] ))  , 
-           color= ggthemes_data$few$colors$Dark[9,2][[1]]
-             )+
+    geom_hline(yintercept=  args.gsd.method$theta, col=color.dash, alpha=0.6, linetype=2, size=0.8)+
     # geom_errorbar(aes(ymin=MODE_GSD_mean-1.96*MODE_GSD_emp_se, ymax=MODE_GSD_mean+1.96*MODE_GSD_emp_se), 
@@
     ) +
-    theme_few() -> MODE_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> MODE_GSD_simplot_thetas
 
@@
     geom_point()+
-    geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-    annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
-           label= expression( paste("ref: ", theta[X] ))  , 
-           color= ggthemes_data$few$colors$Dark[9,2][[1]]
-             )+
+    geom_hline(yintercept=  args.gsd.method$theta, col=color.dash, alpha=0.6, linetype=2, size=0.8)+
     ## ADD rainbow plots
@@
     ) +
-    theme_few() -> CONMIX_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> CONMIX_GSD_simplot_thetas
 
@@
     ) +
-    theme_few() -> MRMIX_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> MRMIX_GSD_simplot_thetas
 
@@
     ) +
-    theme_few() -> MEDIAN_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> MEDIAN_GSD_simplot_thetas
 
@@
     ) +
-    theme_few() -> EGGER_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> EGGER_GSD_simplot_thetas
   ## 7 ROBUST
@@
     ) +
-    theme_few() -> ROBUST_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> ROBUST_GSD_simplot_thetas
   ## 8 PRESSO
@@
     ) +
-    theme_few() -> PRESSO_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> PRESSO_GSD_simplot_thetas
   ## 9 RAPS
@@
     ) +
-    theme_few() -> RAPS_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> RAPS_GSD_simplot_thetas
   ## 10 LASSO
@@
     ) +
-    theme_few() -> LASSO_GSD_simplot_thetas
+    theme_few(base_size_theta_plot) -> LASSO_GSD_simplot_thetas
 
@@
 
+  arrangedplot.thetas.all[[true_method]] <- GSD_simplot_thetas_arrange
+
   if (SAVE.files ==TRUE ){
@@
     ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
-           width = 10, height = 38, device = "svg",
+           width = 4, height = 14, device = "svg",
            units = "in"  # default
@@
 
+names(arrangedplot.thetas.all) <- selected_true_methods
+
+arrangedthetaplot.fto.mrmix <- ggarrange(arrangedplot.thetas.all$FTO_Wald, arrangedplot.thetas.all$MRMix, ncol=2,
+          labels = c("A", "B"), legend = "none", align = "hv")
+
+if (SAVE.files ==TRUE ){
+    filename_arrangedthetaplot.fto.mrmix <- paste0("./output/plots/theta_sims/", Sys.Date(), "_GSD_arranged_thetas_REF_METHOD_FTO_MRMIX_NSIM_",nsim, ".svg")
+    ggsave(filename = filename_arrangedthetaplot.fto.mrmix, plot = arrangedthetaplot.fto.mrmix, 
+           width = 7, height = 24, 
+           units = "in"  # default
+           )
+  }
 ```
@@
     ###
+    if (selected_true_methods[true_method] == "MRMix") {
+      color.dash <- "black"
+    } else if (selected_true_methods[true_method] == "FTO_Wald"){
+      color.dash <- "blue"
+    } else {
+      color.dash <- ggthemes_data$few$colors$Dark[9,2][[1]]
+    }
+
+    base_size_theta_plot <- 6
+
     IVW_BMI_GSD_geom_point_dat <- est_sim %>% 
@@
       geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
-             label= expression( paste("ref: ", theta[X] ))  , 
-             color= ggthemes_data$few$colors$Dark[9,2][[1]]
-               )+
+      # annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
+      #        label= expression( paste("ref: ", theta[X] ))  , 
+      #        color= ggthemes_data$few$colors$Dark[9,2][[1]]
+      #          )+
       # geom_errorbar(aes(ymin=IVW_GSD_mean-1.96*IVW_GSD_emp_se, ymax=IVW_GSD_mean+1.96*IVW_GSD_emp_se), 
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> IVW_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> IVW_GSD_simplot_thetas
 
@@
       geom_point()+
-      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
-             label= expression( paste("ref: ", theta[X] ))  , 
-             color= ggthemes_data$few$colors$Dark[9,2][[1]]
-               )+
+      geom_hline(yintercept=  args.gsd.method$theta, col=color.dash, alpha=0.6, linetype=2, size=0.8)+
       # geom_errorbar(aes(ymin=MODE_GSD_mean-1.96*MODE_GSD_emp_se, ymax=MODE_GSD_mean+1.96*MODE_GSD_emp_se), 
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> MODE_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> MODE_GSD_simplot_thetas
 
@@
       geom_point()+
-      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
-             label= expression( paste("ref: ", theta[X] ))  , 
-             color= ggthemes_data$few$colors$Dark[9,2][[1]]
-               )+
+      geom_hline(yintercept=  args.gsd.method$theta, col=color.dash, alpha=0.6, linetype=2, size=0.8)+
       ## ADD rainbow plots
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> CONMIX_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> CONMIX_GSD_simplot_thetas
 
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> MRMIX_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> MRMIX_GSD_simplot_thetas
 
@@
       geom_point()+
-      geom_hline(yintercept=  args.gsd.method$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
-      annotate(geom = "text", x = 50000, y = args.gsd.method$theta+0.05,
-             label= expression( paste("ref: ", theta[X] ))  , 
-             color= ggthemes_data$few$colors$Dark[9,2][[1]]
-               )+
-      ## ADD rainbow plots
+      geom_hline(yintercept=  args.gsd.method$theta, col=color.dash, alpha=0.6, linetype=2, size=0.8)+
       geom_boxplot(data = MEDIAN_BMI_GSD_geom_point_dat, aes(x=theta_sim_numeric, y=value, group=theta_sim_numeric), 
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> MEDIAN_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> MEDIAN_GSD_simplot_thetas
 
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> EGGER_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> EGGER_GSD_simplot_thetas
     ## 7 ROBUST
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> ROBUST_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> ROBUST_GSD_simplot_thetas
     ## 8 PRESSO
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> PRESSO_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> PRESSO_GSD_simplot_thetas
     ## 9 RAPS
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> RAPS_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> RAPS_GSD_simplot_thetas
     ## 10 LASSO
@@
                              " estimate calculated with total sample size\nGenetic Architecture: ", name_GenArch))+
-      theme_few() -> LASSO_GSD_simplot_thetas
+      theme_few(base_size = base_size_theta_plot) -> LASSO_GSD_simplot_thetas
 
@@
       ggsave(filename = filename_GSD_simplot_thetas_arrange, plot = GSD_simplot_thetas_arrange, 
-             width = 10, height = 38, 
+             width = 4, height = 14, 
              units = "in"  # default
@@
 # same for all references
+est.wald <- est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "FTO_Wald"]
+est.mrmix <- est.bmi.gsd$estimate[est.bmi.gsd$mr_methods == "MRMix"]
+
 df.temp.long[[1]] %>% 
@@
   mutate(N = paste0("Sample Size = ", format(N, scientific=FALSE))) %>% 
+  group_by(N) %>% 
+  ## add columns for colouring closest to FTO
+  mutate(closest_to_FTO_boolean_all = abs(Mean_ALL_INSTR - est.wald) == min(abs(Mean_ALL_INSTR - est.wald)),
+         closest_to_FTO_boolean_strongcommon = abs(Mean_STRONG_COMMON - est.wald) == min(abs(Mean_STRONG_COMMON - est.wald)),
+         closest_to_FTO_boolean_strongrare = abs(Mean_STRONG_RARE - est.wald) == min(abs(Mean_STRONG_RARE - est.wald)),
+         closest_to_FTO_boolean_weakcommon = abs(Mean_WEAK_COMMON - est.wald) == min(abs(Mean_WEAK_COMMON - est.wald)),
+         closest_to_FTO_boolean_weakrare = abs(Mean_WEAK_RARE - est.wald) == min(abs(Mean_WEAK_RARE - est.wald))
+         ) %>% 
+  ## add columns for colouring closest to MRMix
+  mutate(closest_to_MRMIX_boolean_all = abs(Mean_ALL_INSTR - est.mrmix) == min(abs(Mean_ALL_INSTR - est.mrmix)),
+         closest_to_MRMIX_boolean_strongcommon = abs(Mean_STRONG_COMMON - est.mrmix) == min(abs(Mean_STRONG_COMMON - est.mrmix)),
+         closest_to_MRMIX_boolean_strongrare = abs(Mean_STRONG_RARE - est.mrmix) == min(abs(Mean_STRONG_RARE - est.mrmix)),
+         closest_to_MRMIX_boolean_weakcommon = abs(Mean_WEAK_COMMON - est.mrmix) == min(abs(Mean_WEAK_COMMON - est.mrmix)),
+         closest_to_MRMIX_boolean_weakrare = abs(Mean_WEAK_RARE - est.mrmix) == min(abs(Mean_WEAK_RARE - est.mrmix))
+  ) %>% 
   gt(groupname_col="N") %>% 
@@
   ) %>% 
+  ## colour closest values ----------------------------------------------
+  # FTO
+  tab_style(
+    style = cell_text(color = "blue", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_ALL_INSTR"),
+      rows = closest_to_FTO_boolean_all
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(color = "blue", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_STRONG_COMMON"),
+      rows = closest_to_FTO_boolean_strongcommon
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(color = "blue", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_STRONG_RARE"),
+      rows = closest_to_FTO_boolean_strongrare
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(color = "blue", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_WEAK_COMMON"),
+      rows = closest_to_FTO_boolean_weakcommon
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(color = "blue", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_WEAK_RARE"),
+      rows = closest_to_FTO_boolean_weakrare
+    )
+  ) %>% 
+  # MRMix -----
+  tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_ALL_INSTR"),
+      rows = closest_to_MRMIX_boolean_all
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_STRONG_COMMON"),
+      rows = closest_to_MRMIX_boolean_strongcommon
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_STRONG_RARE"),
+      rows = closest_to_MRMIX_boolean_strongrare
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_WEAK_COMMON"),
+      rows = closest_to_MRMIX_boolean_weakcommon
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(color = "black", weight = "bold"),
+    locations = cells_body(
+      columns = contains("Mean_WEAK_RARE"),
+      rows = closest_to_MRMIX_boolean_weakrare
+    )
+  ) %>% 
+  #--------
   cols_merge(
@@
   fmt_number(columns = 3:12,decimals = 3, use_seps=TRUE) %>% 
+  cols_hide(columns = c(closest_to_FTO_boolean_all, closest_to_FTO_boolean_strongcommon, closest_to_FTO_boolean_strongrare, closest_to_FTO_boolean_weakcommon, closest_to_FTO_boolean_weakrare,
+                        closest_to_MRMIX_boolean_all, closest_to_MRMIX_boolean_strongcommon, closest_to_MRMIX_boolean_strongrare, closest_to_MRMIX_boolean_weakcommon, closest_to_MRMIX_boolean_weakrare)) %>%
   tab_source_note(
@@
   # TABLE 02: BIAS, MSE, COVERAGE, TYPE1ERROR ---------------------------------
+  if(selected_true_methods[ref.method] == "FTO_Wald") {
+    highlightcol <- "blue"
+  } else if (selected_true_methods[ref.method] == "MRMix") {
+    highlightcol <- "black"
+  } else {highlightcol <- ggthemes_data$few$colors$Dark[9,2][[1]]}
+
   df.temp.long[[ref.method]] %>%
@@
     mutate(N = paste0("Sample Size = ", format(N, scientific=FALSE))) %>% 
+    group_by(N) %>% 
+    ## add columns for colouring bias
+    mutate(minbias_all = abs(bias_ALL_INSTR) ==  min(abs(bias_ALL_INSTR)),
+           minbias_strongcommon = abs(bias_STRONG_COMMON) == min(abs(bias_STRONG_COMMON)),
+           minbias_strongrare = abs(bias_STRONG_RARE) == min(abs(bias_STRONG_RARE)),
+           minbias_weakcommon = abs(bias_WEAK_COMMON) == min(abs(bias_WEAK_COMMON)),
+           minbias_weakrare = abs(bias_WEAK_RARE) == min(abs(bias_WEAK_RARE)),
+           # cols for mse
+           minmse_all = abs(mse_ALL_INSTR) == min(abs(mse_ALL_INSTR)),
+           minmse_strongcommon = abs(mse_STRONG_COMMON) == min(abs(mse_STRONG_COMMON)),
+           minmse_strongrare = abs(mse_STRONG_RARE) == min(abs(mse_STRONG_RARE)),
+           minmse_weakcommon = abs(mse_WEAK_COMMON) == min(abs(mse_WEAK_COMMON)),
+           minmse_weakrare = abs(mse_WEAK_RARE) == min(abs(mse_WEAK_RARE)),
+           # cols for coverage
+           mincoverage_all = coverage_ALL_INSTR == max(coverage_ALL_INSTR),
+           mincoverage_strongcommon = coverage_STRONG_COMMON == max(coverage_STRONG_COMMON),
+           mincoverage_strongrare = coverage_STRONG_RARE == max(coverage_STRONG_RARE),
+           mincoverage_weakcommon = coverage_WEAK_COMMON == max(coverage_WEAK_COMMON),
+           mincoverage_weakrare = coverage_WEAK_RARE == max(coverage_WEAK_RARE),
+           # cols for type1error
+           mintype1error_all = type1error_ALL_INSTR == max(type1error_ALL_INSTR),
+           mintype1error_strongcommon = type1error_STRONG_COMMON == max(type1error_STRONG_COMMON),
+           mintype1error_strongrare = type1error_STRONG_RARE == max(type1error_STRONG_RARE),
+           mintype1error_weakcommon = type1error_WEAK_COMMON == max(type1error_WEAK_COMMON),
+           mintype1error_weakrare = type1error_WEAK_RARE == max(type1error_WEAK_RARE),
+    ) -> gtperf.dat
+    # make {gt} table
+  gtperf.dat %>% 
     gt(groupname_col="N") %>% 
@@
     ) %>% 
+    ## colour best values ----------------------------------------------
+    # bias
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("bias_ALL_INSTR"),
+        rows = minbias_all
+      )
+    ) %>% 
+      tab_style(
+        style = cell_text(color = highlightcol, weight = "bold"),
+        locations = cells_body(
+          columns = contains("bias_STRONG_COMMON"),
+          rows = minbias_strongcommon
+        )
+      ) %>% 
+      tab_style(
+        style = cell_text(color = highlightcol, weight = "bold"),
+        locations = cells_body(
+          columns = contains("bias_STRONG_RARE"),
+          rows = minbias_strongrare
+        )
+      ) %>% 
+      tab_style(
+        style = cell_text(color = highlightcol, weight = "bold"),
+        locations = cells_body(
+          columns = contains("bias_WEAK_COMMON"),
+          rows = minbias_weakcommon
+        )
+      ) %>% 
+      tab_style(
+        style = cell_text(color = highlightcol, weight = "bold"),
+        locations = cells_body(
+          columns = contains("bias_WEAK_RARE"),
+          rows = minbias_weakrare
+        )
+      ) %>% 
+    # mse
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("mse_ALL_INSTR"),
+        rows = minmse_all
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("mse_STRONG_COMMON"),
+        rows = minmse_strongcommon
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("mse_STRONG_RARE"),
+        rows = minmse_strongrare
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("mse_WEAK_COMMON"),
+        rows = minmse_weakcommon
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("mse_WEAK_RARE"),
+        rows = minmse_weakrare
+      )
+    ) %>% 
+    # coverage
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("coverage_ALL_INSTR"),
+        rows = mincoverage_all
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("coverage_STRONG_COMMON"),
+        rows = mincoverage_strongcommon
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("coverage_STRONG_RARE"),
+        rows = mincoverage_strongrare
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("coverage_WEAK_COMMON"),
+        rows = mincoverage_weakcommon
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("coverage_WEAK_RARE"),
+        rows = mincoverage_weakrare
+      )
+    ) %>% 
+    # type1error
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("type1error_ALL_INSTR"),
+        rows = mintype1error_all
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("type1error_STRONG_COMMON"),
+        rows = mintype1error_strongcommon
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("type1error_STRONG_RARE"),
+        rows = mintype1error_strongrare
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("type1error_WEAK_COMMON"),
+        rows = mintype1error_weakcommon
+      )
+    ) %>% 
+    tab_style(
+      style = cell_text(color = highlightcol, weight = "bold"),
+      locations = cells_body(
+        columns = contains("type1error_WEAK_RARE"),
+        rows = mintype1error_weakrare
+      )
+    ) %>% 
+    # --------------------------------------------------------------------------
+    cols_hide(columns = colnames(gtperf.dat)[23:42]) %>% 
     fmt_number(columns = 3:22,decimals = 3, use_seps=TRUE) %>% 
@@
 
+  if (names(df_metrics_all_scenarios)[ref.method] == "ref_MRMix") {
+    color.bar <- "black"
+  } else if (names(df_metrics_all_scenarios)[ref.method] == "ref_FTO_Wald"){
+    color.bar <- "blue"
+  } else {
+    color.bar <- ggthemes_data$few$colors$Dark[9,2][[1]]
+  }
+
   ## MSE PLOT FACET WRAP ------------------------------------------------------
@@
       drop_na(Mean) %>%  
+      group_by(N) %>% 
+      mutate(smallest_mse = mse == min(mse)) %>% # add a new column to identify the rows with the smallest mse value within each group
+      ungroup() %>% 
+      # PLOT
       ggplot(aes(x=Method, y=mse)) +
-      geom_bar(stat="identity", position=position_dodge2(reverse=TRUE),
+      geom_bar(
+        aes(fill = smallest_mse), # add fill aesthetics to highlight the bars with the smallest mse value within each group
+        stat="identity", position=position_dodge2(reverse=TRUE),
         #color=ggthemes_data$few$colors$Dark[2,2][[1]],

### 2023-05-09 scripts/MSc.Rmd
highlighted best methods in barplots

@@
       ylab(paste0("Mean Squared Error with nsim=", nsim))+
-      labs(title = paste0("Selected Reference Method: ", 
-                              selected_true_methods[ref.method]),
-           subtitle = paste0("Genetic Architecture: ", scenarios[scenario])
+      labs(#title = paste0("Selected Reference Method: ", 
+          #                selected_true_methods[ref.method]),
+           #subtitle = paste0("Genetic Architecture: ", scenarios[scenario])
       )+
@@
       ggthemes::theme_few()+  # remove legend
-      ggthemes::scale_fill_few() -> dodged.barplot.mse.scenario[[scenario]]
+      scale_fill_manual(values = c("gray80", "blue"), guide = "none") -> dodged.barplot.mse.scenario[[scenario]]
 
@@
   }
-  dodged.barplot.mse.references[ref.method] <- dodged.barplot.mse.scenario[1]
-
-  # scenarios without all instruments in arranged plot
+  dodged.barplot.mse.references[ref.method] <- dodged.barplot.mse.scenario[1]   # save scenario 1 "ALL INSTR"
+
+  # scenarios WITHOUT all instruments in arranged plot
   arrangedplot.mse.scenarios <- 

### 2023-05-09 scripts/MSc.Rmd
highlighted best methods in barplots 2

@@
 p1.gsd+
+  xlab("SNP effect on BMI")+
+  ylab("SNP effect on GSD")+
   ggthemes::theme_few() -> p1.gsd
@@
       ggthemes::theme_few()+  # remove legend
-      scale_fill_manual(values = c("gray80", "blue"), guide = "none") -> dodged.barplot.mse.scenario[[scenario]]
+      scale_fill_manual(values = c("gray80", color.bar), guide = "none") -> dodged.barplot.mse.scenario[[scenario]]
 
@@
   # scenarios WITHOUT all instruments in arranged plot
+
+  # make smaller plots
+  base_theme <- theme_few(base_size = 8)
+  for (smallplot in seq_along(dodged.barplot.mse.scenario)) {
+    dodged.barplot.mse.scenario[[smallplot]] <- dodged.barplot.mse.scenario[[smallplot]] +
+      base_theme
+  }
+
   arrangedplot.mse.scenarios <- 
-    ggarrange(plotlist = dodged.barplot.mse.scenario[-1], labels = c("A", "B", "C", "D", "E", "F"))
+    ggarrange(plotlist = dodged.barplot.mse.scenario[-1], labels = c("Scenario 2", "Scenario 3", "Scenario 4", "Scenario 5"))
   plot(arrangedplot.mse.scenarios)
@@
       drop_na(Mean) %>%  
+      group_by(N) %>% 
+      mutate(smallest_bias = abs(bias) == min(abs(bias))) %>% # add a new column to identify the rows with the smallest bias value
+      ungroup() %>% 
+      # PLOT
       ggplot(aes(x=Method, y=bias)) +
-      geom_bar(stat="identity", position=position_dodge2(reverse=TRUE),
+      geom_bar(
+        aes(fill = smallest_bias), # add fill aesthetics to highlight the bars with the smallest mse value within each group
+        stat="identity", position=position_dodge2(reverse=TRUE),
         #color=ggthemes_data$few$colors$Dark[2,2][[1]],
@@
       ylab(paste0("Bias with nsim=", nsim))+
-      labs(title = paste0("Selected Reference Method: ", 
-                              selected_true_methods[ref.method]),
-           subtitle = paste0("Genetic Architecture: ", scenarios[scenario])
+      labs(#title = paste0("Selected Reference Method: ", 
+           #               selected_true_methods[ref.method]),
+           #subtitle = paste0("Genetic Architecture: ", scenarios[scenario])
       )+
@@
       ggthemes::theme_few()+  # remove legend
-      ggthemes::scale_fill_few() -> dodged.barplot.bias.scenario[[scenario]]
+      scale_fill_manual(values = c("gray80", color.bar), guide = "none") -> dodged.barplot.bias.scenario[[scenario]]
 
@@
    # scenarios without all instruments in arranged plot
+
+  # make smaller plots
+  base_theme <- theme_few(base_size = 8)
+  for (smallplot in seq_along(dodged.barplot.bias.scenario)) {
+    dodged.barplot.bias.scenario[[smallplot]] <- dodged.barplot.bias.scenario[[smallplot]] +
+      base_theme
+  }
+
   arrangedplot.bias.scenarios <- 
-    ggarrange(plotlist = dodged.barplot.bias.scenario[-1], labels = c("A", "B", "C", "D", "E", "F"))
+    ggarrange(plotlist = dodged.barplot.bias.scenario[-1], labels = c("Scenario 2", "Scenario 3", "Scenario 4", "Scenario 5"))
 
@@
 
+## all methods compared 
 arrangedplot.bias.references <- 
@@
 
+# Figure 10/11 MSE and BIAS in Scenario 1 with ref estimate FTO Wald
+names(dodged.barplot.mse.references) <- names(df_metrics_all_scenarios)
+names(dodged.barplot.bias.references) <- names(df_metrics_all_scenarios)
+arranged.msebias.scenario1.fto <- ggarrange(dodged.barplot.mse.references$ref_FTO_Wald, dodged.barplot.bias.references$ref_FTO_Wald, ncol=2, labels = c("A", "B"))
+arranged.msebias.scenario1.mrmix <- ggarrange(dodged.barplot.mse.references$ref_MRMix, dodged.barplot.bias.references$ref_MRMix, ncol=2, labels = c("A", "B"))
+
+if (SAVE.files ==TRUE ){
+    ## arranged plots are toooo small 
+    filename.arranged.msebias.scenario1.fto <- paste0("./output/plots/summary-plots/figure10/", Sys.Date(), "_bias_mse_barplot_REF_METHOD_FTO_WALD_NSIM_", nsim)
+    filename.arranged.msebias.scenario1.mrmix <- paste0("./output/plots/summary-plots/figure10/", Sys.Date(), "_bias_mse_barplot_REF_METHOD_MRMIX_NSIM_", nsim)
+
+    ggsave(filename = paste0(filename.arranged.msebias.scenario1.fto, ".svg"), plot =  arranged.msebias.scenario1.fto, 
+             width = 8, height = 7, 
+             units = "in"  # default
+             )
+    ggsave(filename = paste0(filename.arranged.msebias.scenario1.mrmix, ".svg"), plot =  arranged.msebias.scenario1.mrmix, 
+             width = 8, height = 7, 
+             units = "in"  # default
+             )
+}
+
 ```
